### 2020-04-18 main.tex
initial commit

\documentclass[twocolumn]{article}

\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage[backend=biber]{biblatex}
\usepackage{hyperref}

\renewcommand{\baselinestretch}{1.2}

\bibliography{literature.bib}
\title{Forschungsprojekt}
\author{Simon König (3344789), Leon Matzner (3315161),\\ Felix Rollbühler (3310069), Jakob Schmid (??????)}
\date{\today}

\begin{document}
\maketitle

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\nocite{*}

\printbibliography

\end{document}

### 2020-06-09 scripts/Galois_setup.sh
slight changes to Galois_setup.sh

@@ -5,13 +5,20 @@
 # Arguments (supplied arguments must be ordered according to this listing):
 # nd -> doesn't set up the D-Galois applications
 # od -> installs only the D-Galois applications
+# dep -> installs only the dependencies
 # nj -> disables the -j flag in the building process. This option can be necessary if RAM is very limited. 
 
-sudo apt-get install g++
-sudo apt-get install cmake
-sudo apt-get install libboost-all-dev
-sudo apt-get install llvm
-sudo apt-get install libnuma-dev
+sudo apt-get install g++ cmake libboost-all-dev llvm libnuma-dev
+
+if [[ $1 != "nd" ]];
+then
+    sudo apt-get install libopenmpi-dev
+fi
+
+if [[ $1 == "dep" || $2 == "dep" ]];
+then
+    exit 0
+fi
 
 # required for both D-Galois and regular Galois
 mkdir Galois Galois/src Galois/build Galois/bin
@@ -46,7 +53,6 @@ fi
 # for D-Galois
 if [[ $1 != "nd" ]];
 then
-    sudo apt-get install libopenmpi-dev
     cmake build -DGALOIS_ENABLE_DIST=1
     $make_build"sssp_pull"
     cp build/lonestardist/sssp/sssp_pull bin/d-galois-sssp-pull

### 2020-06-12 ganalyze.c
ganalyze

@@ -1,0 +1,74 @@
+#include <stdio.h>
+#include <string.h>
+#include <stdbool.h>
+
+int main(int argc, char *argv[]) {
+
+  char c;
+  int src, dest, weight, lsrc = 0, ldest = 0, edges = 0, hsrc = 0, hdest = 0, hnode;
+  bool weighted = false, src_sorted = true, dest_sorted = true, src_gaps = false;
+
+  // skip commentaries
+  while ((c = getc(stdin)) == '#')
+    while ((c = getc(stdin)) != '\n');
+  fseek(stdin, -1, SEEK_CUR);
+
+  // is graph weighted
+  weighted = argc > 1 && !strcmp(argv[1], "weighted");
+
+  // analyze
+  while (true) {
+    if (weighted) {
+      if (scanf("%d %d %d\n", &src, &dest, &weight) == EOF)
+        break;
+    } else {
+      if (scanf("%d %d\n", &src, &dest) == EOF)
+        break;
+    }
+
+    if (lsrc > src) { // are the source nodes sorted
+      printf("src %d after %d\n", src, lsrc);
+      src_sorted = false;
+    } else if (lsrc == src && ldest > dest) { // are the destination nodes sorted
+      printf("dest %d after %d\n", dest, ldest);
+      dest_sorted = false;
+    } else {
+      if (src > lsrc+1) { // found gap
+        printf("gap %d to %d\n", lsrc, src);
+        src_gaps = true;
+      }
+      if (src > hsrc) { // next biggest src
+        hsrc = src;
+      }
+    }
+    if (dest > hdest) { // next biggest dest
+      hdest = dest;
+    }
+    lsrc = src;
+    ldest = dest;
+    ++edges;
+  }
+
+  // print stats
+  fprintf(stderr, "nodes: %d\n", hsrc > hdest ? hsrc: hdest);
+  fprintf(stderr, "edges: %d\n", edges);
+  fprintf(stderr, "hsrc: %d\n", hsrc);
+  fprintf(stderr, "hdest: %d\n", hdest);
+  if (src_sorted) {
+    fprintf(stderr, "src_sorted: true\n");
+  } else {
+    fprintf(stderr, "src_sorted: false\n");
+  }
+  if (dest_sorted) {
+    fprintf(stderr, "dest_sorted: true\n");
+  } else {
+    fprintf(stderr, "dest_sorted: false\n");
+  }
+  if (src_gaps) {
[…]

### 2020-06-16 documentation/paper/main.tex
dings
ieee template mit bib-Support.. wenn auch hacky

--- 2020-04-18 main.tex
@@
-\documentclass[twocolumn]{article}
+\documentclass[conference,a4paper]{IEEEtran}
 
+\usepackage[noadjust]{cite}
 \usepackage[utf8]{inputenc}
-\usepackage{csquotes}
-\usepackage[backend=biber]{biblatex}
+%\usepackage{csquotes}
 \usepackage{hyperref}
-
-\renewcommand{\baselinestretch}{1.2}
 
@@
 \title{Forschungsprojekt}
-\author{Simon König (3344789), Leon Matzner (3315161),\\ Felix Rollbühler (3310069), Jakob Schmid (??????)}
+\author{Simon König (3344789) \\ \and Leon Matzner (3315161) \\ \and Felix Rollbühler (3310069) \\ \and Jakob Schmid (??????)}
 \date{\today}
@@
 \begin{document}
+
 \maketitle
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\begin{abstract}
+In this paper we will analyze and compare the graph frameworks Galois, Ligra and Polymer in their parformance. We will also set up and compare these frameworks in a distributed scenario.
+\end{abstract}
 
-\nocite{*}
+\begin{IEEEkeywords}
+Galois, Ligra, Polymer, distributed computing, Gluon
+\end{IEEEkeywords}
 
-\printbibliography
+\section{Introduction}
+My intro... blah blah \cite{Polymer}.
+
+\section{Things}
+\input{galois.tex}
+
+\input{polymer.tex}
+
+\section{Conclusion}
+The conclusion goes here.
+
+\section*{Acknowledgment}
+We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Lira} and Polymer~\cite{Polymer}.
+
+Also we use Gluon~\cite{vertGalois} for the distributed setups.
+
+% Das hier ist mega beschissen: 
+% Beim Hinzufügen neuer Referenzen
+% 1. die beiden Zeilen nach diesem Kommentar auskommentieren.
+%    (gemeint ist \bibliographystyle... und \bibliography..)
+%    Dann alle cache files im Ordner löschen.
+% 2. Die folgenden Commands ausführen
+% pdflatex main.tex
+% bibtex main
+% pdflatex main.tex 
+%    das letzte wirft dann Fehler aber das stimmt so..
+% 3. Jetzt main.bbl öffnen und den Inhalt hier drunter rein kopieren.
+%    Man kann kein \input benutzen, weil main.bbl leer ist, wenn die beiden
+%    Zeilen auskommentiert sind. Aber wenn sie nicht auskommentiert sind
+%    kann kein PDF erzeugt werden.. Was zur Hölle denken die sich.
+% 4. Die folgenden beiden Zeilen auskommentieren und pdflatex 
+%    ausführen wie gewohnt.
+
+% \bibliographystyle{IEEEtran}
+% \bibliography{literature}
+
+% Alles ab hier kommt aus main.bbl
+
+% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
+\begin{thebibliography}{1}
+\providecommand{\url}[1]{#1}
+\csname url@samestyle\endcsname
+\providecommand{\newblock}{\relax}
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\BIBentrySTDinterwordspacing}{\spaceskip=0pt\relax}
+\providecommand{\BIBentryALTinterwordstretchfactor}{4}
+\providecommand{\BIBentryALTinterwordspacing}{\spaceskip=\fontdimen2\font plus
+\BIBentryALTinterwordstretchfactor\fontdimen3\font minus
+  \fontdimen4\font\relax}
+\providecommand{\BIBforeignlanguage}[2]{{%
+\expandafter\ifx\csname l@#1\endcsname\relax
+\typeout{** WARNING: IEEEtran.bst: No hyphenation pattern has been}%
+\typeout{** loaded for the language `#1'. Using the pattern for}%
+\typeout{** the default language instead.}%
+\else
+\language=\csname l@#1\endcsname
+\fi
+#2}}
+\providecommand{\BIBdecl}{\relax}
+\BIBdecl
+
+\bibitem{Polymer}
+\BIBentryALTinterwordspacing
+K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
+  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
+  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
+  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
+  2015, p. 183–193. [Online]. Available:
+  \url{https://doi.org/10.1145/2688500.2688507}
+\BIBentrySTDinterwordspacing
+
+\bibitem{Galois}
+\BIBentryALTinterwordspacing
+D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
+  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
+  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
+  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
+  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
+\BIBentrySTDinterwordspacing
+
+\bibitem{Lira}
+\BIBentryALTinterwordspacing
+J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
+  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
+  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
+  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
+  Machinery, 2013, p. 135–146. [Online]. Available:
+  \url{https://doi.org/10.1145/2442516.2442530}
+\BIBentrySTDinterwordspacing
+
+\bibitem{vertGalois}
+\BIBentryALTinterwordspacing
+R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
+  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
+  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
+  Conference on Programming Language Design and Implementation}, ser. PLDI
+  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
+  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
+  \url{https://doi.org/10.1145/3192366.3192404}
+\BIBentrySTDinterwordspacing
+
+\end{thebibliography}
+
+% Alles bis hier kommt aus main.bbl
 

### 2020-06-22 sortEdgeList.cpp
Binary edge list support

@@ -21,7 +21,7 @@ struct edge {
 struct edgeList {
     edge * edges;
     unsigned int length;
-    types type; 
+    types type;
 };
 
 inline void skipWhitespaces(char * line, size_t * i, size_t len) {
@@ -84,7 +84,7 @@ edgeList readEdgeList(char* filename, int isWeighted) {
     while (fgets(line, lineLength, f) != NULL) {
         i = 0;
         len = strlen(line);
-        
+
         skipWhitespaces(line, &i, len);
         // ignore if not a number
         if (i < len && !(line[i] >= '0' && line[i] <= '9')) {
@@ -93,14 +93,14 @@ edgeList readEdgeList(char* filename, int isWeighted) {
         graph.edges[graph.length].from = parseInteger(line, &i, len);
         skipWhitespaces(line, &i, len);
         graph.edges[graph.length].to = parseInteger(line, &i, len);
-        
+
         if (isWeighted) {
             skipWhitespaces(line, &i, len);
             graph.edges[graph.length].weigth = parseInteger(line, &i, len);
         }
         graph.length++;
     }
-    
+
     fclose(f);
     return graph;
 }
@@ -122,6 +122,32 @@ void writeEdgeList(char * filename, edgeList graph) {
     else {
         for (size_t i = 0; i < graph.length; i++) {
             fprintf(f, "%d %d\n", graph.edges[i].from, graph.edges[i].to);
+        }
+    }
+    fclose(f);
+}
+
+void writeEdgeListBin(char * filename, edgeList graph) {
+    FILE* f;
+    f = fopen(filename, "wb");
+    if (f == NULL) {
+        perror("Cannot open file\n");
+        exit(EXIT_FAILURE);
+    }
+
+    if(graph.type == WEIGHTED) {
+        for (int i = 0; i < graph.length; i++) {
+            fwrite(&graph.edges[i].from, sizeof(int), 1, f);
+            fwrite(&graph.edges[i].to, sizeof(int), 1, f);
+            float weight = (float) graph.edges[i].weigth;
+            fwrite(&weight, sizeof(float), 1, f);
+        }
+    }
+    else {
+        for (int i = 0; i < graph.length; i++)
+        {
+            fwrite(&graph.edges[i].from, sizeof(int), 1, f);
+            fwrite(&graph.edges[i].to, sizeof(int), 1, f);
         }
     }
     fclose(f);
[…]

### 2020-06-22 documentation/paper/polymer.tex

%!TEX root=main.tex

\subsection{Polymer}
Installing Polymer is straight forward.

### 2020-06-24 documentation/paper/main.tex

@@
 \section*{Acknowledgment}
-We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Lira} and Polymer~\cite{Polymer}.
+We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Ligra} and Polymer~\cite{Polymer}.
 
 Also we use Gluon~\cite{vertGalois} for the distributed setups.
+
+Gemini~\cite{Gemini}
 
@@
 
+% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
 % Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
@@
 
-\bibitem{Lira}
+\bibitem{Ligra}
 \BIBentryALTinterwordspacing
@@
 
+\bibitem{Gemini}
+\BIBentryALTinterwordspacing
+X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
+  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
+  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
+  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
+  301--316. [Online]. Available:
+  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
+\BIBentrySTDinterwordspacing
+
 \end{thebibliography}

### 2020-06-29 documentation/paper/graphFormats.tex
doc

%!TEX root = ./main.tex

\section{Graph Formats}
A rather big portion of our time was invested in figuring out which graph framework requires which graph formats. We thus decided to give an overview over all the formats we encountered, with explanation on how they represent the graph.

Additionally, to make life in the future a little bit easier, we wrote multiple tools to convert graphs acquired from Snap or Konect to the required formats. Information on how to use those tools is available in our repository.

\subsection{AdjacencyList}
The AdjacencyList and WeightedAdjacencyList formats\cite{adjListFormat} are used by Ligra and Polymer. They represent the directed edges of a graph as a number of offsets that point to a set of target nodes in the file.
First the file contains the number of vertices $n$ and edges $m$, followed by an offset for each vertex. This offset specifies at what point in the following list of numbers the information for a node begins.
Lastly the file format contains a list of target nodes.
The numbers are all separated by newlines.
\begin{gather*}
n\\
m\\
o_1\\
o_2\\
\vdots\\
o_n\\
t_1\\
t_2\\
\vdots\\
t_m
\end{gather*}
The offsets $o_i=k$ and $o_{i+1}=k+j$ mean that vertex $i$ has $j$ outgoing edges, these edges are
\begin{equation*}
	(i,t_k),(i,t_{k+1}),\ldots,(i,t_{k+j-1})
\end{equation*}

For the WeightedAdjacencyList format, the weights are just appended to the end of the file in the same order as the edges.

\subsection{EdgeList}
The EdgeList format is probably the easiest to understand and is one of the most commonly used in the online graph repositories. The directed eges $(s_1,t_1),(s_2,t_2),\ldots$ are represented in the following way.
\begin{align*}
	&s_1,t_1\\
	&s_2,t_2\\
	&\quad\vdots\\
	&s_m,t_m
\end{align*}

\subsection{Binary EdgeList}
The binary EdgeList format is used by Gemini. Finding information on this format required reverse engineering of the Gemini code.

We found that Gemini requires the following input format
\begin{equation*}
	s_1t_1w_1s_2t_2w_2\ldots
\end{equation*}
where $s_i,t_i$ have \texttt{uint32} ?? data type and the weights are \texttt{float32}. 
Gemini will derive the number of edges from the file size, so there is no file header or anything similar allowed.

\subsection{?}

### 2020-06-29 documentation/paper/main.tex
doc

@@
 \usepackage{hyperref}
+\usepackage{amsmath}
 
@@
 \title{Forschungsprojekt}
-\author{Simon König (3344789) \\ \and Leon Matzner (3315161) \\ \and Felix Rollbühler (3310069) \\ \and Jakob Schmid (??????)}
+\author{Simon König (3344789) \and Leon Matzner (3315161) \and Felix Rollbühler (3310069) \and Jakob Schmid (??????)}
 \date{\today}
@@
 
+\input{graphFormats}
 \section{Things}
-\input{galois.tex}
+\input{galois}
 
-\input{polymer.tex}
+\input{polymer}
 
@@
 
+\bibitem{adjListFormat}
+\BIBentryALTinterwordspacing
+J.~Shun, G.~Blelloch, J.~Fineman, P.~Gibbons, A.~Kyrola, K.~Tangwonsan, and
+  H.~V. Simhadri. (2020, Jun.) {Problem Based Benchmark Suite}. graphIO.html.
+  [Online]. Available: \url{http://www.cs.cmu.edu/~pbbs/benchmarks/}
+\BIBentrySTDinterwordspacing
+
 \bibitem{Galois}

### 2020-06-30 scripts/benchmark.sh
improved benchmark.sh

@@ -3,7 +3,7 @@
 # This script requires a correctly set up /etc/hosts file and the moreutils package
 # This script expects two parameter and one files as input.
 # The first parameter is the path to a directory containing all of the graphs. 
-# The directory mustn't contain anything but the graphs in every repuired format (.gr, .adj, .bin) including weighted versions (the filenames for weighted graphs require a leading "w_").
+# The directory mustn't contain anything but the graphs in every repuired format (.gr, .adj, .bin) including weighted versions (the filenames for unweighted graphs require a leading "u_").
 # The second parameter is the path to a directory containing only the binaries (the names of the binaries must be the original name with the name of the framework und a minus prepended). 
 # The file will contain the results of the benchmarks.
 
@@ -12,7 +12,7 @@
 graphs=$(ls $1)
 path_to_bins=$2
 result_file=$3
-page_rank_number_of_iterations=1000
+page_rank_number_of_iterations=5
 echo "# benchmark results (elapsed time in seconds )" > $result_file
 
 for i in $graphs; do
@@ -35,30 +35,55 @@ for i in $graphs; do
         echo "Galois-pagerank-push on $i" >> $result_file
         echo $(( $finished - $started )) >> $result_file
         # pagerank pull
-        # TODO transposed graph
+        bin="${path_to_bins}/galois-pagerank-pull-cpu"
+        result=$(timeout 3h ./$bin --transposedGraph $i | ts '%s')
+        finished=$(grep 'STAT_TYPE' $result | awk '{print $1}')
+        started=$(grep 'Read ' $result | awk '{print $1}')
+        echo "Galois-pagerank-pull on $i" >> $result_file
+        echo $(( $finished - $started )) >> $result_file
         # distributed
     fi
     # polymer and ligra
     if [[ $i == *.adj ]];
     then
-        # shared memory
-        # polymer
         # sssp
-        bin="${path_to_bins}/numa-BellmandFord"
-        result=$(timeout 3h ./$bin $i 0 | ts '%s')
-        # TODO stop time
+        if [[ $i != +(u_) ]];
+        then
+            # polymer sssp
+            bin="${path_to_bins}/polymer-numa-BellmandFord"
+            result=$(timeout 3h ./$bin $i 0)
+            time=$(grep 'BellmanFord' $result)
+            time="${time//BellmanFord : }"
+            time="${time//.*}"
[…]

### 2020-07-01 scripts/benchmark.sh
rewrote benchmark script

@@ -1,93 +1,100 @@
 #!/bin/bash
 
-# This script requires a correctly set up /etc/hosts file and the moreutils package
-# This script expects two parameter and one files as input.
+# This script requires a correctly set up /etc/hosts file and the moreutils package (needed for ts)
+# This script expects two parameter and two files as input.
 # The first parameter is the path to a directory containing all of the graphs. 
 # The directory mustn't contain anything but the graphs in every repuired format (.gr, .adj, .bin) including weighted versions (the filenames for unweighted graphs require a leading "u_").
-# The second parameter is the path to a directory containing only the binaries (the names of the binaries must be the original name with the name of the framework und a minus prepended). 
-# The file will contain the results of the benchmarks.
+# The second parameter is a file containing the base names of all graphs, the respective number of nodes, edges and the biggest usable sssp sourcenode delimited by spaces
+# The third parameter is the path to a directory containing only the binaries (the names of the binaries must be the original name with the name of the framework und a minus prepended). 
+# The final mandatory parameter is file which will contain the results of the benchmarks (previous content will be overwritten).
+# Optionally a log file can be supplied
 
-# TODO decide if unweighted graphs are to be supported
+galois-sssp () {
+    bin="${path_to_bins}/galois-sssp-cpu"
+    result=$(timeout 3h ./$bin $i | ts '%s')
+    finished=$(grep 'Verification' $result | awk '{print $1}')
+    started=$(grep 'Read ' $result | awk '{print $1}')
+    echo "Galois-sssp on $i" >> $result_file
+    echo $(( $finished - $started )) >> $result_file
+}
+galois-pagerank-push () {
+    # pagerank push
+    bin="${path_to_bins}/galois-pagerank-push-cpu"
+    result=$(timeout 3h ./$bin $i | ts '%s')
+    finished=$(grep 'STAT_TYPE' $result | awk '{print $1}')
+    started=$(grep 'Read ' $result | awk '{print $1}')
+    echo "Galois-pagerank-push on $i" >> $result_file
+    echo $(( $finished - $started )) >> $result_file
+}
[…]

### 2020-07-01 scripts/benchmark.sh
benchmark galois

@@ -8,32 +8,56 @@
 # The third parameter is the path to a directory containing only the binaries (the names of the binaries must be the original name with the name of the framework und a minus prepended). 
 # The final mandatory parameter is file which will contain the results of the benchmarks (previous content will be overwritten).
 # Optionally a log file can be supplied
+log() { echo "$@" 1>&2; }
 
 galois-sssp () {
-    bin="${path_to_bins}/galois-sssp-cpu"
-    result=$(timeout 3h ./$bin $i | ts '%s')
-    finished=$(grep 'Verification' $result | awk '{print $1}')
-    started=$(grep 'Read ' $result | awk '{print $1}')
-    echo "Galois-sssp on $i" >> $result_file
-    echo $(( $finished - $started )) >> $result_file
+    bin="$path_to_bins/galois-sssp-cpu"
+    graph="$path_to_graphs/$1.gr"
+    start=$((`date +%s`*1000+`date +%-N`/1000000))
+    result=$(timeout 3h $bin --startNode=$2 $graph 2> /dev/null | ts '%.s')
+    log "$result"
+    finished=$(echo "$result" | grep 'Verification' | awk '{print $1}')
+    started=$(echo "$result" | grep 'Read ' | awk '{print $1}')
+    finished=${finished//.}
+    finished=$(($finished/1000))
+    started=${started//.}
+    started=$(($started/1000))
+    time_read=$(($started-$start))
+    time_calc=$(($finished-$started))
+    echo "galois-sssp $1 $time_read $time_calc"
 }
+
 galois-pagerank-push () {
-    # pagerank push
-    bin="${path_to_bins}/galois-pagerank-push-cpu"
-    result=$(timeout 3h ./$bin $i | ts '%s')
-    finished=$(grep 'STAT_TYPE' $result | awk '{print $1}')
-    started=$(grep 'Read ' $result | awk '{print $1}')
-    echo "Galois-pagerank-push on $i" >> $result_file
-    echo $(( $finished - $started )) >> $result_file
+    bin="$path_to_bins/galois-pagerank-push-cpu"
+    graph="$path_to_graphs/$1.gr"
+    start=$((`date +%s`*1000+`date +%-N`/1000000))
+    result=$(timeout 3h $bin --maxIterations=$pagerank_number_of_iterations $graph 2> /dev/null | ts '%.s')
[…]

### 2020-07-14 documentation/paper/galois.tex

%!TEX root=main.tex

Galois \cite{Galois} is a general purpose library designed for parallel programming. The Galois system supports fine grain tasks, allows for autonomous, speculative execution of these tasks and grants control over the task scheduling policies to the application. It also simplifies the implementation of parallel applications by providing an implicitly parallel unordered-set iterator.\\
For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented. Galois includes applications for for many graph analytics problems, among these are single-source shorthest-paths (sssp) and pagerank. Both of these applications can be executed in shared memory systems and due to the Gluon integration in a distributed environment.\\
Gluon \cite{vertGalois} reduces the communication overhead needed in distributed systems for graph analysis by exploiting structural and temporal invariants.

### 2020-07-15 scripts/benchmark/benchmark.sh
benchmark cases

@@ -11,11 +11,11 @@
 # $1 graph $2 sssp startnode $3 number of nodes
 
 logv () { # writes to stdout
-    echo "$@" 1>&2; 
+    echo "$@" 1>&2;
 }
 log () { # writes to stdout and stderr
     echo "$@"
-    echo "#$@" 1>&2; 
+    echo "#$@" 1>&2;
 }
 
 convert_time () { # from float or exponential seconds to integer microsceconds
@@ -235,13 +235,6 @@ gemini-pagerank-dist () {
     log "gemini-pagerank-dist $1 $pagerank_number_of_iterations - $dur_calc $dur_exec"
 }
 
-path_to_graphs=/home/ubuntu/graph
-graph_info=/home/ubuntu/graph/graph_info.txt
-path_to_bins=/home/ubuntu/bin
-host_file=/home/ubuntu/host_file
-#result_file=$4
-pagerank_number_of_iterations=1
-
 benchmark_graph () { # graph $1, startnode $2, number_of_nodes $3, maxstartnode $4
 	local startnode=$2
 	if [[ "$startnode" == "random" ]]; then
@@ -266,6 +259,3 @@ benchmark_graph () { # graph $1, startnode $2, number_of_nodes $3, maxstartnode 
 	galois-pagerank-pull-dist $1
 }
 
-
-benchmark_graph test random 3 3
-benchmark_graph orkut random 3072442 2724230

### 2020-07-15 scripts/benchmark/benchmark_cases.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+source benchmark.sh
+
+path_to_graphs=/home/ubuntu/graph
+graph_info=/home/ubuntu/graph/graph_info.txt
+path_to_bins=/home/ubuntu/bin
+host_file=/home/ubuntu/host_file
+#result_file=$4
+pagerank_number_of_iterations=1
+
+get_time

### 2020-07-15 scripts/benchmark/benchmark_cases.sh
benchmark first test cases

@@ -8,4 +8,92 @@ host_file=/home/ubuntu/host_file
 #result_file=$4
 pagerank_number_of_iterations=1
 
-get_time
+#Name,           Nodes,      Edges,      MaxSourceNodeId   RandomStartNodes
+#flickrEdges,    105939,     2416948,    74591             5114,73690,23892,30641,38088,16717,68726,25185,31011,42329
+#friendster,     68349467,   2586147869, 56110198          39797743,35222816,1794578,43131729,20457996,490332,4274039,10253002,30789169,54514110
+#orkut,          3072442,    117184899,  2724230           1345950,1962322,2692187,1198261,163962,1407061,1613980,1976485,2350837,1834214
+#twitterMpi,     52579683,   1963263821, 43983853          9226150,28694596,11344459,8641427,34012760,43840792,106660,26370335,38559695,19217596
+#wikipedia,      12150977,   378142420,  12114967          6712096,1761449,4340524,4857604,7825079,6241517,10821044,3259457,8277471,7539859
+
+startnodes_flickr=(5114 73690 23892 30641 38088 16717 68726 25185 31011 42329)
+startnodes_friendster=(39797743 35222816 1794578 43131729 20457996 490332 4274039 10253002 30789169 54514110)
+startnodes_orkut=(1345950 1962322 2692187 1198261 163962 1407061 1613980 1976485 2350837 1834214)
+startnodes_twitter=(9226150 28694596 11344459 8641427 34012760 43840792 106660 26370335 38559695 19217596)
+startnodes_wikipedia=(6712096 1761449 4340524 4857604 7825079 6241517 10821044 3259457 8277471 7539859)
+
+benchmark_polymer_sssp() {
+  for startnode in ${startnodes_flickr[@]}; do
+    polymer-sssp flickr $startnode
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    polymer-sssp friendster $startnode
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    polymer-sssp orkut $startnode
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    polymer-sssp twitter $startnode
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    polymer-sssp wikipedia $startnode
+  done
+}
+
+benchmark_polymer_pagerank() {
[…]

### 2020-07-15 scripts/benchmark/benchmark_cases.sh
benchmark first test cases

@@ -1,12 +1,12 @@
 #!/bin/bash
 source benchmark.sh
 
-path_to_graphs=/home/ubuntu/graph
-graph_info=/home/ubuntu/graph/graph_info.txt
-path_to_bins=/home/ubuntu/bin
-host_file=/home/ubuntu/host_file
+path_to_graphs=/home/fp-ss20/graph
+#graph_info=/home/ubuntu/graph/graph_info.txt
+path_to_bins=/home/fp-ss20/bin
+#host_file=/home/ubuntu/host_file
 #result_file=$4
-pagerank_number_of_iterations=1
+pagerank_number_of_iterations=5
 
 #Name,           Nodes,      Edges,      MaxSourceNodeId   RandomStartNodes
 #flickrEdges,    105939,     2416948,    74591             5114,73690,23892,30641,38088,16717,68726,25185,31011,42329

### 2020-07-15 scripts/benchmark/benchmark.sh
created giraph benchmark functions (not yet usable)

@@ -235,6 +235,36 @@ gemini-pagerank-dist () {
     log "gemini-pagerank-dist $1 $pagerank_number_of_iterations - $dur_calc $dur_exec"
 }
 
+giraph-sssp () {
+    # local bin="${path_to_bins}/giraph-sssp"
+    #local graph="${path_to_graphs}/$1.adj"
+    local time_start=$(get_time)
+    #local result=$(timeout 3h $bin -r $2 -rounds 1 $graph)
+    local dur_exec=$(($(get_time)-$time_start))
+    logv "$result"
+    local dur_init=$(echo $result | grep 'Initialize (ms)')
+    local dur_init="${dur_init//.*Initialize (ms)=}"
+    local dur_total="${result//.*Total (ms)=}" 
+    local dur_calc=$(($dur_total-$dur_init))
+    local dur_calc=$(convert_time $dur_calc)
+    log "giraph-sssp $1 $2 - $dur_calc $dur_exec"
+}
+
+giraph-pagerank () {
+    # local bin="${path_to_bins}/giraph-sssp"
+    #local graph="${path_to_graphs}/$1.adj"
+    local time_start=$(get_time)
+    #local result=$(timeout 3h $bin -r $2 -rounds 1 $graph)
+    local dur_exec=$(($(get_time)-$time_start))
+    logv "$result"
+    local dur_init=$(echo $result | grep 'Initialize (ms)')
+    local dur_init="${dur_init//.*Initialize (ms)=}"
+    local dur_total="${result//.*Total (ms)=}" 
+    local dur_calc=$(($dur_total-$dur_init))
+    local dur_calc=$(convert_time $dur_calc)
+    log "giraph-pagerank $1 $2 - $dur_calc $dur_exec"
+}
+
 benchmark_graph () { # graph $1, startnode $2, number_of_nodes $3, maxstartnode $4
 	local startnode=$2
 	if [[ "$startnode" == "random" ]]; then

### 2020-07-16 scripts/benchmark/benchmark_cases.sh
benchmark_cases gemini

@@ -20,6 +20,11 @@ startnodes_friendster=(39797743 35222816 1794578 43131729 20457996 490332 427403
 startnodes_orkut=(1345950 1962322 2692187 1198261 163962 1407061 1613980 1976485 2350837 1834214)
 startnodes_twitter=(9226150 28694596 11344459 8641427 34012760 43840792 106660 26370335 38559695 19217596)
 startnodes_wikipedia=(6712096 1761449 4340524 4857604 7825079 6241517 10821044 3259457 8277471 7539859)
+nodes_flickr=105939
+nodes_friendster=68349467
+nodes_orkut=3072442
+nodes_twitter=52579683
+nodes_wikipedia=12150977
 
 benchmark_polymer_sssp() {
   for startnode in ${startnodes_flickr[@]}; do
@@ -93,7 +98,43 @@ benchmark_ligra_pagerank() {
   done
 }
 
-benchmark_polymer_sssp
+benchmark_gemini_sssp() {
+  for startnode in ${startnodes_flickr[@]}; do
+    gemini-sssp flickr $startnode $nodes_flickr
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    gemini-sssp friendster $startnode $nodes_friendster
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    gemini-sssp orkut $startnode $nodes_orkut
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    gemini-sssp twitter $startnode $nodes_twitter
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    gemini-sssp wikipedia $startnode $nodes_wikipedia
+  done
+}
+
+benchmark_gemini_pagerank() {
+  for startnode in ${startnodes_flickr[@]}; do
+    gemini-pagerank flickr $nodes_flickr
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    gemini-pagerank friendster $nodes_friendster
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    gemini-pagerank orkut $nodes_orkut
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    gemini-pagerank twitter $nodes_twitter
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    gemini-pagerank wikipedia $nodes_wikipedia
+  done
+}
+
+#benchmark_polymer_sssp
 #benchmark_polymer_pagerank
 #benchmark_ligra_sssp
 #benchmark_ligra_pagerank

### 2020-07-20 sortEdgeList.cpp
added conversion to a Giraph Format

@@ -73,7 +73,7 @@ edgeList readEdgeList(char* filename, int isWeighted) {
     // allocate edgeList
     graph.edges = (edge*) malloc(graph.length * sizeof(edge));
     if (graph.edges == NULL) {
-        fprintf(stderr, "Failed to malloc array for graph with length %d\n", graph.length * sizeof(edge));
+        fprintf(stderr, "Failed to malloc array for graph with length %llu\n", graph.length * sizeof(edge));
         fclose(f);
         return graph;
     }
@@ -268,8 +268,45 @@ void writeAdjacencyGraph(char* filename, edgeList graph){
     fclose(f);
 }
 
+
+
+// writes the given graph to the specified file in the
+// AdjacencyListTextVertexInputFormat format
+// requires a sorted and reindexed input graph
+void writeAdjacencyListTextVertexFormat(char* filename, edgeList graph){
+    FILE* f = fopen(filename, "w");
+    if (f == NULL) {
+        perror("Cannot open file\n");
+        exit(EXIT_FAILURE);
+    }
+
+    //prints <nodeID> <nodeWeight> (<targetNodeId> <edgeWeight>)*\n
+    //where the nodeWeight is set as 1.
+    node_id currentNode = graph.edges[0].from;
+    fprintf(f, "%d 1", currentNode);
+
+    for(size_t i = 0; i < graph.length; i++) {
+        if(currentNode != graph.edges[i].from) {
+            fprintf(f, "\n");//end line
+
+            currentNode = graph.edges[i].from;
+            fprintf(f, "%d 1", currentNode); //begin new line
+        }
+        fprintf(f, " %d", graph.edges[i].to);
+        if(graph.type == WEIGHTED) {
+            fprintf(f, " %d", graph.edges[i].weigth);
+        }
+        else {
+            fprintf(f, " 1");
+        }
+    }
+    fclose(f);
+}
+
+
+
 int main(int argc, char* argv[]) {
[…]

### 2020-07-20 scripts/Giraph_setup.sh
a small script to take over part of the giraph installation, this is unfinished

@@ -1,0 +1,34 @@
+
+# This will install hadoop and giraph,
+# however it will NOT configure:
+# * /etc/hosts
+
+
+#prerequisites
+sudo apt-get update
+sudo apt-get upgrade
+sudo apt-get install default-jdk git maven net-tools
+
+#install hadoop
+cd /usr/local
+sudo wget http://archive.apache.org/dist/hadoop/core/hadoop-0.20.205.0/hadoop-0.20.205.0.tar.gz
+sudo tar xzf hadoop-0.20.205.0.tar.gz
+sudo mv hadoop-0.20.205.0 hadoop
+sudo rm hadoop-0.20.205.0.tar.gz
+
+#permissions to hadoop folder
+sudo chown -R $USER hadoop
+
+#.bashrc, except not
+export HADOOP_HOME=/usr/local/hadoop
+export JAVA_HOME=/usr/lib/jvm/java-11-openjdk-amd64
+export GIRAPH_HOME=/usr/local/giraph
+
+#hadoop-env
+echo "export JAVA_HOME=/usr/lib/jvm/default-java
+export HADOOP_OPTS=-Djava.net.preferIPv4Stack=true" >> /usr/local/hadoop/conf/hadoop-env.sh
+
+#hadoop tmp folder
+sudo mkdir -p /app/hadoop/tmp
+sudo chown $USER /app/hadoop/tmp
+sudo chmod 750 /app/hadoop/tmp

### 2020-07-20 documentation/paper/graphFormats.tex
a bit of progress on the paper, mainly polymer and reordering plus placehoders

@@
 
-\section{Graph Formats}
+\subsection{An overview of some graph formats}
 A rather big portion of our time was invested in figuring out which graph framework requires which graph formats.
@@
 Additionally, to make life in the future a little bit easier, we wrote multiple tools to convert graphs acquired from Snap or Konect to the required formats.
-Information on how to use those tools is available in our repository.
+Additional information on this is available in the section \nameref{supplementaryData} at the end.
 
-\subsection{AdjacencyList}
+\subsubsection{AdjacencyList}
 The AdjacencyList and WeightedAdjacencyList formats\cite{adjListFormat} are used by Ligra and Polymer.
@@
 
-\subsection{EdgeList}
+\subsubsection{EdgeList}
 The EdgeList format is probably the easiest to understand and is one of the most commonly used in the online graph repositories.
@@
 
-\subsection{Binary EdgeList}
+\subsubsection{Binary EdgeList}
 The binary EdgeList format is used by Gemini. Finding information on this format required reverse engineering of the Gemini code.
@@
 
-\subsection{?}
+\subsubsection{Giraph's numerous I/O formats}

### 2020-07-20 documentation/paper/main.tex
a bit of progress on the paper, mainly polymer and reordering plus placehoders

@@
 \usepackage{amsmath}
+
+\newcommand{\todo}{{\color{red}TODO}}
 
@@
 \begin{abstract}
-In this paper we will analyze and compare the graph frameworks Galois, Ligra and Polymer in their parformance. We will also set up and compare these frameworks in a distributed scenario.
+In this paper we will analyze and compare the graph frameworks Galois, Ligra, Polymer, Gemini and Giraph in their parformance.
+All the frameworks will be tested in shared memory and Galois, Gemini and Giraph are tested on a distributed cluster as well.
+Furthermore we will give some insight on the complexity of writing custom applications based on these frameworks.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Galois, Ligra, Polymer, distributed computing, Gluon
+graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon
 \end{IEEEkeywords}
@@
 \section{Introduction}
-My intro... blah blah \cite{Polymer}.
+%\todo
 
-\input{graphFormats}
-\section{Things}
+This paper makes the following contributions:
+\begin{itemize}
+  \item 
+  \item 
+  \item 
+  \item 
+\end{itemize}
+
+\section{Overview of the framworks}
 \input{galois}
@@
 \input{polymer}
+
+\input{giraph}
+
+\section{And}
+\input{graphFormats}%hier oder supplementary data
+
+\section{Our Testing methods}
+
+\section{Results}
+
+\subsection{Pure Performance-Ergebnisse}
+
+\subsection{Komplexität Aufsetzen}
+
+\subsection{Komplexität eigene Apps schreiben}
+
+\section{Discussion}
 
@@
 
-% Das hier ist mega beschissen: 
 % Beim Hinzufügen neuer Referenzen
@@
 
+\section*{Supplementary Data}\label{supplementaryData}
+We have written a number of conversion tools and installation guides to help
+users or develeopers with the use of the tested frameworks.
+
+Everything can be retrieved on our GitHub repository
+\url{http://www.github.com/serengti/Forschungsprojekt}.
+
+For each Framework, there is a
+
 \end{document}

### 2020-07-20 documentation/paper/polymer.tex
a bit of progress on the paper, mainly polymer and reordering plus placehoders

@@
 
-\subsection{Polymer}
-Installing Polymer is straight forward.
+Polymer is very similar to Ligra, in fact Polymer inherits the programming interface from Ligra\cite{Polymer}.
+
+Polymer aims to minimize both random and remote memory accesses by implementing NUMA- and graph-aware data layout and memory access strategies.
+Specifically, Polymer co-locates graph data and the computation within NUMA-nodes to reduce remote memory accesses.
+For example, Polymer eliminates remote accesses by letting threads allocate memory in their local memory node for graph topology data like vertices and edges that are only accessed by one thread.
+Application-defined data with static memory locations which gets dynamically updated during computation is allocated with virtual addresses that make for a seamless cross-node data access.
+Other mutable runtime states (e.g. active vertices) might be dynamically allocated in each iteration. This data is allocated in a distributed way but only accessed through a global lookup table.
+
+%Graph analytics has been long recognized to have many random access and poor data locality.
+Based on the observation that sequential inter-node (i.e., remote) memory accesses have much higher bandwidth than both intra- and inter-node random ones, Polymer borrows the idea from distributed graph systems by replicating vertex data across NUMA-nodes in a lightweight way.
+(This essentially follows the philosophy of current multi-core OS designs by treating a large-scale machine as a distributed system).
+Specifically, a vertex only conducts computation on edges within the local NUMA-node and uses its replicas in other NUMA-node to initiate the computation on other edges.
+Unlike distributed graph systems, Polymer does not distribute application- defined data but still applies all updates to a vertex within a single copy of application-defined data.
+Polymer has three optimizations to improve scheduling on NUMA machines and handle different properties of graphs.
+First, being aware of the hierarchical parallelism and locality in NUMA machines, Polymer is extended with a hierarchical scheduler to reduce the cost for synchronizing threads on multiple NUMA-nodes.
+Second, inspired by vertex-cuts from distributed graph systems, Polymer improves the balance of workload among NUMA-nodes by evenly partitioning edges rather than vertices for skewed graphs.
+Finally, as most graph algorithms converge asymmetrically, using a single data structure for runtime states may drastically degrade the performance, especially for traversal algorithms on high-diameter graphs, Polymer adopts adaptive data structures boost performance.

### 2020-07-21 sortEdgeList.cpp
changed graph conversion to JsonLongDoubleFloatDoubleVertexInputFormat because the other format does not seem to work for some reason

@@ -271,7 +271,7 @@ void writeAdjacencyGraph(char* filename, edgeList graph){
 
 
 // writes the given graph to the specified file in the
-// AdjacencyListTextVertexInputFormat format
+// org.apache.giraph.io.formats.JsonLongDoubleFloatDoubleVertexInputFormat
 // requires a sorted and reindexed input graph
 void writeAdjacencyListTextVertexFormat(char* filename, edgeList graph){
     FILE* f = fopen(filename, "w");
@@ -280,33 +280,61 @@ void writeAdjacencyListTextVertexFormat(char* filename, edgeList graph){
         exit(EXIT_FAILURE);
     }
 
-    //prints <nodeID> <nodeWeight> (<targetNodeId> <edgeWeight>)*\n
-    //where the nodeWeight is set as 1.
-    node_id currentNode = graph.edges[0].from;
-    fprintf(f, "%d 1", currentNode);
-
-    for(size_t i = 0; i < graph.length; i++) {
-        if(currentNode != graph.edges[i].from) {
-            fprintf(f, "\n");//end line
-
-            currentNode = graph.edges[i].from;
-            fprintf(f, "%d 1", currentNode); //begin new line
-        }
-        fprintf(f, " %d", graph.edges[i].to);
-        if(graph.type == WEIGHTED) {
-            fprintf(f, " %d", graph.edges[i].weigth);
-        }
-        else {
-            fprintf(f, " 1");
+    node_id biggestNodeId = graph.edges[graph.length-1].from;
+    for (size_t i = 0; i < graph.length; i++){
+        if (graph.edges[i].to > biggestNodeId){
+            biggestNodeId = graph.edges[i].to;
         }
     }
+
+    //prints [<nodeID>, <nodeWeight>, [[<targetNodeId> <edgeWeight>]*]]\n
+    //where the nodeWeight is set to 0 and
+    //the edge weights are 1 if the graph is not weighted.
+    node_id currentNode = graph.edges[0].from;
+    fprintf(f, "[%d,0,[", currentNode);
+    bool first = true;
+    for(size_t i = 0; i < graph.length; i++) {
+        if(currentNode != graph.edges[i].from) {
+            fprintf(f, "]]\n");//end line
+
+            //this is for nodes that have no outgoing edges
+            while(graph.edges[i].from != currentNode + 1) {
+                currentNode += 1;
+                fprintf(f, "[%d,0,[]]\n", currentNode);
+            }
+
[…]

### 2020-07-21 SimplePageRankComputationWithAggregator.java
PageRank Application for giraph

@@ -1,0 +1,88 @@
+package org.apache.giraph.examples;
+
+/*
+http://buttercola.blogspot.com/2015/04/giraph-page-rank.html
+ */
+ 
+import org.apache.giraph.aggregators.DoubleMaxAggregator;
+import org.apache.giraph.aggregators.DoubleMinAggregator;
+import org.apache.giraph.aggregators.LongSumAggregator;
+import org.apache.giraph.edge.Edge;
+import org.apache.giraph.edge.EdgeFactory;
+import org.apache.giraph.graph.BasicComputation;
+import org.apache.giraph.graph.Vertex;
+import org.apache.giraph.io.VertexReader;
+import org.apache.giraph.io.formats.GeneratedVertexInputFormat;
+import org.apache.giraph.io.formats.TextVertexOutputFormat;
+import org.apache.giraph.master.DefaultMasterCompute;
+import org.apache.giraph.worker.WorkerContext;
+import org.apache.hadoop.io.DoubleWritable;
+import org.apache.hadoop.io.FloatWritable;
+import org.apache.hadoop.io.LongWritable;
+import org.apache.hadoop.io.Text;
+import org.apache.hadoop.mapreduce.InputSplit;
+import org.apache.hadoop.mapreduce.TaskAttemptContext;
+import org.apache.log4j.Logger;
+ 
+import com.google.common.collect.Lists;
+ 
+import java.io.IOException;
+import java.util.List;
+public class SimplePageRankComputationWithAggregator extends BasicComputation<LongWritable,
+    DoubleWritable, FloatWritable, DoubleWritable> {
+  /** Number of supersteps for this test */
+  public static final int MAX_SUPERSTEPS = 30;
+  /** Logger */
+  private static final Logger LOG = 
+      Logger.getLogger(SimplePageRankComputationWithAggregator.class);
+  /** Sum aggregator name */
+  private static String SUM_AGG = "sum";
+  /** Min aggregator name */
+  private static String MIN_AGG = "min";
+  /** Max aggregator name */
+  private static String MAX_AGG = "max";
+ 
+  @Override
+  public void compute(
+      Vertex<LongWritable, DoubleWritable, FloatWritable> vertex, 
+      Iterable<DoubleWritable> messages) throws IOException {
+    if (getSuperstep() > 0) {
+      double sum = 0;
+      for (DoubleWritable message : messages) {
+        sum += message.get();
+      }
[…]

### 2020-07-21 GeneralShortestPathsComputation.java
SSSP with customizable start node ids finished

@@ -1,0 +1,93 @@
+/*
+ * Licensed to the Apache Software Foundation (ASF) under one
+ * or more contributor license agreements.  See the NOTICE file
+ * distributed with this work for additional information
+ * regarding copyright ownership.  The ASF licenses this file
+ * to you under the Apache License, Version 2.0 (the
+ * "License"); you may not use this file except in compliance
+ * with the License.  You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
+package org.apache.giraph.examples;
+
+import org.apache.giraph.graph.BasicComputation;
+import org.apache.giraph.conf.LongConfOption;
+import org.apache.giraph.edge.Edge;
+import org.apache.giraph.graph.Vertex;
+import org.apache.hadoop.io.DoubleWritable;
+import org.apache.hadoop.io.FloatWritable;
+import org.apache.hadoop.io.LongWritable;
+import org.apache.log4j.Logger;
+
+import java.io.IOException;
+
+/**
+ * Demonstrates the basic Pregel shortest paths implementation.
+ */
+@Algorithm(
+    name = "Shortest paths",
+    description = "Finds all shortest paths from a selected vertex"
+)
+public class GeneralShortestPathsComputation extends BasicComputation<
+    LongWritable, DoubleWritable, FloatWritable, DoubleWritable> {
+
+  private static long startVertexId;
+  public static void main(String[] args) {
+    startVertexId = Long.valueOf(args[0]);
+  }
+
+  /** The shortest paths id */
+  public static final LongConfOption SOURCE_ID =
+      new LongConfOption("SimpleShortestPathsVertex.sourceId", startVertexId,
+          "The shortest paths id");
+  /** Class logger */
+  private static final Logger LOG =
+      Logger.getLogger(GeneralShortestPathsComputation.class);
+
+  /**
+   * Is this vertex the source id?
+   *
+   * @param vertex Vertex
+   * @return True if the source id
+   */
[…]

### 2020-07-21 SimplePageRankComputationWithAggregator.java
own PageRank Computation not neccessary, maybe

@@ -1,88 +1,0 @@
-package org.apache.giraph.examples;
-
-/*
-http://buttercola.blogspot.com/2015/04/giraph-page-rank.html
- */
- 
-import org.apache.giraph.aggregators.DoubleMaxAggregator;
-import org.apache.giraph.aggregators.DoubleMinAggregator;
-import org.apache.giraph.aggregators.LongSumAggregator;
-import org.apache.giraph.edge.Edge;
-import org.apache.giraph.edge.EdgeFactory;
-import org.apache.giraph.graph.BasicComputation;
-import org.apache.giraph.graph.Vertex;
-import org.apache.giraph.io.VertexReader;
-import org.apache.giraph.io.formats.GeneratedVertexInputFormat;
-import org.apache.giraph.io.formats.TextVertexOutputFormat;
-import org.apache.giraph.master.DefaultMasterCompute;
-import org.apache.giraph.worker.WorkerContext;
-import org.apache.hadoop.io.DoubleWritable;
-import org.apache.hadoop.io.FloatWritable;
-import org.apache.hadoop.io.LongWritable;
-import org.apache.hadoop.io.Text;
-import org.apache.hadoop.mapreduce.InputSplit;
-import org.apache.hadoop.mapreduce.TaskAttemptContext;
-import org.apache.log4j.Logger;
- 
-import com.google.common.collect.Lists;
- 
-import java.io.IOException;
-import java.util.List;
-public class SimplePageRankComputationWithAggregator extends BasicComputation<LongWritable,
-    DoubleWritable, FloatWritable, DoubleWritable> {
-  /** Number of supersteps for this test */
-  public static final int MAX_SUPERSTEPS = 30;
-  /** Logger */
-  private static final Logger LOG = 
-      Logger.getLogger(SimplePageRankComputationWithAggregator.class);
-  /** Sum aggregator name */
-  private static String SUM_AGG = "sum";
-  /** Min aggregator name */
-  private static String MIN_AGG = "min";
-  /** Max aggregator name */
-  private static String MAX_AGG = "max";
- 
-  @Override
-  public void compute(
-      Vertex<LongWritable, DoubleWritable, FloatWritable> vertex, 
-      Iterable<DoubleWritable> messages) throws IOException {
-    if (getSuperstep() > 0) {
-      double sum = 0;
-      for (DoubleWritable message : messages) {
-        sum += message.get();
-      }
[…]

### 2020-07-22 scripts/benchmark/benchmark.sh
galois cpu benchmark cases

@@ -33,7 +33,7 @@ get_time () { # timestamp in microsceconds
     echo $((`date +%s`*1000000+`date +%-N`/1000))
 }
 
-galois-sssp () {
+galois-sssp-cpu () {
     local bin="$path_to_bins/galois-sssp-cpu"
     local graph="$path_to_graphs/$1.gr"
     local time_start=$(get_time)
@@ -44,7 +44,7 @@ galois-sssp () {
     local time_started_calc=$(echo "$result" | grep 'Read ' | awk '{print $1}')
     local dur_read=$((${time_started_calc//.}-$time_start))
     local dur_calc=$((${time_finished_calc//.}-${time_started_calc//.}))
-    log "galois-sssp $1 $2 $dur_read $dur_calc $dur_exec"
+    log "galois-sssp-cpu $1 $2 $dur_read $dur_calc $dur_exec"
 }
 
 galois-sssp-push-dist () {
@@ -75,7 +75,7 @@ galois-sssp-pull-dist () {
     log "galois-sssp-pull-dist $1 $2 $dur_read $dur_calc $dur_exec"
 }
 
-galois-pagerank-push () {
+galois-pagerank-push-cpu () {
     local bin="$path_to_bins/galois-pagerank-push-cpu"
     local graph="$path_to_graphs/$1.gr"
     local time_start=$(get_time)
@@ -86,7 +86,7 @@ galois-pagerank-push () {
     local time_started_calc=$(echo "$result" | grep 'Read ' | awk '{print $1}')
     local dur_read=$((${time_started_calc//.}-$time_start))
     local dur_calc=$((${time_finished_calc//.}-${time_started_calc//.}))
-    log "galois-pagerank-push $1 $pagerank_number_of_iterations $dur_read $dur_calc $dur_exec"
+    log "galois-pagerank-push-cpu $1 $pagerank_number_of_iterations $dur_read $dur_calc $dur_exec"
 }
 
 galois-pagerank-push-dist () {
@@ -103,7 +103,7 @@ galois-pagerank-push-dist () {
     log "galois-pagerank-push-dist $1 $pagerank_number_of_iterations $dur_read $dur_calc $dur_exec"
 }
 
-galois-pagerank-pull () {
+galois-pagerank-pull-cpu () {
     local bin="$path_to_bins/galois-pagerank-pull-cpu"
     local graph="$path_to_graphs/$1.gr"
     local time_start=$(get_time)
@@ -114,7 +114,7 @@ galois-pagerank-pull () {
[…]

### 2020-07-22 scripts/benchmark/benchmark_cases.sh

@@ -242,9 +242,79 @@ benchmark_polymer_pagerank_write() {
   done
 }
 
-#benchmark_gemini_bfs
-#benchmark_ligra_bfs
-#benchmark_polymer_bfs
-#benchmark_ligra_pagerank_delta
-#benchmark_polymer_pagerank_delta
-#benchmark_polymer_pagerank_write
+benchmark_galois_sssp_cpu() {
+  for startnode in ${startnodes_flickr[@]}; do
+    galois-sssp-cpu flickr $startnode
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    galois-sssp-cpu friendster $startnode
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    galois-sssp-cpu orkut $startnode
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    galois-sssp-cpu twitter $startnode
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    galois-sssp-cpu wikipedia $startnode
+  done
+}
+
+benchmark_galois_pagerank_push_cpu() {
+  for startnode in ${startnodes_flickr[@]}; do
+    galois-pagerank-push-cpu flickr
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    galois-pagerank-push-cpu friendster
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    galois-pagerank-push-cpu orkut
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    galois-pagerank-push-cpu twitter
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    galois-pagerank-push-cpu wikipedia
+  done
+}
+
+benchmark_galois_pagerank_pull_cpu() {
+  for startnode in ${startnodes_flickr[@]}; do
+    galois-pagerank-pull-cpu flickr
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    galois-pagerank-pull-cpu friendster
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    galois-pagerank-pull-cpu orkut
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    galois-pagerank-pull-cpu twitter
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    galois-pagerank-pull-cpu wikipedia
+  done
+}
+
[…]

### 2020-08-10 documentation/paper/galois.tex
little paper work

@@
 
-Galois \cite{Galois} is a general purpose library designed for parallel programming.
+Galois\cite{Galois} is a general purpose library designed for parallel programming.
 The Galois system supports fine grain tasks, allows for autonomous, speculative execution of these tasks and grants control over the task scheduling policies to the application.
-It also simplifies the implementation of parallel applications by providing an implicitly parallel unordered-set iterator.\\
+It also simplifies the implementation of parallel applications by providing an implicitly parallel unordered-set iterator.
+
 For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
-Galois includes applications for for many graph analytics problems, among these are single-source shorthest-paths (sssp) and pagerank.
-Both of these applications can be executed in shared memory systems and due to the Gluon integration in a distributed environment.\\
-Gluon \cite{vertGalois} reduces the communication overhead needed in distributed systems for graph analysis by exploiting structural and temporal invariants.
+Galois includes applications for for many graph analytics problems, among these are single-source shorthest-paths (sssp), breath-first-search (bfs) and pagerank.
+For most of these applications Galois offers a huge amount of algorithms to perform these analytics problems and many setting options, like the amount of threads used or policies for splitting the graph.
+All of these applications can be executed in shared memory systems and due to the Gluon integration in a distributed environment.
+
+Gluon\cite{vertGalois} is a framework written by the Galois team as a middleware to write distributed systems for graph analysis.
+It reduces the communication overhead needed in distributed by exploiting structural and temporal invariants.
+
+The code of Gluon is embedded into the project of Galois and already integrated with it.
+It is possible to integrate Gluon with other frameworks too, which the Galois team showed in their paper\cite{vertGalois}.

### 2020-08-10 documentation/paper/graphFormats.tex
little paper work

@@
 The EdgeList format is probably the easiest to understand and is one of the most commonly used in the online graph repositories.
-The directed eges $(s_1,t_1),(s_2,t_2),\ldots$ are represented in the following way.
+The directed eges $(s_1,t_1),(s_2,t_2),\ldots$ or $(s_1,t_1, w_1),(s_2,t_2, w_2),\ldots$ are represented in the following way.
 \begin{align*}
-	&s_1,t_1\\
-	&s_2,t_2\\
+  &s_1\ \ t_1\ \ w_1\\
+  &s_2\ \ t_2\ \ w_m\\
 	&\quad\vdots\\
-	&s_m,t_m
+  &s_m\ \ t_m\ \ w_m
 \end{align*}
+The weights are optional, everything is ASCI encoded and the inline delimiter is a variable amount of any whitespace.
 
@@
 \end{equation*}
-where $s_i,t_i$ have \texttt{uint32} ?? data type and the weights are \texttt{float32}. 
+where $s_i,t_i$ have \texttt{uint32} data type and the optional weights are \texttt{float32}.
 Gemini will derive the number of edges from the file size, so there is no file header or anything similar allowed.

### 2020-08-10 documentation/paper/main.tex
little paper work
some layouting in main, begin testing methods and some work on giraph setup guide

@@
 \section{Our Testing methods}
+\input{testingMethods}
 
 \section{Results}
+\subsection{Komplexität Aufsetzen}
+
+\subsection{Komplexität eigene Apps schreiben}
 
@@
 
-\subsection{Komplexität Aufsetzen}
+\subsection{Ergebnisse Calc time}
+der offensichtliche, wichtige Vergleich
 
-\subsection{Komplexität eigene Apps schreiben}
+\subsection{Ergebnisse exec time}
+hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
+Aber vielleicht ist auch die setup time bei gleichen frameworks zwischen verteilt und shared memory ganz interessant zu vergleichen.
 
@@
 \section*{Acknowledgment}
-We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Ligra} and Polymer~\cite{Polymer}.
+We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Ligra}, Polymer~\cite{Polymer}, Gemini~\cite{Gemini} as well as Apache Giraph~\cite{Giraph}.
 
@@
 % Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
+% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
 \begin{thebibliography}{1}
@@
 
+\bibitem{Galois}
+\BIBentryALTinterwordspacing
+D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
+  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
+  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
+  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
+  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
+\BIBentrySTDinterwordspacing
+
+\bibitem{vertGalois}
+\BIBentryALTinterwordspacing
+R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
+  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
+  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
+  Conference on Programming Language Design and Implementation}, ser. PLDI
+  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
+  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
+  \url{https://doi.org/10.1145/3192366.3192404}
+\BIBentrySTDinterwordspacing
+
 \bibitem{Polymer}
@@
 
-\bibitem{Galois}
-\BIBentryALTinterwordspacing
-D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
-  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
-  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
-  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
-  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
-\BIBentrySTDinterwordspacing
-
 \bibitem{Ligra}
@@
 
-\bibitem{vertGalois}
-\BIBentryALTinterwordspacing
-R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
-  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
-  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
-  Conference on Programming Language Design and Implementation}, ser. PLDI
-  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
-  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
-  \url{https://doi.org/10.1145/3192366.3192404}
-\BIBentrySTDinterwordspacing
-
 \bibitem{Gemini}
@@
   \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
+\BIBentrySTDinterwordspacing
+
+\bibitem{Giraph}
+\BIBentryALTinterwordspacing
+A.~S. Foundation. (2020, Jun.) {Apache Giraph}. [Online]. Available:
+  \url{https://giraph.apache.org}
 \BIBentrySTDinterwordspacing

### 2020-08-10 documentation/paper/testingMethods.tex
some layouting in main, begin testing methods and some work on giraph setup guide

%!TEX root=main.tex
\subsection{Hardware and Software}
For testing all systems we used 5 machines with 96 CPU cores each (48 physical) and 256 GB of RAM. One of those machines only had 128 GB of RAM, this one was only used as part of the calculation cluster for the distributed systems.
All servers were running Ubuntu 19.x.
Setup of each framework was performed according to our provided installation guides.

\subsection{Benchmark setup}
We measured the total runtime of each process as well as the actual calculation times. The console output of each frameworks was used as an indicator for when loading the graph data was finished. This proved to be more difficult than expected in some cases.

Each test case consisting of graph, framework and in some cases even the algorithm was run 10 times, allowing us to smooth slight variations in the measured times. This also helps to reduce the error we introduced by measuring time through the console log.

All benchmarks were initiated by our benchmark script that is available in our repository.

We will compare both the computation times as well es the execution times. This way we compare how much setup time each framework has.

### 2020-08-10 documentation/paper/main.tex
clean up

@@
 \title{Forschungsprojekt}
-\author{Simon König (3344789) \and Leon Matzner (3315161) \and Felix Rollbühler (3310069) \and Jakob Schmid (??????)}
+\author{\IEEEauthorblockN{Simon König}
+\IEEEauthorblockA{(3344789)\\
+st156571@stud.uni-stuttgart.de}
+\and
+\IEEEauthorblockN{Leon Matzner}
+\IEEEauthorblockA{(3315161)\\
+@stud.uni-stuttgart.de}
+\and
+\IEEEauthorblockN{Felix Rollbühler}
+\IEEEauthorblockA{(3310069)\\
+@stud.uni-stuttgart.de}
+\and
+\IEEEauthorblockN{Jakob Schmid}
+\IEEEauthorblockA{()\\
+@stud.uni-stuttgart.de}}
 \date{\today}
@@
 \begin{itemize}
-  \item 
+  \item Comparison of some of the most widely used graph based calculation frameworks
   \item 
@@
 \section{Overview of the framworks}
-\input{galois}
+\subsection{Galois and Gluon}
+\input{sections/overview/galois}
 
-\input{polymer}
+\subsection{Ligra}
+\input{sections/overview/ligra}
 
-\input{giraph}
+\subsection{Polymer}
+\input{sections/overview/polymer}
+
+\subsection{Gemini}
+\input{sections/overview/gemini}
+
+\subsection{Giraph}
+\input{sections/overview/giraph}
 
 \section{And}
-\input{graphFormats}%hier oder supplementary data
+\input{sections/graphFormats}%hier oder supplementary data?
 
-\section{Our Testing methods}
-\input{testingMethods}
+\section{Testing methods}
+\input{sections/testingMethods}
+
+Hier sollten wir auch evtl Vergleichsgrundlagen für Aufsetzen und Apps schreiben einführen, wenn wir das mit aufnehmen.
 
@@
 \subsection{Komplexität Aufsetzen}
+Systeme wie distr. Giraph bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden. \emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
 
 \subsection{Komplexität eigene Apps schreiben}
+Schwierig objektiv zu vergleichen.
 
@@
 Gemini~\cite{Gemini}
+
+\section*{Supplementary Data}\label{supplementaryData}
+We have written a number of conversion tools and installation guides to help
+users or develeopers with the use of the tested frameworks.
+
+Everything can be retrieved on our GitHub repository
+\url{http://www.github.com/serengti/Forschungsprojekt}.
+
+For each Framework, there is a
 
@@
 
-\section*{Supplementary Data}\label{supplementaryData}
-We have written a number of conversion tools and installation guides to help
-users or develeopers with the use of the tested frameworks.
-
-Everything can be retrieved on our GitHub repository
-\url{http://www.github.com/serengti/Forschungsprojekt}.
-
-For each Framework, there is a
-
 \end{document}

### 2020-08-11 documentation/paper/sections/overview/galois.tex
clean up
improved galois writeup

--- 2020-07-14 documentation/paper/galois.tex
@@
-%!TEX root=main.tex
+%!TEX root=../../main.tex
 
-Galois \cite{Galois} is a general purpose library designed for parallel programming.
+Galois\cite{Galois} is a general purpose library designed for parallel programming.
 The Galois system supports fine grain tasks, allows for autonomous, speculative execution of these tasks and grants control over the task scheduling policies to the application.
-It also simplifies the implementation of parallel applications by providing an implicitly parallel unordered-set iterator.\\
+It also simplifies the implementation of parallel applications by providing an implicitly parallel unordered-set iterators.
+
 For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
-Galois includes applications for for many graph analytics problems, among these are single-source shorthest-paths (sssp) and pagerank.
-Both of these applications can be executed in shared memory systems and due to the Gluon integration in a distributed environment.\\
-Gluon \cite{vertGalois} reduces the communication overhead needed in distributed systems for graph analysis by exploiting structural and temporal invariants.
+Galois includes applications for many graph analytics problems, among these are single-source shorthest-paths (sssp), breath-first-search (bfs) and pagerank.
+For most of these applications Galois offers several different algorithms to perform these analytics problems and many setting options like the amount of threads used or policies for splitting the graph.
+All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment.
+
+Gluon\cite{vertGalois} is a framework written by the Galois team as a middleware to write graph analysis applications for distributed systems.
+It reduces the communication overhead needed in distributed environments by exploiting structural and temporal invariants.
+
+The code of Gluon is embedded in Galois. It is possible to integrate Gluon in other frameworks too, which the Galois team showed in their paper\cite{vertGalois}.

### 2020-08-11 plots/plot_results.py
Plots

@@ -1,0 +1,100 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+plt.rcParams["figure.figsize"] = (10,10)
+
+colors = ['tab:blue', 'tab:orange', 'tab:green', 'tab:red', 'tab:purple', 'tab:brown', 'tab:pink', 'tab:gray', 'tab:olive', 'tab:cyan', 'navy', 'indigo', 'black']
+scale = 'log'
+
+def toNumpyArrays(array):
+    graph = np.empty(len(array), dtype=object)
+    algo = np.empty(len(array), dtype=object)
+    calcTime = np.empty(len(array))
+    totalTime = np.empty(len(array))
+
+    for i in range(len(array)):
+        graph[i] = array[i][0]
+        algo[i] = array[i][1]
+        try:
+            calcTime[i] = float(array[i][2])
+        except ValueError:
+            calcTime[i] = float('nan')
+        try:
+            totalTime[i] = float(array[i][3])
+        except ValueError:
+            totalTime[i] = float('nan')
+    
+    return graph, algo, calcTime, totalTime
+
+def plotAlgosOnePlot(plot, x, y, color):
+    index = 0
+    for g in np.unique(color):
+        i = np.where(color == g)
+        plot.scatter(x[i], y[i], color=colors[index], label=g)
+        index += 1
+    plot.legend()
+    plt.yscale(scale)
+
+def plotAlgosTwoPlots(plot, x, y, color):
+    fig, (ax1, ax2) = plot.subplots(2)
+    
+    index = 0
+    i = np.where(np.logical_or(x == 'flickr', np.logical_or(x == 'orkut', x == 'wikipedia')))
+    x1 = x[i]
+    c1 = color[i]
+    y1 = y[i]
+    for g in np.unique(c1):
+        i = np.where(c1 == g)
+        ax1.scatter(x1[i], y1[i], color=colors[index], label=g)
+        index += 1
+    ax1.legend()
+    ax1.set_yscale(scale)
+
+    index = 0
+    i = np.where(np.logical_or(x == 'rMat28', np.logical_or(x == 'rMat27', np.logical_or(x == 'twitter', x == 'friendster'))))
+    x1 = x[i]
+    c1 = color[i]
+    y1 = y[i]
+    for g in np.unique(c1):
+        i = np.where(c1 == g)
+        ax2.scatter(x1[i], y1[i], color=colors[index], label=g)
+        index += 1
+    ax1.legend()
[…]

### 2020-08-11 scripts/benchmark/benchmark.sh
updated benchmark for giraph

@@ -368,7 +368,8 @@ giraph-sssp () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    log "giraph-sssp $1 $2 $dur_init $dur_calc $dur_exec"
+    local dur_calc_giraph=$(grep 'Superstep ' $result | awk -F '=' '{s += $2} END {print s*1000}')
+    log "giraph-sssp $1 $2 $dur_init $dur_calc $dur_exec $dur_calc_giraph"
 }
 
 giraph-sssp-dist () {
@@ -386,7 +387,8 @@ giraph-sssp-dist () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    log "giraph-sssp-dist $1 $2 $dur_init $dur_calc $dur_exec"
+    local dur_calc_giraph=$(grep 'Superstep ' $result | awk -F '=' '{s += $2} END {print s*1000}')
+    log "giraph-sssp-dist $1 $2 $dur_init $dur_calc $dur_exec $dur_calc_giraph"
 }
 
 giraph-pagerank () {
@@ -401,7 +403,8 @@ giraph-pagerank () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    log "giraph-pagerank $1 $2 $pagerank_number_of_iterations $dur_calc $dur_exec"
+    local dur_calc_giraph=$(grep 'Superstep ' $result | awk -F '=' '{s += $2} END {print s*1000}')
+    log "giraph-pagerank $1 $2 $pagerank_number_of_iterations $dur_calc $dur_exec $dur_calc_giraph"
 }
 
 giraph-pagerank-dist () {
@@ -416,5 +419,6 @@ giraph-pagerank-dist () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    log "giraph-pagerank-dist $1 $2 $pagerank_number_of_iterations $dur_calc $dur_exec"
[…]

### 2020-08-11 scripts/benchmark/benchmark.sh
fixed giraph benchmark

@@ -368,7 +368,7 @@ giraph-sssp () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    local dur_calc_giraph=$(grep 'Superstep ' $result | awk -F '=' '{s += $2} END {print s*1000}')
+    local dur_calc_giraph=$(grep 'Superstep ' "$result" | awk -F '=' '{s += $2} END {print s*1000}')
     log "giraph-sssp $1 $2 $dur_init $dur_calc $dur_exec $dur_calc_giraph"
 }
 
@@ -377,7 +377,7 @@ giraph-sssp-dist () {
     $($HADOOP_HOME/bin/hadoop dfs -rmr /output/$graph)
     local time_start=$(get_time)
     # hardcoded for our current setup
-    local result=$(timeout 3h $HADOOP_HOME/bin/hadoop jar $GIRAPH_HOME/giraph-examples/target/giraph-examples-1.3.0-SNAPSHOT-for-hadoop-1.2.1-jar-with-dependencies.jar org.apache.giraph.GiraphRunner org.apache.giraph.examples.GeneralShortestPathsComputation $2 -vif org.apache.giraph.io.formats.JsonLongDoubleFloatDoubleVertexInputFormat -vip /input/$graph -vof org.apache.giraph.io.formats.IdWithValueTextOutputFormat -op /output/$graph -w 4 2>&1 | ts '%.s')
+    local result=$(timeout 3h $HADOOP_HOME/bin/hadoop jar $GIRAPH_HOME/giraph-examples/target/giraph-examples-1.3.0-SNAPSHOT-for-hadoop-1.2.1-jar-with-dependencies.jar org.apache.giraph.GiraphRunner org.apache.giraph.examples.GeneralShortestPathsComputation $2 -vif org.apache.giraph.io.formats.JsonLongDoubleFloatDoubleVertexInputFormat -vip /input/$graph -vof org.apache.giraph.io.formats.IdWithValueTextOutputFormat -op /output/$graph -w 5 2>&1 | ts '%.s')
     local time_finish=$(get_time)
     local dur_exec=$(($time_finish-$time_start))
     logv "$result"
@@ -387,7 +387,7 @@ giraph-sssp-dist () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
[…]

### 2020-08-11 scripts/benchmark/benchmark.sh
.

@@ -368,7 +368,7 @@ giraph-sssp () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    local dur_calc_giraph=$(grep 'Superstep ' "$result" | awk -F '=' '{s += $2} END {print s*1000}')
+    local dur_calc_giraph=$(echo "$result" | grep 'Superstep ' | awk -F '=' '{s += $2} END {print s*1000}')
     log "giraph-sssp $1 $2 $dur_init $dur_calc $dur_exec $dur_calc_giraph"
 }
 
@@ -387,7 +387,7 @@ giraph-sssp-dist () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    local dur_calc_giraph=$(grep 'Superstep ' "$result" | awk -F '=' '{s += $2} END {print s*1000}')
+    local dur_calc_giraph=$(echo "$result" | grep 'Superstep ' | awk -F '=' '{s += $2} END {print s*1000}')
     log "giraph-sssp-dist $1 $2 $dur_init $dur_calc $dur_exec $dur_calc_giraph"
 }
 
@@ -403,7 +403,7 @@ giraph-pagerank () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
-    local dur_calc_giraph=$(grep 'Superstep ' "$result" | awk -F '=' '{s += $2} END {print s*1000}')
+    local dur_calc_giraph=$(echo "$result" | grep 'Superstep ' | awk -F '=' '{s += $2} END {print s*1000}')
     log "giraph-pagerank $1 $2 $pagerank_number_of_iterations $dur_calc $dur_exec $dur_calc_giraph"
 }
 
@@ -419,6 +419,6 @@ giraph-pagerank-dist () {
     local start_calc=$(echo "$result" | grep 'Running job' | awk '{print $1}')
     local start_calc=$(convert_time $start_calc)
     local dur_calc=$(($time_finish-$start_calc))
[…]

### 2020-08-12 scripts/benchmark/benchmark_evaluation.sh
more benchmark results

@@ -23,6 +23,8 @@ for graph in ${graphs[@]}; do
   for algorithm in ${algorithms[@]}; do
     if [[ "$algorithm" == *gemini* ]]; then
       result=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " (s2-5*s1)/10000000}')
+    elif [[ "$algorithm" == *giraph* ]]; then
+      result=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$7} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
     else
       result=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
     fi

### 2020-08-13 documentation/paper/sections/graphFormats.tex
clean up
giraph overview and input format

--- 2020-08-10 documentation/paper/graphFormats.tex
@@
-%!TEX root = ./main.tex
+%!TEX root = ../main.tex
 
@@
 The numbers are all separated by newlines.
-\begin{gather*}
-n\\
-m\\
-o_1\\
-o_2\\
-\vdots\\
-o_n\\
-t_1\\
-t_2\\
-\vdots\\
+\begin{gather*}%with hacked line spacing to save some space..
+n\\[-1\jot]
+m\\[-1\jot]
+o_1\\[-1\jot]
+o_2\\[-1\jot]
+\vdots\\[-1\jot]
+o_n\\[-1\jot]
+t_1\\[-1\jot]
+t_2\\[-1\jot]
+\vdots\\[-1\jot]
 t_m
@@
 
-\subsubsection{Giraph's numerous I/O formats}
+\subsubsection{Giraph's I/O formats}
+Giraph is capable of parsing many different input and output formats. All of those are explained in Giraph's JavaDoc.
+Both edge- and vertex-centric input formats are possible.
+
+One can even define their own input graph representation or output format.
+For the purposes of this paper, we used the existing \texttt{JsonLongDoubleFloatDoubleVertexInputFormat}\cite{GiraphInputFormat}.
+Our graph conversion tool outputs this format.
+
+In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values, \texttt{float} out-edge weights and the messages are of type \texttt{double} (all are Java data types).
+All this is in a JSON-like format, so each line in the graph file looks as follows
+\begin{equation*}
+	[s,v_s,[[t_1, w_{t_1}], [t_2, w_{t_2}]...]]
+\end{equation*}
+with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.
+
+There is surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

### 2020-08-13 documentation/paper/sections/overview/giraph.tex
giraph overview and input format

%!TEX root=../../main.tex

Apache Giraph\cite{Giraph} is an iterative graph processing framework, built on top of Apache Hadoop. Because of this, expanding shared memory processing to a multi-node cluster is almost seamless.

Giraph is inspired by the Bulk Synchronous Parallel model of distributed computation introduced by Leslie Valiant\cite{BSPforGiraph}. Meaning that it is based on computation units that communicate using messages and are synchonized with barriers.

The input to a Giraph computation is always a directed graph. Not only the edges but also the vertices have a value attached to them. The graph topology is thus not only defined by the vertices and edges but also their initial values.
Furthermore, one can mutate the graph by adding or removing vertices and edges during computation.

Computation is vertex oriented and iterative.
For each iteration step called superstep, the Compute method implementing the algorithm is invoked on each active vertex, with every vertex being active in the beginning.
This method receives messages sent in the previous superstep as well as its vertex value and the values of outgoing edges.
With this data the values are modified and messages to other vertices are sent.
Communication between vertices is only performed via messages, so a vertex has no access to values of other vertices or edges other than it's own outgoing ones.

Supersteps are synchronized with barriers, meaning that all messages only get delivered in the following superstep and computation for the next superstep can only begin after every vertex has finished computing the current superstep.
Edge and vertex values are retained across supersteps.

Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will make the vertex active again.
To end computation, a vote-to-halt method is applied. Each vertex outputs some local information (e.g. the final vertex value) as result.

Giraph is next to Galois probably the most actively developed framework in our comparison.

### 2020-08-14 scripts/benchmark/benchmark_cases.sh
more benchmark results and cases

@@ -488,10 +488,20 @@ benchmark_giraph_sssp() {
   done
 }
 
+benchmark_giraph_pagerank_klein() {
+  for startnode in ${startnodes_flickr[@]}; do
+    giraph-pagerank-dist flickr
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    giraph-pagerank-dist orkut
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    giraph-pagerank-dist wikipedia
+  done
+}
+
 
 time_start=$(get_time)
-#benchmark_galois_dist_rMat27
-#benchmark_galois_dist_rMat28
-benchmark_giraph_sssp
+benchmark_giraph_pagerank_klein
 dur_exec=$((($(get_time)-$time_start)/1000000))
 echo "benchmark took $dur_exec seconds"

### 2020-08-14 documentation/paper/main.tex
giraph overview and input format
added measurement methods for times

@@
 \bibliography{literature.bib}
-\title{Forschungsprojekt}
+\title{A Comparison of Graph Processing Systems}
 \author{\IEEEauthorblockN{Simon König}
@@
 \begin{IEEEkeywords}
-graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon
+graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon, Gemini
 \end{IEEEkeywords}
@@
 
+\subsection{Gemini}
+\input{sections/overview/gemini}
+
+\subsection{Giraph}
+\input{sections/overview/giraph}
+
 \subsection{Ligra}
@@
 \input{sections/overview/polymer}
-
-\subsection{Gemini}
-\input{sections/overview/gemini}
-
-\subsection{Giraph}
-\input{sections/overview/giraph}
 
@@
 \subsection{Komplexität Aufsetzen}
-Systeme wie distr. Giraph bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden. \emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
+Systeme wie distr.
+Galois, bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden.
+\emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
 
@@
 \subsection{Ergebnisse exec time}
+
 hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
@@
 
-Also we use Gluon~\cite{vertGalois} for the distributed setups.
+Also we use Gluon~\cite{vertGalois} for the distributed Galois setups.
 
 Gemini~\cite{Gemini}
+
+%We would like to thank our supervisor Heiko Geppert for continued guidance and support.
 
@@
 
-Everything can be retrieved on our GitHub repository
-\url{http://www.github.com/serengti/Forschungsprojekt}.
+Our GitHub repository: \url{http://www.github.com/serengti/Forschungsprojekt}.
 
-For each Framework, there is a
+The graphs were downloaded from the graph database associated with the Koblenz Network Collection (KONECT)\cite{konect}.
 
@@
 
+\bibitem{konect}
+J.~Kunegis, ``Konect: the koblenz network collection,'' 05 2013, pp.
+  1343--1350.
+
 \end{thebibliography}

### 2020-08-14 documentation/paper/sections/testingMethods.tex
clean up
added measurement methods for times

--- 2020-08-10 documentation/paper/testingMethods.tex
@@
-%!TEX root=main.tex
+%!TEX root=../main.tex
+
 \subsection{Hardware and Software}
 For testing all systems we used 5 machines with 96 CPU cores each (48 physical) and 256 GB of RAM.
-One of those machines only had 128 GB of RAM, this one was only used as part of the calculation cluster for the distributed systems.
+One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
 All servers were running Ubuntu 19.x.
@@
 
-\subsection{Benchmark setup}
-We measured the total runtime of each process as well as the actual calculation times.
-The console output of each frameworks was used as an indicator for when loading the graph data was finished.
-This proved to be more difficult than expected in some cases.
-
-Each test case consisting of graph, framework and in some cases even the algorithm was run 10 times, allowing us to smooth slight variations in the measured times.
-This also helps to reduce the error we introduced by measuring time through the console log.
-
 All benchmarks were initiated by our benchmark script that is available in our repository.
 
-We will compare both the computation times as well es the execution times. This way we compare how much setup time each framework has.
+The complete log files and extracted raw results will be available in our repository as well.
+
+\subsection{Measurements}
+For every framework, we measured the \emph{execution time} as the time from start to finish of the console command.
+
+For the \emph{calculation time}, we tried to extract only the time the framework actually executed the algorithm. We came up with the following:
+\begin{itemize}
+	\item For Galois, we resulted to extracting console log time stamps. Galois outputs \texttt{Reading graph complete.}. Calculation time is the time from this output to the end of execution.
+	\item Polymer outputs the name of the algorithm followed by a measured time.
+	\item Gemini outputs a line \texttt{exec\_time=x}, which was used to measure the calculation time. 
+	\item Ligra outputs a measured time with \texttt{Running time : x}.
+	\item Giraph has built in timers for the iterations (supersteps) which we used to extract the computation time.
+\end{itemize}
+Each test case consisting of graph, framework and algorithm was run 10 times, allowing us to smooth slight variations in the measured times.

### 2020-08-14 scripts/benchmark/benchmark_cases.sh
more benchmark results and cases

@@ -488,7 +488,7 @@ benchmark_giraph_sssp() {
   done
 }
 
-benchmark_giraph_pagerank_klein() {
+benchmark_giraph_pagerank_dist_klein() {
   for startnode in ${startnodes_flickr[@]}; do
     giraph-pagerank-dist flickr
   done
@@ -500,8 +500,20 @@ benchmark_giraph_pagerank_klein() {
   done
 }
 
+benchmark_giraph_sssp_dist_klein() {
+  for startnode in ${startnodes_flickr[@]}; do
+    giraph-sssp-dist flickr $startnode
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    giraph-sssp-dist orkut $startnode
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    giraph-sssp-dist wikipedia $startnode
+  done
+}
+
 
 time_start=$(get_time)
-benchmark_giraph_pagerank_klein
+benchmark_giraph_sssp_dist_klein
 dur_exec=$((($(get_time)-$time_start)/1000000))
 echo "benchmark took $dur_exec seconds"

### 2020-08-14 documentation/paper/main.tex
testing methods, giraph overview

@@
 %\usepackage{csquotes}
+\usepackage{textcomp}
+\usepackage{pifont}
 \usepackage{hyperref}
 \usepackage{amsmath}
-
-\newcommand{\todo}{{\color{red}TODO}}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{multirow}
+\usepackage{booktabs}
+\usepackage{graphics}
+\newcommand*\rot{\rotatebox{90}}
+\newcommand*\OK{\ding{51}}
 
 \bibliography{literature.bib}
+
 \title{A Comparison of Graph Processing Systems}
+
 \author{\IEEEauthorblockN{Simon König}
@@
 @stud.uni-stuttgart.de}}
+
 \date{\today}
@@
 \section{Introduction}
-%\todo
 
@@
 
-Hier sollten wir auch evtl Vergleichsgrundlagen für Aufsetzen und Apps schreiben einführen, wenn wir das mit aufnehmen.
-
 \section{Results}
 \subsection{Komplexität Aufsetzen}
-Systeme wie distr.
-Galois, bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden.
-\emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
+Systeme wie Galois/Gluon, bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden. \emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
 
@@
 We have written a number of conversion tools and installation guides to help
-users or develeopers with the use of the tested frameworks.
+users or developers with the use of the tested frameworks.
 
 Our GitHub repository: \url{http://www.github.com/serengti/Forschungsprojekt}.
-
-The graphs were downloaded from the graph database associated with the Koblenz Network Collection (KONECT)\cite{konect}.
 

### 2020-08-14 documentation/paper/sections/overview/giraph.tex
testing methods, giraph overview

@@
 With this data the values are modified and messages to other vertices are sent.
-Communication between vertices is only performed via messages, so a vertex has no access to values of other vertices or edges other than it's own outgoing ones.
+Communication between vertices is only performed via messages, so a vertex has no access to values of other vertices or edges other than its own outgoing ones.
 

### 2020-08-14 documentation/paper/sections/testingMethods.tex
testing methods, giraph overview

@@
 
-All benchmarks were initiated by our benchmark script that is available in our repository.
+All benchmarks were initiated by our benchmark script that is available in our repository. Galois, Gemini and Giraph were benchmarked on both the 5-node cluster and a singe machine.
+Furthermore, -- since Galois supports this parameter -- we ran multiple tests comparing Galois' performance with different core counts on a single machine.
 
@@
 Each test case consisting of graph, framework and algorithm was run 10 times, allowing us to smooth slight variations in the measured times.
+
+\subsection{Graphs}
+The graphs used in our testing can be seen in detail in \autoref{tbl:graphs}.
+We included a variaty of different graph sizes, from relatively small graphs like the flickr graph up to an rMat28 with $4.2$ billion edges.
+All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
+Both the rMat27 and rMat28 were created with a modified version (we changed the output format to EdgeList) of a graph generator provided by Ligra.
+\begin{table}
+	\caption{Size comparison of the used graphs}
+	\label{tbl:graphs}
+	\centering
+	\begin{tabular}{crr}
+		\hline
+		\bf{Graph}&\# Vertices (M)&\# Edges (M)\\\hline
+		flickr&    		0.1&  2\\
+		orkut&          3&    117\\
+		wikipedia&      12&   378\\
+		twitter&     	52&   1963\\
+		rMat27&         63&   2147\\
+		friendster&     68&   2586\\
+		rMat28&         121&  4294\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{Algorithms}
+We ran Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) algorithms on each framework and graph.
+For frameworks that support multiple implementations (e.g. PR in push and pull modes), we included the alternatives in our testing.
+
+\begin{table}
+	\caption{Natively supported algorithms of the Frameworks}
+	\label{tbl:algorithms}
+	\centering
+	\begin{tabular}{ccccccc}
+		\multicolumn{2}{c}{\bf{Algorithm}}&\rot{Galois}&\rot{Gemini}&\rot{Giraph}&\rot{Ligra}&\rot{Polymer}\\\hline
+		\multirow{3}{*}{\rot{PageRank}}	&Push&\OK&&&?&\OK\\
+										&Pull&\OK&&&?&\\
+										&Delta&?&&&\OK&\OK\\
+		\hline
+	\end{tabular}
+\end{table}
+
+\subsection{Comparison of setup}
+
+\subsection{Implementing new application}

### 2020-08-15 documentation/paper/sections/overview/polymer.tex
clean up
polymer overview

--- 2020-06-22 documentation/paper/polymer.tex
@@
-%!TEX root=main.tex
+%!TEX root=../../main.tex
 
-\subsection{Polymer}
-Installing Polymer is straight forward.
+Polymer is very similar to Ligra, in fact Polymer inherits the programming interfaces \texttt{EdgeMap} and \texttt{VertexMap} from Ligra as its main interface.
+
+Polymer is a vertex-centric framework, that tries to circumvent some of the random memory access drawbacks of such a design.
+It treats a NUMA machine as a distributed cluster and splits work and graph data accordingly between the nodes.\cite[p.187]{Polymer}
+Application-defined data is not distributed. Other runtime state data is allocated in a distributed way but only accessed through a global lookup table.\cite[p.184]{Polymer}

### 2020-08-15 scripts/benchmark/benchmark_cases.sh
more benchmark results and cases

@@ -4,9 +4,9 @@ source benchmark.sh
 path_to_graphs=/home/fp-ss20/graph
 #graph_info=/home/ubuntu/graph/graph_info.txt
 path_to_bins=/home/fp-ss20/bin
-#host_file=/home/ubuntu/host_file
+host_file=/home/ubuntu/host_file
 #result_file=$4
-threads=1
+threads=48
 #hosts=129.69.210.223,129.69.210.224
 hosts=129.69.210.223,129.69.210.224,129.69.210.225,129.69.210.226,129.69.210.227
 pagerank_number_of_iterations=5
@@ -512,8 +512,46 @@ benchmark_giraph_sssp_dist_klein() {
   done
 }
 
+benchmark_gemini_dist() {
+  for startnode in ${startnodes_flickr[@]}; do
+    gemini-sssp-dist flickr $startnode $nodes_flickr
+    gemini-bfs-dist flickr $startnode $nodes_flickr
+    gemini-pagerank-dist flickr $nodes_flickr
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    gemini-sssp-dist orkut $startnode $nodes_orkut
+    gemini-bfs-dist orkut $startnode $nodes_orkut
+    gemini-pagerank-dist orkut $nodes_orkut
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    gemini-sssp-dist wikipedia $startnode $nodes_wikipedia
+    gemini-bfs-dist wikipedia $startnode $nodes_wikipedia
+    gemini-pagerank-dist wikipedia $nodes_wikipedia
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    gemini-sssp-dist twitter $startnode $nodes_twitter
+    gemini-bfs-dist twitter $startnode $nodes_twitter
+    gemini-pagerank-dist twitter $nodes_twitter
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    gemini-sssp-dist friendster $startnode $nodes_friendster
+    gemini-bfs-dist friendster $startnode $nodes_friendster
+    gemini-pagerank-dist friendster $nodes_friendster
+  done
+  for startnode in ${startnodes_rMat27[@]}; do
+    gemini-sssp-dist rMat27 $startnode $nodes_rMat27
+    gemini-bfs-dist rMat27 $startnode $nodes_rMat27
[…]

### 2020-08-16 documentation/paper/main.tex
polymer overview
grouped bar plot, style for plots

@@
 \usepackage[utf8]{inputenc}
-%\usepackage{csquotes}
+\usepackage{csquotes}
 \usepackage{textcomp}
@@
 \usepackage{booktabs}
-\usepackage{graphics}
+\usepackage{graphicx}
 \newcommand*\rot{\rotatebox{90}}
@@
 \section{Results}
-\subsection{Komplexität Aufsetzen}
-Systeme wie Galois/Gluon, bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden. \emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
-
-\subsection{Komplexität eigene Apps schreiben}
-Schwierig objektiv zu vergleichen.
-
-\subsection{Pure Performance-Ergebnisse}
-
-\subsection{Ergebnisse Calc time}
-der offensichtliche, wichtige Vergleich
-
-\subsection{Ergebnisse exec time}
-
-hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
-Aber vielleicht ist auch die setup time bei gleichen frameworks zwischen verteilt und shared memory ganz interessant zu vergleichen.
+\input{sections/results/results}
 

### 2020-08-16 documentation/paper/sections/results/results.tex
grouped bar plot, style for plots

%!TEX root=../../main.tex

\subsection{Komplexität Aufsetzen}
Systeme wie Galois/Gluon, bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden. \emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}

\subsection{Komplexität eigene Apps schreiben}
Schwierig objektiv zu vergleichen.

\subsection{Pure Performance-Ergebnisse}

\subsection{Ergebnisse Calc time}
\begin{figure}
	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP.png}
	\caption{Calculation times for SSSP on a single node}
	\label{fig:ssspSingleNode}
\end{figure}
der offensichtliche, wichtige Vergleich

\subsection{Ergebnisse exec time}

hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
Aber vielleicht ist auch die setup time bei gleichen frameworks zwischen verteilt und shared memory ganz interessant zu vergleichen.

### 2020-08-16 plots/makeplots.py

@@ -1,0 +1,64 @@
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+
+plt.style.use('./style.mplstyle')
+
+#[graph, algo, calc, exec]
+data = []
+with open('../results.txt') as f:
+    for line in f:
+        data.append([])
+        for entry in line.split():
+            data[-1].append(entry)
+
+graph = np.empty(len(data), dtype=object)
+algo = np.empty(len(data), dtype=object)
+calcTime = np.empty(len(data))
+totalTime = np.empty(len(data))
+
+for i in range(len(data)):
+    graph[i] = data[i][0]
+    algo[i] = data[i][1]
+    try:
+        calcTime[i] = float(data[i][2])
+    except ValueError:
+        calcTime[i] = float('nan')
+    try:
+        totalTime[i] = float(data[i][3])
+    except ValueError:
+        totalTime[i] = float('nan')
+
+
+
+colors = ['tab:blue', 'tab:orange', 'tab:green', 'tab:red', 'tab:purple']
+
+#sorted by size of graph
+graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
+#sorted alphabetically
+frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
+
+
+calcTimesSSSPByFramework = []
+execTimesSSSPByFramework = []
+
+for f in frameworks:
+	tmpExec = []
+	tmpCalc = []
+	fname = f.lower() + "-sssp"
+	if f == 'Galois':
+		fname += "-cpu-96thread"
+	for g in graphs:
+		for i in range(len(graph)):
+			if g == graph[i] and fname == algo[i]:
+				tmpCalc.append(calcTime[i])
+				tmpExec.append(totalTime[i])
+	calcTimesSSSPByFramework.append(tmpCalc)
+	execTimesSSSPByFramework.append(tmpExec)
+
+
+from plotFunctions import grouped_bar_plot
+
+grouped_bar_plot(graphs, frameworks, calcTimesSSSPByFramework, title='Single-source Shortest-path on one calculation node', yLabel='Calculation times (s)', saveToFile="singleNodeSSSP.png")
+

### 2020-08-16 plots/plotFunctions.py

@@ -1,0 +1,46 @@
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+
+
+#
+# values = [[f1_g1,f1_g2,...],[f2_g1,f2_g2,...],...]
+def grouped_bar_plot(group_names, legend_names, values, colors=None, width=0.35, scaleFactor=None, title='', yLabel='', yScale='log', saveToFile=None):
+	group_count = len(group_names)
+
+	if len(legend_names) != len(values):
+		raise "Size mismatch: Expected " + len(legend_names) + " categories, got " + len(values)
+
+
+	if scaleFactor == None:
+		scaleFactor = group_count / len(legend_names) + 1
+
+	fig, ax = plt.subplots()
+
+	ind = np.arange(group_count)
+
+	ps = []
+	i = 0
+	for element in values: 
+		tmp = element.copy()
+		while group_count > len(tmp):
+			tmp.append(float('nan'))
+		ps.append(ax.bar(scaleFactor * ind + i * width, tmp, width, bottom=0))
+		i += 1
+
+	ax.set_xticks(scaleFactor * ind + (len(legend_names)-1) / 2 * width)
+	ax.set_xticklabels(group_names)
+	ax.set_yscale(yScale)
+	ax.set_title(title)
+	ax.set_ylabel(yLabel)
+
+	ax.legend(ps, legend_names)
+
+	if saveToFile != None:
+		plt.savefig(saveToFile)
+	else:
+		plt.show()
+
+
+

### 2020-08-17 plots/makeplots.py
line plot for galois threads

@@ -2,8 +2,22 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib
-
 plt.style.use('./style.mplstyle')
+
+
+def insertAt(idx, val, list):
+    if idx < 0:
+        return
+    if idx >= len(list):
+        while idx > len(list):
+            list.append(float('nan'))
+        list.append(val)
+        return
+    else:
+        list[idx] = val
+        return 
+
+
 
 #[graph, algo, calc, exec]
 data = []
@@ -32,8 +46,6 @@ for i in range(len(data)):
 
 
 
-colors = ['tab:blue', 'tab:orange', 'tab:green', 'tab:red', 'tab:purple']
-
 #sorted by size of graph
 graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
 #sorted alphabetically
@@ -48,7 +60,7 @@ for f in frameworks:
 	tmpCalc = []
 	fname = f.lower() + "-sssp"
 	if f == 'Galois':
-		fname += "-cpu-96thread"
+		fname += "-cpu-96thread" # select Galois thread count here
 	for g in graphs:
 		for i in range(len(graph)):
 			if g == graph[i] and fname == algo[i]:
@@ -58,7 +70,43 @@ for f in frameworks:
 	execTimesSSSPByFramework.append(tmpExec)
 
 
-from plotFunctions import grouped_bar_plot
 
-grouped_bar_plot(graphs, frameworks, calcTimesSSSPByFramework, title='Single-source Shortest-path on one calculation node', yLabel='Calculation times (s)', saveToFile="singleNodeSSSP.png")
+calcTimesGaloisByCPUCount = {}
+x = []
+for i in range(len(graph)):
+    if not 'galois-sssp' in algo[i]:
+        continue
+    if '-dist' in algo[i]:
+        continue
 
+    if not graph[i] in calcTimesGaloisByCPUCount:
+        calcTimesGaloisByCPUCount[graph[i]] = []
+
+    xVal = int(algo[i][16:-6]) # galois-sssp- und thread abschneiden
+    if not xVal in x:
+        x.append(xVal)
+
+x.sort()
+for i in range(len(graph)):
+    if not 'galois-sssp' in algo[i]:
+        continue
+    if '-dist' in algo[i]:
+        continue
+
+    xVal = int(algo[i][16:-6])
+
+    idx = -1
+    for j in range(len(x)):
+        if xVal == x[j]:
[…]

### 2020-08-17 plots/plotFunctions.py

@@ -1,16 +1,16 @@
 
 import numpy as np
+import matplotlib.patches as mpatches
 import matplotlib.pyplot as plt
 import matplotlib
 
-
 #
 # values = [[f1_g1,f1_g2,...],[f2_g1,f2_g2,...],...]
-def grouped_bar_plot(group_names, legend_names, values, colors=None, width=0.35, scaleFactor=None, title='', yLabel='', yScale='log', saveToFile=None):
+def grouped_bar_plot(group_names, legend_names, values, width=0.35, scaleFactor=None, title='', yLabel='', yScale='log', saveToFile=None):
 	group_count = len(group_names)
 
 	if len(legend_names) != len(values):
-		raise "Size mismatch: Expected " + len(legend_names) + " categories, got " + len(values)
+		raise "Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values))
 
 
 	if scaleFactor == None:
@@ -35,6 +35,38 @@ def grouped_bar_plot(group_names, legend_names, values, colors=None, width=0.35,
 	ax.set_title(title)
 	ax.set_ylabel(yLabel)
 
+	ax.legend(ps, legend_names)
+
+	if saveToFile != None:
+		plt.savefig(saveToFile)
+	else:
+		plt.show()
+
+
+#
+# values is a dictionary
+def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=None, saveToFile=None):
+
+	if len(legend_names) != len(values):
+		raise "Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values))
+
+	fig, ax = plt.subplots()
+
+	ps = []
+	for element in legend_names: 
+		tmp = values[element].copy()
+		while len(x) > len(tmp):
+			tmp.append(float('nan'))
+		l, = ax.plot(x, tmp, 'o--')
+		ps.append(l)
+
+	if yScale != None:
+		ax.set_yscale(yScale)
+	ax.set_title(title)
+	ax.set_ylabel(yLabel)
+	ax.set_xlabel(xLabel)
+
+	
 	ax.legend(ps, legend_names)
 
 	if saveToFile != None:

### 2020-08-18 documentation/paper/main.tex

--- 2020-08-14 documentation/paper/main.tex
@@
 \bibliography{literature.bib}
-\title{A Comparison of Graph Processing Systems}
+\title{Forschungsprojekt}
 \author{\IEEEauthorblockN{Simon König}
@@
 \begin{IEEEkeywords}
-graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon, Gemini
+graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon
 \end{IEEEkeywords}
@@
 
+\subsection{Ligra}
+\input{sections/overview/ligra}
+
+\subsection{Polymer}
+\input{sections/overview/polymer}
+
 \subsection{Gemini}
@@
 \input{sections/overview/giraph}
-
-\subsection{Ligra}
-\input{sections/overview/ligra}
-
-\subsection{Polymer}
-\input{sections/overview/polymer}
 
@@
 \subsection{Komplexität Aufsetzen}
-Systeme wie distr.
-Galois, bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden.
-\emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
+Systeme wie distr. Giraph bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden. \emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
 
@@
 \subsection{Ergebnisse exec time}
-
 hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
@@
 
-Also we use Gluon~\cite{vertGalois} for the distributed Galois setups.
+Also we use Gluon~\cite{vertGalois} for the distributed setups.
 
 Gemini~\cite{Gemini}
-
-%We would like to thank our supervisor Heiko Geppert for continued guidance and support.
 
@@
 
-Our GitHub repository: \url{http://www.github.com/serengti/Forschungsprojekt}.
+Everything can be retrieved on our GitHub repository
+\url{http://www.github.com/serengti/Forschungsprojekt}.
 
-The graphs were downloaded from the graph database associated with the Koblenz Network Collection (KONECT)\cite{konect}.
+For each Framework, there is a
+
+The graphs are downloaded from \cite{konect}.
 

### 2020-08-19 documentation/paper/sections/overview/gemini.tex
add gemini writeup (needs review)

%!TEX root=../../main.tex

Gemini\cite{Gemini} is a framework for parallel graph processing. In comparrison to the other NUMA-aware frameworks we discuss here Gemini does's not support shared memory parallel processing, it is completely message based using MPI. It seems to be very lightweight and has only the very basically needed functionallity implemented, which leeds to clear source code and small binarys. A basic API hides data and computation distribution details from the user writing applications. It comes with five algorithms already implemented included the three we are testing in this paper. While setting it up and even while benchmarking we had various problems with bugs in the source code, like non zero terminated strings and missing return statements. While debugging we forked the origional repository and made our changes there. DODO link repository.

### 2020-08-19 scripts/benchmark/benchmark_evaluation.sh
evaluation pipeline and variance of results

@@ -22,14 +22,23 @@ done
 for graph in ${graphs[@]}; do
   for algorithm in ${algorithms[@]}; do
     if [[ "$algorithm" == *gemini* ]]; then
-      result=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " (s2-5*s1)/10000000}')
+      means=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " (s2-5*s1)/10000000}')
+      mean1=$(echo "$means" | awk '{print $1}')
+      mean2=$(echo "$means" | awk '{print $2}')
+      vars=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk -v m1="$mean1" -v m2="$mean2" '{s1+=($5/1000000-m1)**2} {s2+=(($6-5*$5)/1000000-m2)**2} END {print s1/9 " " s2/9}')
     elif [[ "$algorithm" == *giraph* ]]; then
-      result=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$7} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
+      means=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$7} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
+      mean1=$(echo "$means" | awk '{print $1}')
+      mean2=$(echo "$means" | awk '{print $2}')
+      vars=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk -v m1="$mean1" -v m2="$mean2" '{s1+=($7/1000000-m1)**2} {s2+=($6/1000000-m2)**2} END {print s1/9 " " s2/9}')
     else
-      result=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
+      means=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
+      mean1=$(echo "$means" | awk '{print $1}')
[…]

### 2020-08-20 documentation/paper/sections/overview/polymer.tex
kleine Änderungen

--- 2020-06-22 documentation/paper/polymer.tex
@@
-%!TEX root=main.tex
+%!TEX root=../../main.tex
 
-\subsection{Polymer}
-Installing Polymer is straight forward.
+Polymer is very similar to Ligra, in fact Polymer inherits the programming interfaces \texttt{EdgeMap} and \texttt{VertexMap} from Ligra as its main interface.\cite{Polymer}
+
+Polymer is a vertex-centric framework, that tries to circumvent some of the random memory access drawbacks of such a design.
+It treats a NUMA machine as a distributed cluster and splits work and graph data accordingly between the nodes.
+Application-defined data is not distributed. Other runtime state data is allocated in a distributed way but only accessed through a global lookup table.

### 2020-08-20 documentation/paper/sections/testingMethods.tex
grouped bar plot, style for plots
kleine Änderungen

@@
 
-The complete log files and extracted raw results will be available in our repository as well.
+The complete benchmark log files and extracted raw results will be available in our repository as well.
 
@@
 We ran Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) algorithms on each framework and graph.
-For frameworks that support multiple implementations (e.g. PR in push and pull modes), we included the alternatives in our testing.
+For frameworks that support multiple implementations (e.g. PageRank in push and pull modes), we included the alternatives in our testing.
 

### 2020-08-20 plots/makeplots.py

@@ -105,8 +105,8 @@ for i in range(len(graph)):
 
 from plotFunctions import *
 
-grouped_bar_plot(graphs, frameworks, calcTimesSSSPByFramework, title='Single-source Shortest-path on one calculation node', yLabel='Calculation time (s)', saveToFile="singleNodeSSSP.png")
+#grouped_bar_plot(graphs, frameworks, calcTimesSSSPByFramework, title='Single-source Shortest-path on one calculation node', yLabel='Calculation time (s)', saveToFile="singleNodeSSSP.png")
 
 #grouped_bar_plot(graphs, frameworks, execTimesSSSPByFramework, title='Single-source Shortest-path on one calculation node', yLabel='Execution time (s)', saveToFile="singleNodeSSSP_execTime.png")
 
-#line_plot(graphs, x, calcTimesGaloisByCPUCount, title='Single-source Shortest-path times by thread count', yLabel='Calculation times (s)', xLabel='Thread count', yScale='log')
+line_plot(graphs, x, calcTimesGaloisByCPUCount, title='Single-source Shortest-path times by thread count', yLabel='Calculation times (s)', xLabel='Thread count', yScale='log', saveToFile="singleNodeSSSPGiraphThreads.png")

### 2020-08-20 plots/makeplots.py
new plots with std deviation

@@ -2,6 +2,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib
+from math import sqrt
 plt.style.use('./style.mplstyle')
 
 
@@ -21,7 +22,7 @@ def insertAt(idx, val, list):
 
 #[graph, algo, calc, exec]
 data = []
-with open('../results.txt') as f:
+with open('../results/results.txt') as f:
     for line in f:
         data.append([])
         for entry in line.split():
@@ -31,6 +32,8 @@ graph = np.empty(len(data), dtype=object)
 algo = np.empty(len(data), dtype=object)
 calcTime = np.empty(len(data))
 totalTime = np.empty(len(data))
+yErrCalc = np.empty(len(data))
+yErrExec = np.empty(len(data))
 
 for i in range(len(data)):
     graph[i] = data[i][0]
@@ -43,6 +46,14 @@ for i in range(len(data)):
         totalTime[i] = float(data[i][3])
     except ValueError:
         totalTime[i] = float('nan')
+    try:
+        yErrCalc[i] = float(data[i][4])
+    except ValueError:
+        yErrCalc[i] = float('nan')
+    try:
+        yErrExec[i] = float(data[i][5])
+    except ValueError:
+        yErrExec[i] = float('nan')
 
 
 
@@ -53,11 +64,15 @@ frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
 
 
 calcTimesSSSPByFramework = []
-execTimesSSSPByFramework = []
+yErrCalcSSSPByFramework = []
 
+execTimesSSSPByFramework = []
+yErrExecSSSPByFramework = []
 for f in frameworks:
 	tmpExec = []
 	tmpCalc = []
+	tmpYErrCalc = []
+	tmpYErrExec = []
 	fname = f.lower() + "-sssp"
 	if f == 'Galois':
 		fname += "-cpu-96thread" # select Galois thread count here
@@ -65,13 +80,19 @@ for f in frameworks:
 		for i in range(len(graph)):
 			if g == graph[i] and fname == algo[i]:
 				tmpCalc.append(calcTime[i])
+				tmpYErrCalc.append(sqrt(yErrCalc[i]))
+
 				tmpExec.append(totalTime[i])
+				tmpYErrExec.append(sqrt(yErrExec[i]))
 	calcTimesSSSPByFramework.append(tmpCalc)
+	yErrCalcSSSPByFramework.append(tmpYErrCalc)
[…]

### 2020-08-20 plots/plotFunctions.py

@@ -6,7 +6,7 @@ import matplotlib
 
 #
 # values = [[f1_g1,f1_g2,...],[f2_g1,f2_g2,...],...]
-def grouped_bar_plot(group_names, legend_names, values, width=0.35, scaleFactor=None, title='', yLabel='', yScale='log', saveToFile=None):
+def grouped_bar_plot(group_names, legend_names, values, yErrs=None, width=0.35, scaleFactor=None, title='', yLabel='', yScale='log', saveToFile=None):
 	group_count = len(group_names)
 
 	if len(legend_names) != len(values):
@@ -22,16 +22,23 @@ def grouped_bar_plot(group_names, legend_names, values, width=0.35, scaleFactor=
 
 	ps = []
 	i = 0
-	for element in values: 
+	for element in values:
+		yErr = None
+		if not yErrs == None:
+			yErr = yErrs[i].copy()
+			while group_count > len(yErr):
+				yErr.append(float('nan'))
 		tmp = element.copy()
 		while group_count > len(tmp):
 			tmp.append(float('nan'))
-		ps.append(ax.bar(scaleFactor * ind + i * width, tmp, width, bottom=0))
+
+		ps.append(ax.bar(scaleFactor * ind + i * width, tmp, width, yerr=yErr, bottom=0))
 		i += 1
 
 	ax.set_xticks(scaleFactor * ind + (len(legend_names)-1) / 2 * width)
 	ax.set_xticklabels(group_names)
-	ax.set_yscale(yScale)
+	if not yScale == None:
+		ax.set_yscale(yScale)
 	ax.set_title(title)
 	ax.set_ylabel(yLabel)
 
@@ -45,7 +52,7 @@ def grouped_bar_plot(group_names, legend_names, values, width=0.35, scaleFactor=
 
 #
 # values is a dictionary
-def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=None, saveToFile=None):
+def line_plot(legend_names, x, values, yErrs=None, title='', yLabel='', xLabel='', yScale=None, saveToFile=None):
 
 	if len(legend_names) != len(values):
 		raise "Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values))
@@ -55,10 +62,20 @@ def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=No
 	ps = []
[…]

### 2020-08-21 scripts/benchmark/benchmark_cases.sh
more benchmark cases

@@ -6,7 +6,7 @@ path_to_graphs=/home/fp-ss20/graph
 path_to_bins=/home/fp-ss20/bin
 host_file=/home/ubuntu/host_file
 #result_file=$4
-threads=48
+threads=2
 #hosts=129.69.210.223,129.69.210.224
 hosts=129.69.210.223,129.69.210.224,129.69.210.225,129.69.210.226,129.69.210.227
 pagerank_number_of_iterations=5
@@ -550,8 +550,54 @@ benchmark_gemini_dist() {
   done
 }
 
+benchmark_galois_cpu() {
+  for startnode in ${startnodes_flickr[@]}; do
+    galois-sssp-cpu flickr $startnode
+    galois-bfs-cpu flickr $startnode
+    galois-pagerank-push-cpu flickr
+    galois-pagerank-pull-cpu flickr
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    galois-sssp-cpu orkut $startnode
+    galois-bfs-cpu orkut $startnode
+    galois-pagerank-push-cpu orkut
+    galois-pagerank-pull-cpu orkut
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    galois-sssp-cpu wikipedia $startnode
+    galois-bfs-cpu wikipedia $startnode
+    galois-pagerank-push-cpu wikipedia
+    galois-pagerank-pull-cpu wikipedia
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    galois-sssp-cpu twitter $startnode
+    galois-bfs-cpu twitter $startnode
+    galois-pagerank-push-cpu twitter
+    galois-pagerank-pull-cpu twitter
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    galois-sssp-cpu friendster $startnode
+    galois-bfs-cpu friendster $startnode
+    galois-pagerank-push-cpu friendster
+    galois-pagerank-pull-cpu friendster
+  done
+  for startnode in ${startnodes_rMat27[@]}; do
+    galois-sssp-cpu rMat27 $startnode
+    galois-bfs-cpu rMat27 $startnode
+    galois-pagerank-push-cpu rMat27
+    galois-pagerank-pull-cpu rMat27
+  done
+  for startnode in ${startnodes_rMat28[@]}; do
+    galois-sssp-cpu rMat28 $startnode
+    galois-bfs-cpu rMat28 $startnode
[…]

### 2020-08-21 scripts/benchmark/benchmark.sh
fixed benchmark script

@@ -37,7 +37,7 @@ galois-sssp-cpu () {
     local bin="$path_to_bins/galois-sssp-cpu"
     local graph="$path_to_graphs/$1.gr"
     local time_start=$(get_time)
-    local result=$(timeout 3h $bin -t=$threads --runs=1 --startNode=$2 $graph 2> /dev/null | ts '%.s')
+    local result=$(timeout 3h $bin -t=$threads --startNode=$2 $graph 2> /dev/null | ts '%.s')
     local dur_exec=$(($(get_time)-$time_start))
     logv "$result"
     local time_finished_calc=$(echo "$result" | grep 'Verification' | awk '{print $1}')
@@ -79,7 +79,7 @@ galois-pagerank-push-cpu () {
     local bin="$path_to_bins/galois-pagerank-push-cpu"
     local graph="$path_to_graphs/$1.gr"
     local time_start=$(get_time)
-    local result=$(timeout 3h $bin -t=$threads --runs=1 --maxIterations=$pagerank_number_of_iterations $graph 2> /dev/null | ts '%.s')
+    local result=$(timeout 3h $bin -t=$threads --maxIterations=$pagerank_number_of_iterations $graph 2> /dev/null | ts '%.s')
     local dur_exec=$(($(get_time)-$time_start))
     logv "$result"
     local time_finished_calc=$(echo "$result" | grep 'STAT_TYPE' | awk '{print $1}')
@@ -107,7 +107,7 @@ galois-pagerank-pull-cpu () {
     local bin="$path_to_bins/galois-pagerank-pull-cpu"
     local graph="$path_to_graphs/$1.gr"
     local time_start=$(get_time)
-    local result=$(timeout 3h $bin -t=$threads --runs=1 --maxIterations=$pagerank_number_of_iterations --transposedGraph  $graph 2> /dev/null | ts '%.s')
+    local result=$(timeout 3h $bin -t=$threads --maxIterations=$pagerank_number_of_iterations --transposedGraph  $graph 2> /dev/null | ts '%.s')
     local dur_exec=$(($(get_time)-$time_start))
     logv "$result"
     local time_finished_calc=$(echo "$result" | grep 'STAT_TYPE' | awk '{print $1}')
@@ -135,7 +135,7 @@ galois-bfs-cpu () {
     local bin="$path_to_bins/galois-bfs-cpu"
     local graph="$path_to_graphs/$1.gr"
     local time_start=$(get_time)
[…]

### 2020-08-21 documentation/paper/sections/overview/ligra.tex
ligra writeup

%!TEX root=../../main.tex

Ligra\cite{Ligra} is a lightweight parallel graph processing framework for shared memory machines. It offers a programming inteface that allows expressing graph traversal algorithms in a simple way.

Algorithms can use the \texttt{EdgeMap} to make computations based on edges or the \texttt{VertexMap} to make computations based on vertices. Those mappings can be applied only to a subset of vertices. Based on the size of the vertex subset the framework automatically switches between a sparse and a dense representation to optimize speed and memory.

### 2020-08-21 documentation/paper/main.tex
Introduction
add matrikelnumber and mail to paper
ligra writeup
bfs plots and some text for testing methods/results

--- 2020-08-16 documentation/paper/main.tex
@@
 \usepackage{graphicx}
-\newcommand*\rot{\rotatebox{90}}
-\newcommand*\OK{\ding{51}}
+
+\usepackage{tabularx}
+\def\tabularxcolumn#1{m{#1}} % vertical centering in tabularx cells
+\newcolumntype{Y}{>{\centering\arraybackslash}X} % cell type Y for centered text with automatic line breaks
+\newcommand*\rot{\rotatebox{90}} % rotates content 90deg ccw, mainly for table headers
+\newcommand*\OK{\ding{51}} % a checkmark-style sign
 
@@
 \IEEEauthorblockA{(3310069)\\
-@stud.uni-stuttgart.de}
+st154960@stud.uni-stuttgart.de}
 \and
 \IEEEauthorblockN{Jakob Schmid}
-\IEEEauthorblockA{()\\
+\IEEEauthorblockA{(3341630)\\
 @stud.uni-stuttgart.de}}
@@
 \section{Introduction}
+%\todo
+In recent years graph sizes have increased significantly, thus performance and memory efficiency of the graph analysis applications is now more important than ever.
 

### 2020-08-21 documentation/paper/sections/results/results.tex
new plots with std deviation
bfs plots and some text for testing methods/results

@@
 
-\subsection{Komplexität Aufsetzen}
-Systeme wie Galois/Gluon, bei denen man erstmal ewig suchen muss bis man überhaupt einen Guide findet sollten hier schlechter abschneiden. \emph{vorausgesetzt, wir nehmen das hier überhaupt mit auf..}
+\subsection{The frameworks during setup and benchmark}
+We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.
 
-\subsection{Komplexität eigene Apps schreiben}
-Schwierig objektiv zu vergleichen.
+\begin{enumerate}
+	\item During setup of Gemini, we realized that the cloned repository contained several bugs, rendering the code as-is unable to perform any calculations.
+We had to fork and modify the source code.
+TODO(Was genau ging nicht, was haben wir geändert?)
+
+	The repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
+	\item Furthermore, we would like to address the setup of Hadoop for Giraph.
+It requires multiple edits in \texttt{xml} files that aren't easily automized.
+This makes the setup rather time consuming, especially if reconfiguration is needed later on.
+	\item Also in order for Giraph to run, several Java tasks have to be constantly running in the background.
+While we don't expect this to have a noticeable performance impact on other tasks, it is still not optimal.
+	\item The Hadoop distributed file system (HDFS) ran us into disk space problems on multiple occasions.
+First, unless configured otherwise, the standard implementation replicates all data three times, distributed over all participating nodes.
+Second, deleting files on the HDFS does not immediately free up disk space because the files are moved to a \emph{recycling bin}-like location.
+
+	\item 
+\end{enumerate}
+
+On a plus side, setup of frameworks like Polymer or Ligra was straight forward and did not require any special treatment.
 
@@
 \begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP.png}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_calcTime.png}
 	\caption{Calculation times for SSSP on a single node}
-	\label{fig:ssspSingleNode}
+	\label{fig:singleNodeSSSP_calc}
 \end{figure}
-der offensichtliche, wichtige Vergleich
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_calcTime.png}
+	\caption{Calculation times for BFS on a single node}
+	\label{fig:singleNodeBFS_calc}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSPGaloisThreads.png}
+	\caption{Calculation times of Galois' SSSP over the thread count}
+	\label{fig:singleNodeSSSPGaloisThreads}
+\end{figure}
 
 \subsection{Ergebnisse exec time}
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_execTime.png}
+	\caption{Execution times for SSSP on a single node}
+	\label{fig:singleNodeSSSP_exec}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
+	\caption{Normalized Overhead SSSP single node}
+	\label{fig:singleNodeSSSP_overheadNormalized}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTime.png}
+	\caption{Overhead SSSP single node}
+	\label{fig:singleNodeSSSP_overhead}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_execTime.png}
+	\caption{Execution times for BFS on a single node}
+	\label{fig:singleNodeBFS_exec}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTime.png}
+	\caption{Overhead BFS single node}
+	\label{fig:singleNodeBFS_overhead}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
+	\caption{Normalized Overhead BFS single node}
+	\label{fig:singleNodeBFS_overheadNormalized}
+\end{figure}
 

### 2020-08-21 documentation/paper/sections/testingMethods.tex
new plots with std deviation
bfs plots and some text for testing methods/results

@@
 All benchmarks were initiated by our benchmark script that is available in our repository. Galois, Gemini and Giraph were benchmarked on both the 5-node cluster and a singe machine.
-Furthermore, -- since Galois supports this parameter -- we ran multiple tests comparing Galois' performance with different core counts on a single machine.
+Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
 
@@
 For frameworks that support multiple implementations (e.g. PageRank in push and pull modes), we included the alternatives in our testing. 
-
+\autoref{tbl:algorithms} shows the tested algorithms in detail.
+\newcommand{\txtdag}{\textsuperscript{\textdagger}}
 \begin{table}
-	\caption{Natively supported algorithms of the Frameworks}
+	\caption{Tested algorithms of the frameworks}
 	\label{tbl:algorithms}
 	\centering
-	\begin{tabular}{ccccccc}
-		\multicolumn{2}{c}{\bf{Algorithm}}&\rot{Galois}&\rot{Gemini}&\rot{Giraph}&\rot{Ligra}&\rot{Polymer}\\\hline
-		\multirow{3}{*}{\rot{PageRank}}	&Push&\OK&&&?&\OK\\
-										&Pull&\OK&&&?&\\
-										&Delta&?&&&\OK&\OK\\
+	\begin{tabularx}{\columnwidth}{cYYYYY}
+		\bf{Algorithm}&\rot{Galois}&\rot{Gemini}&\rot{Giraph}&\rot{Ligra}&\rot{Polymer}\\\hline
+		PageRank&$\Delta$Push, $\Delta$Pull&&&&\\
+		SSSP&?, Push, Pull&&&&\\
+		BFS&Yes&&*&&\\
 		\hline
-	\end{tabular}
+	\end{tabularx}
+
+	* Algorithm not natively supported
+
 \end{table}
@@
 \subsection{Comparison of setup}
-
-\subsection{Implementing new application}

### 2020-08-21 plots/makeplots.py

@@ -20,7 +20,7 @@ def insertAt(idx, val, list):
 
 
 
-#[graph, algo, calc, exec]
+#[graph, algo, calc, exec, varCalc, varExec]
 data = []
 with open('../results/results.txt') as f:
     for line in f:
@@ -59,23 +59,35 @@ for i in range(len(data)):
 
 #sorted by size of graph
 graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
+#graph sizes in million edges
+graphSize = (2, 117, 378, 1963, 2147, 2586, 4294)
 #sorted alphabetically
 frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
 
+
+
+
+#### SSSP TIMES
 
 calcTimesSSSPByFramework = []
 yErrCalcSSSPByFramework = []
 
 execTimesSSSPByFramework = []
 yErrExecSSSPByFramework = []
+
+overheadTimesSSSPByFramework = []
+overheadTimesSSSPByFrameworkNormalized = []
 for f in frameworks:
 	tmpExec = []
 	tmpCalc = []
 	tmpYErrCalc = []
 	tmpYErrExec = []
+	overhead = []
+	overheadNormalized = []
 	fname = f.lower() + "-sssp"
 	if f == 'Galois':
 		fname += "-cpu-96thread" # select Galois thread count here
+	k = 0
 	for g in graphs:
 		for i in range(len(graph)):
 			if g == graph[i] and fname == algo[i]:
@@ -84,12 +96,62 @@ for f in frameworks:
 
 				tmpExec.append(totalTime[i])
 				tmpYErrExec.append(sqrt(yErrExec[i]))
+
+				overhead.append(totalTime[i] - calcTime[i])
+				overheadNormalized.append((totalTime[i] - calcTime[i])/graphSize[k])
+		k += 1 # graph index
+
 	calcTimesSSSPByFramework.append(tmpCalc)
 	yErrCalcSSSPByFramework.append(tmpYErrCalc)
 	execTimesSSSPByFramework.append(tmpExec)
 	yErrExecSSSPByFramework.append(tmpYErrExec)
+	overheadTimesSSSPByFramework.append(overhead)
+	overheadTimesSSSPByFrameworkNormalized.append(overheadNormalized)
+
+#### SSSP TIMES
+
+calcTimesBFSByFramework = []
[…]

### 2020-08-21 plots/makeplots.py
updated plots with new results

@@ -194,23 +194,23 @@ from plotFunctions import *
 
 ## CALC TIME
 #SSSP
-#grouped_bar_plot(graphs, frameworks, calcTimesSSSPByFramework, yErrs=yErrCalcSSSPByFramework, title='Single-source Shortest-path on one calculation node with standard deviation', yLabel='Calculation time (s)', saveToFile="singleNodeSSSP_calcTime.png")
+grouped_bar_plot(graphs, frameworks, calcTimesSSSPByFramework, yErrs=yErrCalcSSSPByFramework, title='Single-source Shortest-path on one calculation node with standard deviation', yLabel='Calculation time (s)', saveToFile="singleNodeSSSP_calcTime.png")
 
 #BFS
 grouped_bar_plot(graphs, frameworks, calcTimesBFSByFramework, yErrs=yErrCalcBFSByFramework, title='Breadth-first search on one calculation node with standard deviation', yLabel='Calculation time (s)', saveToFile="singleNodeBFS_calcTime.png")
 
 ## EXEC TIME
 #SSSP
-#grouped_bar_plot(graphs, frameworks, execTimesSSSPByFramework, yErrs=yErrExecSSSPByFramework, title='Single-source Shortest-path on one calculation node with standard deviation', yLabel='Execution time (s)', saveToFile="singleNodeSSSP_execTime.png")
+grouped_bar_plot(graphs, frameworks, execTimesSSSPByFramework, yErrs=yErrExecSSSPByFramework, title='Single-source Shortest-path on one calculation node with standard deviation', yLabel='Execution time (s)', saveToFile="singleNodeSSSP_execTime.png")
 
 #BFS
 grouped_bar_plot(graphs, frameworks, execTimesBFSByFramework, yErrs=yErrExecBFSByFramework, title='Breadth-first search on one calculation node with standard deviation', yLabel='Execution time (s)', saveToFile="singleNodeBFS_execTime.png")
 
 ## OVERHEAD
 #SSSP
-#grouped_bar_plot(graphs, frameworks, overheadTimesSSSPByFramework, title='Overhead time of each framework during SSSP', yLabel='Overhead time (s)', saveToFile="singleNodeSSSP_overheadTime.png")
+grouped_bar_plot(graphs, frameworks, overheadTimesSSSPByFramework, title='Overhead time of each framework during SSSP', yLabel='Overhead time (s)', saveToFile="singleNodeSSSP_overheadTime.png")
 
[…]

### 2020-08-22 documentation/paper/main.tex
new plots and some work on testingMethods and results. Also cleaned up Gemini overview

@@
 \usepackage{graphicx}
+\usepackage{subcaption}
 
@@
 \IEEEauthorblockA{(3341630)\\
-@stud.uni-stuttgart.de}}
+st157100@stud.uni-stuttgart.de}}
 
@@
 \begin{abstract}
-In this paper we will analyze and compare the graph frameworks Galois, Ligra, Polymer, Gemini and Giraph in their parformance.
-All the frameworks will be tested in shared memory and Galois, Gemini and Giraph are tested on a distributed cluster as well.
-Furthermore we will give some insight on the complexity of writing custom applications based on these frameworks.
+In this paper we will analyze and compare the non-uniform memory access (NUMA) aware graph frameworks Galois, Ligra, Polymer and Gemini in their parformance.
+
+Giraph
+
+All the frameworks will be tested in shared memory. Galois, Gemini and Giraph are tested on a distributed cluster as well.
+We will give some insight on the calculation performance and associated time overhead of each framework.
+For this we compare the frameworks in their performance with Single-source shortest-paths, Breadth-first search and PageRank algorithms.
 \end{abstract}
@@
 \section{Introduction}
-%\todo
 In recent years graph sizes have increased significantly, thus performance and memory efficiency of the graph analysis applications is now more important than ever.
+
+We will compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
+The comparison is performed on both real world data sets and synthetic graphs.
+
+To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
+
+%One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
+%This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.
+
+To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.
 
@@
 The conclusion goes here.
+
+\clearpage
 

### 2020-08-22 documentation/paper/sections/overview/gemini.tex
new plots and some work on testingMethods and results. Also cleaned up Gemini overview

@@
 
-Gemini\cite{Gemini} is a framework for parallel graph processing.
-In comparrison to the other NUMA-aware frameworks we discuss here Gemini does's not support shared memory parallel processing, it is completely message based using MPI.
-It seems to be very lightweight and has only the very basically needed functionallity implemented, which leeds to clear source code and small binarys.
+Gemini is a framework for parallel graph processing\cite{Gemini}.
+In comparison to the other NUMA-aware frameworks we discuss here, Gemini does not support shared memory parallel processing.
+Computation is entirely message based using MPI.
+Gemini is lightweight and only provides basic functionality.
 A basic API hides data and computation distribution details from the user writing applications.
-It comes with five algorithms already implemented included the three we are testing in this paper.
-While setting it up and even while benchmarking we had various problems with bugs in the source code, like non zero terminated strings and missing return statements.
-While debugging we forked the origional repository and made our changes there.
-DODO link repository.
+There are five algorithms (which ones?) already implemented.

### 2020-08-22 documentation/paper/sections/testingMethods.tex
new plots and some work on testingMethods and results. Also cleaned up Gemini overview

@@
 \begin{table}
-	\caption{Size comparison of the used graphs}
-	\label{tbl:graphs}
 	\centering
@@
 	\end{tabular}
+	\caption{Size comparison of the used graphs}
+	\label{tbl:graphs}
 \end{table}
@@
 \begin{table}
-	\caption{Tested algorithms of the frameworks}
-	\label{tbl:algorithms}
 	\centering
-	\begin{tabularx}{\columnwidth}{cYYYYY}
-		\bf{Algorithm}&\rot{Galois}&\rot{Gemini}&\rot{Giraph}&\rot{Ligra}&\rot{Polymer}\\\hline
+	\begin{tabularx}{0.9\columnwidth}{cYYYYY}
+		\bf{Algorithm}&Galois&Gemini&Giraph&Ligra&Polymer\\\hline
 		PageRank&$\Delta$Push, $\Delta$Pull&&&&\\
@@
 	\end{tabularx}
+	\caption{Tested algorithms of the frameworks}
+	\label{tbl:algorithms}
 

### 2020-08-22 plots/makeplots.py

@@ -3,7 +3,15 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib
 from math import sqrt
+import re
 plt.style.use('./style.mplstyle')
+
+#sorted by size of graph
+graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
+#graph sizes in million edges
+graphSize = (2, 117, 378, 1963, 2147, 2586, 4294)
+#sorted alphabetically
+frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
 
 
 def insertAt(idx, val, list):
@@ -19,7 +27,7 @@ def insertAt(idx, val, list):
         return 
 
 
-
+## RAW DATA
 #[graph, algo, calc, exec, varCalc, varExec]
 data = []
 with open('../results/results.txt') as f:
@@ -47,28 +55,17 @@ for i in range(len(data)):
     except ValueError:
         totalTime[i] = float('nan')
     try:
-        yErrCalc[i] = float(data[i][4])
+        yErrCalc[i] = sqrt(float(data[i][4]))
     except ValueError:
         yErrCalc[i] = float('nan')
     try:
-        yErrExec[i] = float(data[i][5])
+        yErrExec[i] = sqrt(float(data[i][5]))
     except ValueError:
         yErrExec[i] = float('nan')
 
 
 
-#sorted by size of graph
-graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
-#graph sizes in million edges
-graphSize = (2, 117, 378, 1963, 2147, 2586, 4294)
-#sorted alphabetically
-frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
-
-
-
-
 #### SSSP TIMES
-
 calcTimesSSSPByFramework = []
 yErrCalcSSSPByFramework = []
 
@@ -92,10 +89,10 @@ for f in frameworks:
 		for i in range(len(graph)):
 			if g == graph[i] and fname == algo[i]:
 				tmpCalc.append(calcTime[i])
-				tmpYErrCalc.append(sqrt(yErrCalc[i]))
+				tmpYErrCalc.append(yErrCalc[i])
 
 				tmpExec.append(totalTime[i])
-				tmpYErrExec.append(sqrt(yErrExec[i]))
+				tmpYErrExec.append(yErrExec[i])
 
[…]

### 2020-08-24 documentation/paper/sections/overview/giraph.tex

@@
 
-Apache Giraph\cite{Giraph} is an iterative graph processing framework, built on top of Apache Hadoop. Because of this, expanding shared memory processing to a multi-node cluster is almost seamless.
+Apache Giraph is an iterative graph processing framework, built on top of Apache Hadoop\cite{Giraph}.
+Hadoop as a large MapReduce infrastructure provides a reliable (fault tolerant) basis for large scale graph processing.
+Because of the underlying Hadoop MapReduce infrastructure, expanding single-node processing to a multi-node cluster is almost seamless.
 
-Giraph is inspired by the Bulk Synchronous Parallel model of distributed computation introduced by Leslie Valiant\cite{BSPforGiraph}.
-Meaning that it is based on computation units that communicate using messages and are synchonized with barriers.
+Giraph is based on computation units that communicate using messages and are synchonized with barriers.
 

### 2020-08-24 documentation/paper/sections/results/results.tex

@@
 \begin{enumerate}
-	\item During setup of Gemini, we realized that the cloned repository contained several bugs, rendering the code as-is unable to perform any calculations.
-We had to fork and modify the source code.
-TODO(Was genau ging nicht, was haben wir geändert?)
+	\item During setup and benchmark of Gemini, we encountered several bugs in the cloned repository. These include non zero-terminated strings or even missing return statements.
 
-	The repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
+	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code.
+A repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
+
 	\item Furthermore, we would like to address the setup of Hadoop for Giraph.
@@
 This makes the setup rather time consuming, especially if reconfiguration is needed later on.
-	\item Also in order for Giraph to run, several Java tasks have to be constantly running in the background.
-While we don't expect this to have a noticeable performance impact on other tasks, it is still not optimal.
-	\item The Hadoop distributed file system (HDFS) ran us into disk space problems on multiple occasions.
+	\item In order for Giraph to run, several Java tasks (the Hadoop infrastructure) have to be constantly running in the background.
+While we don't expect this to have a significant performance impact on other tasks, it is still suboptimal.
+	%\item The Hadoop distributed file system (HDFS) ran us into disk space problems on multiple occasions.
 First, unless configured otherwise, the standard implementation replicates all data three times, distributed over all participating nodes.
@@
 
-\subsection{Pure Performance-Ergebnisse}
-
-\subsection{Ergebnisse Calc time}
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_calcTime.png}
+\subsection{Single-source Shortest-paths}
+\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
+In these figures, Galois with 96 threads is shown.
+We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
+\begin{figure*}[t!]
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
 		\caption{Calculation times for SSSP on a single node}
 		\label{fig:singleNodeSSSP_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
+		\caption{Execution times for SSSP on a single node}
+		\label{fig:singleNodeSSSP_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
+		\caption{Overhead time (execution time - calculation time) normalized by the graph size in million edges}
+		\label{fig:singleNodeSSSP_overheadNormalized}
+	\end{subfigure}
+	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.}
+	\label{fig:singleNodeSSSP}
+\end{figure*}
+
+\subsection{Breadth-first search}
+In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
+\begin{figure*}
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
+		\caption{Calculation times for BFS on a single node}
+		\label{fig:singleNodeBFS_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
+		\caption{Execution times for BFS on a single node}
+		\label{fig:singleNodeBFS_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
+		\caption{Overhead time (execution time - calculation time) normalized by the graph size in million edges}
+		\label{fig:singleNodeBFS_overheadNormalized}
+	\end{subfigure}
+
+	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing}
+\end{figure*}
+
+\subsection{PageRank}
+
+\subsection{Galois speedup}
+\label{sec:galois_speedup}
+For Galois we
+
+\begin{figure}
+	\begin{subfigure}{\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
+		\caption{Single-source Shortest-paths}
+		\label{fig:singleNodeSSSPGaloisThreads}
+	\end{subfigure}
+	\begin{subfigure}{\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
+		\caption{Breadth-first search}
+		\label{fig:singleNodeBFSGaloisThreads}
+	\end{subfigure}
+	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths and Breadth-first search algorithms.}
 \end{figure}
@@
 \begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_calcTime.png}
-	\caption{Calculation times for BFS on a single node}
-	\label{fig:singleNodeBFS_calc}
+	\begin{subfigure}{\columnwidth}
+		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPullGaloisThreads.png}
+		\caption{PageRank Pull}
+		\label{fig:singleNodePRPullGaloisThreads}
+	\end{subfigure}
+	\begin{subfigure}{\columnwidth}
+		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPushGaloisThreads.png}
+		\caption{PageRank Push}
+		\label{fig:singleNodePRPushGaloisThreads}
+	\end{subfigure}
+	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms.}
 \end{figure}
 
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSPGaloisThreads.png}
-	\caption{Calculation times of Galois' SSSP over the thread count}
-	\label{fig:singleNodeSSSPGaloisThreads}
-\end{figure}
-
-\subsection{Ergebnisse exec time}
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_execTime.png}
-	\caption{Execution times for SSSP on a single node}
-	\label{fig:singleNodeSSSP_exec}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
-	\caption{Normalized Overhead SSSP single node}
-	\label{fig:singleNodeSSSP_overheadNormalized}
-\end{figure}
+\subsection{Execution times and overhead}
 
@@
 \begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_execTime.png}
-	\caption{Execution times for BFS on a single node}
-	\label{fig:singleNodeBFS_exec}
-\end{figure}
-
-\begin{figure}
 	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTime.png}
@@
 
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
-	\caption{Normalized Overhead BFS single node}
-	\label{fig:singleNodeBFS_overheadNormalized}
-\end{figure}
-
 hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.

### 2020-08-24 documentation/paper/sections/results/results.tex
distributed plots

@@
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time (execution time - calculation time) normalized by the graph size in million edges}
+		\caption{Overhead time normalized by the graph size in million edges}
 		\label{fig:singleNodeSSSP_overheadNormalized}
@@
 	\label{fig:singleNodeSSSP}
+\end{figure*}
+
+\begin{figure*}[t!]
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
+		\caption{Calculation times for distributed SSSP}
+		\label{fig:distributedSSSP_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
+		\caption{Execution times for distributed SSSP}
+		\label{fig:distributedSSSP_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalized.png}
+		\caption{Overhead time normalized by the graph size in million edges}
+		\label{fig:distributedSSSP_overheadNormalized}
+	\end{subfigure}
+	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
+	\label{fig:distributedSSSP}
 \end{figure*}
@@
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
-		\caption{Overhead time (execution time - calculation time) normalized by the graph size in million edges}
+		\caption{Overhead time normalized by the graph size in million edges}
 		\label{fig:singleNodeBFS_overheadNormalized}
@@
 	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing}
+\end{figure*}
+
+\begin{figure*}
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
+		\caption{Calculation times for distributed BFS}
+		\label{fig:distributedBFS_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
+		\caption{Execution times for distributed BFS}
+		\label{fig:distributedBFS_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
+		\caption{Overhead time normalized by the graph size in million edges}
+		\label{fig:distributedBFS_overheadNormalized}
+	\end{subfigure}
+
+	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
 \end{figure*}

### 2020-08-24 documentation/paper/sections/testingMethods.tex
distributed plots

@@
 One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
-All servers were running Ubuntu 19.x.
-Setup of each framework was performed according to our provided installation guides.
+All five machines were running Ubuntu 18.04.2 LTS.
+Setup of each framework was performed according to our provided installation guides\footnote{The setup guides are available at \url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}.
 
@@
 	\item For Galois, we resulted to extracting console log time stamps. Galois outputs \texttt{Reading graph complete.}. Calculation time is the time from this output to the end of execution.
-	\item Polymer outputs the name of the algorithm followed by a measured time.
+
+	We know that this is not the most realiable way for measuring the calculation times. 
+	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.
+	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time that is used for cleanup after calculation in the measurement.
+
+	\item Polymer outputs the name of the algorithm followed by an internally measured time.
+
 	\item Gemini outputs a line \texttt{exec\_time=x}, which was used to measure the calculation time.
-	\item Ligra outputs a measured time with \texttt{Running time : x}.
-	\item Giraph has built in timers for the iterations (supersteps) which we used to extract the computation time.
+
+	\item Ligra outputs its time measurement with \texttt{Running time : x}.
+
+	\item Giraph has built in timers for the iterations (supersteps) which we summed up to extract the computation time.
 \end{itemize}
+Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.
+
 Each test case consisting of graph, framework and algorithm was run 10 times, allowing us to smooth slight variations in the measured times.
+Later on, we provide the mean values of the individual times as well as the standard deviation where meaningful.
 
@@
 \subsection{Algorithms}
-We ran Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) algorithms on each framework and graph.
+The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark framework with every graph.
 For frameworks that support multiple implementations (e.g. PageRank in push and pull modes), we included the alternatives in our testing. 
 \autoref{tbl:algorithms} shows the tested algorithms in detail.
+
 \newcommand{\txtdag}{\textsuperscript{\textdagger}}
 \begin{table}
+TODO
 	\centering

### 2020-08-24 plots/makedata.py

@@ -1,0 +1,325 @@
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+from math import sqrt
+import re
+plt.style.use('./style.mplstyle')
+
+#sorted by size of graph
+graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
+#graph sizes in million edges
+graphSize = (2, 117, 378, 1963, 2147, 2586, 4294)
+#sorted alphabetically
+frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
+
+
+def insertAt(idx, val, list):
+    if idx < 0:
+        return
+    if idx >= len(list):
+        while idx > len(list):
+            list.append(float('nan'))
+        list.append(val)
+        return
+    else:
+        list[idx] = val
+        return 
+
+
+## RAW DATA
+#[graph, algo, calc, exec, varCalc, varExec]
+data = []
+with open('../results/results.txt') as f:
+    for line in f:
+        data.append([])
+        for entry in line.split():
+            data[-1].append(entry)
+
+graph = np.empty(len(data), dtype=object)
+algo = np.empty(len(data), dtype=object)
+calcTime = np.empty(len(data))
+totalTime = np.empty(len(data))
+yErrCalc = np.empty(len(data))
+yErrExec = np.empty(len(data))
+
+for i in range(len(data)):
+
+    graph[i] = data[i][0]
+    algo[i] = data[i][1]
+    try:
+        calcTime[i] = float(data[i][2])
+    except ValueError:
+        calcTime[i] = float('nan')
+    try:
+        totalTime[i] = float(data[i][3])
+    except ValueError:
+        totalTime[i] = float('nan')
+    try:
+        yErrCalc[i] = sqrt(float(data[i][4]))
+    except ValueError:
+        yErrCalc[i] = float('nan')
+    try:
+        yErrExec[i] = sqrt(float(data[i][5]))
+    except ValueError:
+        yErrExec[i] = float('nan')
+
+
+
+
+
+
+
+
+
+
+#### SSSP TIMES SINGLE NODE
+calcTimeSSSP_singleNode = []
+yErrCalcSSSP_singleNode = []
+
+execTimeSSSP_singleNode = []
+yErrExecSSSP_singleNode = []
+
+overheadSSSP_singleNode = []
[…]

### 2020-08-24 plots/makeplots.py

@@ -1,251 +1,101 @@
 
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib
-from math import sqrt
-import re
-plt.style.use('./style.mplstyle')
-
-#sorted by size of graph
-graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
-#graph sizes in million edges
-graphSize = (2, 117, 378, 1963, 2147, 2586, 4294)
-#sorted alphabetically
-frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
-
-
-def insertAt(idx, val, list):
-    if idx < 0:
-        return
-    if idx >= len(list):
-        while idx > len(list):
-            list.append(float('nan'))
-        list.append(val)
-        return
-    else:
-        list[idx] = val
-        return 
-
-
-## RAW DATA
-#[graph, algo, calc, exec, varCalc, varExec]
-data = []
-with open('../results/results.txt') as f:
-    for line in f:
-        data.append([])
-        for entry in line.split():
-            data[-1].append(entry)
-
-graph = np.empty(len(data), dtype=object)
-algo = np.empty(len(data), dtype=object)
-calcTime = np.empty(len(data))
-totalTime = np.empty(len(data))
-yErrCalc = np.empty(len(data))
-yErrExec = np.empty(len(data))
-
-for i in range(len(data)):
-    graph[i] = data[i][0]
-    algo[i] = data[i][1]
-    try:
-        calcTime[i] = float(data[i][2])
-    except ValueError:
-        calcTime[i] = float('nan')
-    try:
-        totalTime[i] = float(data[i][3])
-    except ValueError:
-        totalTime[i] = float('nan')
-    try:
-        yErrCalc[i] = sqrt(float(data[i][4]))
-    except ValueError:
-        yErrCalc[i] = float('nan')
-    try:
-        yErrExec[i] = sqrt(float(data[i][5]))
-    except ValueError:
-        yErrExec[i] = float('nan')
+from makedata import *
+from plotFunctions import *
 
 
 
-#### SSSP TIMES
-calcTimesSSSPByFramework = []
[…]

### 2020-08-24 plots/plotFunctions.py

@@ -10,7 +10,7 @@ def grouped_bar_plot(group_names, legend_names, values, yErrs=None, width=0.35, 
 	group_count = len(group_names)
 
 	if len(legend_names) != len(values):
-		raise "Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values))
+		raise Exception("Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values)))
 
 
 	if scaleFactor == None:
@@ -55,7 +55,7 @@ def grouped_bar_plot(group_names, legend_names, values, yErrs=None, width=0.35, 
 def line_plot(legend_names, x, values, yErrs=None, title='', yLabel='', xLabel='', yScale=None, saveToFile=None):
 
 	if len(legend_names) != len(values):
-		raise "Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values))
+		raise Exception("Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values)))
 
 	fig, ax = plt.subplots()
 

### 2020-08-25 plots/makedata.py
Pagerank Plots. Die müssen dringend nochmal überarbeitet werden

@@ -269,6 +269,113 @@ for f in dist_frameworks_bfs:
 
 
 
+
+#### PAGERANK SINGLE NODE
+singleNode_frameworks_pr = {
+    "galois-pagerank-push-cpu-96thread":"Galois Push",
+    "galois-pagerank-pull-cpu-96thread":"Galois Pull",
+    "gemini-pagerank":"Gemini",
+    "giraph-pagerank":"Giraph",
+    "ligra-pagerank":"Ligra",
+    "ligra-pagerank-delta":"Ligra Delta",
+    "polymer-pagerank":"Polymer",
+    "polymer-pagerank-delta":"Polymer Delta"}
+
+calcTimePR_singleNode = []
+yErrCalcPR_singleNode = []
+
+execTimePR_singleNode = []
+yErrExecPR_singleNode = []
+
+overheadPR_singleNode = []
+overheadPRNormalized_singleNode = []
+for f in singleNode_frameworks_pr:
+    tmpExec = []
+    tmpCalc = []
+    tmpYErrCalc = []
+    tmpYErrExec = []
+    overhead = []
+    overheadNormalized = []
+    
+    k = 0 # graph index
+    for g in graphs:
+        for i in range(len(graph)):
+            if g == graph[i] and f == algo[i]:
+                tmpCalc.append(calcTime[i])
+                tmpYErrCalc.append(sqrt(yErrCalc[i]))
+
+                tmpExec.append(totalTime[i])
+                tmpYErrExec.append(sqrt(yErrExec[i]))
+
+                overhead.append(totalTime[i] - calcTime[i])
+                overheadNormalized.append((totalTime[i] - calcTime[i])/graphSize[k])
+        k += 1 # graph index
+
+
+    calcTimePR_singleNode.append(tmpCalc)
+    yErrCalcPR_singleNode.append(tmpYErrCalc)
+    execTimePR_singleNode.append(tmpExec)
+    yErrExecPR_singleNode.append(tmpYErrExec)
+    overheadPR_singleNode.append(overhead)
+    overheadPRNormalized_singleNode.append(overheadNormalized)
+
+
+
+### PR DISTRIBUTED
+dist_frameworks_pr = {
+    "galois-pagerank-push-dist":"Galois Push",
+    "galois-pagerank-pull-dist":"Galois Pull",
+    "gemini-pagerank-dist":"Gemini",
+    "giraph-pagerank-dist":"Giraph"}
+
+calcTimePR_distributed = []
[…]

### 2020-08-25 plots/makeplots.py

@@ -5,97 +5,136 @@ from plotFunctions import *
 
 
 distributed = True
-singleNode = True
+singleNode = False
 
-galoisSpeedup = True
-
-sssp = True
-bfs = True
+sssp = False
+bfs = False
 pr = True
 
+calcTimes = False
+execTimes = False
+overheads = False
+
+
+galoisSpeedup = False
 
 ### PLOTS
-
-
-
 #SINGLE NODE
-if singleNode:
+if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 	## CALC TIME
 	#SSSP
-	if sssp:
+	if sssp or calcTimes:
 		grouped_bar_plot(graphs, frameworks, calcTimeSSSP_singleNode, yErrs=yErrCalcSSSP_singleNode, title='SSSP single node', yLabel='Calculation time (s)', saveToFile="singleNodeSSSP_calcTime.png")
 
 	#BFS
-	if bfs:
+	if bfs or calcTimes:
 		grouped_bar_plot(graphs, frameworks, calcTimeBFS_singleNode, yErrs=yErrCalcBFS_singleNode, title='BFS single node', yLabel='Calculation time (s)', saveToFile="singleNodeBFS_calcTime.png")
+
+	if pr or calcTimes:
+		grouped_bar_plot(graphs, singleNode_frameworks_pr.values(), calcTimePR_singleNode, yErrs=yErrCalcPR_singleNode, title='PR single node (ugly)', yLabel='Calculation time (s)', saveToFile="singleNodePR_calcTime.png", scaleFactor=3.25)
+
+
+
 
 	## EXEC TIME
 	#SSSP
-	if sssp:
+	if sssp or execTimes:
 		grouped_bar_plot(graphs, frameworks, execTimeSSSP_singleNode, yErrs=yErrExecSSSP_singleNode, title='SSSP single node', yLabel='Execution time (s)', saveToFile="singleNodeSSSP_execTime.png")
 
 	#BFS
-	if bfs:
+	if bfs or execTimes:
 		grouped_bar_plot(graphs, frameworks, execTimeBFS_singleNode, yErrs=yErrExecBFS_singleNode, title='BFS single node', yLabel='Execution time (s)', saveToFile="singleNodeBFS_execTime.png")
+
+	if pr or execTimes:
[…]

### 2020-08-26 scripts/benchmark/benchmark_evaluation.sh
giraph benchmark results

@@ -21,27 +21,35 @@ done
 # evaluate
 for graph in ${graphs[@]}; do
   for algorithm in ${algorithms[@]}; do
-    if [[ "$algorithm" == *gemini* ]]; then
-      means=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " (s2-5*s1)/10000000}')
-      mean1=$(echo "$means" | awk '{print $1}')
-      mean2=$(echo "$means" | awk '{print $2}')
-      vars=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk -v m1="$mean1" -v m2="$mean2" '{s1+=($5/1000000-m1)**2} {s2+=(($6-5*$5)/1000000-m2)**2} END {print s1/9 " " s2/9}')
-    elif [[ "$algorithm" == *giraph* ]]; then
-      means=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$7} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
-      mean1=$(echo "$means" | awk '{print $1}')
-      mean2=$(echo "$means" | awk '{print $2}')
-      vars=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk -v m1="$mean1" -v m2="$mean2" '{s1+=($7/1000000-m1)**2} {s2+=($6/1000000-m2)**2} END {print s1/9 " " s2/9}')
+    error=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | grep "ERROR" | wc -l)
+    timeout=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | grep "TIMEOUT" | wc -l)
+    if [[ $error -gt 0 ]]; then
+      echo "$graph $algorithm ERROR ERROR ERROR ERROR"
+    elif [[ $timeout -gt 0 ]]; then
+      echo "$graph $algorithm TIMEOUT TIMEOUT TIMEOUT TIMEOUT"
     else
-      means=$(cat ${@} | grep "^#" | grep "$graph" | grep "$algorithm " | awk '{s1+=$5} {s2+=$6} END {print s1/10000000 " " s2/10000000}')
-      mean1=$(echo "$means" | awk '{print $1}')
-      mean2=$(echo "$means" | awk '{print $2}')
[…]

### 2020-08-27 plots/makeplots.py
Plots updated with giraph results

@@ -11,9 +11,9 @@ sssp = False
 bfs = False
 pr = True
 
-calcTimes = False
-execTimes = False
-overheads = False
+calcTimes = True
+execTimes = True
+overheads = True
 
 
 galoisSpeedup = False

### 2020-08-27 documentation/paper/sections/results/results.tex
Pagerank Plots. Die müssen dringend nochmal überarbeitet werden
Results distributed SSSP

@@
 \subsection{Single-source Shortest-paths}
+
+\subsubsection{Single-node}
+
 \autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
@@
 	\end{subfigure}
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.}
+	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.
+	The runs on rMat28 for Ligra and Polymer failed and the frameworks were unable to complete the task.}
 	\label{fig:singleNodeSSSP}
@@
 
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
+\subsubsection{Distributed}
+\begin{figure}[t!]
+	\begin{subfigure}{\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
@@
 	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedSSSP_overheadNormalized}
-	\end{subfigure}
 	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
 	\label{fig:distributedSSSP}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalizedToGalois.png}
+	\caption{Overhead times of each framework normalized by the overhead time of Galois Push}
+	\label{fig:distributedSSSP_overhead}
+\end{figure}
+
+For the distributed scenario, \autoref{fig:distributedSSSP} shows the benchmark results as calculation and execution times.
+
+Results are especially interesting for Giraph since it seems to not cope well with synthetic graphs.
+Analyzing the computation times in \autoref{fig:distributedSSSP_calc}, we see that it is the fastest framework on our real-world graphs.
+And that with a considerable margin of other frameworks always taking at least 50\% longer (Gemini on flickr) up to Galois Pull needing 18$\times$ longer on wikipedia.
+On both synthetic graphs however, Giraph is the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.
+
+While Giraph's computation times are very competitive, when comparing the execution times in \autoref{fig:distributedSSSP_exec} we see that Giraph is actually the slowest framework on 5 out of 7 graphs.
+For the other two, namely twitter and friendster, Giraph is second slowest with only Gemini taking longer to complete.
+
+Giraph and Gemini's very long execution times are only due to their overhead being many orders of magnitude larger than Galois overhead (\autoref{fig:distributedSSSP_overhead}).
+Overhead for Gemini is greater than that of Galois on every graph. From just a 20\% increase on flickr up to friendster, where the overhead is 90$\times$ that of Galois Push.
+For Giraph the overhead times are not as extreme but still generally worse.
+Even on flickr, Giraph's overhead time is already 23$\times$ that of Galois.
+On friendster, where Gemini was worst, Giraph \emph{only} requires 73$\times$ the overhead time of Galois.
 
@@
 
+\begin{figure*}
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
+		\caption{Calculation times for PR on a single node}
+		\label{fig:singleNodePR_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
+		\caption{Execution times for PR on a single node}
+		\label{fig:singleNodePR_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTimeNormalized.png}
+		\caption{Overhead time normalized by the graph size in million edges}
+		\label{fig:singleNodePR_overheadNormalized}
+	\end{subfigure}
+
+	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing}
+\end{figure*}
+
+\begin{figure*}
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
+		\caption{Calculation times for distributed PR}
+		\label{fig:distributedPR_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
+		\caption{Execution times for distributed PR}
+		\label{fig:distributedPR_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.3\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTimeNormalized.png}
+		\caption{Overhead time normalized by the graph size in million edges}
+		\label{fig:distributedPR_overheadNormalized}
+	\end{subfigure}
+
+	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
+\end{figure*}
+
 \subsection{Galois speedup}

### 2020-08-27 plots/makedata.py

@@ -160,10 +160,17 @@ for f in dist_frameworks_sssp:
     overheadSSSP_distributed.append(overhead)
     overheadSSSPNormalized_distributed.append(overheadNormalized)
 
+overheadSSSP_distributed_normalizedToGalois = []
 
-
-
-
+print("Overhead comparison distributed SSSP")
+i = 0
+for k in dist_frameworks_sssp:
+    overheadSSSP_distributed_normalizedToGalois.append([])
+    print(dist_frameworks_sssp[k])
+    for j in range(len(overheadSSSP_distributed[i])):
+        print(graphs[j] + ": ", calcTimeSSSP_distributed[i][j]/calcTimeSSSP_distributed[2][j])
+        overheadSSSP_distributed_normalizedToGalois[i].append(overheadSSSP_distributed[i][j]/overheadSSSP_distributed[0][j]) 
+    i += 1
 
 
 

### 2020-08-27 plots/makeplots.py

@@ -11,8 +11,8 @@ sssp = False
 bfs = False
 pr = True
 
-calcTimes = True
-execTimes = True
+calcTimes = False
+execTimes = False
 overheads = True
 
 
@@ -125,6 +125,8 @@ if distributed and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), overheadSSSP_distributed, title='Overhead time of each framework during SSSP', yLabel='Overhead time (s)', saveToFile="distributedSSSP_overheadTime.png")
 
 		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), overheadSSSPNormalized_distributed, title='SSSP distributed', yLabel='Overhead time (s) (normalized)', saveToFile="distributedSSSP_overheadTimeNormalized.png", yScale='linear')
+
+		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), overheadSSSP_distributed_normalizedToGalois, title='SSSP distributed', yLabel='Galois Push overhead times', saveToFile="distributedSSSP_overheadTimeNormalizedToGalois.png", yScale='linear')
 
 	#BFS
 	if bfs or overheads:

### 2020-08-27 SimpleBFSStructureComputation.java
Giraph BFS Algorithm, untested

@@ -1,0 +1,142 @@
+/*
+ * Licensed to the Apache Software Foundation (ASF) under one
+ * or more contributor license agreements.  See the NOTICE file
+ * distributed with this work for additional information
+ * regarding copyright ownership.  The ASF licenses this file
+ * to you under the Apache License, Version 2.0 (the
+ * "License"); you may not use this file except in compliance
+ * with the License.  You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
+
+/*
+	Retrieved from https://github.com/MarcoLotz/GiraphBFSSO
+ */
+
+//package uk.co.qmul.giraph.structurebfs; //removed
+
+import org.apache.giraph.Algorithm;
+import org.apache.giraph.graph.BasicComputation;
+import org.apache.giraph.conf.LongConfOption;
+import org.apache.giraph.edge.Edge;
+import org.apache.giraph.graph.Vertex;
+import org.apache.hadoop.io.DoubleWritable;
+import org.apache.hadoop.io.FloatWritable;
+import org.apache.hadoop.io.LongWritable;
+import org.apache.log4j.Logger;
+
+import java.io.IOException;
+
+/**
+ * Demonstrates the basic Pregel Breadth First Search applied to graph
+ * structure.
+ * 
+ * SimpleBFSComputation is the basic class with the configurations for a BFS
+ * search and the basic configurations for it. The "Structure" means that it
+ * computes the whole structure in order to check the depth from the start
+ * vertex
+ * 
+ * @author Marco Aurelio Lotz
+ */
+@Algorithm(name = "Breadth First Search Structural oriented", description = "Uses Breadth First Search from a source vertex to calculate depths")
+public class SimpleBFSStructureComputation
+		extends
+		BasicComputation<LongWritable, DoubleWritable, FloatWritable, DoubleWritable> {
+
+	// added the following for dynamic start vertices
+	private static long startVertexId;
+	public static void main(String[] args) {
[…]

### 2020-08-27 SimplePageRankComputation.java
Pagerank iteration down to 5..

@@ -52,7 +52,7 @@ import java.util.List;
 public class SimplePageRankComputation extends BasicComputation<LongWritable,
     DoubleWritable, FloatWritable, DoubleWritable> {
   /** Number of supersteps for this test */
-  public static final int MAX_SUPERSTEPS = 30;
+  public static final int MAX_SUPERSTEPS = 5;
   /** Logger */
   private static final Logger LOG =
       Logger.getLogger(SimplePageRankComputation.class);

### 2020-08-27 SimpleBFSStructureComputation.java
Giraph BFS Algorithm added to install steps and benchmark script

@@ -22,6 +22,7 @@
  */
 
 //package uk.co.qmul.giraph.structurebfs; //removed
+package org.apache.giraph.examples;
 
 import org.apache.giraph.Algorithm;
 import org.apache.giraph.graph.BasicComputation;

### 2020-08-27 scripts/benchmark/benchmark.sh

@@ -355,7 +355,7 @@ gemini-bfs-dist () {
 
 giraph-sssp () {
     local graph="$1.gir"
-    $($HADOOP_HOME/bin/hadoop dfs -rmr /output/$graph)
+    $($HADOOP_HOME/bin/hadoop dfs -rmr -skipTrash /output/$graph)
     local time_start=$(get_time)
     # hardcoded for our current setup
     local result=$(timeout 3h $HADOOP_HOME/bin/hadoop jar $GIRAPH_HOME/giraph-examples/target/giraph-examples-1.3.0-SNAPSHOT-for-hadoop-1.2.1-jar-with-dependencies.jar org.apache.giraph.GiraphRunner org.apache.giraph.examples.GeneralShortestPathsComputation $2 -vif org.apache.giraph.io.formats.JsonLongDoubleFloatDoubleVertexInputFormat -vip /input/$graph -vof org.apache.giraph.io.formats.IdWithValueTextOutputFormat -op /output/$graph -w 1 2>&1 | ts '%.s')
@@ -374,7 +374,7 @@ giraph-sssp () {
 
 giraph-sssp-dist () {
     local graph="$1.gir"
-    $($HADOOP_HOME/bin/hadoop dfs -rmr /output/$graph)
+    $($HADOOP_HOME/bin/hadoop dfs -rmr -skipTrash /output/$graph)
     local time_start=$(get_time)
     # hardcoded for our current setup
     local result=$(timeout 3h $HADOOP_HOME/bin/hadoop jar $GIRAPH_HOME/giraph-examples/target/giraph-examples-1.3.0-SNAPSHOT-for-hadoop-1.2.1-jar-with-dependencies.jar org.apache.giraph.GiraphRunner org.apache.giraph.examples.GeneralShortestPathsComputation $2 -vif org.apache.giraph.io.formats.JsonLongDoubleFloatDoubleVertexInputFormat -vip /input/$graph -vof org.apache.giraph.io.formats.IdWithValueTextOutputFormat -op /output/$graph -w 5 2>&1 | ts '%.s')
@@ -393,7 +393,7 @@ giraph-sssp-dist () {
 
 giraph-pagerank () {
     local graph="$1.gir"
-    $($HADOOP_HOME/bin/hadoop dfs -rmr /output/$graph)
+    $($HADOOP_HOME/bin/hadoop dfs -rmr -skipTrash /output/$graph)
     local time_start=$(get_time)
     # hardcoded for our current setup
[…]

### 2020-08-27 scripts/benchmark/benchmark.sh
giraph pagerank cases

@@ -439,7 +439,7 @@ giraph-bfs-dist () {
     log "giraph-bfs-dist $1 $pagerank_number_of_iterations - $dur_calc $dur_exec $dur_calc_giraph"
 }
 
-giraph-bfs-dist () {
+giraph-bfs () {
     local graph="$1.gir"
     $($HADOOP_HOME/bin/hadoop dfs -rmr -skipTrash /output/$graph)
     local time_start=$(get_time)

### 2020-08-27 scripts/benchmark/benchmark_cases.sh

@@ -488,7 +488,7 @@ benchmark_giraph_sssp() {
   done
 }
 
-benchmark_giraph_pagerank_dist_klein() {
+benchmark_giraph_pagerank_dist() {
   for startnode in ${startnodes_flickr[@]}; do
     giraph-pagerank-dist flickr
   done
@@ -497,6 +497,18 @@ benchmark_giraph_pagerank_dist_klein() {
   done
   for startnode in ${startnodes_wikipedia[@]}; do
     giraph-pagerank-dist wikipedia
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    giraph-pagerank-dist twitter
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    giraph-pagerank-dist friendster
+  done
+  for startnode in ${startnodes_rMat27[@]}; do
+    giraph-pagerank-dist rMat27
+  done
+  for startnode in ${startnodes_rMat28[@]}; do
+    giraph-pagerank-dist rMat28
   done
 }
 
@@ -598,6 +610,6 @@ benchmark_galois_cpu() {
 
 
 time_start=$(get_time)
-benchmark_galois_cpu
+benchmark_giraph_pagerank_dist
 dur_exec=$((($(get_time)-$time_start)/1000000))
 echo "benchmark took $dur_exec seconds"

### 2020-08-27 plots/makedata.py
new speedup plots, bugfix in plot script

@@ -14,17 +14,6 @@ graphSize = (2, 117, 378, 1963, 2147, 2586, 4294)
 frameworks = ('Galois', 'Gemini', 'Giraph', 'Ligra', 'Polymer')
 
 
-def insertAt(idx, val, list):
-    if idx < 0:
-        return
-    if idx >= len(list):
-        while idx > len(list):
-            list.append(float('nan'))
-        list.append(val)
-        return
-    else:
-        list[idx] = val
-        return 
 
 
 ## RAW DATA
@@ -162,13 +151,13 @@ for f in dist_frameworks_sssp:
 
 overheadSSSP_distributed_normalizedToGalois = []
 
-print("Overhead comparison distributed SSSP")
+#print("Overhead comparison distributed SSSP")
 i = 0
 for k in dist_frameworks_sssp:
     overheadSSSP_distributed_normalizedToGalois.append([])
-    print(dist_frameworks_sssp[k])
+    #print(dist_frameworks_sssp[k])
     for j in range(len(overheadSSSP_distributed[i])):
-        print(graphs[j] + ": ", calcTimeSSSP_distributed[i][j]/calcTimeSSSP_distributed[2][j])
+        #print(graphs[j] + ": ", calcTimeSSSP_distributed[i][j]/calcTimeSSSP_distributed[2][j])
         overheadSSSP_distributed_normalizedToGalois[i].append(overheadSSSP_distributed[i][j]/overheadSSSP_distributed[0][j]) 
     i += 1
 
@@ -390,10 +379,10 @@ speedupGaloisPRPush = {}
 speedupGaloisPRPull = {}
 x = []
 for g in graphs:
-    speedupGaloisSSSP[g] = []
-    speedupGaloisBFS[g] = []
-    speedupGaloisPRPush[g] = []
-    speedupGaloisPRPull[g] = []
+    speedupGaloisSSSP[g] = {}
+    speedupGaloisBFS[g] = {}
+    speedupGaloisPRPush[g] = {}
+    speedupGaloisPRPull[g] = {}
 
 for i in range(len(graph)):
     if not 'galois' in algo[i]:
@@ -414,26 +403,27 @@ for i in range(len(graph)):
     
     xVal = int(re.sub('\D', '', algo[i]))
 
-    for j in range(len(x)):
-        if xVal == x[j]:
[…]

### 2020-08-27 plots/makeplots.py

@@ -4,7 +4,7 @@ from plotFunctions import *
 
 
 
-distributed = True
+distributed = False
 singleNode = False
 
 sssp = False
@@ -16,7 +16,7 @@ execTimes = False
 overheads = True
 
 
-galoisSpeedup = False
+galoisSpeedup = True
 
 ### PLOTS
 #SINGLE NODE

### 2020-08-27 plots/plotFunctions.py

@@ -52,7 +52,7 @@ def grouped_bar_plot(group_names, legend_names, values, yErrs=None, width=0.35, 
 
 #
 # values is a dictionary
-def line_plot(legend_names, x, values, yErrs=None, title='', yLabel='', xLabel='', yScale=None, saveToFile=None):
+def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=None, saveToFile=None):
 
 	if len(legend_names) != len(values):
 		raise Exception("Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values)))
@@ -61,21 +61,13 @@ def line_plot(legend_names, x, values, yErrs=None, title='', yLabel='', xLabel='
 
 	ps = []
 	for element in legend_names: 
-		tmp = values[element].copy()
-		yErr = None
-		if not yErrs == None:
-			yErr = yErrs[element].copy()
-			while len(x) > len(yErr):
-				yErr.append(float('nan'))
-		while len(x) > len(tmp):
-			tmp.append(float('nan'))
+		x.sort()
+		tmp = []
+		for x_ in x:
+			tmp.append(values[element][x_])
 
-		if not yErr == None:
-			l = ax.errorbar(x, tmp, yerr=yErr, fmt='--o')
-			ps.append(l)
-		else:
-			l, = ax.plot(x, tmp, 'o--')
-			ps.append(l)
+		l, = ax.plot(x, tmp, 'o--')
+		ps.append(l)
 
 	if yScale != None:
 		ax.set_yscale(yScale)

### 2020-08-27 plots/makedata.py
updated std dev. in the plots

@@ -191,10 +191,10 @@ for f in frameworks:
         for i in range(len(graph)):
             if g == graph[i] and fname == algo[i]:
                 tmpCalc.append(calcTime[i])
-                tmpYErrCalc.append(sqrt(yErrCalc[i]))
+                tmpYErrCalc.append(yErrCalc[i])
 
                 tmpExec.append(totalTime[i])
-                tmpYErrExec.append(sqrt(yErrExec[i]))
+                tmpYErrExec.append(yErrExec[i])
 
                 overhead.append(totalTime[i] - calcTime[i])
                 overheadNormalized.append((totalTime[i] - calcTime[i])/graphSize[k])
@@ -238,10 +238,10 @@ for f in dist_frameworks_bfs:
         for i in range(len(graph)):
             if g == graph[i] and f == algo[i]:
                 tmpCalc.append(calcTime[i])
-                tmpYErrCalc.append(sqrt(yErrCalc[i]))
+                tmpYErrCalc.append(yErrCalc[i])
 
                 tmpExec.append(totalTime[i])
-                tmpYErrExec.append(sqrt(yErrExec[i]))
+                tmpYErrExec.append(yErrExec[i])
 
                 overhead.append(totalTime[i] - calcTime[i])
                 overheadNormalized.append((totalTime[i] - calcTime[i])/graphSize[k])
@@ -298,10 +298,10 @@ for f in singleNode_frameworks_pr:
         for i in range(len(graph)):
             if g == graph[i] and f == algo[i]:
                 tmpCalc.append(calcTime[i])
-                tmpYErrCalc.append(sqrt(yErrCalc[i]))
+                tmpYErrCalc.append(yErrCalc[i])
 
                 tmpExec.append(totalTime[i])
-                tmpYErrExec.append(sqrt(yErrExec[i]))
+                tmpYErrExec.append(yErrExec[i])
 
                 overhead.append(totalTime[i] - calcTime[i])
                 overheadNormalized.append((totalTime[i] - calcTime[i])/graphSize[k])
@@ -345,10 +345,10 @@ for f in dist_frameworks_pr:
         for i in range(len(graph)):
             if g == graph[i] and f == algo[i]:
                 tmpCalc.append(calcTime[i])
-                tmpYErrCalc.append(sqrt(yErrCalc[i]))
+                tmpYErrCalc.append(yErrCalc[i])
 
                 tmpExec.append(totalTime[i])
[…]

### 2020-08-27 plots/makeplots.py

@@ -4,15 +4,15 @@ from plotFunctions import *
 
 
 
-distributed = False
-singleNode = False
+distributed = True
+singleNode = True
 
-sssp = False
-bfs = False
+sssp = True
+bfs = True
 pr = True
 
-calcTimes = False
-execTimes = False
+calcTimes = True
+execTimes = True
 overheads = True
 
 

### 2020-08-27 documentation/paper/sections/results/results.tex
results for galois speedup sssp

@@
 
+\clearpage
+
 \subsection{Single-source Shortest-paths}
@@
 \subsubsection{Single-node}
-
 \autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
@@
 We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}[t!]
+\begin{figure*}[ht!]
 	\begin{subfigure}{0.3\textwidth}
@@
 \subsubsection{Distributed}
-\begin{figure}[t!]
+\begin{figure}
 	\begin{subfigure}{\columnwidth}
@@
 \label{sec:galois_speedup}
-For Galois we
+Analyzing the calculation time speedups for Galois, we can compare how or if the different algorithms benefit from increasing thread numbers.
 
+\subsubsection{Single-source Shortest-path}
 \begin{figure}
-	\begin{subfigure}{\columnwidth}
 	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
-		\caption{Single-source Shortest-paths}
-		\label{fig:singleNodeSSSPGaloisThreads}
-	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
+	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths}
+	\label{fig:galoisSpeedupSSSP}
+\end{figure}
+Starting with SSSP which is the algorithm that really is at an advantage when using many threads in \autoref{fig:galoisSpeedupSSSP}.
+
+For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads. 
+Twitter has the best speedup, requiring only 38\%\ the calculation time with 2 threads compared to one, 25\%\ using 4 and 12.9\%\ using 8 threads.
+Behaviour on friendster is similar with 52\%\ at 2 threads, 29\%\ at 4 and 16\%\ at 8 threads compared to one.
+
+Anything above 32 threads however no longer helps decrease the computation time, in some cases even the opposite e.g.
+calculation on rMat28 is actually slower with 48 (13\% slower) or 96 threads (28\% slower) compared to 40 threads.
+
+Small graphs like flickr or orkut, neither benefit from more threads nor is the performance held up by synchronization overhead.
+
+\subsubsection{Breadth-first search}
+\begin{figure}
 	\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
-		\caption{Breadth-first search}
-		\label{fig:singleNodeBFSGaloisThreads}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths and Breadth-first search algorithms.}
+	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search}
+	\label{fig:galoisSpeedupBFS}
 \end{figure}
@@
 		\caption{PageRank Pull}
-		\label{fig:singleNodePRPullGaloisThreads}
+		\label{fig:galoisSpeedupPRPull}
 	\end{subfigure}
@@
 		\caption{PageRank Push}
-		\label{fig:singleNodePRPushGaloisThreads}
+		\label{fig:galoisSpeedupPRPush}
 	\end{subfigure}

### 2020-08-27 plots/makedata.py

@@ -402,7 +402,6 @@ for i in range(len(graph)):
         continue
     
     xVal = int(re.sub('\D', '', algo[i]))
-
     
     if 'sssp' in algo[i]:
         speedupGaloisSSSP[graph[i]][xVal] = calcTime[i]
@@ -412,6 +411,7 @@ for i in range(len(graph)):
         speedupGaloisPRPush[graph[i]][xVal] = calcTime[i]
     elif 'pagerank-pull' in algo[i]:
         speedupGaloisPRPull[graph[i]][xVal] = calcTime[i]
+
 
 
 #print(speedupGaloisPRPush)
@@ -427,3 +427,18 @@ for g in graphs:
         speedupGaloisPRPush[g][x_] = tPRPush / speedupGaloisPRPush[g][x_]
         speedupGaloisPRPull[g][x_] = tPRPull / speedupGaloisPRPull[g][x_]
 
+
+if False:
+    for g in graphs:
+        if g == 'flickr' or g == 'orkut':
+            continue
+        tmp = speedupGaloisBFS[g]
+        print(g)
+        last = 1
+        xs = []
+        for k in tmp:
+            xs.append(k)
+        xs.sort()
+        for k in xs:
+            print(k, tmp[k], tmp[k] / tmp[last])
+            #last = k

### 2020-08-27 plots/makeplots.py

@@ -4,16 +4,16 @@ from plotFunctions import *
 
 
 
-distributed = True
-singleNode = True
+distributed = False
+singleNode = False
 
 sssp = True
-bfs = True
-pr = True
+bfs = False
+pr = False
 
-calcTimes = True
-execTimes = True
-overheads = True
+calcTimes = False
+execTimes = False
+overheads = False
 
 
 galoisSpeedup = True
@@ -77,7 +77,9 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 ## GALOIS THREAD COUNT
 if galoisSpeedup:
-	line_plot(graphs, x, speedupGaloisSSSP, title='Speedup SSSP', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads.png")
+	line_plot(graphs, x, speedupGaloisSSSP, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads.png")
+
+	line_plot(graphs, x[:-5], speedupGaloisSSSP, title='Speedup SSSP', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads_short.png")
 
 	line_plot(graphs, x, speedupGaloisBFS, title='SpeedupBFS', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeBFSGaloisThreads.png")
 
@@ -95,7 +97,7 @@ if distributed and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 	#SSSP
 
 	if sssp or calcTimes:
-		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), calcTimeSSSP_distributed, yErrs=yErrCalcSSSP_distributed, title='SSSP distributed', yLabel='Calculation time (s)', saveToFile="distributedSSSP_calcTime.png")
+		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), calcTimeSSSP_distributed, yErrs=yErrCalcSSSP_distributed, yLabel='Calculation time (s)', saveToFile="distributedSSSP_calcTime.png")
 
 	#BFS
 	if bfs or calcTimes:
[…]

### 2020-08-28 documentation/paper/main.tex
Plots updated with giraph results
erste fixes der Rückmeldung

@@
 \usepackage{subcaption}
+\usepackage{amssymb}
 
@@
 \IEEEauthorblockA{(3315161)\\
-@stud.uni-stuttgart.de}
+st155698@stud.uni-stuttgart.de}
 \and
@@
 \begin{abstract}
-In this paper we will analyze and compare the non-uniform memory access (NUMA) aware graph frameworks Galois, Ligra, Polymer and Gemini in their parformance.
-
-Giraph
-
-All the frameworks will be tested in shared memory. Galois, Gemini and Giraph are tested on a distributed cluster as well.
-We will give some insight on the calculation performance and associated time overhead of each framework.
-For this we compare the frameworks in their performance with Single-source shortest-paths, Breadth-first search and PageRank algorithms.
+TODO
 \end{abstract}
@@
 \section{Introduction}
-In recent years graph sizes have increased significantly, thus performance and memory efficiency of the graph analysis applications is now more important than ever.
+In recent years graph sizes have increased significantly [?] .
+Thus performance and memory efficiency of the graph analysis applications is now more important than ever.
 
-We will compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
+Many applications like
+
+Wachsende graphen,
+benötigt für folgende Applikationen,
+verwenden graph algorithmen,
+diese müssen schnell und verteilt laufen
+dafür wurden Frameworks entwickelt...
+
+und das noch schön formulieren
+
+We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
 The comparison is performed on both real world data sets and synthetic graphs.
@@
 \begin{itemize}
-  \item Comparison of some of the most widely used graph based calculation frameworks
+  \item Comparison of several state-of-the-art graph processing frameworks
   \item
@@
 \end{itemize}
+
+\section{Preliminaries}
+
+\paragraph{Graphs and Paths}
+An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
+The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set becomes 
+\begin{equation*}
+  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
+\end{equation*}
+and in the \emph{undirected} case $E$ is a set of two-sets 
+\begin{equation*}
+  E\subseteq\{\{x,y\}\,|\, x,y\in V, x\neq y\}.
+\end{equation*}
+
+The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.
+
+Independently of the graph being directed or not, a graph can be \emph{weighted}.
+In this case the edge set is expanded by a numerical value $E_{\text{weighted}}=E_{\text{unweighted}}\times \mathbb R$ further describing the relation.
+
+The size of a graph is often described in the number of edges $|E|$ because this number is typically much larger than the number of vertices $|V|$.
+
+A \emph{Path} from start $s$ to target $t$ is a set of edges $P\subseteq E^n$ with
+\begin{equation*}
+   P=((x_0,x_1),(x_1,x_2),\ldots, (x_{n-1},x_n))
+\end{equation*}
+where all $(x_i,x_{i+1})\in E$ and $x_0=s, x_n=t$.
+
+Thus we call a target $t$ \emph{accessible} from $s$ if a Path from $s$ to $t$ exists.
+
+\paragraph{Single-Source Shortest-Paths}
+Single-Source Shortest-Paths (SSSP) describes the problem of finding a path along some edges of the input graph from a given start node to a target.
+
+Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
+The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
+In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.
+
+The most common implementations are Dijkstra's algorithm or BellmanFord.
+
+\paragraph{Breadth-First Search}
+Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input. In some cases a target vertex is also given.
+
+The output is usually a set of vertices that are accessible from the start vertex. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.
+
+\paragraph{PageRank}
+
+\paragraph{Pregel Model}
+
+% Google, to overcome, these challenges came up with Pregel.
+% Provides scalability
+% Fault-tolerance
+% Flexibility to express arbitrary algorithms
+% The high level organization of Pregel programs
+% is inspired by Valiant’s Bulk Synchronous
+%  Parallel model 
+\paragraph{Congest Model}
 

### 2020-08-28 documentation/paper/sections/graphFormats.tex
erste fixes der Rückmeldung

--- 2020-08-10 documentation/paper/graphFormats.tex
@@
-%!TEX root = ./main.tex
+%!TEX root = ../main.tex
 
-\subsection{An overview of some graph formats}
-A rather big portion of our time was invested in figuring out which graph framework requires which graph formats.
-We thus decided to give an overview over all the formats we encountered, with explanation on how they represent the graph.
+\subsection{An Overview of Important Graph Formats}
+Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.
+Data Sets retreived from KONECT can be directly read and translated.
 
-Additionally, to make life in the future a little bit easier, we wrote multiple tools to convert graphs acquired from Snap or Konect to the required formats.
-Additional information on this is available in the section \nameref{supplementaryData} at the end.
+The following sections explain the output formats of our conversion tool.
+\subsubsection{AdjacencyList}
+The AdjacencyList and WeightedAdjacencyList formats are used by Ligra and Polymer.
+The format was initially specified for the Problem Based Benchmark Suite, an open source repository to compare different parallel programming methodologies in terms of performance and code quality \cite{pbbs}.
 
-\subsubsection{AdjacencyList}
-The AdjacencyList and WeightedAdjacencyList formats\cite{adjListFormat} are used by Ligra and Polymer.
-They represent the directed edges of a graph as a number of offsets that point to a set of target nodes in the file.
-First the file contains the number of vertices $n$ and edges $m$, followed by an offset for each vertex.
-This offset specifies at what point in the following list of numbers the information for a node begins.
-Lastly the file format contains a list of target nodes.
-The numbers are all separated by newlines.
-\begin{gather*}
-n\\
-m\\
-o_1\\
-o_2\\
-\vdots\\
-o_n\\
-t_1\\
-t_2\\
-\vdots\\
-t_m
-\end{gather*}
-The offsets $o_i=k$ and $o_{i+1}=k+j$ mean that vertex $i$ has $j$ outgoing edges, these edges are
+The file looks as follows
 \begin{equation*}
-	(i,t_k),(i,t_{k+1}),\ldots,(i,t_{k+j-1})
+	n, m, o_1, \ldots, o_n, t_1, \ldots, t_m
+\end{equation*}
+where commas are \texttt{\textbackslash n}. First, $n$ is the number of vertices and $m$ the number of edges in the graph.
+
+The $o_k$ are the so-called offsets. Each vertex $k$ has an offset $o_k$, that describes an index in the following list of the $t_i$.
+The $t_i$ are vertex IDs describing target nodes of a directed edge. 
+The index $o_k$ in the list of target nodes is the point where edges outgoing from vertex $k$ begin to be declared. So vertex $k$ has the outgoing edges
+\begin{equation*}
+	(k, t_{o_k}), (k, t_{o_k+1}),\ldots, (k, t_{o_{k+1}-1}).
 \end{equation*}
 
-For the WeightedAdjacencyList format, the weights are just appended to the end of the file in the same order as the edges.
+For the WeightedAdjacencyList format, the weights are appended to the end of the file in an order corresponding to the target nodes.
 
 \subsubsection{EdgeList}
-The EdgeList format is probably the easiest to understand and is one of the most commonly used in the online graph repositories.
-The directed eges $(s_1,t_1),(s_2,t_2),\ldots$ or $(s_1,t_1, w_1),(s_2,t_2, w_2),\ldots$ are represented in the following way.
-\begin{align*}
-  &s_1\ \ t_1\ \ w_1\\
-  &s_2\ \ t_2\ \ w_m\\
-	&\quad\vdots\\
-  &s_m\ \ t_m\ \ w_m
-\end{align*}
-The weights are optional, everything is ASCI encoded and the inline delimiter is a variable amount of any whitespace.
+The EdgeList format is the most intuitive and one of the most commonly used in online data set repositories.
+The KONECT database uses this format and thus it is the input format for our conversion tool.
+
+An edge list is a set of directed eges $(s_1,t_1),(s_2,t_2),\ldots$ where $s_i$ is a vertex ID representing the start vertex and $t_i$ is a vertex ID representing the target vertex.
+In the format, there is one edges per line and the vertex IDs $s_i, t_i$ are separated with any whitespace character.
+
+For a WeightedEdgeList, the edge weights are appended to each line, again separated by a whitespace character.
 
 \subsubsection{Binary EdgeList}
-The binary EdgeList format is used by Gemini. Finding information on this format required reverse engineering of the Gemini code.
+The binary EdgeList format is used by Gemini.
 
-We found that Gemini requires the following input format
+For $s_i, t_i$ some vertex IDs and $w_i$ the weight of a directed edge $(s_i,t_i, w_i)$, Gemini requires the following input format
 \begin{equation*}
@@
 
-\subsubsection{Giraph's numerous I/O formats}
+\subsubsection{Giraph's I/O formats}
+Giraph is capable of parsing many different input and output formats. All of those are explained in Giraph's JavaDoc\footnote{\url{http://giraph.apache.org/apidocs/index.html}}.
+Both edge- and vertex-centric input formats are possible.
+
+One can even define their own input graph representation or output format. For the purposes of this paper, we used an existing format similar to AdjacencyList but represented in a JSON-like manner.
+
+In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values, \texttt{float} out-edge weights.
+Each line in the graph file looks as follows
+\begin{equation*}
+	[s,v_s,[[t_1, w_{t_1}], [t_2, w_{t_2}]...]]
+\end{equation*}
+with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.
+
+There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

### 2020-08-28 documentation/paper/sections/overview/giraph.tex
erste fixes der Rückmeldung

@@
 Apache Giraph is an iterative graph processing framework, built on top of Apache Hadoop\cite{Giraph}.
-Hadoop as a large MapReduce infrastructure provides a reliable (fault tolerant) basis for large scale graph processing.
+Hadoop as a large MapReduce infrastructure providing a reliable (fault tolerant) basis for large scale graph processing.
 Because of the underlying Hadoop MapReduce infrastructure, expanding single-node processing to a multi-node cluster is almost seamless.
@@
 
-Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will make the vertex active again.
-To end computation, a vote-to-halt method is applied. Each vertex outputs some local information (e.g. the final vertex value) as result.
+Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will reactivate the vertex again.
+A vote-to-halt method is applied, i.e. if all vertices are inactive or if a user defined superstep is reached the computation ends.
 
-Giraph is next to Galois probably the most actively developed framework in our comparison.
+Each vertex outputs some local information (e.g. the final vertex value) as result.
+
+Giraph being an Apache project makes it the most actively maintained and tested project in our comparison.
+Over the course of this project, several new updates were pushed to Giraph's source repository\footnote{https://gitbox.apache.org/repos/asf?p=giraph.git}.
+
+TODO: Facebook?

### 2020-08-28 documentation/paper/sections/results/bfs.tex
erste fixes der Rückmeldung

--- 2020-08-24 documentation/paper/sections/results/results.tex
@@
 %!TEX root=../../main.tex
-
-\subsection{The frameworks during setup and benchmark}
-We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.
-
-\begin{enumerate}
-	\item During setup and benchmark of Gemini, we encountered several bugs in the cloned repository. These include non zero-terminated strings or even missing return statements.
-
-	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code.
-A repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
-
-	\item Furthermore, we would like to address the setup of Hadoop for Giraph.
-It requires multiple edits in \texttt{xml} files that aren't easily automized.
-This makes the setup rather time consuming, especially if reconfiguration is needed later on.
-	\item In order for Giraph to run, several Java tasks (the Hadoop infrastructure) have to be constantly running in the background.
-While we don't expect this to have a significant performance impact on other tasks, it is still suboptimal.
-	%\item The Hadoop distributed file system (HDFS) ran us into disk space problems on multiple occasions.
-First, unless configured otherwise, the standard implementation replicates all data three times, distributed over all participating nodes.
-Second, deleting files on the HDFS does not immediately free up disk space because the files are moved to a \emph{recycling bin}-like location.
-
-	\item 
-\end{enumerate}
-
-On a plus side, setup of frameworks like Polymer or Ligra was straight forward and did not require any special treatment.
-
-\subsection{Single-source Shortest-paths}
-\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
-In these figures, Galois with 96 threads is shown.
-We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
-		\caption{Calculation times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
-		\caption{Execution times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.}
-	\label{fig:singleNodeSSSP}
-\end{figure*}
-
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
-		\caption{Calculation times for distributed SSSP}
-		\label{fig:distributedSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{Execution times for distributed SSSP}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
-	\label{fig:distributedSSSP}
-\end{figure*}
 
@@
 \end{figure*}
-
-\subsection{PageRank}
-
-\subsection{Galois speedup}
-\label{sec:galois_speedup}
-For Galois we
-
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
-		\caption{Single-source Shortest-paths}
-		\label{fig:singleNodeSSSPGaloisThreads}
-	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
-		\caption{Breadth-first search}
-		\label{fig:singleNodeBFSGaloisThreads}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths and Breadth-first search algorithms.}
-\end{figure}
-
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPullGaloisThreads.png}
-		\caption{PageRank Pull}
-		\label{fig:singleNodePRPullGaloisThreads}
-	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPushGaloisThreads.png}
-		\caption{PageRank Push}
-		\label{fig:singleNodePRPushGaloisThreads}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms.}
-\end{figure}
-
-\subsection{Execution times and overhead}
-
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTime.png}
-	\caption{Overhead SSSP single node}
-	\label{fig:singleNodeSSSP_overhead}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTime.png}
-	\caption{Overhead BFS single node}
-	\label{fig:singleNodeBFS_overhead}
-\end{figure}
-
-hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
-Aber vielleicht ist auch die setup time bei gleichen frameworks zwischen verteilt und shared memory ganz interessant zu vergleichen.

### 2020-08-28 documentation/paper/sections/results/galoisSpeedup.tex
erste fixes der Rückmeldung

--- 2020-08-24 documentation/paper/sections/results/results.tex
@@
 %!TEX root=../../main.tex
-
-\subsection{The frameworks during setup and benchmark}
-We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.
-
-\begin{enumerate}
-	\item During setup and benchmark of Gemini, we encountered several bugs in the cloned repository. These include non zero-terminated strings or even missing return statements.
-
-	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code.
-A repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
-
-	\item Furthermore, we would like to address the setup of Hadoop for Giraph.
-It requires multiple edits in \texttt{xml} files that aren't easily automized.
-This makes the setup rather time consuming, especially if reconfiguration is needed later on.
-	\item In order for Giraph to run, several Java tasks (the Hadoop infrastructure) have to be constantly running in the background.
-While we don't expect this to have a significant performance impact on other tasks, it is still suboptimal.
-	%\item The Hadoop distributed file system (HDFS) ran us into disk space problems on multiple occasions.
-First, unless configured otherwise, the standard implementation replicates all data three times, distributed over all participating nodes.
-Second, deleting files on the HDFS does not immediately free up disk space because the files are moved to a \emph{recycling bin}-like location.
-
-	\item 
-\end{enumerate}
-
-On a plus side, setup of frameworks like Polymer or Ligra was straight forward and did not require any special treatment.
-
-\subsection{Single-source Shortest-paths}
-\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
-In these figures, Galois with 96 threads is shown.
-We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
-		\caption{Calculation times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
-		\caption{Execution times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.}
-	\label{fig:singleNodeSSSP}
-\end{figure*}
-
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
-		\caption{Calculation times for distributed SSSP}
-		\label{fig:distributedSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{Execution times for distributed SSSP}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
-	\label{fig:distributedSSSP}
-\end{figure*}
-
-\subsection{Breadth-first search}
-In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
-		\caption{Calculation times for BFS on a single node}
-		\label{fig:singleNodeBFS_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
-		\caption{Execution times for BFS on a single node}
-		\label{fig:singleNodeBFS_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeBFS_overheadNormalized}
-	\end{subfigure}
-
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing}
-\end{figure*}
-
-\begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
-		\caption{Calculation times for distributed BFS}
-		\label{fig:distributedBFS_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
-		\caption{Execution times for distributed BFS}
-		\label{fig:distributedBFS_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedBFS_overheadNormalized}
-	\end{subfigure}
-
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
-\end{figure*}
-
-\subsection{PageRank}
 
@@
 \label{sec:galois_speedup}
-For Galois we
+Analyzing the calculation time speedups for Galois, we can compare how or if the different algorithms benefit from increasing thread numbers.
 
+\subsubsection{Single-source Shortest-path}
 \begin{figure}
-	\begin{subfigure}{\columnwidth}
 	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
-		\caption{Single-source Shortest-paths}
-		\label{fig:singleNodeSSSPGaloisThreads}
-	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
+	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths}
+	\label{fig:galoisSpeedupSSSP}
+\end{figure}
+Starting with SSSP which is the algorithm that really is at an advantage when using many threads in \autoref{fig:galoisSpeedupSSSP}.
+
+For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads. 
+Twitter has the best speedup, requiring only 38\%\ the calculation time with 2 threads compared to one, 25\%\ using 4 and 12.9\%\ using 8 threads.
+Behaviour on friendster is similar with 52\%\ at 2 threads, 29\%\ at 4 and 16\%\ at 8 threads compared to one.
+
+Anything above 32 threads however no longer helps decrease the computation time, in some cases even the opposite e.g.
+calculation on rMat28 is actually slower with 48 (13\% slower) or 96 threads (28\% slower) compared to 40 threads.
+
+Small graphs like flickr or orkut, neither benefit from more threads nor is the performance held up by synchronization overhead.
+
+\subsubsection{Breadth-first search}
+\begin{figure}
 	\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
-		\caption{Breadth-first search}
-		\label{fig:singleNodeBFSGaloisThreads}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths and Breadth-first search algorithms.}
+	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search}
+	\label{fig:galoisSpeedupBFS}
 \end{figure}
 
+Flickr is sped up by about 5\%\ with 4 threads, any more than that will actually decrease performance, making computation time up to 15\%\ (96 threads) longer.
+
+On all graphs, the speedup never exceeds 6$\times$ even when using 96 thread.
+The initial speedup when switching from one to two threads is actually smaller than one, thus decreasing performance, on 4 of 7 graphs.
+Only flickr, twitter and rMat28 can benefit slightly by a speedup of 2\%, 41\%\ and 5\%\ respectively.
+
+When comparing four threads to one, BFS on all but two graphs can be sped up by anywhere from 5\%\ (flickr) to 68\% (rMat28). The speedup is thus only possible to a very small degree.
+For the other two graphs, computation on friendster with 8 threads is just as fast as one thread and computation on rMat27 is actually 19\% slower.
+
+\subsubsection{PageRank}
 \begin{figure}
@@
 \end{figure}
-
-\subsection{Execution times and overhead}
-
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTime.png}
-	\caption{Overhead SSSP single node}
-	\label{fig:singleNodeSSSP_overhead}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTime.png}
-	\caption{Overhead BFS single node}
-	\label{fig:singleNodeBFS_overhead}
-\end{figure}
-
-hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
-Aber vielleicht ist auch die setup time bei gleichen frameworks zwischen verteilt und shared memory ganz interessant zu vergleichen.

### 2020-08-28 documentation/paper/sections/results/pr.tex
erste fixes der Rückmeldung

--- 2020-08-24 documentation/paper/sections/results/results.tex
@@
 
-\subsection{The frameworks during setup and benchmark}
-We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.
+\subsection{PageRank}
 
-\begin{enumerate}
-	\item During setup and benchmark of Gemini, we encountered several bugs in the cloned repository. These include non zero-terminated strings or even missing return statements.
-
-	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code.
-A repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
-
-	\item Furthermore, we would like to address the setup of Hadoop for Giraph.
-It requires multiple edits in \texttt{xml} files that aren't easily automized.
-This makes the setup rather time consuming, especially if reconfiguration is needed later on.
-	\item In order for Giraph to run, several Java tasks (the Hadoop infrastructure) have to be constantly running in the background.
-While we don't expect this to have a significant performance impact on other tasks, it is still suboptimal.
-	%\item The Hadoop distributed file system (HDFS) ran us into disk space problems on multiple occasions.
-First, unless configured otherwise, the standard implementation replicates all data three times, distributed over all participating nodes.
-Second, deleting files on the HDFS does not immediately free up disk space because the files are moved to a \emph{recycling bin}-like location.
-
-	\item 
-\end{enumerate}
-
-On a plus side, setup of frameworks like Polymer or Ligra was straight forward and did not require any special treatment.
-
-\subsection{Single-source Shortest-paths}
-\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
-In these figures, Galois with 96 threads is shown.
-We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}[t!]
+\begin{figure*}
 	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
-		\caption{Calculation times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_calc}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
+		\caption{Calculation times for PR on a single node}
+		\label{fig:singleNodePR_calc}
 	\end{subfigure}
@@
 	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
-		\caption{Execution times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_exec}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
+		\caption{Execution times for PR on a single node}
+		\label{fig:singleNodePR_exec}
 	\end{subfigure}
@@
 	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTimeNormalized.png}
 		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.}
-	\label{fig:singleNodeSSSP}
-\end{figure*}
-
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
-		\caption{Calculation times for distributed SSSP}
-		\label{fig:distributedSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{Execution times for distributed SSSP}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
-	\label{fig:distributedSSSP}
-\end{figure*}
-
-\subsection{Breadth-first search}
-In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
-		\caption{Calculation times for BFS on a single node}
-		\label{fig:singleNodeBFS_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
-		\caption{Execution times for BFS on a single node}
-		\label{fig:singleNodeBFS_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeBFS_overheadNormalized}
+		\label{fig:singleNodePR_overheadNormalized}
 	\end{subfigure}
@@
 	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
-		\caption{Calculation times for distributed BFS}
-		\label{fig:distributedBFS_calc}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
+		\caption{Calculation times for distributed PR}
+		\label{fig:distributedPR_calc}
 	\end{subfigure}
@@
 	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
-		\caption{Execution times for distributed BFS}
-		\label{fig:distributedBFS_exec}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
+		\caption{Execution times for distributed PR}
+		\label{fig:distributedPR_exec}
 	\end{subfigure}
@@
 	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTimeNormalized.png}
 		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedBFS_overheadNormalized}
+		\label{fig:distributedPR_overheadNormalized}
 	\end{subfigure}

### 2020-08-28 documentation/paper/sections/results/results.tex
erste fixes der Rückmeldung

--- 2020-08-24 documentation/paper/sections/results/results.tex
@@
 
-\subsection{Single-source Shortest-paths}
-\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
-In these figures, Galois with 96 threads is shown.
-We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
-		\caption{Calculation times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
-		\caption{Execution times for SSSP on a single node}
-		\label{fig:singleNodeSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.}
-	\label{fig:singleNodeSSSP}
-\end{figure*}
+\clearpage
 
-\begin{figure*}[t!]
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
-		\caption{Calculation times for distributed SSSP}
-		\label{fig:distributedSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{Execution times for distributed SSSP}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedSSSP_overheadNormalized}
-	\end{subfigure}
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
-	\label{fig:distributedSSSP}
-\end{figure*}
+\input{sections/results/sssp}
 
-\subsection{Breadth-first search}
-In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
-		\caption{Calculation times for BFS on a single node}
-		\label{fig:singleNodeBFS_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
-		\caption{Execution times for BFS on a single node}
-		\label{fig:singleNodeBFS_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeBFS_overheadNormalized}
-	\end{subfigure}
+\input{sections/results/bfs}
 
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing}
-\end{figure*}
+\input{sections/results/pr}
 
-\begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
-		\caption{Calculation times for distributed BFS}
-		\label{fig:distributedBFS_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
-		\caption{Execution times for distributed BFS}
-		\label{fig:distributedBFS_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedBFS_overheadNormalized}
-	\end{subfigure}
+\input{sections/results/galoisSpeedup}
 
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
-\end{figure*}
+% \begin{figure}
+% 	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTime.png}
+% 	\caption{Overhead SSSP single node}
+% 	\label{fig:singleNodeSSSP_overhead}
+% \end{figure}
 
-\subsection{PageRank}
-
-\subsection{Galois speedup}
-\label{sec:galois_speedup}
-For Galois we
-
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
-		\caption{Single-source Shortest-paths}
-		\label{fig:singleNodeSSSPGaloisThreads}
-	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
-		\caption{Breadth-first search}
-		\label{fig:singleNodeBFSGaloisThreads}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths and Breadth-first search algorithms.}
-\end{figure}
-
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPullGaloisThreads.png}
-		\caption{PageRank Pull}
-		\label{fig:singleNodePRPullGaloisThreads}
-	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPushGaloisThreads.png}
-		\caption{PageRank Push}
-		\label{fig:singleNodePRPushGaloisThreads}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms.}
-\end{figure}
-
-\subsection{Execution times and overhead}
-
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTime.png}
-	\caption{Overhead SSSP single node}
-	\label{fig:singleNodeSSSP_overhead}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTime.png}
-	\caption{Overhead BFS single node}
-	\label{fig:singleNodeBFS_overhead}
-\end{figure}
-
-hier erhoffe ich mir einen Vergleich der Ladezeiten und erwarte, dass Systeme wie Giraph, die erstmal auf irgendwas warten schlecht abschneiden.
-Aber vielleicht ist auch die setup time bei gleichen frameworks zwischen verteilt und shared memory ganz interessant zu vergleichen.
+% \begin{figure}
+% 	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTime.png}
+% 	\caption{Overhead BFS single node}
+% 	\label{fig:singleNodeBFS_overhead}
+% \end{figure}

### 2020-08-28 documentation/paper/sections/results/sssp.tex
erste fixes der Rückmeldung

%!TEX root=../../main.tex

\subsection{Single-source Shortest-paths}

\subsubsection{Single-node}
\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks. In these figures, Galois with 96 threads is shown. We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
\begin{figure*}[ht!]
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
		\caption{Calculation times for SSSP on a single node}
		\label{fig:singleNodeSSSP_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
		\caption{Execution times for SSSP on a single node}
		\label{fig:singleNodeSSSP_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
		\caption{Overhead time normalized by the graph size in million edges}
		\label{fig:singleNodeSSSP_overheadNormalized}
	\end{subfigure}
	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.
	The runs on rMat28 for Ligra and Polymer failed and the frameworks were unable to complete the task.}
	\label{fig:singleNodeSSSP}
\end{figure*}

\subsubsection{Distributed}
\begin{figure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
		\caption{Calculation times for distributed SSSP}
		\label{fig:distributedSSSP_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
		\caption{Execution times for distributed SSSP}
		\label{fig:distributedSSSP_exec}
	\end{subfigure}
	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
	\label{fig:distributedSSSP}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalizedToGalois.png}
	\caption{Overhead times of each framework normalized by the overhead time of Galois Push}
	\label{fig:distributedSSSP_overhead}
\end{figure}

For the distributed scenario, \autoref{fig:distributedSSSP} shows the benchmark results as calculation and execution times.

Results are especially interesting for Giraph since it seems to not cope well with synthetic graphs. Analyzing the computation times in \autoref{fig:distributedSSSP_calc}, we see that it is the fastest framework on our real-world graphs. And that with a considerable margin of other frameworks always taking at least 50\% longer (Gemini on flickr) up to Galois Pull needing 18$\times$ longer on wikipedia.  
On both synthetic graphs however, Giraph is the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.

While Giraph's computation times are very competitive, when comparing the execution times in \autoref{fig:distributedSSSP_exec} we see that Giraph is actually the slowest framework on 5 out of 7 graphs. For the other two, namely twitter and friendster, Giraph is second slowest with only Gemini taking longer to complete.

Giraph and Gemini's very long execution times are only due to their overhead being many orders of magnitude larger than Galois overhead (\autoref{fig:distributedSSSP_overhead}).
Overhead for Gemini is greater than that of Galois on every graph. From just a 20\% increase on flickr up to friendster, where the overhead is 90$\times$ that of Galois Push.
For Giraph the overhead times are not as extreme but still generally worse. Even on flickr, Giraph's overhead time is already 23$\times$ that of Galois. On friendster, where Gemini was worst, Giraph \emph{only} requires 73$\times$ the overhead time of Galois.

### 2020-08-28 documentation/paper/sections/testingMethods.tex
erste fixes der Rückmeldung

@@
 
-\subsection{Hardware and Software}
-For testing all systems we used 5 machines with 96 CPU cores each (48 physical) and 256 GB of RAM.
+\subsection{Hardware}
+For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each.
 One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
@@
 
-All benchmarks were initiated by our benchmark script that is available in our repository. Galois, Gemini and Giraph were benchmarked on both the 5-node cluster and a singe machine.
+\subsection{Benchmark Software}
+All benchmarks were initiated by our benchmark script, that is available in our repository. Galois, Gemini and Giraph were benchmarked on both the 5-node cluster and a singe machine.
 Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
 
-The complete benchmark log files and extracted raw results will be available in our repository as well.
+The complete benchmark log files and extracted raw results are available in our repository.
+
+\subsection{Data Sets}
+The graphs used in our testing can be seen in detail in \autoref{tbl:graphs}.
+We included a variaty of different graph sizes, from relatively small graphs like the flickr graph with 100 thousand edges up to an rMat28 with $4.2$ billion edges.
+All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
+Both the rMat27 and rMat28 were created with a modified version (we changed the output format to EdgeList) of a graph generator provided by Ligra.
+\begin{table}
+	\centering
+	\caption{Size Comparison of the Used Graphs}
+	\begin{tabular}{crr}
+		\hline
+		\bf{Graph}&\# Vertices (M)&\# Edges (M)\\\hline
+		flickr&    		0.1&  2\\
+		orkut&          3&    117\\
+		wikipedia&      12&   378\\
+		twitter&     	52&   1963\\
+		rMat27&         63&   2147\\
+		friendster&     68&   2586\\
+		rMat28&         121&  4294\\
+		\hline
+	\end{tabular}
+	\label{tbl:graphs}
+\end{table}
 
@@
 
-\subsection{Graphs}
-The graphs used in our testing can be seen in detail in \autoref{tbl:graphs}.
-We included a variaty of different graph sizes, from relatively small graphs like the flickr graph up to an rMat28 with $4.2$ billion edges.
-All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
-Both the rMat27 and rMat28 were created with a modified version (we changed the output format to EdgeList) of a graph generator provided by Ligra.
-\begin{table}
-	\centering
-	\begin{tabular}{crr}
-		\hline
-		\bf{Graph}&\# Vertices (M)&\# Edges (M)\\\hline
-		flickr&    		0.1&  2\\
-		orkut&          3&    117\\
-		wikipedia&      12&   378\\
-		twitter&     	52&   1963\\
-		rMat27&         63&   2147\\
-		friendster&     68&   2586\\
-		rMat28&         121&  4294\\
-		\hline
-	\end{tabular}
-	\caption{Size comparison of the used graphs}
-	\label{tbl:graphs}
-\end{table}
-
 \subsection{Algorithms}
@@
 \begin{table}
-TODO
+	\renewcommand{\arraystretch}{1.3}
 	\centering
+	\caption{Tested Algorithms of Each Framework}
 	\begin{tabularx}{0.9\columnwidth}{cYYYYY}
 		\bf{Algorithm}&Galois&Gemini&Giraph&Ligra&Polymer\\\hline
-		PageRank&$\Delta$Push, $\Delta$Pull&&&&\\
-		SSSP&?, Push, Pull&&&&\\
-		BFS&Yes&&*&&\\
+		PageRank&$\Delta$Push, $\Delta$Pull&&&&Regular, $\Delta$\\\hline
+		SSSP&?, Push, Pull&&&&\\\hline
+		BFS&Yes&&Yes*&&\\
 		\hline
 	\end{tabularx}
-	\caption{Tested algorithms of the frameworks}
 	\label{tbl:algorithms}
 
-	* Algorithm not natively supported
+	{\vspace*{1ex}* Algorithm not natively supported
+
+	\txtdag{} Modified algorithm
+
+	\textsuperscript{d} Single node and distributed
+	}
 

### 2020-08-28 scripts/benchmark/benchmark_cases.sh
more benchmark cases

@@ -607,9 +607,33 @@ benchmark_galois_cpu() {
   done
 }
 
+benchmark_giraph_bfs_dist() {
+  for startnode in ${startnodes_flickr[@]}; do
+    giraph-bfs-dist flickr $startnode
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    giraph-bfs-dist friendster $startnode
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    giraph-bfs-dist orkut $startnode
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    giraph-bfs-dist twitter $startnode
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    giraph-bfs-dist wikipedia $startnode
+  done
+  for startnode in ${startnodes_rMat27[@]}; do
+    giraph-bfs-dist rMat27 $startnode
+  done
+  for startnode in ${startnodes_rMat28[@]}; do
+    giraph-bfs-dist rMat28 $startnode
+  done
+}
+
 
 
 time_start=$(get_time)
-benchmark_giraph_pagerank_dist
+benchmark_giraph_bfs_dist
 dur_exec=$((($(get_time)-$time_start)/1000000))
 echo "benchmark took $dur_exec seconds"

### 2020-08-29 documentation/paper/main.tex
preliminaries, more results, giraph and testingMethods reworked

--- 2020-08-21 documentation/paper/main.tex
@@
-\documentclass[conference,a4paper]{IEEEtran}
+\documentclass[a4paper]{IEEEtran}
 
@@
 \usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{amssymb}
 
@@
 \newcolumntype{Y}{>{\centering\arraybackslash}X} % cell type Y for centered text with automatic line breaks
-\newcommand*\rot{\rotatebox{90}} % rotates content 90deg ccw, mainly for table headers
+\newcommand*\rot{\rotatebox{45}} % rotates content 90deg ccw, mainly for table headers
 \newcommand*\OK{\ding{51}} % a checkmark-style sign
@@
 \author{\IEEEauthorblockN{Simon König}
-\IEEEauthorblockA{(3344789)\\
+\IEEEauthorblockA{(3344789)
 st156571@stud.uni-stuttgart.de}
+
 \and
+
 \IEEEauthorblockN{Leon Matzner}
-\IEEEauthorblockA{(3315161)\\
-@stud.uni-stuttgart.de}
+\IEEEauthorblockA{(3315161)
+st155698@stud.uni-stuttgart.de}
+
 \and
+
 \IEEEauthorblockN{Felix Rollbühler}
-\IEEEauthorblockA{(3310069)\\
+\IEEEauthorblockA{(3310069)
 st154960@stud.uni-stuttgart.de}
+
 \and
+
 \IEEEauthorblockN{Jakob Schmid}
-\IEEEauthorblockA{(3341630)\\
-@stud.uni-stuttgart.de}}
+\IEEEauthorblockA{(3341630)
+st157100@stud.uni-stuttgart.de}}
 
@@
 \begin{abstract}
-In this paper we will analyze and compare the graph frameworks Galois, Ligra, Polymer, Gemini and Giraph in their parformance.
-All the frameworks will be tested in shared memory and Galois, Gemini and Giraph are tested on a distributed cluster as well.
-Furthermore we will give some insight on the complexity of writing custom applications based on these frameworks.
+TODO
 \end{abstract}
@@
 
+\section*{TODOs}
+\begin{itemize}
+  \item Appendix für Setup Guides - md zu pdf
+\end{itemize}
+
 \section{Introduction}
-%\todo
-In recent years graph sizes have increased significantly, thus performance and memory efficiency of the graph analysis applications is now more important than ever.
+% In recent years graph sizes have increased significantly [?] .
+% Thus performance and memory efficiency of the graph analysis applications is now more important than ever.
+
+% Many applications like
+
+% Wachsende graphen,
+% benötigt für folgende Applikationen,
+% verwenden graph algorithmen,
+% diese müssen schnell und verteilt laufen
+% dafür wurden Frameworks entwickelt...
+
+% und das noch schön formulieren
+
+% We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
+% The comparison is performed on both real world data sets and synthetic graphs.
+
+% To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
+
+%One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
+%This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.
+
+% To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.
 
@@
 
-\section{Overview of the framworks}
-\subsection{Galois and Gluon}
-\input{sections/overview/galois}
+\input{sections/preliminiaries}
 
-\subsection{Gemini}
-\input{sections/overview/gemini}
+\input{sections/overview/overview}
 
-\subsection{Giraph}
-\input{sections/overview/giraph}
-
-\subsection{Ligra}
-\input{sections/overview/ligra}
-
-\subsection{Polymer}
-\input{sections/overview/polymer}
-
-\section{And}
-\input{sections/graphFormats}%hier oder supplementary data?
-
-\section{Testing methods}
 \input{sections/testingMethods}
 
-\section{Results}
 \input{sections/results/results}
@@
 \section{Conclusion}
-The conclusion goes here.
 
-\section*{Acknowledgment}
+\clearpage
+
+\section*{Acknowledgments}
 We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Ligra}, Polymer~\cite{Polymer}, Gemini~\cite{Gemini} as well as Apache Giraph~\cite{Giraph}.
@@
 
-\section*{Supplementary Data}\label{supplementaryData}
+\appendices
+\section{Installation Guides}\label{app:installationGuides}
+The complete installation guides with steps on how to run the same algorithms we used is available in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}
+
+In case of frameworks like Giraph, where for example no BFS algorithm is provided by the framework itself, the algorithm and a guide on how to include it in the framework can be found in the repository as well.
+
+\section{Conversion Tool}\label{app:conversionTool}
+In order to
+
 We have written a number of conversion tools and installation guides to help
@@
 % Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
-% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
-% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
 \begin{thebibliography}{1}
@@
 
+\bibitem{bsp}
+\BIBentryALTinterwordspacing
+L.~G. Valiant, ``A bridging model for parallel computation,'' \emph{Commun.
+  ACM}, vol.~33, no.~8, p. 103–111, Aug. 1990. [Online]. Available:
+  \url{https://doi.org/10.1145/79173.79181}
+\BIBentrySTDinterwordspacing
+
 \bibitem{Galois}
@@
 
-\bibitem{Polymer}
+\bibitem{Gemini}
 \BIBentryALTinterwordspacing
-K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
-  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
-  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
-  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
-  2015, p. 183–193. [Online]. Available:
-  \url{https://doi.org/10.1145/2688500.2688507}
+X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
+  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
+  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
+  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
+  301--316. [Online]. Available:
+  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
 \BIBentrySTDinterwordspacing
 
-\bibitem{adjListFormat}
+\bibitem{Giraph}
 \BIBentryALTinterwordspacing
-J.~Shun, G.~Blelloch, J.~Fineman, P.~Gibbons, A.~Kyrola, K.~Tangwonsan, and
-  H.~V. Simhadri. (2020, Jun.) {Problem Based Benchmark Suite}. graphIO.html.
-  [Online]. Available: \url{http://www.cs.cmu.edu/~pbbs/benchmarks/}
+{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
+  Available: \url{https://giraph.apache.org}
 \BIBentrySTDinterwordspacing
@@
 
-\bibitem{Gemini}
+\bibitem{Polymer}
 \BIBentryALTinterwordspacing
-X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
-  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
-  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
-  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
-  301--316. [Online]. Available:
-  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
+K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
+  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
+  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
+  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
+  2015, p. 183–193. [Online]. Available:
+  \url{https://doi.org/10.1145/2688500.2688507}
 \BIBentrySTDinterwordspacing
 
-\bibitem{Giraph}
+\bibitem{pbbs}
 \BIBentryALTinterwordspacing
-A.~S. Foundation. (2020, Jun.) {Apache Giraph}. [Online]. Available:
-  \url{https://giraph.apache.org}
+J.~Shun, G.~Blelloch, J.~Fineman, P.~Gibbons, A.~Kyrola, K.~Tangwonsan, and
+  H.~V. Simhadri. (2020, Aug.) {Problem Based Benchmark Suite }. [Online].
+  Available: \url{http://www.cs.cmu.edu/~pbbs/}
 \BIBentrySTDinterwordspacing

### 2020-08-29 documentation/paper/sections/graphFormats.tex
preliminaries, more results, giraph and testingMethods reworked

@@
 
-\subsection{An Overview of Important Graph Formats}
+\subsection{Important Graph Formats}
 Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.

### 2020-08-29 plots/makedata.py

@@ -155,14 +155,14 @@ overheadSSSP_distributed_normalizedToGalois = []
 i = 0
 for k in dist_frameworks_sssp:
     overheadSSSP_distributed_normalizedToGalois.append([])
-    #print(dist_frameworks_sssp[k])
+    #print(dist_frameworks_sssp[k], end='')
     for j in range(len(overheadSSSP_distributed[i])):
-        #print(graphs[j] + ": ", calcTimeSSSP_distributed[i][j]/calcTimeSSSP_distributed[2][j])
+        #print("&", round(execTimeSSSP_distributed[i][j]/execTimeSSSP_distributed[0][j] * 100) / 100, end='')
         overheadSSSP_distributed_normalizedToGalois[i].append(overheadSSSP_distributed[i][j]/overheadSSSP_distributed[0][j]) 
     i += 1
+    #print("\\\\")
 
-
-
+#print(graphs)
 
 
 

### 2020-08-29 documentation/paper/sections/overview/giraph.tex

@@
 
-Apache Giraph is an iterative graph processing framework, built on top of Apache Hadoop\cite{Giraph}.
-Hadoop as a large MapReduce infrastructure providing a reliable (fault tolerant) basis for large scale graph processing.
-Because of the underlying Hadoop MapReduce infrastructure, expanding single-node processing to a multi-node cluster is almost seamless.
+Apache Giraph is an iterative graph processing framework, built on top of Apache Hadoop \cite{Giraph}.
+Hadoop is a large MapReduce infrastructure providing a reliable (fault tolerant) basis for large scale graph processing.
 
-Giraph is based on computation units that communicate using messages and are synchonized with barriers.
+TODO // MapReduce hier oder wo anders erklären?
+MapReduce is a in three phases
+\begin{enumerate}
+	\item Map phase: The input data is distributed between a set of map processes, the map functionality is specified by the user.
+Ideally all map processes run in parallel so the map processes need to be independent.
+Results from this phase are written into (multiple) intermediate storage points.
+	\item Shuffle phase: The results are grouped according to a key provided by the map algorithm. Each set of results is then handed to one system for the next phase.
+	\item Reduce phase: Every set of intermediate results is input to exactly one reduce process. The reduce functionality is again specified by the user and ideally runs in parallel.
+\end{enumerate}
+Giraph uses the Map functionality of Hadoop to run the algorithms, Reduce is only used as an identity function.
+
+Giraph's computation model is closely related to the BSP model discussed in \autoref{sec:bsp}.
+This means that Giraph is based on computation units that communicate using messages and are synchonized with barriers.
 
@@
 A vote-to-halt method is applied, i.e. if all vertices are inactive or if a user defined superstep is reached the computation ends.
+Once calculation is finished, each vertex outputs some local information (e.g. the final vertex value) as result.
 
-Each vertex outputs some local information (e.g. the final vertex value) as result.
+Because of the underlying Hadoop infrastructure, expanding single-node processing to a multi-node cluster is seamless.
+Hadoop supplies a distributed file system (HDFS), on which all computation is performed.
+Giraph is thus, even when only using a single node, running in a distributed manner.
 
 Giraph being an Apache project makes it the most actively maintained and tested project in our comparison.
-Over the course of this project, several new updates were pushed to Giraph's source repository\footnote{https://gitbox.apache.org/repos/asf?p=giraph.git}.
+Over the course of this project, several new updates were pushed to Giraph's source repository\footnote{\url{https://gitbox.apache.org/repos/asf?p=giraph.git}}.
 

### 2020-08-29 documentation/paper/sections/overview/overview.tex

%!TEX root=../../main.tex

\section{Overview}
We start with a short overview describing functionality and characteristics of each framework, followed by their required input data formats.

//
\subsection{Galois and Gluon}
\input{sections/overview/galois}

\subsection{Gemini}
\input{sections/overview/gemini}

\subsection{Giraph}
\input{sections/overview/giraph}

\subsection{Ligra}
\input{sections/overview/ligra}

\subsection{Polymer}
\input{sections/overview/polymer}

\input{sections/graphFormats}

### 2020-08-29 documentation/paper/sections/preliminiaries.tex

%!TEX root=../main.tex

\section{Preliminaries}
TODO //

\subsection{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set becomes 
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case $E$ is a set of two-sets 
\begin{equation*}
  E\subseteq\{\{x,y\}\,|\, x,y\in V, x\neq y\}.
\end{equation*}

The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.

Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case the edge set is expanded by a numerical value $E_{\text{weighted}}=E_{\text{unweighted}}\times \mathbb R$ further describing the relation.

The size of a graph is often described in the number of edges $|E|$ because this number is typically much larger than the number of vertices $|V|$.

A \emph{Path} from start $s$ to target $t$ is a set of edges $P\subseteq E^n$ with
\begin{equation*}
   P=((x_0,x_1),(x_1,x_2),\ldots, (x_{n-1},x_n))
\end{equation*}
where all $(x_i,x_{i+1})\in E$ and $x_0=s, x_n=t$.

Thus we call a target $t$ \emph{accessible} from $s$ if a Path from $s$ to $t$ exists.

\subsection{Single-Source Shortest-Paths}
Single-Source Shortest-Paths (SSSP) describes the problem of finding a path along some edges of the input graph from a given start node to a target.

Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.

The most common implementations are Dijkstra's algorithm or BellmanFord.

\subsection{Breadth-First Search}
Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input. In some cases a target vertex is also given.

The output is usually a set of vertices that are accessible from the start vertex. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.

It is called breadth-first search because the algorithm searches in a path length-based way. First all paths of length 1 i.e. all neighbors of the start node are checked before checking paths of length 2 and so on.

The search algorithm, where the paths of maximum length are checked first is called Depth-First Search.

\subsection{PageRank}
\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph.

The analogy is that the graph represents website pages of the Word Wide Web, that are hyperlinked between one another. A website that is more important is likely to receive more links from other webiste.
PageRank counts the number and quality of links to a page to estimate the importance of a page.

For example, Google Search uses PageRank to rank web pages in their search engine results.

The output of PageRank is a percentage for each vertex. This percentage, called the PageRank of the vertex, is the probability with which a web surfer starting at a random web page reaches this webpage (vertex). 
With a high probability the web surfer uses a random link from the web page they are currently on and with a smaller probability (called damping factor) they jump to a completely random web page.

\paragraph{Delta-PageRank}
?

\subsection{Bulk-Synchronous Parallel Model}
\label{sec:bsp}
The Bulk-Synchronous Parallel (BSP) model is a computation model developed by Leslie Valiant \cite{bsp}. It is commonly used in computation environment with large amounts of synchronous computation, for example Giraph is based on this model.

This model describes components, a communication network between those components and a method of synchronization.
The components are capable of performing computations and transactions on \emph{local} memory. Pairs of components can only communicate using messages, thus remote memory access is also only possible in this way.
Synchronization is realized through barriers for some or all processes.

BSP algorithms are performed in a series of global supersteps. These consist of three steps, beginng with the processors performing local computations concurrently.
This step can overlap with the second, the communication between components. Processes can exchange information to access remote data.
Lastly, processes reaching a barrier wait until all other processes have reached the same barrier.

\subsection{Pregel Model}

% Google, to overcome, these challenges came up with Pregel.
% Provides scalability
% Fault-tolerance
% Flexibility to express arbitrary algorithms
% The high level organization of Pregel programs
% is inspired by Valiant’s Bulk Synchronous
%  Parallel model 
\subsection{Congest Model}

\subsection{Push/Pull Differences}

### 2020-08-29 documentation/paper/sections/results/sssp.tex

@@
 \subsection{Single-source Shortest-paths}
+For the results of our SSSP runs, we first analyze the single-node and distributed performances separately before comparing the two.
 
 \subsubsection{Single-node}
-\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks.
-In these figures, Galois with 96 threads is shown.
-We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
-\begin{figure*}[ht!]
+Beginning with the single-node performance, \autoref{fig:singleNodeSSSP} shows the average calculation and execution times for SSSP on the different frameworks.
+
+Because we did not see the perfomance going down when increasing thread counts on Galois, we always refer to Galois with 96 threads in this section (including the figures). 
+For a more detailed analysis of Galois' behaviour with different thread counts, see \autoref{sec:galois_speedup}.
+
+TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.
+\begin{figure}
 	\begin{subfigure}{0.3\textwidth}
@@
 	\label{fig:singleNodeSSSP}
-\end{figure*}
+\end{figure}
 
 \subsubsection{Distributed}
+
 \begin{figure}
+	\centering
 	\begin{subfigure}{\columnwidth}
@@
 	\end{subfigure}
-	\hfil
 	\begin{subfigure}{\columnwidth}
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalizedToGalois.png}
@@
 
-For the distributed scenario, \autoref{fig:distributedSSSP} shows the benchmark results as calculation and execution times.
+Moving on to the distributed cluster, \autoref{fig:distributedSSSP} shows the benchmark results as calculation and execution times.
 
-Results are especially interesting for Giraph since it seems to not cope well with synthetic graphs.
+Comparing the two Galois implementations, we find the calculation and execution times to be similar on smaller graphs and Push being the superior implementation for SSSP on larger data sets.
+Galois Pull being anywhere from just as fast to 3.5$\times$ slower on real-world data sets compared to the Push variant.
+The synthetic graphs are more extreme, execution times are close to 4$\times$ (rMat27) and 5$\times$ (rMat28) longer.
+
+Both Galois implementations have significantly smaller execution times compared to Gemini or Giraph on all graphs, see \autoref{tbl:ssspexec}.
+You can see Gemini being worse by at least a factor of 4 compared to Galois Push on all graphs except flickr.
+Giraph's execution times in comparison to this are even worse, taking at least 7$\times$ longer than Galois Push on all graphs.
+
+Evidently, Galois Push is the fastest algorithm in our lineup on 6 out of 7 graphs. With the exception being flickr, where Galois Push takes negligibly longer than the Pull counterpart.
+
+\begin{table}
+\renewcommand{\arraystretch}{1.3}
+\centering
+\caption{Distributed SSSP Execution Times Relative to Galois Push}
+\begin{tabular}{crrrr}
+\hline
+\bf{Data Set}&Galois Push&Galois Pull&Gemini&Giraph\\\hline
+flickr&1.0&0.97&1.02&17.46\\
+orkut&1.0&1.28&4.12&14.38\\
+wikipedia&1.0&1.88&5.26&7.11\\
+twitter&1.0&3.56&9.79&8.76\\
+rMat27&1.0&3.98&9.91&14.52\\
+friendster&1.0&2.32&9.72&7.59\\
+rMat28&1.0&5.70&11.07&16.5\\\hline
+\end{tabular}
+\label{tbl:ssspexec}
+\end{table}
+
+When taking a closer look at Giraph, it seems to not cope well with synthetic data sets.
 Analyzing the computation times in \autoref{fig:distributedSSSP_calc}, we see that it is the fastest framework on our real-world graphs.
-And that with a considerable margin of other frameworks always taking at least 50\% longer (Gemini on flickr) up to Galois Pull needing 18$\times$ longer on wikipedia.
-On both synthetic graphs however, Giraph is the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.
+And that with a considerable margin of other frameworks always taking at least 50\% longer (lower bound here is Gemini on flickr) up to Galois Pull needing 18$\times$ more time on wikipedia.
+On both synthetic graphs however, Giraph is actually the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.
 
@@
 On friendster, where Gemini was worst, Giraph \emph{only} requires 73$\times$ the overhead time of Galois.
+
+\subsubsection{Single-Node vs. Distributed}
+TODO
+\begin{itemize}
+	\item GaloisPush vs Galois CPU // Galois Pull verliert bei Dist.
+	\item Giraph vs. Giraph
+	\item Gemini vs Gemini
+\end{itemize}

### 2020-08-29 documentation/paper/sections/testingMethods.tex

@@
 
-\subsection{Hardware}
+\section{Testing Methods}
 For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each.
@@
 All five machines were running Ubuntu 18.04.2 LTS.
-Setup of each framework was performed according to our provided installation guides\footnote{The setup guides are available at \url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}.
 
-\subsection{Benchmark Software}
-All benchmarks were initiated by our benchmark script, that is available in our repository. Galois, Gemini and Giraph were benchmarked on both the 5-node cluster and a singe machine.
+Setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
+All benchmark cases were initiated by our benchmark script available in our repository. 
+All five frameworks are tested on a single server. 
+Galois, Gemini and Giraph were benchmarked in on the distributed 5-node cluster as well.
 Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
@@
 The graphs used in our testing can be seen in detail in \autoref{tbl:graphs}.
-We included a variaty of different graph sizes, from relatively small graphs like the flickr graph with 100 thousand edges up to an rMat28 with $4.2$ billion edges.
+We included a variaty of different graph sizes, from relatively small graphs like the flickr graph with 2 million edges up to an rMat28 with 4.2 billion edges.
 All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
@@
 	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.
-	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time that is used for cleanup after calculation in the measurement.
+	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time in the measurement that is used for cleanup after calculation.
 
@@
 
-	\item Giraph has built in timers for the iterations (supersteps) which we summed up to extract the computation time.
+	\item Giraph has built in timers for the iterations (supersteps), the sum of those is the computation time.
 \end{itemize}
@@
 The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark framework with every graph.
-For frameworks that support multiple implementations (e.g. PageRank in push and pull modes), we included the alternatives in our testing. 
-\autoref{tbl:algorithms} shows the tested algorithms in detail.
+We always show the results of PageRank with a maximum of five iterations.
+For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included the alternatives in our testing.
 
-\newcommand{\txtdag}{\textsuperscript{\textdagger}}
-\begin{table}
-	\renewcommand{\arraystretch}{1.3}
-	\centering
-	\caption{Tested Algorithms of Each Framework}
-	\begin{tabularx}{0.9\columnwidth}{cYYYYY}
-		\bf{Algorithm}&Galois&Gemini&Giraph&Ligra&Polymer\\\hline
-		PageRank&$\Delta$Push, $\Delta$Pull&&&&Regular, $\Delta$\\\hline
-		SSSP&?, Push, Pull&&&&\\\hline
-		BFS&Yes&&Yes*&&\\
-		\hline
-	\end{tabularx}
-	\label{tbl:algorithms}
+In detail, the algorithms for each framework are:
+\begin{itemize}
+	\item Galois: CPU, Push, Pull für PR und SSSP..
+	\item Gemini:
+	\item Giraph: Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
+	\item Ligra:
+	\item Polymer: Polymer supports two implementations of Pagerank
+\end{itemize}
 
-	{\vspace*{1ex}* Algorithm not natively supported
+% \autoref{tbl:algorithms} shows the tested algorithms in detail.
 
-	\txtdag{} Modified algorithm
+% \newcommand{\txtdag}{\textsuperscript{\textdagger}}
+% \newcommand{\singleNode}{\textsuperscript{s}}
+% \newcommand{\distributed}{\textsuperscript{d}}
+% \begin{table}
+% 	\renewcommand{\arraystretch}{1.3}
+% 	\centering
+% 	\caption{REMOVE THIS // Tested Algorithms of Each Framework}
+% 	\begin{tabularx}{0.9\columnwidth}{cYYYYY}
+% 		\bf{Algorithm}&Galois&Gemini&Giraph&Ligra&Polymer\\\hline
+% 		PageRank&Yes\singleNode{}, $\Delta$Push\distributed{}, $\Delta$Pull\distributed{}&&Yes\singleNode\distributed&&Regular\singleNode, $\Delta$\singleNode\\\hline
+% 		SSSP&?, Push, Pull&&Yes\singleNode\distributed\txtdag&&Yes\singleNode{}\\\hline
+% 		BFS&Yes&&Yes\singleNode\distributed{}*&&\\
+% 		\hline
+% 		\multicolumn{6}{p{0.5\columnwidth}}{* Algorithm not natively supported
 
-	\textsuperscript{d} Single node and distributed
-	}
+% 		\txtdag{} Modified algorithm
 
-\end{table}
+% 		\distributed{} distributed
+
+% 		\singleNode{} single-node}
+% 	\end{tabularx}
+% 	\label{tbl:algorithms}
+
+% \end{table}
 

### 2020-08-29 documentation/paper/sections/overview/gemini.tex
rewrote gemini writeup

@@
 
-Gemini is a framework for parallel graph processing\cite{Gemini}.
-In comparison to the other NUMA-aware frameworks we discuss here, Gemini does not support shared memory parallel processing.
-Computation is entirely message based using MPI.
-Gemini is lightweight and only provides basic functionality.
-A basic API hides data and computation distribution details from the user writing applications.
-There are five algorithms (which ones?) already implemented.
+Gemini is a framework for parallel graph processing \cite{Gemini}.
+It was developed with the goal to deliver a generally better performance through efficient communication.
+While most other distributed graph processing systems achieve very good results in the shared-memory area, they often deliver unsatisfactory results in distributed computing.
+Furthermore, a well optimized single-threaded implementation often outperforms a distributed system.
+Therefore it is necessary to not only focus on the performance of the computation but also of the performance of the communication.
+Gemini tries to bridge the gap between efficient shared-memory and scalable distributed systems.
+To achieve this goal, Gemini, in contrast to the other NUMA-aware frameworks discussed here, does not support shared-memory calculation, but chooses the distributed message-based approach from scratch.
+
+Gemini is fairly lightweight and seems well structured with a clearly defined API between the core framework and the implementations of the individual algorithms.
+The five already implemented algorithms are single source shortest path (sssp), breath first search (bfs), pagerank, connected components (cc) and biconnected components (bc).

### 2020-09-01 scripts/benchmark/benchmark_cases.sh
more benchmark cases

@@ -631,9 +631,82 @@ benchmark_giraph_bfs_dist() {
   done
 }
 
+benchmark_giraph_pagerank() {
+  for startnode in ${startnodes_flickr[@]}; do
+    giraph-pagerank flickr
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    giraph-pagerank orkut
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    giraph-pagerank wikipedia
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    giraph-pagerank twitter
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    giraph-pagerank friendster
+  done
+  for startnode in ${startnodes_rMat27[@]}; do
+    giraph-pagerank rMat27
+  done
+  for startnode in ${startnodes_rMat28[@]}; do
+    giraph-pagerank rMat28
+  done
+}
+
+benchmark_giraph_bfs() {
+  for startnode in ${startnodes_flickr[@]}; do
+    giraph-bfs flickr $startnode
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    giraph-bfs friendster $startnode
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    giraph-bfs orkut $startnode
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    giraph-bfs twitter $startnode
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    giraph-bfs wikipedia $startnode
+  done
+  for startnode in ${startnodes_rMat27[@]}; do
+    giraph-bfs rMat27 $startnode
+  done
+  for startnode in ${startnodes_rMat28[@]}; do
+    giraph-bfs rMat28 $startnode
+  done
+}
+
+benchmark_giraph_sssp() {
+  for startnode in ${startnodes_flickr[@]}; do
+    giraph-sssp flickr $startnode
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    giraph-sssp friendster $startnode
+  done
+  for startnode in ${startnodes_orkut[@]}; do
+    giraph-sssp orkut $startnode
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    giraph-sssp twitter $startnode
+  done
+  for startnode in ${startnodes_wikipedia[@]}; do
+    giraph-sssp wikipedia $startnode
[…]

### 2020-09-01 documentation/paper/sections/graphFormats.tex
adjList to adjGraph with compressed sparse rows

@@
 The following sections explain the output formats of our conversion tool.
-\subsubsection{AdjacencyList}
-The AdjacencyList and WeightedAdjacencyList formats are used by Ligra and Polymer.
-The format was initially specified for the Problem Based Benchmark Suite, an open source repository to compare different parallel programming methodologies in terms of performance and code quality \cite{pbbs}.
+\subsubsection{AdjacencyGraph}
+The AdjacencyGraph and WeightedAdjacencyGraph formats used by Ligra and Polymer are similar to the more popular \emph{compressed sparse rows} format.
+
+The  The format was initially specified for the Problem Based Benchmark Suite, an open source repository to compare different parallel programming methodologies in terms of performance and code quality \cite{pbbs}.
 
@@
 
-For the WeightedAdjacencyList format, the weights are appended to the end of the file in an order corresponding to the target nodes.
+For the WeightedAdjacencyGraph format, the weights are appended to the end of the file in an order corresponding to the target nodes.
+
+The files always start with the name of the format i.e. AdjacencyGraph or WeightedAdjacencyGraph in the first line.
 

### 2020-09-01 scripts/benchmark/benchmark_cases.sh
more benchmark cases

@@ -679,34 +679,46 @@ benchmark_giraph_bfs() {
   done
 }
 
-benchmark_giraph_sssp() {
+benchmark_giraph() {
   for startnode in ${startnodes_flickr[@]}; do
     giraph-sssp flickr $startnode
-  done
-  for startnode in ${startnodes_friendster[@]}; do
-    giraph-sssp friendster $startnode
+    giraph-bfs flickr $startnode
+    giraph-pagerank flickr
   done
   for startnode in ${startnodes_orkut[@]}; do
     giraph-sssp orkut $startnode
-  done
-  for startnode in ${startnodes_twitter[@]}; do
-    giraph-sssp twitter $startnode
+    giraph-bfs orkut $startnode
+    giraph-pagerank orkut
   done
   for startnode in ${startnodes_wikipedia[@]}; do
     giraph-sssp wikipedia $startnode
+    giraph-bfs wikipedia $startnode
+    giraph-pagerank wikipedia
+  done
+  for startnode in ${startnodes_twitter[@]}; do
+    giraph-sssp twitter $startnode
+    giraph-bfs twitter $startnode
+    giraph-pagerank twitter
+  done
+  for startnode in ${startnodes_friendster[@]}; do
+    giraph-sssp friendster $startnode
+    giraph-bfs friendster $startnode
+    giraph-pagerank friendster
   done
   for startnode in ${startnodes_rMat27[@]}; do
     giraph-sssp rMat27 $startnode
+    giraph-bfs rMat27 $startnode
+    giraph-pagerank rMat27
   done
   for startnode in ${startnodes_rMat28[@]}; do
     giraph-sssp rMat28 $startnode
+    giraph-bfs rMat28 $startnode
+    giraph-pagerank rMat28
   done
 }
 
 
 time_start=$(get_time)
-benchmark_giraph_sssp
-benchmark_giraph_bfs
-benchmark_giraph_pagerank
+benchmark_giraph
 dur_exec=$((($(get_time)-$time_start)/1000000))
 echo "benchmark took $dur_exec seconds"

### 2020-09-01 documentation/paper/sections/results/bfs.tex
plots with new Giraph data, some text for dist bfs

@@
 
-\begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
+\subsubsection{Distributed}
+\begin{figure}
+	\begin{subfigure}{\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
@@
 	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
+	% \hfil
+	\begin{subfigure}{\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
@@
 	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:distributedBFS_overheadNormalized}
-	\end{subfigure}
+	% \hfil
+	% \begin{subfigure}{0.3\textwidth}
+	% 	\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
+	% 	\caption{Overhead time normalized by the graph size in million edges}
+	% 	\label{fig:distributedBFS_overheadNormalized}
+	% \end{subfigure}
+	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
+	\label{fig:distributedBFS}
+\end{figure}
+For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case. All measurements can be seen in \autoref{fig:distributedBFS}.
 
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
-\end{figure*}
+First, the calculation times in \autoref{fig:distributedBFS_calc}.
+It shows Giraph having the shortest calculation times on the real-world graphs, while Giraph's calculatin times on both rMat27 and rMat28 are the worst of all frameworks.
+
+Comparing the execution times in \autoref{fig:distributedBFS_exec} results in the same findings as with SSSP. 
+While Gemini can compete with Galois on the small flickr graph, moving to larger data sets shows the worse performance of Gemini compared to Galois.
+
+Much like when running SSSP, Giraph is slowest on all but one graph. Only on friendster is Gemini marginally slower, which was also the case for SSSP.
+
+Both Galois implementations are again similar to the behaviour on SSSP.
+Galois Push is generally faster than the Pull alternative while both Push and Pull versions are faster than Gemini and Giraph across all graphs.
+This makes Galois Push the clear winner for distributed BFS.

### 2020-09-01 plots/makeplots.py

@@ -4,19 +4,19 @@ from plotFunctions import *
 
 
 
-distributed = False
+distributed = True
 singleNode = False
 
-sssp = True
-bfs = False
-pr = False
+sssp = False
+bfs = True
+pr = True
 
 calcTimes = False
 execTimes = False
 overheads = False
 
 
-galoisSpeedup = True
+galoisSpeedup = False
 
 ### PLOTS
 #SINGLE NODE

### 2020-09-02 plots/makedata.py
updated plots, made the bars more pronounced and increased readability

@@ -429,16 +429,23 @@ for g in graphs:
 
 
 if False:
+    maxs = []
     for g in graphs:
-        if g == 'flickr' or g == 'orkut':
+        if g == "flickr":
             continue
-        tmp = speedupGaloisBFS[g]
+        tmp = speedupGaloisPRPull[g]
         print(g)
         last = 1
         xs = []
         for k in tmp:
             xs.append(k)
         xs.sort()
+        tmps = []
         for k in xs:
-            print(k, tmp[k], tmp[k] / tmp[last])
+            tmps.append(tmp[k])
+            if(tmp[k] > 1):
+                print(k, tmp[k])
+        maxs.append(max(tmps))
+        #print("inter: ", (tmp[32] - tmp[16])/(32-16) * (24-16) + tmp[16])
             #last = k
+    print(maxs)

### 2020-09-02 plots/makeplots.py

@@ -7,7 +7,7 @@ from plotFunctions import *
 distributed = True
 singleNode = False
 
-sssp = False
+sssp = True
 bfs = True
 pr = True
 
@@ -101,11 +101,11 @@ if distributed and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 	#BFS
 	if bfs or calcTimes:
-		grouped_bar_plot(graphs, dist_frameworks_bfs.values(), calcTimeBFS_distributed, yErrs=yErrCalcBFS_distributed, title='BFS distributed', yLabel='Calculation time (s)', saveToFile="distributedBFS_calcTime.png")
+		grouped_bar_plot(graphs, dist_frameworks_bfs.values(), calcTimeBFS_distributed, yErrs=yErrCalcBFS_distributed, yLabel='Calculation time (s)', saveToFile="distributedBFS_calcTime.png")
 
 	#PR
 	if pr or calcTimes:
-		grouped_bar_plot(graphs, dist_frameworks_pr.values(), calcTimePR_distributed, yErrs=yErrCalcPR_distributed, title='PR distributed', yLabel='Calculation time (s)', saveToFile="distributedPR_calcTime.png")
+		grouped_bar_plot(graphs, dist_frameworks_pr.values(), calcTimePR_distributed, yErrs=yErrCalcPR_distributed, yLabel='Calculation time (s)', saveToFile="distributedPR_calcTime.png")
 
 
 	## EXEC TIME
@@ -115,11 +115,11 @@ if distributed and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 	#BFS
 	if bfs or execTimes:
-		grouped_bar_plot(graphs, dist_frameworks_bfs.values(), execTimeBFS_distributed, yErrs=yErrExecBFS_distributed, title='BFS distributed', yLabel='Execution time (s)', saveToFile="distributedBFS_execTime.png")
+		grouped_bar_plot(graphs, dist_frameworks_bfs.values(), execTimeBFS_distributed, yErrs=yErrExecBFS_distributed, yLabel='Execution time (s)', saveToFile="distributedBFS_execTime.png")
 
 	#PR
 	if pr or calcTimes:
[…]

### 2020-09-02 documentation/paper/sections/results/galoisSpeedup.tex
preliminaries, more results, giraph and testingMethods reworked
dist pagerank restults and pr speedup results

@@
 
+For our speedup results on BFS, \autoref{fig:galoisSpeedupBFS} shows the calculation time speedup of Galois' BFS.
+
 Flickr is sped up by about 5\%\ with 4 threads, any more than that will actually decrease performance, making computation time up to 15\%\ (96 threads) longer.
@@
 \subsubsection{PageRank}
+We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.
+This is a perfect example for an algorithm that does not benefit from multithreaded computation.
+Computation is hardly sped up on any graph other than flicker, where the reached maximum is 64\%. This maximum is reached at two threads, with speedup steadily declining above that.
+The rMat28 is the only other graph of one could say computation was sped up. Here we reached a maximum speedup of 31\%\ at 96 threads.
+All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases and if so only by a small margin.
+Computation on Orkut and Twitter reaches a speedup maximum of 12\%\ and 5\%\ at 4 threads, while being less or equal to 1 in all other cases.
+The wikipedia graph is never sped up.
+Friendster and rMat27 can be sped up by 6.5\%\ or 10\%\ respectively on 8 threads.
+
+Speedup results on PageRank show odd behaviour in the Galois implementation.
+There is a significant performance loss on 4, 24 and 40 threads that is far from the expected behaviour.
+This is most visible for the Push variant seen in \autoref{fig:galoisSpeedupPRPush}, we validated the shown results two times.
+Especially, the speedup for 24 threads is (by interpolating between 16 and 32 threads) expected to be anywhere between 25\% and 94\%.
+Actually however, the system does not reach a speedup of more than 4\%\ on any graph, with only rMat27 actually reaching a value greater than 1. 
+On all other graphs, using 24 threads is anywhere from 3\% (flickr) to 9\%\ (wikipedia) slower than using just one thread.
+
+Similar yet less pronounced behaviour is oberservable for Pull in \autoref{fig:galoisSpeedupPRPull}. 
+Here especially the values for 24 and 40 threads show a loss in performance.
+It is most visible on the values for twitter and friendster, where both values drop significantly compared to the neighbouring 32 and 48 thread results.
+
 \begin{figure}
 	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPullGaloisThreads.png}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
 		\caption{PageRank Pull}
@@
 	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\columnwidth]{../../plots/singleNodePRPushGaloisThreads.png}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
 		\caption{PageRank Push}

### 2020-09-02 documentation/paper/sections/results/pr.tex
plots with new Giraph data, some text for dist bfs
dist pagerank restults and pr speedup results

@@
 
+\subsubsection{Distributed}
+The \autoref{fig:distributedPR} shows our results of PageRank on the distributed cluster.
+
+First of all, Giraph was unable to complete the test because it required more than 250GB of RAM for rMat28. Thus this results is missing.
+
+When comparing the calculation times in \autoref{fig:distributedPR_calc} to the execution times in \autoref{fig:distributedPR_exec}, we see similar behaviour of all frameworks.
+This means that unlike with SSSP or BFS, the calculation times and execution times are similar with respect to the relations of the frameworks to one another.
+More specifically, there are no overhead outliers like it was the case with Giraph on SSSP.
+
 \begin{figure*}
@@
 	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
+		\caption{Overhead time}
 		\label{fig:distributedPR_overheadNormalized}
@@
 	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
+	\label{fig:distributedPR}
 \end{figure*}

### 2020-09-03 documentation/paper/sections/overview/giraph.tex
added pregel info, moved Mapreduce from giraph to preliminiaries, giraph is now ugly again

@@
 
-TODO // MapReduce hier oder wo anders erklären?
-MapReduce is a in three phases
-\begin{enumerate}
-	\item Map phase: The input data is distributed between a set of map processes, the map functionality is specified by the user.
-Ideally all map processes run in parallel so the map processes need to be independent.
-Results from this phase are written into (multiple) intermediate storage points.
-	\item Shuffle phase: The results are grouped according to a key provided by the map algorithm. Each set of results is then handed to one system for the next phase.
-	\item Reduce phase: Every set of intermediate results is input to exactly one reduce process. The reduce functionality is again specified by the user and ideally runs in parallel.
-\end{enumerate}
-Giraph uses the Map functionality of Hadoop to run the algorithms, Reduce is only used as an identity function.
+Giraph is an example for an open-source system similar to Pregel.
+Thus, Giraph's computation model is closely related to the BSP model discussed in \autoref{sec:bsp}. 
+//TODO: Hier wiederholt sich jetzt einiges.. ärgerlich.
 
-Giraph's computation model is closely related to the BSP model discussed in \autoref{sec:bsp}.
+// Mapreduce: Giraph uses the Map functionality of Hadoop to run the algorithms, Reduce is only used as an identity function.
+
 This means that Giraph is based on computation units that communicate using messages and are synchonized with barriers.

### 2020-09-03 documentation/paper/sections/preliminiaries.tex
Ligra writeup
added pregel info, moved Mapreduce from giraph to preliminiaries, giraph is now ugly again

@@
 
-\subsection{Pregel Model}
+One of the most famous state-of-the-art graph processing systems is Pregel \cite{pregel}. It is based on the BSP computation model.
+Pregel and many open-source versions similar to it were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures.
 
-% Google, to overcome, these challenges came up with Pregel.
-% Provides scalability
-% Fault-tolerance
-% Flexibility to express arbitrary algorithms
-% The high level organization of Pregel programs
-% is inspired by Valiant’s Bulk Synchronous
-%  Parallel model 
+\subsection{MapReduce}
+The MapReduce model is a computation infrastructure developed by Google to reliably handle large data sets on distributed clusters \cite{mapreduce}.
+
+A user specifies just the two functions Map and Reduce.
+The system hides the details of parallelization, fault-tolerance, data distribution and load balancing away from the application logic.
+All of these features are automatically provided.
+
+Execution is performed in three phases:
+\begin{enumerate}
+	\item Map phase: The input data is distributed between a set of Map processes, the Map functionality is specified by the user.
+Ideally all Map processes run in parallel so the map processes need to be independent.
+Results from this phase are written into (multiple) intermediate storage points.
+	\item Shuffle phase: The results are grouped according to a key provided by the Map algorithm. Each set of results is then handed to one system for the next phase.
+	\item Reduce phase: Every set of intermediate results is input to exactly one reduce process. The Reduce functionality is again specified by the user and ideally runs in parallel.
+\end{enumerate}
+
+We show Giraph \cite{Giraph} as an example of a system using this framework.
+
 \subsection{Congest Model}

### 2020-09-03 documentation/paper/main.tex
giraph overview is less ugly now

@@
 
+\input{sections/relatedWork}
+
 \input{sections/preliminiaries}
@@
 % Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
-\begin{thebibliography}{1}
+\begin{thebibliography}{10}
 \providecommand{\url}[1]{#1}
@@
   \url{https://doi.org/10.1145/79173.79181}
+\BIBentrySTDinterwordspacing
+
+\bibitem{pregel}
+\BIBentryALTinterwordspacing
+G.~Malewicz, M.~H. Austern, A.~J. Bik, J.~C. Dehnert, I.~Horn, N.~Leiser, and
+  G.~Czajkowski, ``Pregel: A system for large-scale graph processing,'' in
+  \emph{Proceedings of the 2010 ACM SIGMOD International Conference on
+  Management of Data}, ser. SIGMOD '10.\hskip 1em plus 0.5em minus 0.4em\relax
+  New York, NY, USA: Association for Computing Machinery, 2010, p. 135–146.
+  [Online]. Available: \url{https://doi.org/10.1145/1807167.1807184}
+\BIBentrySTDinterwordspacing
+
+\bibitem{mapreduce}
+J.~Dean and S.~Ghemawat, ``Mapreduce: Simplified data processing on large
+  clusters,'' in \emph{In Proceedings of Operating Systems Design and
+  Implementation (OSDI}, pp. 137--150.
+
+\bibitem{Giraph}
+\BIBentryALTinterwordspacing
+{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
+  Available: \url{https://giraph.apache.org}
 \BIBentrySTDinterwordspacing
@@
   \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
-\BIBentrySTDinterwordspacing
-
-\bibitem{Giraph}
-\BIBentryALTinterwordspacing
-{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
-  Available: \url{https://giraph.apache.org}
 \BIBentrySTDinterwordspacing

### 2020-09-03 documentation/paper/sections/overview/giraph.tex
giraph overview is less ugly now

@@
 
-Apache Giraph is an iterative graph processing framework, built on top of Apache Hadoop \cite{Giraph}.
-Hadoop is a large MapReduce infrastructure providing a reliable (fault tolerant) basis for large scale graph processing.
-
-Giraph is an example for an open-source system similar to Pregel.
+Apache Giraph is an example for an open-source system similar to Pregel.
 Thus, Giraph's computation model is closely related to the BSP model discussed in \autoref{sec:bsp}. 
-//TODO: Hier wiederholt sich jetzt einiges.. ärgerlich.
-
-// Mapreduce: Giraph uses the Map functionality of Hadoop to run the algorithms, Reduce is only used as an identity function.
-
-This means that Giraph is based on computation units that communicate using messages and are synchonized with barriers.
+This means that Giraph is based on computation units that communicate using messages and are synchonized with barriers \cite{Giraph}.
 
@@
 Communication between vertices is only performed via messages, so a vertex has no access to values of other vertices or edges other than its own outgoing ones.
-
 Supersteps are synchronized with barriers, meaning that all messages only get delivered in the following superstep and computation for the next superstep can only begin after every vertex has finished computing the current superstep.
 Edge and vertex values are retained across supersteps.
-
 Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will reactivate the vertex again.
@@
 
-Because of the underlying Hadoop infrastructure, expanding single-node processing to a multi-node cluster is seamless.
+In order for Giraph a achieve scalability and parallelization, it is built on top of Apache Hadoop \cite{Giraph}.
+Hadoop is a MapReduce infrastructure providing a fault tolerant basis for large scale graph processing.
 Hadoop supplies a distributed file system (HDFS), on which all computation is performed.
 Giraph is thus, even when only using a single node, running in a distributed manner.
+But expanding single-node processing to a multi-node cluster is seamless.
+Giraph uses the Map functionality of Hadoop to run the algorithms, Reduce is only used as the identity function.
 
@@
 Over the course of this project, several new updates were pushed to Giraph's source repository\footnote{\url{https://gitbox.apache.org/repos/asf?p=giraph.git}}.
-
-TODO: Facebook?

### 2020-09-03 documentation/paper/sections/overview/ligra.tex

@@
 
-Ligra\cite{Ligra} is a lightweight parallel graph processing framework for shared memory machines. It offers a programming inteface that allows expressing graph traversal algorithms in a simple way.
+Ligra is a lightweight graph processing framework for shared memory machines \cite{Ligra}.
+It offers a vertex-centric programming interface which can be used to apply a function to each vertex or outgoing edge of a set of vertices in parallel.
+While doing so the framework can generate a set of active vertices for the next iteration.
+This abstraction makes it well suited for writing graph traversal algorithms.
 
-Algorithms can use the \texttt{EdgeMap} to make computations based on edges or the \texttt{VertexMap} to make computations based on vertices.
-Those mappings can be applied only to a subset of vertices.
-Based on the size of the vertex subset the framework automatically switches between a sparse and a dense representation to optimize speed and memory.
+When mapping over edges Ligra optimizes algorithms by switching between a push-based and a pull-based approach based on the size of the set.
+When the size of the set is above a threshold a pull approach is used.
+If the threshold is not reached a push approach is used.

### 2020-09-04 documentation/paper/sections/graphFormats.tex
rückmeldung 2, noch nicht alles umgesetzt

--- 2020-08-29 documentation/paper/sections/graphFormats.tex
@@
 
-\subsection{Important Graph Formats}
+\section{Important Graph Formats}
 Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.
@@
 
+//SNAP
+
 The following sections explain the output formats of our conversion tool.
-\subsubsection{AdjacencyList}
-The AdjacencyList and WeightedAdjacencyList formats are used by Ligra and Polymer.
+\subsection{AdjacencyGraph}
+The AdjacencyGraph and WeightedAdjacencyGraph formats used by Ligra and Polymer are similar to the more popular \emph{compressed sparse rows} format.
 The format was initially specified for the Problem Based Benchmark Suite, an open source repository to compare different parallel programming methodologies in terms of performance and code quality \cite{pbbs}.
@@
 \begin{equation*}
-	n, m, o_1, \ldots, o_n, t_1, \ldots, t_m
+	x, n, m, o_1, \ldots, o_n, t_1, \ldots, t_m
 \end{equation*}
-where commas are \texttt{\textbackslash n}. First, $n$ is the number of vertices and $m$ the number of edges in the graph.
-
+where commas are newlines. The files always start with the name of the format i.e. AdjacencyGraph or WeightedAdjacencyGraph in the first line, here shown as $x$.
+Followed by $n$, the number of vertices and $m$ the number of edges in the graph.
 The $o_k$ are the so-called offsets. Each vertex $k$ has an offset $o_k$, that describes an index in the following list of the $t_i$.
@@
 
-For the WeightedAdjacencyList format, the weights are appended to the end of the file in an order corresponding to the target nodes.
+For the WeightedAdjacencyGraph format, the weights are appended to the end of the file in an order corresponding to the target nodes.
 
-\subsubsection{EdgeList}
-The EdgeList format is the most intuitive and one of the most commonly used in online data set repositories.
-The KONECT database uses this format and thus it is the input format for our conversion tool.
+\subsection{EdgeList}
+The EdgeList format is one of the most commonly used in online data set repositories. The KONECT database uses this format and thus it is the input format for our conversion tool.
 
-An edge list is a set of directed eges $(s_1,t_1),(s_2,t_2),\ldots$ where $s_i$ is a vertex ID representing the start vertex and $t_i$ is a vertex ID representing the target vertex.
-In the format, there is one edges per line and the vertex IDs $s_i, t_i$ are separated with any whitespace character.
+An edge list is a set of directed edges $(s_1,t_1),(s_2,t_2),\ldots$ where $s_i$ is a vertex ID representing the start vertex and $t_i$ is a vertex ID representing the target vertex.
+In the format, there is one edge per line and the vertex IDs $s_i, t_i$ are separated with any whitespace character.
 
-For a WeightedEdgeList, the edge weights are appended to each line, again separated by a whitespace character.
+For a WeightedEdgeList, the edge weights are appended to each line, again separated by any number of whitespace characters.
 
-\subsubsection{Binary EdgeList}
+\subsection{Binary EdgeList}
 The binary EdgeList format is used by Gemini.
-
 For $s_i, t_i$ some vertex IDs and $w_i$ the weight of a directed edge $(s_i,t_i, w_i)$, Gemini requires the following input format
@@
 where $s_i,t_i$ have \texttt{uint32} data type and the optional weights are \texttt{float32}.
-Gemini will derive the number of edges from the file size, so there is no file header or anything similar allowed.
+Gemini derives the number of edges from the file size, so there is no file header or anything similar allowed.
 
-\subsubsection{Giraph's I/O formats}
+\subsection{Giraph's I/O formats}
 Giraph is capable of parsing many different input and output formats. All of those are explained in Giraph's JavaDoc\footnote{\url{http://giraph.apache.org/apidocs/index.html}}.
 Both edge- and vertex-centric input formats are possible.
-
 One can even define their own input graph representation or output format. For the purposes of this paper, we used an existing format similar to AdjacencyList but represented in a JSON-like manner.
 
-In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values, \texttt{float} out-edge weights.
+In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values and \texttt{float} out-edge weights.
 Each line in the graph file looks as follows
@@
 with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.
-
 There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

### 2020-09-04 documentation/paper/sections/overview/overview.tex
rückmeldung 2, noch nicht alles umgesetzt

@@
 \section{Overview}
+In this section we describe the functionality and characteristics of several state-of-the-art graph processing frameworks. 
 We start with a short overview describing functionality and characteristics of each framework, followed by their required input data formats.
+
+Further
 

### 2020-09-04 documentation/paper/sections/preliminiaries.tex
rückmeldung 2, noch nicht alles umgesetzt

@@
 \end{equation*}
+The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.
+The size of a graph is defined as the number of edges $|E|$ // CITATION?.
+Independently of the graph being directed or not, a graph can be \emph{weighted}.
+In this case a function $w:E\rightarrow \mathbb R$ is introduced, that maps an edge to a numerical value, further describing the relation.
 
-The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.
-
-Independently of the graph being directed or not, a graph can be \emph{weighted}.
-In this case the edge set is expanded by a numerical value $E_{\text{weighted}}=E_{\text{unweighted}}\times \mathbb R$ further describing the relation.
-
-The size of a graph is often described in the number of edges $|E|$ because this number is typically much larger than the number of vertices $|V|$.
-
-A \emph{Path} from start $s$ to target $t$ is a set of edges $P\subseteq E^n$ with
+A \emph{Path} from starting vertex $s$ to target vertex $t$ is a sequence of vertices
 \begin{equation*}
-   P=((x_0,x_1),(x_1,x_2),\ldots, (x_{n-1},x_n))
+	P=(x_1,x_2,\ldots,x_n)\subseteq V^n
 \end{equation*}
-where all $(x_i,x_{i+1})\in E$ and $x_0=s, x_n=t$.
-
-Thus we call a target $t$ \emph{accessible} from $s$ if a Path from $s$ to $t$ exists.
+with the condition $(x_i,x_{i+1})\in E$ for each $i\in\{1,\ldots,n-1\}$ and $x_1=s, x_n=t$.
+Thus we call a target $t$ \emph{reachable} from $s$ if a Path from $s$ to $t$ exists.
 
 \subsection{Single-Source Shortest-Paths}
-Single-Source Shortest-Paths (SSSP) describes the problem of finding a path along some edges of the input graph from a given start node to a target.
-
+Single-Source Shortest-Paths (SSSP) describes the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
 Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
@@
 In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.
-
 The most common implementations are Dijkstra's algorithm or BellmanFord.
@@
 \subsection{Breadth-First Search}
-Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input. In some cases a target vertex is also given.
-
-The output is usually a set of vertices that are accessible from the start vertex. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.
-
+Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input.
+The output is a set of vertices that are reachable from the start vertex.
+In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.
 It is called breadth-first search because the algorithm searches in a path length-based way.
@@
 all neighbors of the start node are checked before checking paths of length 2 and so on.
-
 The search algorithm, where the paths of maximum length are checked first is called Depth-First Search.
@@
 
-For example, Google Search uses PageRank to rank web pages in their search engine results.
-
 The output of PageRank is a percentage for each vertex.
@@
 
+For example, Google Search uses PageRank to rank web pages in their search engine results.
 \paragraph{Delta-PageRank}
-?
+This is a optimization of the traditional PageRank algorithm.
+The PageRank score of a vertex is only updated if the relative change of the PageRank is larger than some delta.
+In the next iteration only a subset of PageRanks has to be recalculated.
 
@@
 \label{sec:bsp}
-The Bulk-Synchronous Parallel (BSP) model is a computation model developed by Leslie Valiant \cite{bsp}.
-It is commonly used in computation environment with large amounts of synchronous computation, for example Giraph is based on this model.
+The Bulk-Synchronous Parallel (BSP) model is a computation model developed by Leslie Valiant \cite{bsp}. It is commonly used in computation environments with large amounts of synchronous computation.
 
@@
 
-One of the most famous state-of-the-art graph processing systems is Pregel \cite{pregel}. It is based on the BSP computation model.
-Pregel and many open-source versions similar to it were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures.
+One of the most famous graph processing systems, Pregel \cite{pregel} is based on the BSP computation model.
+Pregel and many open-source versions similar to it were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures \cite{Giraph}.
 
@@
 
-We show Giraph \cite{Giraph} as an example of a system using this framework.
+Giraph \cite{Giraph} is an example of a system using this framework.
 

### 2020-09-04 documentation/paper/sections/results/bfs.tex
rückmeldung 2, noch nicht alles umgesetzt

@@
 \subsection{Breadth-first search}
+\subsubsection{Single-Node}
+TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.
+
 In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
@@
 \end{figure}
-For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case. All measurements can be seen in \autoref{fig:distributedBFS}.
+For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case.
+This is expected since both are graph traversal algorithms starting in one source vertex.
+Hence, calculation complexity for each node and communication overhead is similar.
+All measurements can be seen in \autoref{fig:distributedBFS}.
 

### 2020-09-04 documentation/paper/sections/results/galoisSpeedup.tex
rückmeldung 2, noch nicht alles umgesetzt

@@
 \end{figure}
-Starting with SSSP which is the algorithm that really is at an advantage when using many threads in \autoref{fig:galoisSpeedupSSSP}.
-
+Starting with SSSP, we see an algorithm that benefits from many available threads in \autoref{fig:galoisSpeedupSSSP}.
 For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads. 
@@
 
-On all graphs, the speedup never exceeds 6$\times$ even when using 96 thread.
+On all graphs, the speedup never exceeds 6$\times$ even when using 96 threads.
 The initial speedup when switching from one to two threads is actually smaller than one, thus decreasing performance, on 4 of 7 graphs.

### 2020-09-04 documentation/paper/sections/results/results.tex
rückmeldung 2, noch nicht alles umgesetzt

@@
 %!TEX root=../../main.tex
+
+\section{Results}
 
@@
 	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code.
-A repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
+Our changes can be found in one of our repositories\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
 
@@
 
-On a plus side, setup of frameworks like Polymer or Ligra was straight forward and did not require any special treatment.
-
-\clearpage
+On a plus side, setup of the frameworks Polymer and Ligra was straight forward and did not require any special treatment.
 

### 2020-09-04 documentation/paper/sections/results/sssp.tex
plots with new Giraph data, some text for dist bfs
rückmeldung 2, noch nicht alles umgesetzt

@@
 	\centering
-	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalizedToGalois.png}
-	\caption{Overhead times of each framework normalized by the overhead time of Galois Push}
+	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTime.png}
+	\caption{Distributed Single-Source Shortest-Paths overhead times of each framework}
 	\label{fig:distributedSSSP_overhead}
@@
 
-Both Galois implementations have significantly smaller execution times compared to Gemini or Giraph on all graphs, see \autoref{tbl:ssspexec}.
+Both Galois implementations have significantly smaller execution times compared to Gemini or Giraph on all graphs (cf. \autoref{tbl:ssspexec}).
 You can see Gemini being worse by at least a factor of 4 compared to Galois Push on all graphs except flickr.

### 2020-09-04 documentation/paper/sections/testingMethods.tex
rückmeldung 2, noch nicht alles umgesetzt

@@
 
-\section{Testing Methods}
+\section{Evaluation}
+Einleitung
+
+\subsection{Testing Methods}
 For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each.
@@
 Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
+Unless mentioned otherwise, we always show results of each framework utilizing 96 threads (i.e. the maximum on our machines) for the single-node evaluation.
 
-The complete benchmark log files and extracted raw results are available in our repository.
+The complete benchmark log files and extracted raw results are available in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt} // das ist hier unschön.}.
 
@@
 For every framework, we measured the \emph{execution time} as the time from start to finish of the console command.
+For the \emph{calculation time}, we tried to extract only the time the framework actually executed the algorithm. 
+Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.
+For measuring the calculation time, came up with the following:
+\begin{itemize}
+	\item For Galois, we extract console log time stamps. Galois outputs \enquote{\texttt{Reading graph complete.}}. Calculation time is the time from this output to the end of execution.
 
-For the \emph{calculation time}, we tried to extract only the time the framework actually executed the algorithm. We came up with the following:
-\begin{itemize}
-	\item For Galois, we resulted to extracting console log time stamps. Galois outputs \texttt{Reading graph complete.}. Calculation time is the time from this output to the end of execution.
-
-	We know that this is not the most realiable way for measuring the calculation times. 
+	This is not the most realiable way for measuring the calculation times. 
 	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.
 	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time in the measurement that is used for cleanup after calculation.
+
+	However, this method is the only way of retreiving any measurements without introducing custom modifications to the Galois source code.
 
@@
 \end{itemize}
-Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.
-
-Each test case consisting of graph, framework and algorithm was run 10 times, allowing us to smooth slight variations in the measured times.
+Each evaluation consisting of graph, framework and algorithm was run 10 times, allowing us to smooth slight variations in the measured times.
 Later on, we provide the mean values of the individual times as well as the standard deviation where meaningful.
@@
 We always show the results of PageRank with a maximum of five iterations.
-For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included the alternatives in our testing.
+For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included both in our evaluation.
 

### 2020-09-04 documentation/paper/sections/results/pr.tex
rückmeldung 2, noch nicht alles umgesetzt
results for pr, removed two algos

@@
 \subsection{PageRank}
+\subsubsection{Single-Node}
+
+Ligra and Polymer support both regular and Delta-PageRank variants.
+Ligra's regular PR implementation is faster on 4 of 7 graphs.
+If the regular version is slower than delta, that is only by a small difference.
+Explicitly, regular is slower than delta by 19\%\ on twitter, 17\%\ on rMat27 or 6\%\ for friendster.
+For the other graphs, the delta version is slower by a far greater margin, 50\%\ on flickr, 45\%\ on orkut, 13\%\ on wikipedia and 68\% on rMat28.
+Hence, we only show the results of Ligra's regular PageRank implementation in our evaluation.
+For Polymer we found the delta version to be faster on all graphs except rMat28. Delta-PR is on average 15\%\ faster on the first six graphs, while only being 0.3\%\ slower on rMat28.
+Thus, the following only shows Polymer's faster Delta-PR implementation.
+
+TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.
 
@@
 The \autoref{fig:distributedPR} shows our results of PageRank on the distributed cluster.
-
 First of all, Giraph was unable to complete the test because it required more than 250GB of RAM for rMat28. Thus this results is missing.
-
 When comparing the calculation times in \autoref{fig:distributedPR_calc} to the execution times in \autoref{fig:distributedPR_exec}, we see similar behaviour of all frameworks.

### 2020-09-04 plots/makedata.py

@@ -273,9 +273,10 @@ singleNode_frameworks_pr = {
     "gemini-pagerank":"Gemini",
     "giraph-pagerank":"Giraph",
     "ligra-pagerank":"Ligra",
-    "ligra-pagerank-delta":"Ligra Delta",
-    "polymer-pagerank":"Polymer",
-    "polymer-pagerank-delta":"Polymer Delta"}
+    #"ligra-pagerank-delta":"Ligra Delta",
+    #"polymer-pagerank":"Polymer",
+    "polymer-pagerank-delta":"Polymer Delta"
+    }
 
 calcTimePR_singleNode = []
 yErrCalcPR_singleNode = []
@@ -314,6 +315,44 @@ for f in singleNode_frameworks_pr:
     yErrExecPR_singleNode.append(tmpYErrExec)
     overheadPR_singleNode.append(overhead)
     overheadPRNormalized_singleNode.append(overheadNormalized)
+
+
+if False:
+    k = 0
+    values = {}
+    for f in singleNode_frameworks_pr:
+        if 'ligra' in f or 'polymer' in f:
+            values[f] = execTimePR_singleNode[k]
+        k += 1
+
+    ligra = []
+    polymer = []
+
+    for l in range(len(values["ligra-pagerank"])):
+        v1 = values["ligra-pagerank"][l]
+        v2 = values["ligra-pagerank-delta"][l]
+        v = abs((v1-v2) / max(v1,v2))
+        ligra.append(v)
+    # print(ligra)
+    # print(sum(ligra) / len(ligra))
+
+    for l in range(len(values["polymer-pagerank"])):
+        v1 = values["polymer-pagerank"][l]
+        v2 = values["polymer-pagerank-delta"][l]
+        if v1>v2:
+            print("delta")
+        else:
+            print("reg")
+        v = abs((v1-v2) / max(v1,v2))
+        polymer.append(v)
+
+    print(polymer)
+    print(sum(polymer[:-1]) / len(polymer[:-1])) 
+
+    
+
+
+
 
 
 

### 2020-09-04 plots/makeplots.py

@@ -4,11 +4,11 @@ from plotFunctions import *
 
 
 
-distributed = True
-singleNode = False
+distributed = False
+singleNode = True
 
-sssp = True
-bfs = True
+sssp = False
+bfs = False
 pr = True
 
 calcTimes = False
@@ -46,7 +46,7 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 		grouped_bar_plot(graphs, frameworks, execTimeBFS_singleNode, yErrs=yErrExecBFS_singleNode, title='BFS single node', yLabel='Execution time (s)', saveToFile="singleNodeBFS_execTime.png")
 
 	if pr or execTimes:
-		grouped_bar_plot(graphs, singleNode_frameworks_pr.values(), execTimePR_singleNode, yErrs=yErrExecPR_singleNode, title='PR single node (ugly)', yLabel='Execution time (s)', saveToFile="singleNodePR_execTime.png", scaleFactor=3.25)
+		grouped_bar_plot(graphs, singleNode_frameworks_pr.values(), execTimePR_singleNode, yErrs=yErrExecPR_singleNode, title='PR single node (ugly)', yLabel='Execution time (s)', saveToFile="singleNodePR_execTime.png")
 
 
 

### 2020-09-05 documentation/paper/main.tex
weitere fixes zur Rückmeldung

@@
 \newcommand*\OK{\ding{51}} % a checkmark-style sign
+
+\newcommand{\todo}[1]{{\color{red!90!black}//\bf{TODO:} #1}}
 
@@
 
-\bibitem{bsp}
+\bibitem{Polymer}
 \BIBentryALTinterwordspacing
-L.~G. Valiant, ``A bridging model for parallel computation,'' \emph{Commun.
-  ACM}, vol.~33, no.~8, p. 103–111, Aug. 1990. [Online]. Available:
-  \url{https://doi.org/10.1145/79173.79181}
+K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
+  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
+  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
+  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
+  2015, p. 183–193. [Online]. Available:
+  \url{https://doi.org/10.1145/2688500.2688507}
+\BIBentrySTDinterwordspacing
+
+\bibitem{Ligra}
+\BIBentryALTinterwordspacing
+J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
+  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
+  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
+  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
+  Machinery, 2013, p. 135–146. [Online]. Available:
+  \url{https://doi.org/10.1145/2442516.2442530}
 \BIBentrySTDinterwordspacing
@@
 
-\bibitem{mapreduce}
-J.~Dean and S.~Ghemawat, ``Mapreduce: Simplified data processing on large
-  clusters,'' in \emph{In Proceedings of Operating Systems Design and
-  Implementation (OSDI}, pp. 137--150.
+\bibitem{pagerank}
+\BIBentryALTinterwordspacing
+S.~Brin and L.~Page, ``Reprint of: The anatomy of a large-scale hypertextual
+  web search engine,'' \emph{Computer Networks}, vol.~56, no.~18, pp. 3825 --
+  3833, 2012, the WEB we live in. [Online]. Available:
+  \url{http://www.sciencedirect.com/science/article/pii/S1389128612003611}
+\BIBentrySTDinterwordspacing
+
+\bibitem{bsp}
+\BIBentryALTinterwordspacing
+L.~G. Valiant, ``A bridging model for parallel computation,'' \emph{Commun.
+  ACM}, vol.~33, no.~8, p. 103–111, Aug. 1990. [Online]. Available:
+  \url{https://doi.org/10.1145/79173.79181}
+\BIBentrySTDinterwordspacing
 
@@
 \BIBentrySTDinterwordspacing
+
+\bibitem{graphx}
+J.~E. Gonzalez, R.~S. Xin, A.~Dave, D.~Crankshaw, M.~J. Franklin, and
+  I.~Stoica, ``Graphx: Graph processing in a distributed dataflow framework,''
+  in \emph{Proceedings of the 11th USENIX Conference on Operating Systems
+  Design and Implementation}, ser. OSDI'14.\hskip 1em plus 0.5em minus
+  0.4em\relax USA: USENIX Association, 2014, p. 599–613.
+
+\bibitem{powergraph}
+\BIBentryALTinterwordspacing
+J.~E. Gonzalez, Y.~Low, H.~Gu, D.~Bickson, and C.~Guestrin, ``Powergraph:
+  Distributed graph-parallel computation on natural graphs,'' in \emph{10th
+  {USENIX} Symposium on Operating Systems Design and Implementation ({OSDI}
+  12)}.\hskip 1em plus 0.5em minus 0.4em\relax Hollywood, CA: {USENIX}
+  Association, Oct. 2012, pp. 17--30. [Online]. Available:
+  \url{https://www.usenix.org/conference/osdi12/technical-sessions/presentation/gonzalez}
+\BIBentrySTDinterwordspacing
+
+\bibitem{mapreduce}
+J.~Dean and S.~Ghemawat, ``Mapreduce: Simplified data processing on large
+  clusters,'' in \emph{In Proceedings of Operating Systems Design and
+  Implementation (OSDI}, pp. 137--150.
 
@@
 
-\bibitem{Ligra}
+\bibitem{konect}
+J.~Kunegis, ``Konect: the koblenz network collection,'' 05 2013, pp.
+  1343--1350.
+
+\bibitem{snap}
 \BIBentryALTinterwordspacing
-J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
-  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
-  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
-  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
-  Machinery, 2013, p. 135–146. [Online]. Available:
-  \url{https://doi.org/10.1145/2442516.2442530}
-\BIBentrySTDinterwordspacing
-
-\bibitem{Polymer}
-\BIBentryALTinterwordspacing
-K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
-  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
-  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
-  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
-  2015, p. 183–193. [Online]. Available:
-  \url{https://doi.org/10.1145/2688500.2688507}
+Stanford~University. (2020, Sep.) Stanford network analysis project. [Online].
+  Available: \url{https://snap.stanford.edu}
 \BIBentrySTDinterwordspacing
@@
 
-\bibitem{konect}
-J.~Kunegis, ``Konect: the koblenz network collection,'' 05 2013, pp.
-  1343--1350.
-
 \end{thebibliography}

### 2020-09-05 documentation/paper/sections/overview/overview.tex
weitere fixes zur Rückmeldung

--- 2020-08-29 documentation/paper/sections/overview/overview.tex
@@
 
-\section{Overview}
-We start with a short overview describing functionality and characteristics of each framework, followed by their required input data formats.
+\section{Framework Overview}
+We start with a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
+All of the following frameworks are part of our testing.
 
-//
 \subsection{Galois and Gluon}

### 2020-09-05 documentation/paper/sections/preliminiaries.tex
weitere fixes zur Rückmeldung

@@
 An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
-The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set becomes 
+The edge set describes a number of connections or relations between two vertices.
+Depending on these relations, a graph can be directed or undirected.
+For a \emph{directed graph} the edge set is defined as
 \begin{equation*}
@@
 \end{equation*}
-and in the \emph{undirected} case $E$ is a set of two-sets 
-\begin{equation*}
-  E\subseteq\{\{x,y\}\,|\, x,y\in V, x\neq y\}.
-\end{equation*}
-The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.
-The size of a graph is defined as the number of edges $|E|$ // CITATION?.
+and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E$, it holds $(x,y)=(y,x)$. 
+The size of a graph is defined as the number of edges $|E|$\todo{citation?}.
 Independently of the graph being directed or not, a graph can be \emph{weighted}.
@@
 \begin{equation*}
-	P=(x_1,x_2,\ldots,x_n)\subseteq V^n
+	P=(x_1,x_2,\ldots,x_n)\in V^n
 \end{equation*}
@@
 In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.
-The most common implementations are Dijkstra's algorithm or BellmanFord.
+The most common sequential implementations are Dijkstra's algorithm or BellmanFord \cite{Polymer, Ligra, pregel}.
 
 \subsection{Breadth-First Search}
-Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input.
+Breadth-first search (BFS) is a search problem on a graph.
+It requires an unweighted graph and a start vertex as input.
 The output is a set of vertices that are reachable from the start vertex.
 In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.
-It is called breadth-first search because the algorithm searches in a path length-based way.
-First all paths of length 1 i.e.
-all neighbors of the start node are checked before checking paths of length 2 and so on.
-The search algorithm, where the paths of maximum length are checked first is called Depth-First Search.
+It is called Breadth-First search because the algorithm searches in a path length-based way.
+First all paths of length one i.e.
+all neighbors of the start vertex are checked before checking paths of length two and so on.
+The search algorithm, where the paths of maximum length are checked first is called Depth-First search.
 
 \subsection{PageRank}
-\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph.
-
+\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}.
+This represents a centrality metric of the vertices.
 The analogy is that the graph represents website pages of the Word Wide Web, that are hyperlinked between one another.
@@
 PageRank counts the number and quality of links to a page to estimate the importance of a page.
-
 The output of PageRank is a percentage for each vertex.
@@
 
-For example, Google Search uses PageRank to rank web pages in their search engine results.
-\paragraph{Delta-PageRank}
-This is a optimization of the traditional PageRank algorithm.
-The PageRank score of a vertex is only updated if the relative change of the PageRank is larger than some delta.
-In the next iteration only a subset of PageRanks has to be recalculated.
+The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.
+
+An optimization to the traditional PageRank implementation is called \emph{Delta-PageRank}.
+The PageRank score of a vertex is only updated if the relative change of the PageRank is larger than some user-defined delta.
+This effectively reduces the amount of vertices for which the PageRank has to be recalculated in following iterations.
+
+\subsection{Push and Pull Variants}
+\todo{}
 
@@
 Pairs of components can only communicate using messages, thus remote memory access is also only possible in this way.
+The Messages have a user-defined form and should be as small as possible to keep the network traffic low. \todo{Congest Model, log limit in was?}
 Synchronization is realized through barriers for some or all processes.
-
 BSP algorithms are performed in a series of global supersteps. These consist of three steps, beginng with the processors performing local computations concurrently.
@@
 
-One of the most famous graph processing systems, Pregel \cite{pregel} is based on the BSP computation model.
-Pregel and many open-source versions similar to it were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures \cite{Giraph}.
+One of the most famous graph processing systems, Pregel \cite{pregel} is based on the BSP computation model. We include Giraph, an open-source variant of Pregel in our evaluation.
+Pregel, Giraph and many frameworks similar to those were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures \cite{Giraph,graphx,powergraph}.
 
@@
 All of these features are automatically provided.
-
 Execution is performed in three phases:
@@
 \end{enumerate}
-
 Giraph \cite{Giraph} is an example of a system using this framework.
-
-\subsection{Congest Model}
-
-\subsection{Push/Pull Differences}

### 2020-09-05 plots/makedata.py

@@ -488,3 +488,4 @@ if False:
         #print("inter: ", (tmp[32] - tmp[16])/(32-16) * (24-16) + tmp[16])
             #last = k
     print(maxs)
+

### 2020-09-06 documentation/paper/sections/results/galoisSpeedup.tex
new Hugepage plots and updated speedups

@@
 \end{figure}
+
+\section{Hugepages}
+\begin{figure}
+	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths with Hugepages}
+	\label{fig:galoisHPSpeedupSSSP}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisHPThreads.png}
+	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search with Hugepages}
+	\label{fig:galoisHPSpeedupBFS}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}{\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
+		\caption{PageRank Pull}
+		\label{fig:galoisHPSpeedupPRPull}
+	\end{subfigure}
+	\begin{subfigure}{\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisHPThreads.png}
+		\caption{PageRank Push}
+		\label{fig:galoisHPSpeedupPRPush}
+	\end{subfigure}
+	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms using Hugepages.}
+\end{figure}

### 2020-09-06 plots/makedata.py

@@ -428,6 +428,8 @@ for i in range(len(graph)):
         continue
     if '-dist' in algo[i]:
         continue
+    if '-hp-' in algo[i]:
+        continue
     
     num = int(re.sub('\D', '', algo[i]))
     if not num in x:
@@ -439,7 +441,9 @@ for i in range(len(graph)):
         continue
     if '-dist' in algo[i]:
         continue
-    
+    if '-hp-' in algo[i]:
+        continue
+
     xVal = int(re.sub('\D', '', algo[i]))
     
     if 'sssp' in algo[i]:
@@ -488,4 +492,66 @@ if False:
         #print("inter: ", (tmp[32] - tmp[16])/(32-16) * (24-16) + tmp[16])
             #last = k
     print(maxs)
+
+
+
+
+###SPEEDUP GALOIS HP
+speedupGaloisSSSP_HP = {}
+speedupGaloisBFS_HP = {}
+speedupGaloisPRPush_HP = {}
+speedupGaloisPRPull_HP = {}
+xHP = []
+for g in graphs:
+    speedupGaloisSSSP_HP[g] = {}
+    speedupGaloisBFS_HP[g] = {}
+    speedupGaloisPRPush_HP[g] = {}
+    speedupGaloisPRPull_HP[g] = {}
+
+for i in range(len(graph)):
+    if not 'galois' in algo[i]:
+        continue
+    if '-dist' in algo[i]:
+        continue
+    if not '-hp-' in algo[i]:
+        continue
+    
+    num = int(re.sub('\D', '', algo[i]))
+    if not num in xHP:
+        xHP.append(num)
+xHP.sort()
+
+for i in range(len(graph)):
+    if not 'galois' in algo[i]:
+        continue
+    if '-dist' in algo[i]:
+        continue
+    if not '-hp-' in algo[i]:
+        continue
+    
+    xVal = int(re.sub('\D', '', algo[i]))
+    
+    if 'sssp' in algo[i]:
+        speedupGaloisSSSP_HP[graph[i]][xVal] = calcTime[i]
+    elif 'bfs' in algo[i]:
+        speedupGaloisBFS_HP[graph[i]][xVal] = calcTime[i]
+    elif 'pagerank-push' in algo[i]:
+        speedupGaloisPRPush_HP[graph[i]][xVal] = calcTime[i]
[…]

### 2020-09-06 plots/makeplots.py

@@ -5,18 +5,18 @@ from plotFunctions import *
 
 
 distributed = False
-singleNode = True
+singleNode = False
 
 sssp = False
 bfs = False
-pr = True
+pr = False
 
 calcTimes = False
 execTimes = False
 overheads = False
 
 
-galoisSpeedup = False
+galoisSpeedup = True
 
 ### PLOTS
 #SINGLE NODE
@@ -86,6 +86,19 @@ if galoisSpeedup:
 	line_plot(graphs, x, speedupGaloisPRPush, title='Speedup PR Push', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPushGaloisThreads.png")
 
 	line_plot(graphs, x, speedupGaloisPRPull, title='Speedup PR Pull', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPullGaloisThreads.png")
+
+
+## GALOIS HP THREAD COUNT
+if galoisSpeedup:
+	line_plot(graphs, xHP, speedupGaloisSSSP_HP, yLabel='Average calculation time speedup', xLabel='Thread count', title='Speedup SSSP HP', yScale='linear', saveToFile="singleNodeSSSPGaloisHPThreads.png")
+
+	line_plot(graphs, xHP[:-5], speedupGaloisSSSP_HP, title='Speedup SSSP HP', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisHPThreads_short.png")
+
+	line_plot(graphs, xHP, speedupGaloisBFS_HP, title='Speedup BFS HP', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeBFSGaloisHPThreads.png")
+
+	line_plot(graphs, xHP, speedupGaloisPRPush_HP, title='Speedup PR Push HP',  xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPushGaloisHPThreads.png")
+
+	line_plot(graphs, xHP, speedupGaloisPRPull_HP, title='Speedup PR Pull HP', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPullGaloisHPThreads.png")
 
 
 

### 2020-09-06 documentation/paper/sections/results/galoisSpeedup.tex
galois speedup sssp updated

@@
 For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads. 
-Twitter has the best speedup, requiring only 38\%\ the calculation time with 2 threads compared to one, 25\%\ using 4 and 12.9\%\ using 8 threads.
-Behaviour on friendster is similar with 52\%\ at 2 threads, 29\%\ at 4 and 16\%\ at 8 threads compared to one.
+Twitter has the best speedup overall. It is 2.6$\times$ with 2 threads compared to one, 4$\times$ with 4, 7.7$\times$ with 8 and 9.7$\times$ using 16 threads.
+Behaviour on friendster is similarly good. Here speedup is 1.9$\times$ at 2 threads compared to one, 3.5$\times$ at 4, 6.1$\times$ at 8 threads and 9.7$\times$ at 16 threads.
+Anything above 16 threads however no longer helps decrease the computation time significantly on any graph.
+Speedup above 16 threads is always less than double the speedup of 16 threads.
+The maximum measured speedups are 10$\times$ (96 threads) for wikipedia, 17$\times$ (96 threads) for twitter, 11$\times$ (96 threads) for rMat27, 16$\times$ (48 threads) for friendster and 19$\times$ (40 threads) for rMat28.
+In some cases increasing thread counts even prolongues calculation time.
+For example calculation on rMat28 is actually slower with 48 or 96 threads compared to 40 threads.
+For 40 threads, the speedup is nearly 19$\times$, on 48 threads 17$\times$ and with 96 threads only 15$\times$ compared to one thread.
 
-Anything above 32 threads however no longer helps decrease the computation time, in some cases even the opposite e.g.
-calculation on rMat28 is actually slower with 48 (13\% slower) or 96 threads (28\% slower) compared to 40 threads.
-
-Small graphs like flickr or orkut, neither benefit from more threads nor is the performance held up by synchronization overhead.
+Small graphs, i.e. flickr and orkut neither benefit from more threads nor is the performance significantly held up by synchronization overhead.
+Performance on flickr can not be sped up at all, with speedup on flickr being very close to 1 for 1 to 8 threads and between 0.7$\times$ to 0.9$\times$ from 16 to 96 threads.
+Orkut reaches maximum speedup of 1.6$\times$ at 16 threads. However on orkut, the speedup is always greater or equal to 1.
 

### 2020-09-06 documentation/paper/sections/results/galoisSpeedup.tex
galois speedup bfs updated

@@
 For our speedup results on BFS, \autoref{fig:galoisSpeedupBFS} shows the calculation time speedup of Galois' BFS.
-
-Flickr is sped up by about 5\%\ with 4 threads, any more than that will actually decrease performance, making computation time up to 15\%\ (96 threads) longer.
-
 On all graphs, the speedup never exceeds 6$\times$ even when using 96 threads.
-The initial speedup when switching from one to two threads is actually smaller than one, thus decreasing performance, on 4 of 7 graphs.
-Only flickr, twitter and rMat28 can benefit slightly by a speedup of 2\%, 41\%\ and 5\%\ respectively.
-
-When comparing four threads to one, BFS on all but two graphs can be sped up by anywhere from 5\%\ (flickr) to 68\% (rMat28). The speedup is thus only possible to a very small degree.
-For the other two graphs, computation on friendster with 8 threads is just as fast as one thread and computation on rMat27 is actually 19\% slower.
+For the smaller graphs (flickr, orkut), we have the same behaviour as on SSSP. Speedup is close to 1 in all cases, with orkut reaching a maximum speedup of 1.6$\times$ at 24 threads.
+On the larger graphs, speedup is possible but only to a very small degree.
+Wikipedia has a maximum speedup of 4.9$\times$ at 96 threads, for twitter it is 2.5$\times$ at 96 threads and 5.6$\times$ on friendster.
+The two synthetic data sets reach speedups of 4.2$\times$ and 4.7$\times$ for rMat27 and rMat28 on 96 threads.
 
@@
 Computation is hardly sped up on any graph other than flicker, where the reached maximum is 64\%. This maximum is reached at two threads, with speedup steadily declining above that.
-The rMat28 is the only other graph of one could say computation was sped up. Here we reached a maximum speedup of 31\%\ at 96 threads.
+The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 31\%\ at 96 threads.
 All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases and if so only by a small margin.

### 2020-09-10 documentation/paper/sections/graphFormats.tex
weitere fixes zur Rückmeldung
kleine fixes von der Rückmeldung und Polymer Overview

@@
 
-\section{Important Graph Formats}
+\section{Graph Formats}
 Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.
+
+The two most popular graph databases are those associated with the Koblenz Network Collection (KONECT) \cite{konect} and Stanford Network Analysis Project (SNAP) \cite{snap}.
+
 Data Sets retreived from KONECT can be directly read and translated.
 
-//SNAP
+\todo{SNAP?}
 
@@
 with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.
-There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.
+%There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

### 2020-09-10 documentation/paper/sections/overview/polymer.tex
kleine fixes von der Rückmeldung und Polymer Overview

@@
 
-Polymer is very similar to Ligra, in fact Polymer inherits the programming interfaces \texttt{EdgeMap} and \texttt{VertexMap} from Ligra as its main interface.\cite{Polymer}
+Polymer is a NUMA-aware graph-analytics system that inherits the scatter- gather programming interface from Ligra \cite{Ligra}. 
+Key differences are the data layout and access strategies, Polymer implements.
+The goal is to minimize random and remote memory accesses to improve performance. 
+%Computation is performed in a vertex-centric manner.
 
-Polymer is a vertex-centric framework, that tries to circumvent some of the random memory access drawbacks of such a design.
-It treats a NUMA machine as a distributed cluster and splits work and graph data accordingly between the nodes.
-Application-defined data is not distributed. Other runtime state data is allocated in a distributed way but only accessed through a global lookup table.
+The first optimization Polymer applies is data locality and access methods across NUMA nodes \cite{Polymer}.
+A general design principle for NUMA machines is to partition the input data so that computation can be grouped with the corresponding data on one node.
+Polymer adopts this and allocates graph data according to the access patterns.
+It treats a NUMA machine like a distributed cluster and splits work and graph data accordingly between the nodes. 
+Vertices and Edges are partitioned and then allocated across the corresponding memory nodes of the threads, eliminating most remote memory accesses.
+However, some computation requires vertices to perform computations on edges that are not in the local NUMA-node.
+For this case, Polymer introduces lightweight vertex replicas that are used to initiate computation on remote edges.
+
+Polymer furthermore has custom storage principles for application-defined data.
+For such data, the memory locations are static but the data undergo frequent dynamic updates.
+Due to frequent exchanges of application-defined data between the nodes, remote memory accesses are inevitable.
+Hence, Polymer allocates application-defined data with virtual adresses, while distributing the actual memory locations across the nodes.
+Thus, all updates are applied on a single copy of application-defined data.
+
+Runtime states that are dynamically allocated in each iteration would however create overhead due to repeated construction of a virtual address space.
+These states are thus stored in a custom lock-less (i.e. avoiding contention) lookup table.
+
+Polymer does not only optimize data locality but also its scheduling is custom.
+Time to synchronize threads on different cores increases dramatically with the growing number of involved sockets.
+Inter-node synchronization takes one order of magnitude longer time than intra-node synchronization \cite{Polymer}.
+Thus, Polymer implements a topology-aware hierarchical synchronization barrier.
+A group of threads on the same NUMA-node shares a partition of data.
+This allows them to first only synchronize with threads on the same NUMA-node.
+Only the last thread of each group sychronizes across groups (i.e.
+nodes).
+This behaviour decreases the amount of needed cache coherence broadcasts across the nodes.
+
+Furthermore, Polymer switches between different data structures representing the runtime state, similiar to Ligra's behaviour.
+The main deciding factor to switch is the amount of active vertices relative to an application-defined threshold.
+Polymer uses a lock-less tree structure representing the active vertices.
+The leaves use bitmaps, which are efficient for a large proportion of active vertices.
+When only a small amount of vertices is active, the drawbacks of traversing through sparse bitmaps can be avoided by switching data structures.
+
+Polymer inherits the programming interfaces \texttt{EdgeMap} and \texttt{VertexMap} from Ligra as its main interface.

### 2020-09-10 documentation/paper/sections/preliminiaries.tex
kleine fixes von der Rückmeldung und Polymer Overview

@@
 and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E$, it holds $(x,y)=(y,x)$. 
-The size of a graph is defined as the number of edges $|E|$\todo{citation?}.
+The size of a graph is defined as the number of edges $|E|$ \cite{newman2010networks}.
 Independently of the graph being directed or not, a graph can be \emph{weighted}.
@@
 Single-Source Shortest-Paths (SSSP) describes the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
-Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
+Input to the problem is a weighted graph $G=(V,E)$ and a start vertex $s\in V$. Output is the shortest possible distance from $s$ to each vertex in $V$. 
 The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
@@
 \subsection{PageRank}
+\todo{Aufbau unschön, Formel fehlt}
 \emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}. 
+The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.
+
 This represents a centrality metric of the vertices.
@@
 With a high probability the web surfer uses a random link from the web page they are currently on and with a smaller probability (called damping factor) they jump to a completely random web page.
-
-The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.
 

### 2020-09-10 documentation/paper/sections/testingMethods.tex
results for pr, removed two algos
kleine fixes von der Rückmeldung und Polymer Overview

@@
 
-\subsection{Testing Methods}
+\subsection{Environment}
 For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each.
@@
 Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
+Furthermore, Galois is a framework capable of utilizing hugepages. We include an evaluation using those on the single node as well.
 Unless mentioned otherwise, we always show results of each framework utilizing 96 threads (i.e. the maximum on our machines) for the single-node evaluation.

### 2020-09-10 plots/makeplots.py

@@ -13,7 +13,7 @@ pr = False
 
 calcTimes = False
 execTimes = False
-overheads = False
+overheads = True
 
 
 galoisSpeedup = True
@@ -141,7 +141,7 @@ if distributed and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), overheadSSSPNormalized_distributed, title='SSSP distributed', yLabel='Overhead time (s) (normalized)', saveToFile="distributedSSSP_overheadTimeNormalized.png", yScale='linear')
 
-		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), overheadSSSP_distributed_normalizedToGalois, yLabel='Galois Push overhead times', saveToFile="distributedSSSP_overheadTimeNormalizedToGalois.png", yScale='linear')
+		grouped_bar_plot(graphs, dist_frameworks_sssp.values(), overheadSSSP_distributed_normalizedToGalois, yLabel='Galois Push overhead times', saveToFile="distributedSSSP_overheadTimeNormalizedToGalois.png", yScale='log')
 
 	#BFS
 	if bfs or overheads:

### 2020-09-10 documentation/paper/sections/overview/gemini.tex

@@
 
-Gemini is a framework for parallel graph processing \cite{Gemini}.
+Gemini is a framework for distributed graph processing \cite{Gemini}.
 It was developed with the goal to deliver a generally better performance through efficient communication.
@@
 
-Gemini is fairly lightweight and seems well structured with a clearly defined API between the core framework and the implementations of the individual algorithms.
+The real bottleneck of distributed systems is not the communication itself, but the extra instructions, as well as memory references and a lower usage of multiple cores compared to the shared memory counterparts.
+The main reasons for this are (1) the use of hash maps to convert vertex IDs between local and global states, (2) the maintenance of vertex replicas, (3) the communication-bound apply phase in the GAS abstraction, and (4) the lack of dynamic scheduling \cite{Gemini}.
+Gemini tries to work around all the problems mentioned above, by implemening a message-based system from scratch and getting rid of the extra mapping layer between shared-memory computation and communication.
+Therefore Ligra's push-pull computation model was adopted and applied to the distributed computation.
+Furthermore a chunk-based partitioning scheme was implemented, which allows to partition a graph without a large overhead.
+Gemini also implements a co-scheduling mechanism to connect the computation and inter-node communication.
+
+Gemini is fairly lightweight and has a clearly defined API between the core framework and the implementations of the individual algorithms.
 The five already implemented algorithms are single source shortest path (sssp), breath first search (bfs), pagerank, connected components (cc) and biconnected components (bc).

### 2020-09-11 documentation/paper/sections/overview/galois.tex
more consistency and todos

@@
 %!TEX root=../../main.tex
+\todo{Rückmeldung 2 umsetzen}
 
@@
 For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
-Galois includes applications for many graph analytics problems, among these are single-source shorthest-paths (sssp), breath-first-search (bfs) and pagerank.
+Galois includes applications for many graph analytics problems, among these are Single-Source Shorthest-Paths (SSSP), Breath-First Search (BFS) and PageRank (PR).
 For most of these applications Galois offers several different algorithms to perform these analytics problems and many setting options like the amount of threads used or policies for splitting the graph.

### 2020-09-11 plots/makedata.py
unified colors for each framework

@@ -83,6 +83,8 @@ for f in frameworks:
         fname += "-cpu-96thread" # select Galois thread count here
     k = 0 # graph index
     for g in graphs:
+        if g == 'rMat28':
+            continue
         for i in range(len(graph)):
             if g == graph[i] and fname == algo[i]:
                 tmpCalc.append(calcTime[i])
@@ -102,8 +104,29 @@ for f in frameworks:
     overheadSSSP_singleNode.append(overhead)
     overheadSSSPNormalized_singleNode.append(overheadNormalized)
 
-
-
+if False:
+    k = 0
+    gr = 0
+    s = np.zeros(len(graphs)-1)
+    for f in frameworks:
+        if f == 'Giraph':
+            k += 1
+            continue
+        for i in range(len(calcTimeSSSP_singleNode[k])):
+            s[i] += calcTimeSSSP_singleNode[k][i]
+        k += 1
+    s /= 4
+    print("average calculation time:", s)
+    k = 0
+    for f in frameworks:
+        # if f == 'Giraph':
+        #     k += 1
+        #     continue
+        print(f)
+        for i in range(1, len(execTimeSSSP_singleNode[k])):
+            print(graphs[i], round((execTimeSSSP_singleNode[k][i] / execTimeSSSP_singleNode[0][i]) * 100), "%")
+            print(graphs[i], round(execTimeSSSP_singleNode[k][i] * 100) / 100, "s")
+        k += 1
 
 ## SSSP DISTRIBUTED
 dist_frameworks_sssp = {

### 2020-09-11 plots/makeplots.py

@@ -4,19 +4,29 @@ from plotFunctions import *
 
 
 
-distributed = False
-singleNode = False
+distributed = True
+singleNode = True
 
-sssp = False
-bfs = False
-pr = False
+sssp = True
+bfs = True
+pr = True
 
 calcTimes = False
 execTimes = False
-overheads = True
+overheads = False
 
 
-galoisSpeedup = True
+galoisSpeedup = False
+
+
+colors = {"Galois":"#1f77b4",
+"Galois Push":"#1e2bb5",
+"Galois Pull":"#1eb5a8",
+"Gemini":"#ff7f0e",
+"Giraph":"#2ca02c",
+"Ligra":"#d62728",
+"Polymer":"#9467bd",
+"Polymer Delta":"#bc65bb"}
 
 ### PLOTS
 #SINGLE NODE
@@ -24,14 +34,15 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 	## CALC TIME
 	#SSSP
 	if sssp or calcTimes:
-		grouped_bar_plot(graphs, frameworks, calcTimeSSSP_singleNode, yErrs=yErrCalcSSSP_singleNode, title='SSSP single node', yLabel='Calculation time (s)', saveToFile="singleNodeSSSP_calcTime.png")
+		grouped_bar_plot(graphs[:-1], frameworks, calcTimeSSSP_singleNode, yErrs=yErrCalcSSSP_singleNode, yLabel='Calculation time (s)', saveToFile="singleNodeSSSP_calcTime.png", colors_=colors)
 
 	#BFS
 	if bfs or calcTimes:
-		grouped_bar_plot(graphs, frameworks, calcTimeBFS_singleNode, yErrs=yErrCalcBFS_singleNode, title='BFS single node', yLabel='Calculation time (s)', saveToFile="singleNodeBFS_calcTime.png")
+		grouped_bar_plot(graphs, frameworks, calcTimeBFS_singleNode, yErrs=yErrCalcBFS_singleNode, title='BFS single node', yLabel='Calculation time (s)', saveToFile="singleNodeBFS_calcTime.png", colors_=colors)
 
+	#PR
 	if pr or calcTimes:
-		grouped_bar_plot(graphs, singleNode_frameworks_pr.values(), calcTimePR_singleNode, yErrs=yErrCalcPR_singleNode, title='PR single node (ugly)', yLabel='Calculation time (s)', saveToFile="singleNodePR_calcTime.png", scaleFactor=3.25)
[…]

### 2020-09-11 plots/plotFunctions.py

@@ -6,12 +6,11 @@ import matplotlib
 
 #
 # values = [[f1_g1,f1_g2,...],[f2_g1,f2_g2,...],...]
-def grouped_bar_plot(group_names, legend_names, values, yErrs=None, width=0.35, scaleFactor=None, title='', yLabel='', yScale='log', saveToFile=None):
+def grouped_bar_plot(group_names, legend_names, values, colors_=None, yErrs=None, width=0.35, scaleFactor=None, title='', yLabel='', yScale='log', saveToFile=None):
 	group_count = len(group_names)
 
 	if len(legend_names) != len(values):
 		raise Exception("Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values)))
-
 
 	if scaleFactor == None:
 		scaleFactor = (len(legend_names) + 1.5) * width
@@ -32,7 +31,10 @@ def grouped_bar_plot(group_names, legend_names, values, yErrs=None, width=0.35, 
 		while group_count > len(tmp):
 			tmp.append(float('nan'))
 
-		ps.append(ax.bar(scaleFactor * ind + i * width, tmp, width, yerr=yErr, bottom=0))
+		if not colors_ == None:
+			ps.append(ax.bar(scaleFactor * ind + i * width, tmp, width, yerr=yErr, bottom=0, color=colors_[legend_names[i]]))
+		else:
+			ps.append(ax.bar(scaleFactor * ind + i * width, tmp, width, yerr=yErr, bottom=0))
 		i += 1
 
 	ax.set_xticks(scaleFactor * ind + (len(legend_names)-1) / 2 * width)

### 2020-09-11 documentation/paper/sections/results/sssp.tex
more consistency and todos
single node results for sssp

@@
 \subsubsection{Single-node}
-Beginning with the single-node performance, \autoref{fig:singleNodeSSSP} shows the average calculation and execution times for SSSP on the different frameworks.
-
-Because we did not see the perfomance going down when increasing thread counts on Galois, we always refer to Galois with 96 threads in this section (including the figures). 
-For a more detailed analysis of Galois' behaviour with different thread counts, see \autoref{sec:galois_speedup}.
-
-TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.
-\begin{figure}
-	\begin{subfigure}{0.3\textwidth}
+\begin{figure*}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
-		\caption{Calculation times for SSSP on a single node}
+		\caption{Calculation times}
 		\label{fig:singleNodeSSSP_calc}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
-		\caption{Execution times for SSSP on a single node}
+		\caption{Execution times}
 		\label{fig:singleNodeSSSP_exec}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
-		\label{fig:singleNodeSSSP_overheadNormalized}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTime.png}
+		\caption{Overhead time}
+		\label{fig:singleNodeSSSP_overhead}
 	\end{subfigure}
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.
-	The runs on rMat28 for Ligra and Polymer failed and the frameworks were unable to complete the task.}
+	\caption{Average times for SSSP on a single computation node, black bars represent one standard deviation in our testing.}
 	\label{fig:singleNodeSSSP}
-\end{figure}
+\end{figure*}
+Beginning with the single-node performance, \autoref{fig:singleNodeSSSP} shows the average calculation and execution times for SSSP on the different frameworks.
+Note, that Giraph ran out of memory ($>$250 GB) for all graphs larger than wikipedia (twitter, rMat27, friendster and rMat28).
+Thus, the data points on the larger graphs for Giraph are missing in the figures and our evaluation.
+Also, both Ligra and Polymer failed on rMat28. Because of that, removed rMat28 in this section.
+
+Upon analyzing the calculation time (cf.
+\autoref{fig:singleNodeSSSP_calc}), most obvious is the fact that Giraph is at least one order of magnitude slower than any other framework.
+The only exception is Gemini, where Giraph is \emph{only} 4$\times$ slower on wikipedia.
+In general however, Giraph is on average 24$\times$ slower than the other frameworks on flickr, 17$\times$ on orkut and 12$\times$ on wikipedia.
+
+Comparing the other frameworks (i.e.
+excluding Giraph in the further comparisons), shows them to perform similar on the smaller graphs.
+There, the average calculation times of the four frameworks (Galois, Gemini, Ligra, Polymer) are 35ms on flickr, 77ms on orkut, 538ms on wikipedia.
+The four frameworks are always within $x$ ms of this average. Galois and Gemini being 24ms and 12ms faster than the average. Ligra and Polymer are slower by 27ms and 9ms.
+Orkut is the graph where all the four frameworks are within 20ms of their average, Galois being the outlier at 19ms slower than average.
+For wikipedia, Galois and Ligra are close to each other, while Gemini is 560ms (104\%) slower than the average of the four frameworks.
+And Polymer is the fastest framework on wikipedia, being 330ms (61\%) faster than the others.
+For the larger graphs, Polymer has the shortest computation time as well.
+Gemini is second, taking about 1.9$\times$ times the computation time of Polymer on twitter and friendster while being equally fast (8\% longer) on rMat27.
+Galois and Ligra however have much longer computation times compared to Polymer.
+Galois takes anywhere from 3.3$\times$ (twitter) to 4.2$\times$ (friendster) the computation time of Polymer.
+Ligra requires 9.8$\times$, 7.0$\times$ and 3.2$\times$ the computation time of Polymer on twitter, rMat27 and friendster respectively.
+This makes Ligra the slowest Framework on twitter and rMat27.
+Galois is the slowest on friendster.
+However, while especially Galois is comparably slow to Polymer in the computation time, it is important to keep in mind that Polymer could not finish computation on rMat28.
+Meanwhile both Gemini and Galois managed just fine.
+
+Moving on to the execution times, which can be seen in \autoref{fig:singleNodeBFS_exec}. 
+A general overview shows Galois as the clear winner on most graphs. It has the smalles computation times on all graphs except flickr, where Galois is 25ms (8\%) slower than Ligra. 
+On the other 5 graphs however, Galois has the smaller execution time by one or sometimes even two orders of magnitude. 
+For example, on orkut Galois requires 0.78s to execute, while Gemini takes 13.8s, Giraph takes 135.9s, Ligra takes 2.66s and Polymer takes 26.6s.
+This goes on for the larger graphs, Ligra being second-fastest and Polymer and Gemini close together but both always at least one order of magnitude slower than Galois.
 
 \subsubsection{Distributed}
-
-\begin{figure}
+\begin{figure*}
 	\centering
-	\begin{subfigure}{\columnwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
-		\caption{Calculation times for distributed SSSP}
+		\caption{Calculation times}
 		\label{fig:distributedSSSP_calc}
 	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{Execution times for distributed SSSP}
+		\caption{Execution times}
 		\label{fig:distributedSSSP_exec}
 	\end{subfigure}
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTime.png}
+		\caption{Overhead times}
+		\label{fig:distributedSSSP_overhead}
+	\end{subfigure}
+	\caption{Average times for SSSP on the distributed cluster, black bars represent one standard deviation in our testing.}
 	\label{fig:distributedSSSP}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTime.png}
-	\caption{Distributed Single-Source Shortest-Paths overhead times of each framework}
-	\label{fig:distributedSSSP_overhead}
-\end{figure}
+\end{figure*}
 
@@
 
+\todo{Absolute Zeiten in \autoref{tbl:ssspexec} mit angeben}
 \begin{table}

### 2020-09-12 documentation/paper/sections/preliminiaries.tex
more consistency and todos
clean up repository and one more todo
paper preliminiaries push and pull variants

@@
 \section{Preliminaries}
-TODO //
+\todo{Section beschreiben}
 
@@
 \subsection{Push and Pull Variants}
-\todo{}
+Many parallel graph algorithms, as well as the three we consider here, are implemented by iterating over vertices, or edges \cite{Galois}.
+The respective vertex or edge, which is considered in an iteration, is also called \emph{active} vertex or edge.
+Applying an operator to this \emph{active} vertex or edge is called an \emph{activity}.
+This operator considers always only a local so called \emph{neighborhood} to the \emph{active} vertex or edge.
+To this \emph{neighborhood} belong only vertices and edges, which are in the direct surrounding of the \emph{active} vertex or edge, which allows to parallelise the iteration relatively easy.
+
+In general such an operator can change the structure of the \emph{neighbourhood} too.
+But in the following we assume, that an operator changes only \emph{labels} of vertices or edges in the \emph{neighbourhood}.
+Often this operator can be implemented in two different ways, called \emph{push style} or \emph{pull style}.
+A \emph{push-style} operator reads the value of the \emph{active} vertex or edge and updates the values of its \emph{neighborhood}.
+In contrast the \emph{pull-style} operator reads all values of its \emph{neighborhood} and updates the value of the \emph{active} vertex or edge.
+
+\emph{Pull-style} operators need less synchronization in parallel implementations, because unlike \emph{pull style} there is only one write and many read operations.
+Thus locks can be avoided.
+So \emph{push-style} operators are more efficient, if there are only a few \emph{active} vertices or edges at the same time, or the \emph{neighborhoods} do not overlap, which can not be avoided in general.
+On the other hand \emph{pull-style} operators are more efficient, if there are many \emph{active} vertices or edges at the same time.
 
@@
 Giraph \cite{Giraph} is an example of a system using this framework.
+
+\subsection{NUMA-awareness}
+\todo{Section}

### 2020-09-12 documentation/paper/sections/preliminiaries.tex
paper prelimiaries push and pull

@@
 
-In general such an operator can change the structure of the \emph{neighbourhood} too.
+In general such an operator can change the whole structure of the \emph{neighbourhood}.
 But in the following we assume, that an operator changes only \emph{labels} of vertices or edges in the \emph{neighbourhood}.
+\emph{Labels} of edges also called \emph{weights}.
 Often this operator can be implemented in two different ways, called \emph{push style} or \emph{pull style}.
-A \emph{push-style} operator reads the value of the \emph{active} vertex or edge and updates the values of its \emph{neighborhood}.
+A \emph{push-style} operator reads the \emph{label} of the \emph{active} vertex or edge and updates the \emph{label} of its \emph{neighborhood}.
 In contrast the \emph{pull-style} operator reads all values of its \emph{neighborhood} and updates the value of the \emph{active} vertex or edge.

### 2020-09-14 documentation/paper/sections/overview/galois.tex
Galois writeup complete

--- 2020-08-11 documentation/paper/sections/overview/galois.tex
@@
 %!TEX root=../../main.tex
+%TODO add bibliography for origGalois and cite it here
+Galois \cite{origGalois} is a general purpose library designed for parallel programming.
+The system reduces the complexity of writing parallel applications by providing implicitly parallel (unordered or partially ordered) set iterators.
+These iterators perform operations optimistically, detect arising conflicts and resolve these by invoking inverse methods accordingly.
+The tasks in the sets can be ordered, the tasks may be executed out of order, but the conflict resolution will ensure semantic equivalence to sequential strictly ordered execution.
 
-Galois\cite{Galois} is a general purpose library designed for parallel programming.
+%Galois \cite{Galois} is a general purpose library designed for parallel programming.
 The Galois system supports fine grain tasks, allows for autonomous, speculative execution of these tasks and grants control over the task scheduling policies to the application.
@@
 
-For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
+The graph analysis subsystem of Galois \cite{Galois} provides a library of scalable data structures and a topology aware priority scheduler, including optimizations for distributed execution.
+The scheduler splits the tasks into bags according to a specified partial order, which in turn provide the cores with chunks of tasks.
+A global map manages the various bags.
+Every thread keeps a lazy cache of a portion of the global map, in order to reduce the strain on the global map.
+Galois includes applications for many graph analytics problems, among these are SSSP, BFS and pagerank.
+For most of these applications Galois offers several different algorithms to perform these analytics problems and many options e.g.
+the amount of threads used or policies for splitting the graph.
+All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment
+
+%For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
 Galois includes applications for many graph analytics problems, among these are single-source shorthest-paths (sssp), breath-first-search (bfs) and pagerank.
@@
 
-Gluon\cite{vertGalois} is a framework written by the Galois team as a middleware to write graph analysis applications for distributed systems.
+Gluon \cite{vertGalois} is a framework written for Galois as a middleware for distributed graph analysis applications.
 It reduces the communication overhead needed in distributed environments by exploiting structural and temporal invariants.
+Depending on the used graph partitioning policy only a subset of the messages of a naive approach must be sent (structural invariant).
+Gluon establishes a mapping of local vertex ID's to the order in which the values will be sent/received between the owner and each mirror.
+A message includes a bit vector where a one in the i-th position means that the according vertex of the established order has been updated.
+Thus a message only has to include updated values without the need to state the vertex ID (temporal invariants).
+Gluon is embedded in Galois, but can be integrated in other graph analysis frameworks as well \cite{vertGalois}.
 
-The code of Gluon is embedded in Galois. It is possible to integrate Gluon in other frameworks too, which the Galois team showed in their paper\cite{vertGalois}.
+%The code of Gluon is embedded in Galois. It is possible to integrate Gluon in other frameworks \cite{vertGalois}.

### 2020-09-14 documentation/paper/sections/graphFormats.tex

@@
 The two most popular graph databases are those associated with the Koblenz Network Collection (KONECT) \cite{konect} and Stanford Network Analysis Project (SNAP) \cite{snap}.
-
-Data Sets retreived from KONECT can be directly read and translated.
-
-\todo{SNAP?}
+Data Sets retreived from one of them can be directly read and translated.
 
@@
 The $o_k$ are the so-called offsets. Each vertex $k$ has an offset $o_k$, that describes an index in the following list of the $t_i$.
-The $t_i$ are vertex IDs describing target nodes of a directed edge. 
-The index $o_k$ in the list of target nodes is the point where edges outgoing from vertex $k$ begin to be declared. So vertex $k$ has the outgoing edges
+The $t_i$ are vertex IDs describing target vertices of a directed edge.
+The index $o_k$ in the list of target vertices is the point where edges outgoing from vertex $k$ begin to be declared. So vertex $k$ has the outgoing edges
 \begin{equation*}
@@
 
-For the WeightedAdjacencyGraph format, the weights are appended to the end of the file in an order corresponding to the target nodes.
+For the WeightedAdjacencyGraph format, the weights are appended to the end of the file in an order corresponding to the target vertices.
 
@@
 An edge list is a set of directed edges $(s_1,t_1),(s_2,t_2),\ldots$ where $s_i$ is a vertex ID representing the start vertex and $t_i$ is a vertex ID representing the target vertex.
-In the format, there is one edge per line and the vertex IDs $s_i, t_i$ are separated with any whitespace character.
+In the format, there is one edge per line and the vertex IDs $s_i, t_i$ are separated with an arbitray amount of whitespace characters.
 

### 2020-09-14 documentation/paper/sections/overview/gemini.tex

@@
 Gemini is fairly lightweight and has a clearly defined API between the core framework and the implementations of the individual algorithms.
-The five already implemented algorithms are single source shortest path (sssp), breath first search (bfs), pagerank, connected components (cc) and biconnected components (bc).
+The five already implemented algorithms are Single-Source Shortest-Path (SSSP), Breath-First Search (BFS), PageRank (PR), Connected-Components (CC) and Biconnected-Components (BC).

### 2020-09-14 documentation/paper/sections/overview/giraph.tex

@@
 %!TEX root=../../main.tex
-
 Apache Giraph is an example for an open-source system similar to Pregel.
@@
 
-Computation is vertex oriented and iterative.
-For each iteration step called superstep, the Compute method implementing the algorithm is invoked on each active vertex, with every vertex being active in the beginning.
+The computation is vertex oriented and iterative.
+For each iteration step called superstep, the \emph{compute} method implementing the algorithm is invoked on each active vertex, with every vertex being active in the beginning.
 This method receives messages sent in the previous superstep as well as its vertex value and the values of outgoing edges.
 With this data the values are modified and messages to other vertices are sent.
-Communication between vertices is only performed via messages, so a vertex has no access to values of other vertices or edges other than its own outgoing ones.
+Communication between vertices is only performed via messages, so a vertex has no direct access to values of other vertices.
+The only visible information is the set of attached edges and their weights.
 Supersteps are synchronized with barriers, meaning that all messages only get delivered in the following superstep and computation for the next superstep can only begin after every vertex has finished computing the current superstep.
 Edge and vertex values are retained across supersteps.
-Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will reactivate the vertex again.
-A vote-to-halt method is applied, i.e. if all vertices are inactive or if a user defined superstep is reached the computation ends.
+Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will reactivate the vertex.
+A vote-to-halt method is applied, i.e. if all vertices are inactive or if a user defined superstep number is reached the computation ends.
 Once calculation is finished, each vertex outputs some local information (e.g. the final vertex value) as result.
@@
 Hadoop is a MapReduce infrastructure providing a fault tolerant basis for large scale graph processing.
-Hadoop supplies a distributed file system (HDFS), on which all computation is performed.
+Hadoop supplies a distributed file system (HDFS), on which all computations are performed.
 Giraph is thus, even when only using a single node, running in a distributed manner.
-But expanding single-node processing to a multi-node cluster is seamless.
-Giraph uses the Map functionality of Hadoop to run the algorithms, Reduce is only used as the identity function.
+Hence, expanding single-node processing to a multi-node cluster is seamless.
+Giraph uses the Map functionality of Hadoop to run the algorithms. Reduce is only used as the identity function.
 
 Giraph being an Apache project makes it the most actively maintained and tested project in our comparison.
-Over the course of this project, several new updates were pushed to Giraph's source repository\footnote{\url{https://gitbox.apache.org/repos/asf?p=giraph.git}}.
+While writing this paper, several new updates were pushed to Giraph's source repository\footnote{\url{https://gitbox.apache.org/repos/asf?p=giraph.git}}.

### 2020-09-14 documentation/paper/sections/relatedWork.tex

--- 2020-09-10 documentation/paper/sections/preliminiaries.tex
@@
 
-\section{Preliminaries}
-TODO //
-
-\subsection{Graphs and Paths}
-An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
-The edge set describes a number of connections or relations between two vertices.
-Depending on these relations, a graph can be directed or undirected.
-For a \emph{directed graph} the edge set is defined as
-\begin{equation*}
-  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
-\end{equation*}
-and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E$, it holds $(x,y)=(y,x)$. 
-The size of a graph is defined as the number of edges $|E|$ \cite{newman2010networks}.
-Independently of the graph being directed or not, a graph can be \emph{weighted}.
-In this case a function $w:E\rightarrow \mathbb R$ is introduced, that maps an edge to a numerical value, further describing the relation.
-
-A \emph{Path} from starting vertex $s$ to target vertex $t$ is a sequence of vertices
-\begin{equation*}
-	P=(x_1,x_2,\ldots,x_n)\in V^n
-\end{equation*}
-with the condition $(x_i,x_{i+1})\in E$ for each $i\in\{1,\ldots,n-1\}$ and $x_1=s, x_n=t$.
-Thus we call a target $t$ \emph{reachable} from $s$ if a Path from $s$ to $t$ exists.
-
-\subsection{Single-Source Shortest-Paths}
-Single-Source Shortest-Paths (SSSP) describes the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
-Input to the problem is a weighted graph $G=(V,E)$ and a start vertex $s\in V$. Output is the shortest possible distance from $s$ to each vertex in $V$. 
-The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
-In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.
-The most common sequential implementations are Dijkstra's algorithm or BellmanFord \cite{Polymer, Ligra, pregel}.
-
-\subsection{Breadth-First Search}
-Breadth-first search (BFS) is a search problem on a graph.
-It requires an unweighted graph and a start vertex as input.
-The output is a set of vertices that are reachable from the start vertex.
-In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.
-It is called Breadth-First search because the algorithm searches in a path length-based way.
-First all paths of length one i.e.
-all neighbors of the start vertex are checked before checking paths of length two and so on.
-The search algorithm, where the paths of maximum length are checked first is called Depth-First search.
-
-\subsection{PageRank}
-\todo{Aufbau unschön, Formel fehlt}
-\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}. 
-The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.
-
-This represents a centrality metric of the vertices.
-The analogy is that the graph represents website pages of the Word Wide Web, that are hyperlinked between one another.
-A website that is more important is likely to receive more links from other webiste.
-PageRank counts the number and quality of links to a page to estimate the importance of a page.
-The output of PageRank is a percentage for each vertex.
-This percentage, called the PageRank of the vertex, is the probability with which a web surfer starting at a random web page reaches this webpage (vertex).
-With a high probability the web surfer uses a random link from the web page they are currently on and with a smaller probability (called damping factor) they jump to a completely random web page.
-
-An optimization to the traditional PageRank implementation is called \emph{Delta-PageRank}.
-The PageRank score of a vertex is only updated if the relative change of the PageRank is larger than some user-defined delta.
-This effectively reduces the amount of vertices for which the PageRank has to be recalculated in following iterations.
-
-\subsection{Push and Pull Variants}
-\todo{}
-
-\subsection{Bulk-Synchronous Parallel Model}
-\label{sec:bsp}
-The Bulk-Synchronous Parallel (BSP) model is a computation model developed by Leslie Valiant \cite{bsp}. It is commonly used in computation environments with large amounts of synchronous computation.
-
-This model describes components, a communication network between those components and a method of synchronization.
-The components are capable of performing computations and transactions on \emph{local} memory.
-Pairs of components can only communicate using messages, thus remote memory access is also only possible in this way.
-The Messages have a user-defined form and should be as small as possible to keep the network traffic low. \todo{Congest Model, log limit in was?}
-Synchronization is realized through barriers for some or all processes.
-BSP algorithms are performed in a series of global supersteps. These consist of three steps, beginng with the processors performing local computations concurrently.
-This step can overlap with the second, the communication between components. Processes can exchange information to access remote data.
-Lastly, processes reaching a barrier wait until all other processes have reached the same barrier.
-
-One of the most famous graph processing systems, Pregel \cite{pregel} is based on the BSP computation model. We include Giraph, an open-source variant of Pregel in our evaluation.
-Pregel, Giraph and many frameworks similar to those were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures \cite{Giraph,graphx,powergraph}.
-
-\subsection{MapReduce}
-The MapReduce model is a computation infrastructure developed by Google to reliably handle large data sets on distributed clusters \cite{mapreduce}.
-
-A user specifies just the two functions Map and Reduce.
-The system hides the details of parallelization, fault-tolerance, data distribution and load balancing away from the application logic.
-All of these features are automatically provided.
-Execution is performed in three phases:
-\begin{enumerate}
-	\item Map phase: The input data is distributed between a set of Map processes, the Map functionality is specified by the user.
-Ideally all Map processes run in parallel so the map processes need to be independent.
-Results from this phase are written into (multiple) intermediate storage points.
-	\item Shuffle phase: The results are grouped according to a key provided by the Map algorithm. Each set of results is then handed to one system for the next phase.
-	\item Reduce phase: Every set of intermediate results is input to exactly one reduce process. The Reduce functionality is again specified by the user and ideally runs in parallel.
-\end{enumerate}
-Giraph \cite{Giraph} is an example of a system using this framework.
+\section{Related Work}
+\todo{Section}

### 2020-09-14 documentation/paper/sections/results/bfs.tex

@@
 \subsubsection{Single-Node}
-TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.
+\todo{Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.}
 
@@
 This is expected since both are graph traversal algorithms starting in one source vertex.
-Hence, calculation complexity for each node and communication overhead is similar.
+Hence, calculation complexity for each vertex and communication overhead is similar.
 All measurements can be seen in \autoref{fig:distributedBFS}.

### 2020-09-14 documentation/paper/sections/results/bfs.tex
bfs single node results and images rescaled and moved to be more unified

@@
 \subsection{Breadth-first search}
+As with our results for SSSP, for BFS we will begin with our single-node results before looking at the distributed scenario. Lastly we will compare the two cases.
 \subsubsection{Single-Node}
-\todo{Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.}
-
-In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
 \begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
-		\caption{Calculation times for BFS on a single node}
+		\caption{Calculation time}
 		\label{fig:singleNodeBFS_calc}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
-		\caption{Execution times for BFS on a single node}
+		\caption{Execution time}
 		\label{fig:singleNodeBFS_exec}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTime.png}
+		\caption{Overhead time}
 		\label{fig:singleNodeBFS_overheadNormalized}
 	\end{subfigure}
+	\caption{Average times for BFS on a single computation node, black bars represent one standard deviation in our testing}
+\end{figure*}
+Just like with SSSP, Giraph ran out of memory ($>$250 GB) on any graph larger than wikipedia.
+Also, Polymer failed to complete on rMat28.
+Ligra, that failed during SSSP however completed our benchmark for BFS.
 
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing}
-\end{figure*}
+The calculation times provided in \autoref{fig:singleNodeBFS_calc} show Gemini and Ligra to be comparable in their performance. 
+Their computation times deviate less than 151ms on all graphs except wikipedia and rMat28, wight Ligra being the faster framework on most graphs.
+Ligra is 39ms faster on orkut, 2ms on twitter, 50ms on rMat27 and 151ms on friendster.
+In turn, Gemini is 17 ms faster than Ligra on flickr.
+Only on wikipedia and rMat28, there is a noticeable difference between the two frameworks. Ligra is 1.2s (7.8$\times$) faster than Gemini on wikipedia and 696ms (4.4$\times$) faster on rMat28.
+
+For the remaining frameworks, we compare to Ligra since it is generally slightly faster than Gemini.
+Giraph and Polymer are one to two orders of magnitude slower than both Ligra or Gemini. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
+Giraph takes 13$\times$, 32$\times$ and 27$\times$ longer than Ligra on flickr, orkut and wikipedia respectively.
+As we said, Polymer is comparable to Ligra on flickr, for the other graphs however, its computation times are even longer than those of Giraph.
+The upside is though, that Polymer managed -- contrary to Giraph -- to finish computation on some larger graphs.
+Polymer takes 55$\times$ longer on orkut, 60$\times$ on wikipedia, 129$\times$ on twitter, 530$\times$ on rMat27 and 240$\times$ on friendster. 
+Especially interesting is here the difference between the times of rMat27 and friendster.
+There are 20\% (439M) edges more in friendster, yet the computation for the synthetic rMat27 takes 42\% longer.
+
+Galois calculation performance is comparable to Gemini or Ligra on the smaller three graphs (flickr, orkut and wikipedia).
+On the larger graphs however, Galois is slower than Gemini or Ligra but always faster than Polymer by one order of magnitude.
 
 \subsubsection{Distributed}
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
-		\caption{Calculation times for distributed BFS}
+		\caption{Calculation time}
 		\label{fig:distributedBFS_calc}
 	\end{subfigure}
-	% \hfil
-	\begin{subfigure}{\columnwidth}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
-		\caption{Execution times for distributed BFS}
+		\caption{Execution time}
 		\label{fig:distributedBFS_exec}
 	\end{subfigure}
-	% \hfil
+	\hfil
 	% \begin{subfigure}{0.3\textwidth}
@@
 	% \end{subfigure}
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
+	\caption{Average times for BFS on the distributed cluster, black bars represent one standard deviation in our testing}
 	\label{fig:distributedBFS}
-\end{figure}
+\end{figure*}
 For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case.

### 2020-09-14 documentation/paper/sections/results/pr.tex
more consistency and todos
bfs single node results and images rescaled and moved to be more unified

@@
 
-TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.
+\todo{Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.}
 
 \begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
-		\caption{Calculation times for PR on a single node}
+		\caption{Calculation time}
 		\label{fig:singleNodePR_calc}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
-		\caption{Execution times for PR on a single node}
+		\caption{Execution time}
 		\label{fig:singleNodePR_exec}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTimeNormalized.png}
-		\caption{Overhead time normalized by the graph size in million edges}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTime.png}
+		\caption{Overhead time}
 		\label{fig:singleNodePR_overheadNormalized}
@@
 
-	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing}
+	\caption{Average times for PR on a single computation node, black bars represent one standard deviation in our testing}
 \end{figure*}
@@
 \begin{figure*}
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
-		\caption{Calculation times for distributed PR}
+		\caption{Calculation time}
 		\label{fig:distributedPR_calc}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
-		\caption{Execution times for distributed PR}
+		\caption{Execution time}
 		\label{fig:distributedPR_exec}
@@
 	\hfil
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
@@
 	\end{subfigure}
-
-	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
+	\caption{Average times for PR on the distributed cluster, black bars represent one standard deviation in our testing}
 	\label{fig:distributedPR}

### 2020-09-14 documentation/paper/sections/results/sssp.tex
sssp dist table updated

@@
 
-\todo{Absolute Zeiten in \autoref{tbl:ssspexec} mit angeben}
 \begin{table}
-\renewcommand{\arraystretch}{1.3}
+\renewcommand{\arraystretch}{1.2}
+\tiny
 \centering
-\caption{Distributed SSSP Execution Times Relative to Galois Push}
-\begin{tabular}{crrrr}
+\caption{Distributed SSSP Execution Times and their Realation to Galois Push}
+\begin{tabular}{crrrrrrrr}
 \hline
-\bf{Data Set}&Galois Push&Galois Pull&Gemini&Giraph\\\hline
-flickr&1.0&0.97&1.02&17.46\\
-orkut&1.0&1.28&4.12&14.38\\
-wikipedia&1.0&1.88&5.26&7.11\\
-twitter&1.0&3.56&9.79&8.76\\
-rMat27&1.0&3.98&9.91&14.52\\
-friendster&1.0&2.32&9.72&7.59\\
-rMat28&1.0&5.70&11.07&16.5\\\hline
+\bf{Data Set}&
+\multicolumn{2}{c}{\bf Galois Push}&\multicolumn{2}{c}{\bf Galois Pull}&\multicolumn{2}{c}{\bf Gemini}&\multicolumn{2}{c}{\bf Giraph}\\
+&$\times$&(s)&$\times$&(s)&$\times$&(s)&$\times$&(s)\\\hline
+flickr & 1.0 & (2.4) & 0.97 & (2.4) & 1.02 & (2.5) & 17.46 & (42.6)\\
+orkut & 1.0 & (4.2) & 1.28 & (5.4) & 4.12 & (17.3) & 14.38 & (60.4)\\
+wikipedia & 1.0 & (14.2) & 1.88 & (26.6) & 5.26 & (74.5) & 7.11 & (100.8) \\
+twitter & 1.0 & (40.0) & 3.56 & (142.2) & 9.79 & (391.1) & 8.76 & (349.9)\\
+rMat27 & 1.0 & (39.0) & 3.98 & (154.9) & 9.91 & (386.0) & 14.52 & (565.8)\\
+friendster & 1.0 & (58.5) & 2.32 & (136.0) & 9.72 & (568.9) & 7.59 & (444.0)\\
+rMat28 & 1.0 & (71.5) & 5.7 & (407.9) & 11.07 & (792.0) & 16.5 & (1180.2)\\\hline
 \end{tabular}
@@
 \subsubsection{Single-Node vs. Distributed}
-TODO
+\todo{Vergleich}
 \begin{itemize}

### 2020-09-14 plots/makedata.py

@@ -172,19 +172,22 @@ for f in dist_frameworks_sssp:
     overheadSSSP_distributed.append(overhead)
     overheadSSSPNormalized_distributed.append(overheadNormalized)
 
-overheadSSSP_distributed_normalizedToGalois = []
+execTimeSSSP_distributed_normalizedToGalois = []
 
 #print("Overhead comparison distributed SSSP")
 i = 0
 for k in dist_frameworks_sssp:
-    overheadSSSP_distributed_normalizedToGalois.append([])
-    #print(dist_frameworks_sssp[k], end='')
-    for j in range(len(overheadSSSP_distributed[i])):
-        #print("&", round(execTimeSSSP_distributed[i][j]/execTimeSSSP_distributed[0][j] * 100) / 100, end='')
-        overheadSSSP_distributed_normalizedToGalois[i].append(overheadSSSP_distributed[i][j]/overheadSSSP_distributed[0][j]) 
+    execTimeSSSP_distributed_normalizedToGalois.append([])
+    for j in range(len(execTimeSSSP_distributed[i])):
+        execTimeSSSP_distributed_normalizedToGalois[i].append(execTimeSSSP_distributed[i][j]/execTimeSSSP_distributed[0][j]) 
     i += 1
-    #print("\\\\")
 
+for i in range(len(execTimeSSSP_distributed_normalizedToGalois[0])):
+    print(graphs[i], end=" & ")
+    for j in range(len(execTimeSSSP_distributed_normalizedToGalois)):
+        print(round(execTimeSSSP_distributed_normalizedToGalois[j][i] * 100)/100,end=" & (")
+        print(round(execTimeSSSP_distributed[j][i] * 10)/10 ,end=") & ")
+    print("\\\\")
 #print(graphs)
 
 
@@ -230,6 +233,17 @@ for f in frameworks:
     yErrExecBFS_singleNode.append(tmpYErrExec)
     overheadBFS_singleNode.append(overhead)
     overheadBFSNormalized_singleNode.append(overheadNormalized)
+
+if False:
+    k = 0
+    v = []
+    for f in frameworks:
+        print(f)
+        print(calcTimeBFS_singleNode[k])
+        print([round(calcTimeBFS_singleNode[k][i]/calcTimeBFS_singleNode[3][i] * 100) / 100 for i in range(len(calcTimeBFS_singleNode[k]))])
+        k += 1
+
+
 
 
 

### 2020-09-15 plots/makeplots.py
kleine fixes, normalisierung in sssp von overhead zu exec

@@ -154,7 +154,7 @@ if distributed and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 		grouped_bar_plot(graphs, list(dist_frameworks_sssp.values()), overheadSSSPNormalized_distributed, title='SSSP distributed', yLabel='Overhead time (s) (normalized)', saveToFile="distributedSSSP_overheadTimeNormalized.png", yScale='linear', colors_=colors)
 
-		grouped_bar_plot(graphs, list(dist_frameworks_sssp.values()), overheadSSSP_distributed_normalizedToGalois, yLabel='Galois Push overhead times', saveToFile="distributedSSSP_overheadTimeNormalizedToGalois.png", yScale='log', colors_=colors)
+		grouped_bar_plot(graphs, list(dist_frameworks_sssp.values()), execTimeSSSP_distributed_normalizedToGalois, yLabel='Galois Push execution times', saveToFile="distributedSSSP_executionTimeNormalizedToGalois.png", yScale='log', colors_=colors)
 
 	#BFS
 	if bfs or overheads:

### 2020-09-15 documentation/paper/main.tex
kleine fixes von der Rückmeldung und Polymer Overview
new references

@@
 \begin{abstract}
-TODO
+\todo{Section}
 \end{abstract}
@@
 \section{Introduction}
+\todo{Section}
 % In recent years graph sizes have increased significantly [?] .
@@
 \section{Discussion}
+\todo{Section}
 
 \section{Conclusion}
+\todo{Section}
 
@@
 \section*{Acknowledgments}
-We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Ligra}, Polymer~\cite{Polymer}, Gemini~\cite{Gemini} as well as Apache Giraph~\cite{Giraph}.
-
-Also we use Gluon~\cite{vertGalois} for the distributed Galois setups.
-
-Gemini~\cite{Gemini}
-
-%We would like to thank our supervisor Heiko Geppert for continued guidance and support.
+%We would like to thank our supervisor Heiko Geppert for his continued guidance and support.
 
 \appendices
+\todo{Beide zusammenfassen}
 \section{Installation Guides}\label{app:installationGuides}
@@
 
-% \bibliographystyle{IEEEtran}
-% \bibliography{literature}
+%\bibliographystyle{IEEEtran}
+%\bibliography{literature}
 
@@
 \BIBdecl
+
+\bibitem{newman2010networks}
+\BIBentryALTinterwordspacing
+M.~Newman, \emph{Networks: An Introduction}.\hskip 1em plus 0.5em minus
+  0.4em\relax OUP Oxford, 2010. [Online]. Available:
+  \url{https://books.google.de/books?id=-DgTDAAAQBAJ}
+\BIBentrySTDinterwordspacing
 
@@
 \BIBentryALTinterwordspacing
-S.~Brin and L.~Page, ``Reprint of: The anatomy of a large-scale hypertextual
-  web search engine,'' \emph{Computer Networks}, vol.~56, no.~18, pp. 3825 --
-  3833, 2012, the WEB we live in. [Online]. Available:
-  \url{http://www.sciencedirect.com/science/article/pii/S1389128612003611}
+S.~Brin and L.~Page, ``The anatomy of a large-scale hypertextual web search
+  engine,'' \emph{Computer Networks and ISDN Systems}, vol.~30, no.~1, pp. 107
+  -- 117, 1998, proceedings of the Seventh International World Wide Web
+  Conference. [Online]. Available:
+  \url{http://www.sciencedirect.com/science/article/pii/S016975529800110X}
+\BIBentrySTDinterwordspacing
+
+\bibitem{tao}
+\BIBentryALTinterwordspacing
+K.~Pingali, D.~Nguyen, M.~Kulkarni, M.~Burtscher, M.~A. Hassaan, R.~Kaleem,
+  T.-H. Lee, A.~Lenharth, R.~Manevich, M.~M\'{e}ndez-Lojo, D.~Prountzos, and
+  X.~Sui, ``The tao of parallelism in algorithms,'' in \emph{Proceedings of the
+  32nd ACM SIGPLAN Conference on Programming Language Design and
+  Implementation}, ser. PLDI '11.\hskip 1em plus 0.5em minus 0.4em\relax New
+  York, NY, USA: Association for Computing Machinery, 2011, p. 12–25.
+  [Online]. Available: \url{https://doi.org/10.1145/1993498.1993501}
 \BIBentrySTDinterwordspacing
@@
 
+\bibitem{hugepages}
+\BIBentryALTinterwordspacing
+A.~Panwar, A.~Prasad, and K.~Gopinath, ``Making huge pages actually useful,''
+  in \emph{Proceedings of the Twenty-Third International Conference on
+  Architectural Support for Programming Languages and Operating Systems}, ser.
+  ASPLOS '18.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA:
+  Association for Computing Machinery, 2018, p. 679–692. [Online]. Available:
+  \url{https://doi.org/10.1145/3173162.3173203}
+\BIBentrySTDinterwordspacing
+
+\bibitem{origGalois}
+\BIBentryALTinterwordspacing
+M.~Kulkarni, K.~Pingali, B.~Walter, G.~Ramanarayanan, K.~Bala, and L.~P. Chew,
+  ``Optimistic parallelism requires abstractions,'' in \emph{Proceedings of the
+  28th ACM SIGPLAN Conference on Programming Language Design and
+  Implementation}, ser. PLDI '07.\hskip 1em plus 0.5em minus 0.4em\relax New
+  York, NY, USA: Association for Computing Machinery, 2007, p. 211–222.
+  [Online]. Available: \url{https://doi.org/10.1145/1250734.1250759}
+\BIBentrySTDinterwordspacing
+
 \bibitem{Galois}
@@
 \BIBentryALTinterwordspacing
-Stanford~University. (2020, Sep.) Stanford network analysis project. [Online].
-  Available: \url{https://snap.stanford.edu}
+{Stanford University}. (2020, Sep.) Stanford network analysis project.
+  [Online]. Available: \url{https://snap.stanford.edu}
 \BIBentrySTDinterwordspacing
@@
 
-% Alles bis hier kommt aus main.bbl
-
 \end{document}

### 2020-09-15 documentation/paper/sections/overview/gemini.tex

@@
 Gemini tries to bridge the gap between efficient shared-memory and scalable distributed systems.
-To achieve this goal, Gemini, in contrast to the other NUMA-aware frameworks discussed here, does not support shared-memory calculation, but chooses the distributed message-based approach from scratch.
+To achieve this goal, Gemini, in contrast to the other frameworks discussed here, does not support shared-memory calculation, but chooses the distributed message-based approach from scratch.
 

### 2020-09-15 documentation/paper/sections/preliminiaries.tex

@@
 \subsection{Push and Pull Variants}
-Many parallel graph algorithms, as well as the three we consider here, are implemented by iterating over vertices, or edges \cite{Galois}.
-The respective vertex or edge, which is considered in an iteration, is also called \emph{active} vertex or edge.
+Many parallel graph algorithms, as well as the three we consider here, are implemented by iterating over vertices, or edges \cite{tao}.
+The respective vertex or edge, which is considered in an iteration, is also called \emph{active}.
 Applying an operator to this \emph{active} vertex or edge is called an \emph{activity}.
@@
 
-\subsection{NUMA-awareness}
-\todo{Section}
+\subsection{Hugepages}
+\todo{Section schreiben}
+\cite{hugepages}

### 2020-09-15 documentation/paper/sections/results/galoisSpeedup.tex
bfs single node results and images rescaled and moved to be more unified
kleine änderungen

@@
 
-\subsubsection{Single-source Shortest-path}
-\begin{figure}
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
@@
 		\label{fig:galoisSpeedupSSSP}
-\end{figure}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
+		\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search}
+		\label{fig:galoisSpeedupBFS}
+	\end{subfigure}
+	\hfil
+\end{figure*}
+
+\subsubsection{Single-source Shortest-path}
+
 Starting with SSSP, we see an algorithm that benefits from many available threads in \autoref{fig:galoisSpeedupSSSP}.
@@
 \subsubsection{Breadth-first search}
-\begin{figure}
-	\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
-	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search}
-	\label{fig:galoisSpeedupBFS}
-\end{figure}
 
@@
 
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
@@
 	\end{subfigure}
+	\hfil
 	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms.}
-\end{figure}
+\end{figure*}
 
-\section{Hugepages}
+\subsubsection{Comparsion Hugepages}
+\todo{Section fehlt noch ganz, Bilder sind noch zu groß aber fertig.}
 \begin{figure}

### 2020-09-15 documentation/paper/sections/preliminiaries.tex
weitere kleine Änderungen

@@
 
-\emph{Pull-style} operators need less synchronization in parallel implementations, because unlike \emph{pull style} there is only one write and many read operations.
+\emph{Pull-style} operators need less synchronization in parallel implementations, because unlike \emph{push style} there is only one write and many read operations.
 Thus locks can be avoided.

### 2020-09-16 plots/makedata.py
updated plots to make them more easily comparable

@@ -116,16 +116,17 @@ if False:
             s[i] += calcTimeSSSP_singleNode[k][i]
         k += 1
     s /= 4
-    print("average calculation time:", s)
+    #print("average calculation time:", s)
     k = 0
     for f in frameworks:
         # if f == 'Giraph':
         #     k += 1
         #     continue
-        print(f)
+        #print(f)
         for i in range(1, len(execTimeSSSP_singleNode[k])):
-            print(graphs[i], round((execTimeSSSP_singleNode[k][i] / execTimeSSSP_singleNode[0][i]) * 100), "%")
-            print(graphs[i], round(execTimeSSSP_singleNode[k][i] * 100) / 100, "s")
+            pass
+            #print(graphs[i], round((execTimeSSSP_singleNode[k][i] / execTimeSSSP_singleNode[0][i]) * 100), "%")
+            #print(graphs[i], round(execTimeSSSP_singleNode[k][i] * 100) / 100, "s")
         k += 1
 
 ## SSSP DISTRIBUTED
@@ -182,12 +183,13 @@ for k in dist_frameworks_sssp:
         execTimeSSSP_distributed_normalizedToGalois[i].append(execTimeSSSP_distributed[i][j]/execTimeSSSP_distributed[0][j]) 
     i += 1
 
-for i in range(len(execTimeSSSP_distributed_normalizedToGalois[0])):
-    print(graphs[i], end=" & ")
-    for j in range(len(execTimeSSSP_distributed_normalizedToGalois)):
-        print(round(execTimeSSSP_distributed_normalizedToGalois[j][i] * 100)/100,end=" & (")
-        print(round(execTimeSSSP_distributed[j][i] * 10)/10 ,end=") & ")
-    print("\\\\")
+if False:
+    for i in range(len(execTimeSSSP_distributed_normalizedToGalois[0])):
+        print(graphs[i], end=" & ")
+        for j in range(len(execTimeSSSP_distributed_normalizedToGalois)):
+            print(round(execTimeSSSP_distributed_normalizedToGalois[j][i] * 100)/100,end=" & (")
+            print(round(execTimeSSSP_distributed[j][i] * 10)/10 ,end=") & ")
+        print("\\\\")
 #print(graphs)
 
 
@@ -469,6 +471,7 @@ for i in range(len(graph)):
         continue
     
     num = int(re.sub('\D', '', algo[i]))
+
[…]

### 2020-09-16 plots/makeplots.py

@@ -4,19 +4,19 @@ from plotFunctions import *
 
 
 
-distributed = True
-singleNode = True
+distributed = False
+singleNode = False
 
-sssp = True
-bfs = True
-pr = True
+sssp = False
+bfs = False
+pr = False
 
 calcTimes = False
 execTimes = False
 overheads = False
 
 
-galoisSpeedup = False
+galoisSpeedup = True
 
 
 colors = {"Galois":"#1f77b4",
@@ -89,28 +89,24 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 ## GALOIS THREAD COUNT
 if galoisSpeedup:
-	line_plot(graphs, x, speedupGaloisSSSP, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads.png")
+	line_plot(graphs, x, speedupGaloisSSSP, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads.png", ylim=[0,20])
 
-	line_plot(graphs, x[:-5], speedupGaloisSSSP, title='Speedup SSSP', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads_short.png")
+	line_plot(graphs, x, speedupGaloisBFS, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeBFSGaloisThreads.png", ylim=[0,11])
 
-	line_plot(graphs, x, speedupGaloisBFS, title='SpeedupBFS', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeBFSGaloisThreads.png")
+	line_plot(graphs, x, speedupGaloisPRPush, title='Speedup PR Push', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPushGaloisThreads.png", ylim=[0.85,2.5])
 
-	line_plot(graphs, x, speedupGaloisPRPush, title='Speedup PR Push', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPushGaloisThreads.png")
-
[…]

### 2020-09-16 plots/plotFunctions.py

@@ -54,7 +54,7 @@ def grouped_bar_plot(group_names, legend_names, values, colors_=None, yErrs=None
 
 #
 # values is a dictionary
-def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=None, saveToFile=None):
+def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=None, saveToFile=None, ylim=None):
 
 	if len(legend_names) != len(values):
 		raise Exception("Size mismatch: Expected " + str(len(legend_names)) + " categories, got " + str(len(values)))
@@ -76,6 +76,8 @@ def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=No
 	ax.set_title(title)
 	ax.set_ylabel(yLabel)
 	ax.set_xlabel(xLabel)
+	if ylim != None:
+		ax.set_ylim(ylim)
 
 	
 	ax.legend(ps, legend_names)

### 2020-09-16 documentation/paper/sections/testingMethods.tex
weitere kleine Änderungen
used algorithms update

@@
 \section{Evaluation}
-Einleitung
+\todo{Einleitung}
 
@@
 \begin{itemize}
-	\item Galois: CPU, Push, Pull für PR und SSSP..
-	\item Gemini:
+	\item Galois: Galois supports all of our tested algorithms, with both a Push and a Pull variant for PageRank available.
+In the distributed scenario, there are Push and Pull versions for SSSP and BFS available as well.
+	\item Gemini: \todo{}
 	\item Giraph: Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
-	\item Ligra:
-	\item Polymer: Polymer supports two implementations of Pagerank
+For SSSP, slight variations had to be made to the default implementation, to allow us to use different start vertices.
+For PageRank the supplied implementation is used.
+	\item Ligra: \todo{}
+	\item Polymer: Polymer suppoerts SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
 \end{itemize}

### 2020-09-16 documentation/paper/sections/testingMethods.tex
used algorithms update

@@
 \begin{itemize}
-	\item Galois: Galois supports all of our tested algorithms, with both a Push and a Pull variant for PageRank available.
+	\item Galois supports all of our tested algorithms, with both a Push and a Pull variant for PageRank available.
 In the distributed scenario, there are Push and Pull versions for SSSP and BFS available as well.
-	\item Gemini: \todo{}
-	\item Giraph: Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
+	\item Gemini \todo{}
+	\item Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
 For SSSP, slight variations had to be made to the default implementation, to allow us to use different start vertices.
 For PageRank the supplied implementation is used.
-	\item Ligra: \todo{}
-	\item Polymer: Polymer suppoerts SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
+	\item Ligra \todo{}
+	\item Polymer supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
 \end{itemize}
-
-% \autoref{tbl:algorithms} shows the tested algorithms in detail.
-
-% \newcommand{\txtdag}{\textsuperscript{\textdagger}}
-% \newcommand{\singleNode}{\textsuperscript{s}}
-% \newcommand{\distributed}{\textsuperscript{d}}
-% \begin{table}
-% 	\renewcommand{\arraystretch}{1.3}
-% 	\centering
-% 	\caption{REMOVE THIS // Tested Algorithms of Each Framework}
-% 	\begin{tabularx}{0.9\columnwidth}{cYYYYY}
-% 		\bf{Algorithm}&Galois&Gemini&Giraph&Ligra&Polymer\\\hline
-% 		PageRank&Yes\singleNode{}, $\Delta$Push\distributed{}, $\Delta$Pull\distributed{}&&Yes\singleNode\distributed&&Regular\singleNode, $\Delta$\singleNode\\\hline
-% 		SSSP&?, Push, Pull&&Yes\singleNode\distributed\txtdag&&Yes\singleNode{}\\\hline
-% 		BFS&Yes&&Yes\singleNode\distributed{}*&&\\
-% 		\hline
-% 		\multicolumn{6}{p{0.5\columnwidth}}{* Algorithm not natively supported
-
-% 		\txtdag{} Modified algorithm
-
-% 		\distributed{} distributed
-
-% 		\singleNode{} single-node}
-% 	\end{tabularx}
-% 	\label{tbl:algorithms}
-
-% \end{table}
-
-\subsection{Comparison of setup}

### 2020-09-16 documentation/paper/sections/results/bfs.tex
sssp und bfs nochmal überarbeitet, das fasse ich hoffentlich nicht mehr an

@@
 \subsection{Breadth-first search}
-As with our results for SSSP, for BFS we will begin with our single-node results before looking at the distributed scenario. Lastly we will compare the two cases.
+As with our results for SSSP, for BFS we will begin with our single-node results before looking at the distributed scenario.
 \subsubsection{Single-Node}
@@
 Their computation times deviate less than 151ms on all graphs except wikipedia and rMat28, wight Ligra being the faster framework on most graphs.
-Ligra is 39ms faster on orkut, 2ms on twitter, 50ms on rMat27 and 151ms on friendster.
-In turn, Gemini is 17 ms faster than Ligra on flickr.
+Ligra is between 2ms (twitter) and 151ms (friendster) faster than Gemini.
+In turn, Gemini is 17ms faster than Ligra on flickr.
 Only on wikipedia and rMat28, there is a noticeable difference between the two frameworks. Ligra is 1.2s (7.8$\times$) faster than Gemini on wikipedia and 696ms (4.4$\times$) faster on rMat28.
@@
 Giraph and Polymer are one to two orders of magnitude slower than both Ligra or Gemini. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
-Giraph takes 13$\times$, 32$\times$ and 27$\times$ longer than Ligra on flickr, orkut and wikipedia respectively.
-As we said, Polymer is comparable to Ligra on flickr, for the other graphs however, its computation times are even longer than those of Giraph.
+Giraph takes between 13$\times$ and 32$\times$ longer than Ligra on flickr, orkut or wikipedia.
+As we said, Polymer is comparable to Ligra on flickr, but for the other graphs, its computation times are even longer than those of Giraph.
 The upside is though, that Polymer managed -- contrary to Giraph -- to finish computation on some larger graphs.
-Polymer takes 55$\times$ longer on orkut, 60$\times$ on wikipedia, 129$\times$ on twitter, 530$\times$ on rMat27 and 240$\times$ on friendster. 
+Polymer takes between 55$\times$ and 530$\times$ longer than Ligra on the graphs larger than flickr.
 Especially interesting is here the difference between the times of rMat27 and friendster.
-There are 20\% (439M) edges more in friendster, yet the computation for the synthetic rMat27 takes 42\% longer.
+There are 20\% (439M) more edges in friendster, yet the computation for the synthetic rMat27 takes 42\% longer.
 
 Galois calculation performance is comparable to Gemini or Ligra on the smaller three graphs (flickr, orkut and wikipedia).
-On the larger graphs however, Galois is slower than Gemini or Ligra but always faster than Polymer by one order of magnitude.
+Only on the larger graphs is Galois slower than Gemini or Ligra, meanwhile Galois is always faster than Polymer by one order of magnitude.
 
@@
 	% \end{subfigure}
+	% \hfil
 	\caption{Average times for BFS on the distributed cluster, black bars represent one standard deviation in our testing}
@@
 This is expected since both are graph traversal algorithms starting in one source vertex.
-Hence, calculation complexity for each vertex and communication overhead is similar.
-All measurements can be seen in \autoref{fig:distributedBFS}.
+Hence, calculation complexity for each vertex and communication overhead is similar. All measurements can be seen in \autoref{fig:distributedBFS}.
 
-First, the calculation times in \autoref{fig:distributedBFS_calc}.
-It shows Giraph having the shortest calculation times on the real-world graphs, while Giraph's calculatin times on both rMat27 and rMat28 are the worst of all frameworks.
+The results for calculation time (cf.
+\autoref{fig:distributedBFS_calc}) show Giraph to have very short calculation times on the real-world graphs, while Giraph's calculation times on both rMat27 and rMat28 are the worst of all frameworks.
+Thus, Giraph is fastest on the three smallest graphs and second fastest on twitter and friendster.
+On those two graphs, Gemini is fastest with only a small margin between the two frameworks.
+Just like with SSSP, the Galois implementations have the longest calculation time on the real-world graphs. And Galois is second slowest, on the synthetic graphs.
 
-Comparing the execution times in \autoref{fig:distributedBFS_exec} results in the same findings as with SSSP.
+Comparing the execution times in \autoref{fig:distributedBFS_exec} results again in similar findings to SSSP.
 While Gemini can compete with Galois on the small flickr graph, moving to larger data sets shows the worse performance of Gemini compared to Galois.
-
-Much like when running SSSP, Giraph is slowest on all but one graph. Only on friendster is Gemini marginally slower, which was also the case for SSSP.
-
-Both Galois implementations are again similar to the behaviour on SSSP.
+Similar to SSSP, Giraph is slowest on all but one graph. Only on friendster is Gemini marginally slower, this was also the case for SSSP.
 Galois Push is generally faster than the Pull alternative while both Push and Pull versions are faster than Gemini and Giraph across all graphs.

### 2020-09-16 documentation/paper/sections/results/sssp.tex
kleine fixes, normalisierung in sssp von overhead zu exec
kleine änderungen
sssp und bfs nochmal überarbeitet, das fasse ich hoffentlich nicht mehr an

@@
 There, the average calculation times of the four frameworks (Galois, Gemini, Ligra, Polymer) are 35ms on flickr, 77ms on orkut, 538ms on wikipedia.
-The four frameworks are always within $x$ ms of this average. Galois and Gemini being 24ms and 12ms faster than the average. Ligra and Polymer are slower by 27ms and 9ms.
+The four frameworks are close to this average for the smaller graphs. Galois and Gemini being 24ms and 12ms faster than the average. Ligra and Polymer are slower by 27ms and 9ms.
 Orkut is the graph where all the four frameworks are within 20ms of their average, Galois being the outlier at 19ms slower than average.
@@
 Galois takes anywhere from 3.3$\times$ (twitter) to 4.2$\times$ (friendster) the computation time of Polymer.
-Ligra requires 9.8$\times$, 7.0$\times$ and 3.2$\times$ the computation time of Polymer on twitter, rMat27 and friendster respectively.
-This makes Ligra the slowest Framework on twitter and rMat27.
+Ligra requires between 3.2$\times$ (friendster) and 9.8$\times$ (twitter) the computation time of Polymer on the larger graphs.
+This makes Ligra the slowest Framework on both twitter and rMat27.
 Galois is the slowest on friendster.
@@
 
-Moving on to the execution times, which can be seen in \autoref{fig:singleNodeBFS_exec}. 
-A general overview shows Galois as the clear winner on most graphs. It has the smalles computation times on all graphs except flickr, where Galois is 25ms (8\%) slower than Ligra. 
+The execution times show Galois to be the clear winner on most graphs (cf. \autoref{fig:singleNodeBFS_exec}).
+It has the smallest computation times on all graphs except flickr, where Galois is 25ms (8\%) slower than Ligra. 
 On the other 5 graphs however, Galois has the smaller execution time by one or sometimes even two orders of magnitude. 
-For example, on orkut Galois requires 0.78s to execute, while Gemini takes 13.8s, Giraph takes 135.9s, Ligra takes 2.66s and Polymer takes 26.6s.
-This goes on for the larger graphs, Ligra being second-fastest and Polymer and Gemini close together but both always at least one order of magnitude slower than Galois.
+For example, on orkut Galois requires 0.78s to execute, while the other frameworks require between 2.66s (Ligra) and 26.6s (Polymer).
+This goes on for the larger graphs, Ligra being second-fastest. Polymer and Gemini close together but both always at least one order of magnitude slower than Galois.
 
@@
 \end{figure*}
-
-Moving on to the distributed cluster, \autoref{fig:distributedSSSP} shows the benchmark results as calculation and execution times.
-
-Comparing the two Galois implementations, we find the calculation and execution times to be similar on smaller graphs and Push being the superior implementation for SSSP on larger data sets.
-Galois Pull being anywhere from just as fast to 3.5$\times$ slower on real-world data sets compared to the Push variant.
-The synthetic graphs are more extreme, execution times are close to 4$\times$ (rMat27) and 5$\times$ (rMat28) longer.
-
+On the distributed cluster, we find similar results as on the single node (cf. \autoref{fig:distributedSSSP}).
 Both Galois implementations have significantly smaller execution times compared to Gemini or Giraph on all graphs (cf. \autoref{tbl:ssspexec}).
@@
 Giraph's execution times in comparison to this are even worse, taking at least 7$\times$ longer than Galois Push on all graphs.
-
+Comparing the two Galois implementations, we find the calculation and execution times to be similar on smaller graphs and Push being the superior implementation for SSSP on larger data sets.
+Galois Pull is anywhere from just as fast to 3.5$\times$ slower on real-world data sets compared to the Push variant.
+The synthetic graphs are more extreme.
+Execution times are close to 4$\times$ (rMat27) and 5$\times$ (rMat28) longer on Pull.
 Evidently, Galois Push is the fastest algorithm in our lineup on 6 out of 7 graphs. With the exception being flickr, where Galois Push takes negligibly longer than the Pull counterpart.
@@
 \centering
-\caption{Distributed SSSP Execution Times and their Realation to Galois Push}
+\caption{Distributed SSSP Execution Times and Their Realation to Galois Push}
 \begin{tabular}{crrrrrrrr}
@@
 On both synthetic graphs however, Giraph is actually the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.
-
 While Giraph's computation times are very competitive, when comparing the execution times in \autoref{fig:distributedSSSP_exec} we see that Giraph is actually the slowest framework on 5 out of 7 graphs.
 For the other two, namely twitter and friendster, Giraph is second slowest with only Gemini taking longer to complete.
-
 Giraph and Gemini's very long execution times are only due to their overhead being many orders of magnitude larger than Galois overhead (\autoref{fig:distributedSSSP_overhead}).
@@
 On friendster, where Gemini was worst, Giraph \emph{only} requires 73$\times$ the overhead time of Galois.
-
-\subsubsection{Single-Node vs. Distributed}
-\todo{Vergleich}
-\begin{itemize}
-	\item GaloisPush vs Galois CPU // Galois Pull verliert bei Dist.
-	\item Giraph vs. Giraph
-	\item Gemini vs Gemini
-\end{itemize}

### 2020-09-16 documentation/paper/sections/results/galoisSpeedup.tex
hauptsächlich pr single node auswertung

--- 2020-09-06 documentation/paper/sections/results/galoisSpeedup.tex
@@
 \label{sec:galois_speedup}
-Analyzing the calculation time speedups for Galois, we can compare how or if the different algorithms benefit from increasing thread numbers.
+This section is dedicated to analyzing the speedup behaviour of Galois under the change to two parameters.
+First we change the thread count Galois is using.
+Second, a comparison between Galois using hugepages and Galois without hugepages is made.
+We compare the \emph{calculation speedups} of Galois on the different graphs.
+Thus, we show the calculation time on any thread count normalized by the calculation time in the single-threaded environment.
+Beginning with SSSP, followed by BFS and last the comparison for both PR Push and Pull.
 
-\subsubsection{Single-source Shortest-path}
-\begin{figure}
+\subsubsection{Single-Source Shortest-Paths}
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
-	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths}
+		\caption{without hugepages}
+		\label{fig:galoisSpeedupSSSP_noHP}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+		\caption{with hugepages}
+		\label{fig:galoisSpeedupSSSP_HP}
+	\end{subfigure}
+	\hfil
+	\caption{Calculation time speedups on SSSP for the different graphs with and without hugepages.}
 	\label{fig:galoisSpeedupSSSP}
-\end{figure}
-Starting with SSSP, we see an algorithm that benefits from many available threads in \autoref{fig:galoisSpeedupSSSP}.
+\end{figure*}
+Starting with SSSP, we see an algorithm that benefits from many available threads (cf. \autoref{fig:galoisSpeedupSSSP}).
+We first look at the speedups without hugepages, seen in \autoref{fig:galoisSpeedupSSSP_noHP}.
 For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads.
@@
 
-\subsubsection{Breadth-first search}
-\begin{figure}
+Upon activating the hugepages, we acquired the results seen in \autoref{fig:galoisSpeedupSSSP_HP}. Here, the overall results are similar to the findings without hugepages.
+
+\todo{Hugepages}
+
+\subsubsection{Breadth-First Search}
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
-	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search}
+		\caption{without hugepages}
+		\label{fig:galoisSpeedupBFS_noHP}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisHPThreads.png}
+		\caption{with hugepages}
+		\label{fig:galoisSpeedupBFS_HP}
+	\end{subfigure}
+	\hfil
+	\caption{Calculation time speedups on BFS for the different graphs with and without hugepages.}
 	\label{fig:galoisSpeedupBFS}
-\end{figure}
+\end{figure*}
 
@@
 \subsubsection{PageRank}
+
 We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.
@@
 
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
+\todo{unfertig}
+We assume the reason for this behaviour to be founded in the hugepages. Galois is recommended to be run with hugepages.
+
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
-		\caption{PageRank Pull}
+		\caption{without hugepages}
+		\label{fig:galoisSpeedupPRPull_noHP}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
+		\caption{with hugepages}
+		\label{fig:galoisSpeedupPRPull_HP}
+	\end{subfigure}
+	\hfil
+	\caption{Calculation time speedups on PR Pull for the different graphs with and without hugepages.}
 	\label{fig:galoisSpeedupPRPull}
+\end{figure*}
+
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
+		\caption{without hugepages}
+		\label{fig:galoisSpeedupPRPush_noHP}
 	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
-		\caption{PageRank Push}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisHPThreads.png}
+		\caption{with hugepages}
+		\label{fig:galoisSpeedupPRPush_HP}
+	\end{subfigure}
+	\hfil
+	\caption{Calculation time speedups on PR Pushgi for the different graphs with and without hugepages.}
 	\label{fig:galoisSpeedupPRPush}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms.}
-\end{figure}
-
-\section{Hugepages}
-\begin{figure}
-	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
-	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths with Hugepages}
-	\label{fig:galoisHPSpeedupSSSP}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisHPThreads.png}
-	\caption{Calculation time speedup with increasing thread count for Galois Breadth-first search with Hugepages}
-	\label{fig:galoisHPSpeedupBFS}
-\end{figure}
-
-\begin{figure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
-		\caption{PageRank Pull}
-		\label{fig:galoisHPSpeedupPRPull}
-	\end{subfigure}
-	\begin{subfigure}{\columnwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisHPThreads.png}
-		\caption{PageRank Push}
-		\label{fig:galoisHPSpeedupPRPush}
-	\end{subfigure}
-	\caption{Calculation time speedup with increasing thread count for Galois PageRank Push and Pull algorithms using Hugepages.}
-\end{figure}
+\end{figure*}

### 2020-09-16 documentation/paper/sections/results/pr.tex
kleine änderungen
hauptsächlich pr single node auswertung

@@
 \subsection{PageRank}
+
 \subsubsection{Single-Node}
-
 Ligra and Polymer support both regular and Delta-PageRank variants.
@@
 If the regular version is slower than delta, that is only by a small difference.
-Explicitly, regular is slower than delta by 19\%\ on twitter, 17\%\ on rMat27 or 6\%\ for friendster.
-For the other graphs, the delta version is slower by a far greater margin, 50\%\ on flickr, 45\%\ on orkut, 13\%\ on wikipedia and 68\% on rMat28.
+Explicitly, regular is slower than delta by a range of 6\% to 19\%\ on twitter, rMat27 or friendster.
+For the other graphs, the delta version is slower by a far greater margin of 13\% to 68\%.
 Hence, we only show the results of Ligra's regular PageRank implementation in our evaluation.
@@
 Thus, the following only shows Polymer's faster Delta-PR implementation.
-
-\todo{Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.}
-
+Giraph required more than the available 250 GB of RAM on any graph larger than wikipedia, hence all of Giraph's results for the larger graphs are missing here.
 \begin{figure*}
+	\hfil
 	\begin{subfigure}{0.32\textwidth}
@@
 	\hfil
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTime.png}
-		\caption{Overhead time}
-		\label{fig:singleNodePR_overheadNormalized}
-	\end{subfigure}
+	\caption{Average times for PR on a single computation node, black bars represent one standard deviation in our testing}
+	\label{fig:singleNodePR}
+\end{figure*}
 
-	\caption{Average times for PR on a single computation node, black bars represent one standard deviation in our testing}
-\end{figure*}
+The calculation times show some odd behaviour of Galois Push.
+The required time is less than 1ms, regardless of the graph (cf.
+\autoref{fig:singleNodePR_calc}).
+Meanwhile there was no output produced, that would indicate any kind of error.
+These results would make the calculation times of Galois Push the smallest on all graphs, with a difference of at least one order of magnitude.
+However, we are very suspicious of these results and thus exclude the calculation time for Galois Push in further comparisons.
+Because the execution time of Galois Push is always considerably longer than the execution time of Galois Pull (cf.
+\autoref{fig:singleNodePR_calc}).
+This leads us to believe that the output that we used for our measurements contains an error.
+
+For the three graphs on which Giraph computed successfully, it is the slowest framework in both calculation and execution times.
+And that by a difference of three orders of magnitude in the calculation time and one to two orders of magnitude in execution time (cf \autoref{fig:singleNodePR}).
+On the larger graphs (i.e.
+those, where there is no data for Giraph), Gemini and Ligra are always slowest in execution time.
+Contrary to this, Galois Pull has the smallest execution times on all graphs except flickr (cf.
+\autoref{fig:singleNodePR_exec}).
+Ligra is fastest on flickr, while being second fastest on wikipedia, rMat27 and rMat28.
+Galois Push is second fastest on orkut and friendster.
+Interestingly, the execution time for Ligra is at a maximum for twitter. The required time is steadily decreasing with increasing graph size.
 
@@
 The \autoref{fig:distributedPR} shows our results of PageRank on the distributed cluster.
-First of all, Giraph was unable to complete the test because it required more than 250GB of RAM for rMat28. Thus this results is missing.
+First of all, Giraph was unable to complete the test because it required more than 250GB of RAM for rMat28, thus this result is missing.
 When comparing the calculation times in \autoref{fig:distributedPR_calc} to the execution times in \autoref{fig:distributedPR_exec}, we see similar behaviour of all frameworks.
@@
 More specifically, there are no overhead outliers like it was the case with Giraph on SSSP.
-
 \begin{figure*}
+	\hfil
 	\begin{subfigure}{0.32\textwidth}
@@
 	\hfil
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
-		\caption{Overhead time}
-		\label{fig:distributedPR_overheadNormalized}
-	\end{subfigure}
 	\caption{Average times for PR on the distributed cluster, black bars represent one standard deviation in our testing}
@@
 \end{figure*}
+
+\todo{Auswertung fehlt noch.}

### 2020-09-16 documentation/paper/sections/results/galoisSpeedup.tex
speedup sssp

@@
 	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
@@
 	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
@@
 \end{figure*}
-Starting with SSSP, we see an algorithm that benefits from many available threads (cf. \autoref{fig:galoisSpeedupSSSP}).
+Starting with SSSP, we see an algorithm that benefits from many available threads this can be seen in \autoref{fig:galoisSpeedupSSSP}.
 We first look at the speedups without hugepages, seen in \autoref{fig:galoisSpeedupSSSP_noHP}.
@@
 For 40 threads, the speedup is nearly 19$\times$, on 48 threads 17$\times$ and with 96 threads only 15$\times$ compared to one thread.
-
 Small graphs, i.e. flickr and orkut neither benefit from more threads nor is the performance significantly held up by synchronization overhead.
@@
 
-Upon activating the hugepages, we acquired the results seen in \autoref{fig:galoisSpeedupSSSP_HP}. Here, the overall results are similar to the findings without hugepages.
+Upon activating the hugepages, we acquired the results seen in \autoref{fig:galoisSpeedupSSSP_HP}.
+Here, the overall results are similar to the findings without hugepages.
+\autoref{tbl:ssspMeansVariances} shows the mean speedups and variances over the different graphs.
+We see that the mean speedup is either very similar or slightly reduced by the hugepages.
+But in all cases, the variance is significantly smaller when using hugepages.
+This proves a slightly smaller but more reliable speedup with hugepages.
+Examples for this are on the one hand, orkut that could not reach a speedup above 1.6$\times$.
+Now, with hugepages it is 4.3$\times$ faster with 96 threads compared to one thread.
+On the other hand, twitter reached a speedup of 17$\times$ without hugepages and only 13$\times$ with.
 
-\todo{Hugepages}
+\begin{table}
+\renewcommand{\arraystretch}{1.2}
+%\scriptsize
+\centering
+\caption{Mean Speedups and Variances for SSSP With and Without HugePages}
+\label{tbl:ssspMeansVariances}
+\begin{tabular}{ccccccccc}
+\hline
+{\bf HP}&\multicolumn{2}{c}{\bf 1}&\multicolumn{2}{c}{\bf 2}&\multicolumn{2}{c}{\bf 4}&\multicolumn{2}{c}{\bf 8}\\
+&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
+w/o&1.0&0.0&1.6&0.2&2.5&1.1&4.5&5.3\\
+w/&1.0&0.0&1.6&0.1&2.5&0.9&3.4&2.1\\
+\hline
+\hline
+{\bf HP}&\multicolumn{2}{c}{\bf 16}&\multicolumn{2}{c}{\bf 32}&\multicolumn{2}{c}{\bf 48}&\multicolumn{2}{c}{\bf 96}\\
+&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
+w/o&6.7&13.0&9.6&38.3&10.3&41.4&10.2&38.1\\
+w/&5.8&7.3&7.0&10.8&10.7&31.4&10.8&29.9\\\hline
+\end{tabular}
+\end{table}
 
@@
 	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
@@
 	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisHPThreads.png}
@@
 \end{figure*}
-
 For our speedup results on BFS, \autoref{fig:galoisSpeedupBFS} shows the calculation time speedup of Galois' BFS.
@@
 	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
@@
 	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
@@
 	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
@@
 	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.4\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisHPThreads.png}

### 2020-09-16 plots/makedata.py

@@ -388,6 +388,8 @@ if False:
     print(polymer)
     print(sum(polymer[:-1]) / len(polymer[:-1])) 
 
+if False:
+    print(calcTimePR_singleNode[0])
     
 
 
@@ -609,3 +611,36 @@ for g in graphs:
         speedupGaloisPRPush_HP[g][x_] = tPRPush / speedupGaloisPRPush_HP[g][x_]
         speedupGaloisPRPull_HP[g][x_] = tPRPull / speedupGaloisPRPull_HP[g][x_]
 
+
+if False:
+    for xVal in xHP[:-4]:
+        print(xVal,"&",xVal,end="&")
+    print("\\\\")
+    for xVal in xHP[:-4]:
+        mean = 0
+        for k in graphs:
+            mean += speedupGaloisSSSP[k][xVal]
+        mean /= len(graphs)
+        print(round(mean * 10)/ 10, end="&")
+
+        variance = 0
+        for k in graphs:
+            variance += (speedupGaloisSSSP[k][xVal] - mean)**2
+        variance /= len(graphs)
+        print("(", round(variance * 10)/ 10, ")", end="&")
+
+    print("\\\\")
+    for xVal in xHP[:-4]:
+        meanHP = 0
+        for k in graphs:
+            meanHP += speedupGaloisSSSP_HP[k][xVal]
+        meanHP /= len(graphs)
+        print(round(meanHP * 10)/ 10, end="&")
+
+        variance = 0
+        for k in graphs:
+            variance += (speedupGaloisSSSP_HP[k][xVal] - meanHP)**2
+        variance /= len(graphs)
+        print("(", round(variance * 10)/ 10, ")", end="&")
+
+    print("\\\\")

### 2020-09-16 documentation/paper/sections/results/sssp.tex
table styling

@@
 \begin{tabular}{crrrrrrrr}
-\hline
+\toprule
 \bf{Data Set}&
 \multicolumn{2}{c}{\bf Galois Push}&\multicolumn{2}{c}{\bf Galois Pull}&\multicolumn{2}{c}{\bf Gemini}&\multicolumn{2}{c}{\bf Giraph}\\
-&$\times$&(s)&$\times$&(s)&$\times$&(s)&$\times$&(s)\\\hline
+&$\times$&(s)&$\times$&(s)&$\times$&(s)&$\times$&(s)\\\midrule
 flickr & 1.0 & (2.4) & 0.97 & (2.4) & 1.02 & (2.5) & 17.46 & (42.6)\\
@@
 friendster & 1.0 & (58.5) & 2.32 & (136.0) & 9.72 & (568.9) & 7.59 & (444.0)\\
-rMat28 & 1.0 & (71.5) & 5.7 & (407.9) & 11.07 & (792.0) & 16.5 & (1180.2)\\\hline
+rMat28 & 1.0 & (71.5) & 5.7 & (407.9) & 11.07 & (792.0) & 16.5 & (1180.2)\\\bottomrule
 \end{tabular}

### 2020-09-16 documentation/paper/sections/results/galoisSpeedup.tex
speedup sssp and bfs finished

@@
 \begin{tabular}{ccccccccc}
-\hline
-{\bf HP}&\multicolumn{2}{c}{\bf 1}&\multicolumn{2}{c}{\bf 2}&\multicolumn{2}{c}{\bf 4}&\multicolumn{2}{c}{\bf 8}\\
-&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
+\toprule
+&\multicolumn{8}{c}{\bf Thread count}\\
+\cmidrule(r){2-9}
+&\multicolumn{2}{c}{\bf 1}&\multicolumn{2}{c}{\bf 2}&\multicolumn{2}{c}{\bf 4}&\multicolumn{2}{c}{\bf 8}\\
+{\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
 w/o&1.0&0.0&1.6&0.2&2.5&1.1&4.5&5.3\\
 w/&1.0&0.0&1.6&0.1&2.5&0.9&3.4&2.1\\
-\hline
-\hline
-{\bf HP}&\multicolumn{2}{c}{\bf 16}&\multicolumn{2}{c}{\bf 32}&\multicolumn{2}{c}{\bf 48}&\multicolumn{2}{c}{\bf 96}\\
-&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
+\bottomrule
+\toprule
+&\multicolumn{8}{c}{\bf Thread count}\\
+\cmidrule(r){2-9}
+&\multicolumn{2}{c}{\bf 16}&\multicolumn{2}{c}{\bf 32}&\multicolumn{2}{c}{\bf 48}&\multicolumn{2}{c}{\bf 96}\\
+{\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
 w/o&6.7&13.0&9.6&38.3&10.3&41.4&10.2&38.1\\
-w/&5.8&7.3&7.0&10.8&10.7&31.4&10.8&29.9\\\hline
+w/&5.8&7.3&7.0&10.8&10.7&31.4&10.8&29.9\\\bottomrule
+\multicolumn{9}{c}{* HugePages}
 \end{tabular}
@@
 \end{figure*}
-For our speedup results on BFS, \autoref{fig:galoisSpeedupBFS} shows the calculation time speedup of Galois' BFS.
-On all graphs, the speedup never exceeds 6$\times$ even when using 96 threads.
-For the smaller graphs (flickr, orkut), we have the same behaviour as on SSSP. Speedup is close to 1 in all cases, with orkut reaching a maximum speedup of 1.6$\times$ at 24 threads.
-On the larger graphs, speedup is possible but only to a very small degree.
-Wikipedia has a maximum speedup of 4.9$\times$ at 96 threads, for twitter it is 2.5$\times$ at 96 threads and 5.6$\times$ on friendster.
-The two synthetic data sets reach speedups of 4.2$\times$ and 4.7$\times$ for rMat27 and rMat28 on 96 threads.
+For our speedup results on BFS, \autoref{fig:galoisSpeedupBFS} shows the calculation time speedup of Galois' BFS with and without hugepages.
+If we look at the results without hugepages first, we see most significantly, that the speedup never exceeds 6$\times$ even when using 96 threads (cf. \autoref{fig:galoisSpeedupBFS_noHP}).
+For the smaller graphs (flickr, orkut), we have the same behaviour as on SSSP without hugepages. Speedup is close to 1 in all cases, with orkut reaching a maximum speedup of 1.6$\times$ at 24 threads.
+That said, the larger graphs are not benefitting from more threads as much as they did with SSSP.
+Twitter for example, reaches a speedup of 2$\times$ only with 48 or more threads.
+Meanwhile on SSSP, twitter reached a speedup of around 17$\times$ on those thread counts.
+For the other graphs, the speedup is between 4.2$\times$ (rMat27) and 5.5$\times$ (friendster) at 96 threads.
+So while speedups are possible, not even remotely to the same degree as on SSSP.
+This in turn is not intuitive, one would assume that these two algorithms perform similarly.
+Both algorithms are iterative traversal algorithms with comparable computation and synchronization complexity.
+This behaviour extends even to the case with hugepages (cf.
+\autoref{fig:galoisSpeedupBFS_HP}).
+While the results are generally better, still not to the same degree as SSSP.
+With hugepages, BFS reaches a maximum speedup of 10.5$\times$ on wikipedia.
+The two graphs with largest speedup, namely wikipedia and friendster roughly follow a line with slope 0.125.
+So with every 8 threads, the speedup is increased by about 1.
+Orkut follows the same line up to about 32 threads, slowly flatting off above that.
+The other graphs hardly reach a speedup of 3$\times$, even at 96 threads.
 
 \subsubsection{PageRank}
-
 We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.

### 2020-09-16 documentation/paper/sections/results/galoisSpeedup.tex
speedup pr pull vorerst fertig

@@
 We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.
-This is a perfect example for an algorithm that does not benefit from multithreaded computation.
-Computation is hardly sped up on any graph other than flicker, where the reached maximum is 64\%. This maximum is reached at two threads, with speedup steadily declining above that.
+Without hugepages, computation is hardly sped up on any graph other than flickr, where the reached maximum is 64\% (cf.
+\autoref{fig:galoisSpeedupPRPull_noHP}).
+This maximum is reached at two threads, with speedup steadily declining above that.
 The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 31\%\ at 96 threads.
@@
 Friendster and rMat27 can be sped up by 6.5\%\ or 10\%\ respectively on 8 threads.
+
+Most of this changes on activation of the hugepages, see \autoref{fig:galoisSpeedupPRPull_HP}.
+Here actually all graphs except flickr reach a speedup greater than 1.5$\times$.
+Furthermore, orkut is the only graph that never reaches a speedup of 2$\times$.
+Twitter, rMat27, friendster and rMat28 all reach a speedup of around 2.5$\times$ at 96 threads.
+
+%###
 

### 2020-09-18 documentation/paper/main.tex
kleine änderungen
weitere kleine Änderungen
Appendices merged

@@
 \section{Discussion}
-\todo{Section}
+\todo{Hier zunächst mal ein kurzer Vergleich zwischen Distributed und Single Node wo sinnvoll, dann zwischen den Frameworks und so weiter..}
 
@@
 
-\appendices
-\todo{Beide zusammenfassen}
-\section{Installation Guides}\label{app:installationGuides}
-The complete installation guides with steps on how to run the same algorithms we used is available in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}
-
-In case of frameworks like Giraph, where for example no BFS algorithm is provided by the framework itself, the algorithm and a guide on how to include it in the framework can be found in the repository as well.
-
-\section{Conversion Tool}\label{app:conversionTool}
-In order to
-
-We have written a number of conversion tools and installation guides to help
-users or developers with the use of the tested frameworks.
-
-Our GitHub repository: \url{http://www.github.com/serengti/Forschungsprojekt}.
+\appendix
+\label{app:installationGuides}\label{app:conversionTool}
+We provide a set of installation guides, describing setup and usage of each of the presented frameworks.
+In case of Giraph, where for example no BFS algorithm is provided by the framework itself, the algorithm and a guide on how to include it in the framework is provided as well.
+Furthermore, the conversion tool we described to convert between the different graph input formats is available alongside the guides in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}.
 

### 2020-09-18 documentation/paper/sections/results/galoisSpeedup.tex
table beautify
speedup finished, with conclusions for each subsection

@@
 	\hfil
-	\caption{Calculation time speedups on SSSP for the different graphs with and without hugepages.}
+	\caption{Calculation time speedups on SSSP}
 	\label{fig:galoisSpeedupSSSP}
@@
 
+While the hugepages did not neccessarily help improve the speedup value, they helped bring the different graphs together.
+Thus, the overall speedup benefit might not be as good as without hugepages.
+But the speedup is much more reliable regardless of the graph topology.
+
 \begin{table}
-\renewcommand{\arraystretch}{1.2}
-%\scriptsize
+\renewcommand{\arraystretch}{1.3}
 \centering
@@
 \label{tbl:ssspMeansVariances}
-\begin{tabular}{ccccccccc}
+% \begin{tabular}{ccccccccc}
+% \toprule
+% &\multicolumn{8}{c}{\bf Thread count}\\
+% \cmidrule(r){2-9}
+% &\multicolumn{2}{c}{\bf 1}&\multicolumn{2}{c}{\bf 2}&\multicolumn{2}{c}{\bf 4}&\multicolumn{2}{c}{\bf 8}\\
+% {\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
+% w/o&1.0&0.0&1.6&0.2&2.5&1.1&4.5&5.3\\
+% w/&1.0&0.0&1.6&0.1&2.5&0.9&3.4&2.1\\
+% \midrule
+% \addlinespace
+% %&\multicolumn{8}{c}{\bf Thread count}\\
+% %\cmidrule(r){2-9}
+% &\multicolumn{2}{c}{\bf 16}&\multicolumn{2}{c}{\bf 32}&\multicolumn{2}{c}{\bf 48}&\multicolumn{2}{c}{\bf 96}\\
+% {\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
+% w/o&6.7&13.0&9.6&38.3&10.3&41.4&10.2&38.1\\
+% w/&5.8&7.3&7.0&10.8&10.7&31.4&10.8&29.9\\\bottomrule
+% \multicolumn{9}{c}{* HugePages}
+% \end{tabular}
+\begin{tabular}{c@{\hskip 2\tabcolsep}rrc@{\hskip 2\tabcolsep}rr}
 \toprule
-&\multicolumn{8}{c}{\bf Thread count}\\
-\cmidrule(r){2-9}
-&\multicolumn{2}{c}{\bf 1}&\multicolumn{2}{c}{\bf 2}&\multicolumn{2}{c}{\bf 4}&\multicolumn{2}{c}{\bf 8}\\
-{\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
-w/o&1.0&0.0&1.6&0.2&2.5&1.1&4.5&5.3\\
-w/&1.0&0.0&1.6&0.1&2.5&0.9&3.4&2.1\\
+&\multicolumn{2}{c}{\!\!\!$\mu$}&\enspace&\multicolumn{2}{c}{\!\!$\sigma^2$}\\
+\cmidrule(r){2-3}\cmidrule(r){5-6}
+{\bf\#Threads}&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}&\enspace&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}\\\midrule
+\bf1 & 1.0 & 1.0 &\enspace& 0.0 & 0.0 \\
+\bf2 & 1.6 & 1.6 &\enspace& 0.2 & 0.1 \\
+\bf4 & 2.5 & 2.5 &\enspace& 1.1 & 0.9 \\
+\bf8 & 4.5 & 3.4 &\enspace& 5.3 & 2.1 \\
+\bf16 & 6.7 & 5.8 &\enspace& 13.0 & 7.3 \\
+\bf32 & 9.6 & 7.0 &\enspace& 38.3 & 10.8 \\
+\bf48 & 10.3 & 10.7 &\enspace& 41.4 & 31.4 \\
+\bf96 & 10.2 & 10.8 &\enspace& 38.1 & 29.9 \\
 \bottomrule
-\toprule
-&\multicolumn{8}{c}{\bf Thread count}\\
-\cmidrule(r){2-9}
-&\multicolumn{2}{c}{\bf 16}&\multicolumn{2}{c}{\bf 32}&\multicolumn{2}{c}{\bf 48}&\multicolumn{2}{c}{\bf 96}\\
-{\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
-w/o&6.7&13.0&9.6&38.3&10.3&41.4&10.2&38.1\\
-w/&5.8&7.3&7.0&10.8&10.7&31.4&10.8&29.9\\\bottomrule
-\multicolumn{9}{c}{* HugePages}
 \end{tabular}
@@
 	\hfil
-	\caption{Calculation time speedups on BFS for the different graphs with and without hugepages.}
+	\caption{Calculation time speedups on BFS}
 	\label{fig:galoisSpeedupBFS}
@@
 If we look at the results without hugepages first, we see most significantly, that the speedup never exceeds 6$\times$ even when using 96 threads (cf. \autoref{fig:galoisSpeedupBFS_noHP}).
-For the smaller graphs (flickr, orkut), we have the same behaviour as on SSSP without hugepages. Speedup is close to 1 in all cases, with orkut reaching a maximum speedup of 1.6$\times$ at 24 threads.
+For both flickr and orkut, we have the same behaviour as on SSSP without hugepages. Speedup is close to 1 in all cases, with orkut reaching a maximum speedup of 1.6$\times$ at 24 threads.
 That said, the larger graphs are not benefitting from more threads as much as they did with SSSP.
@@
 
-\subsubsection{PageRank}
-We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.
-Without hugepages, computation is hardly sped up on any graph other than flickr, where the reached maximum is 64\% (cf.
-\autoref{fig:galoisSpeedupPRPull_noHP}).
-This maximum is reached at two threads, with speedup steadily declining above that.
-The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 31\%\ at 96 threads.
-All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases and if so only by a small margin.
-Computation on Orkut and Twitter reaches a speedup maximum of 12\%\ and 5\%\ at 4 threads, while being less or equal to 1 in all other cases.
-The wikipedia graph is never sped up.
-Friendster and rMat27 can be sped up by 6.5\%\ or 10\%\ respectively on 8 threads.
-
-Most of this changes on activation of the hugepages, see \autoref{fig:galoisSpeedupPRPull_HP}.
-Here actually all graphs except flickr reach a speedup greater than 1.5$\times$.
-Furthermore, orkut is the only graph that never reaches a speedup of 2$\times$.
-Twitter, rMat27, friendster and rMat28 all reach a speedup of around 2.5$\times$ at 96 threads.
-
-%###
-
-Speedup results on PageRank show odd behaviour in the Galois implementation.
-There is a significant performance loss on 4, 24 and 40 threads that is far from the expected behaviour.
-This is most visible for the Push variant seen in \autoref{fig:galoisSpeedupPRPush}, we validated the shown results two times.
-Especially, the speedup for 24 threads is (by interpolating between 16 and 32 threads) expected to be anywhere between 25\% and 94\%.
-Actually however, the system does not reach a speedup of more than 4\%\ on any graph, with only rMat27 actually reaching a value greater than 1.
-On all other graphs, using 24 threads is anywhere from 3\% (flickr) to 9\%\ (wikipedia) slower than using just one thread.
-
-Similar yet less pronounced behaviour is oberservable for Pull in \autoref{fig:galoisSpeedupPRPull}.
-Here especially the values for 24 and 40 threads show a loss in performance.
-It is most visible on the values for twitter and friendster, where both values drop significantly compared to the neighbouring 32 and 48 thread results.
-
-\todo{unfertig}
-We assume the reason for this behaviour to be founded in the hugepages. Galois is recommended to be run with hugepages.
-
+\subsubsection{PageRank Pull}
 \begin{figure*}
@@
 	\hfil
-	\caption{Calculation time speedups on PR Pull for the different graphs with and without hugepages.}
+	\caption{Calculation time speedups on PR Pull}
 	\label{fig:galoisSpeedupPRPull}
 \end{figure*}
+We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.
+Without hugepages, computation time is hardly reduced on any graph other than flickr, where the reached maximum is 1.6$\times$ (cf.
+\autoref{fig:galoisSpeedupPRPull_noHP}).
+This maximum is reached at two threads, with speedup steadily declining above that.
+The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 1.3$\times$ at 96 threads.
+All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases and if so only by a small margin.
+Computation on Orkut and Twitter reaches a speedup maximum of 12\%\ and 5\%\ at 4 threads, while being less or equal to 1 in all other cases.
+Speedup on the wikipedia graph is never grater than one.
+Friendster and rMat27 can be sped up by 6.5\%\ or 10\%\ respectively on 8 threads.
 
+Most of this changes with the activation of hugepages, our data can be seen in \autoref{fig:galoisSpeedupPRPull_HP}. 
+Here actually all graphs except flickr reach a speedup greater than 1.5$\times$. Remember, that 1.6$\times$ was the maximum possible speedup without hugepages.
+Furthermore, orkut is the only graph that never reaches a speedup of 2$\times$.
+Twitter, rMat27, friendster and rMat28 all reach a speedup of around 2.5$\times$ at 96 threads.
+Thus, we conclude that hugepages are very much recommended when using Galois PageRank Pull or a similar algorithm.
+Only with hugepages enabled do we observe a considerable increase in performance when using many threads, rather than a slowdown compared to the single threaded performance.
+
+\subsubsection{PageRank Push}
 \begin{figure*}
@@
 	\hfil
-	\caption{Calculation time speedups on PR Pushgi for the different graphs with and without hugepages.}
+	\caption{Calculation time speedups on PR Push}
 	\label{fig:galoisSpeedupPRPush}
 \end{figure*}
+The speedup results on the Push variant show odd behaviour in the Galois implementation, both with and without hugepages.
+Especially without hugepages, there is a significant performance loss on 4, 24 and 40 threads that is far from the expected behaviour.
+This is most visible in \autoref{fig:galoisSpeedupPRPush_noHP}, we validated the shown results with multiple samples each.
+The speedup for 24 threads is, by interpolating between 16 and 32 threads, expected to be anywhere between 1.3$\times$ and 1.9$\times$.
+Actually however, the system does not reach a speedup of more than 1.04$\times$ on any graph, with only rMat27 actually reaching a value greater than 1.
+On all other graphs, using 24 threads is anywhere from 3\% (flickr) to 9\%\ (wikipedia) slower than when using just one thread.
+We intially assumed that this is due to the missing hugepages, Galois is recommended to run with.
+This seems to be only partly true, since the results with hugepages are generally better, but still show some of this behaviour (cf.
+\autoref{fig:galoisSpeedupPRPush_HP}).
+There is only one significant anomaly at 32 threads, while the points where we observed the performance drop before look more \emph{reasonable}\todo{das ist nicht das wort, das ich suche..}.
+At 32 threads, speedup for 5 of 7 graphs drops to a value between 1.4$\times$ and 1.5$\times$.
+Those five graphs reach a speedup larger than 1.8$\times$ both at 16 and 48 threads (i.e.
+immediately before and after 32).
+
+If we disregard these anomalies however, we see a significant performance increase with hugepages compared to without them.
+Actually, all graphs show an improvement in their maximum reached spedups.
+The largest improvements are from 1.8$\times$ to 2.2$\times$ for flickr, 1.5$\times$ to 2.1$\times$ for rMat27, 1.5$\times$ to 2.4$\times$ for friendster and 1.4$\times$ to 2.2$\times$ for rMat28, just by switching hugepages on.
+Interestingly, with hugepages all graphs reach a speedup with 4 threads that is very similar, if not equal to the speedup at 96 threads.
+The largest discrepancy between the two thread counts is rMat27, with a performance uplift from 1.96$\times$ to 2.14$\times$ at 4 and 96 threads.
+
+Using hugepages shows very effective, while it does not completely remove the anomalies we observed, it definitely increases performance by a considerable margin.

### 2020-09-18 plots/makedata.py

@@ -611,36 +611,63 @@ for g in graphs:
         speedupGaloisPRPush_HP[g][x_] = tPRPush / speedupGaloisPRPush_HP[g][x_]
         speedupGaloisPRPull_HP[g][x_] = tPRPull / speedupGaloisPRPull_HP[g][x_]
 
+# for g in graphs:
+#     print(g, max([speedupGaloisPRPush[g][k] for k in xHP]))
+    # print(g, max([speedupGaloisPRPush_HP[g][k] for k in xHP]))
 
 if False:
-    for xVal in xHP[:-4]:
+    means_noHP = {}
+    means_HP = {}
+    vars_noHP = {}
+    vars_HP = {}
+    for xVal in xHP:
         print(xVal,"&",xVal,end="&")
     print("\\\\")
-    for xVal in xHP[:-4]:
+    for xVal in xHP:
         mean = 0
         for k in graphs:
             mean += speedupGaloisSSSP[k][xVal]
         mean /= len(graphs)
-        print(round(mean * 10)/ 10, end="&")
+        val = round(mean * 10)/ 10
+        means_noHP[xVal] = val
+        print(val, end="&")
 
         variance = 0
         for k in graphs:
             variance += (speedupGaloisSSSP[k][xVal] - mean)**2
         variance /= len(graphs)
-        print("(", round(variance * 10)/ 10, ")", end="&")
+        variance = round(variance * 10) / 10
+        vars_noHP[xVal] = variance
+        print(variance, end="&")
 
     print("\\\\")
-    for xVal in xHP[:-4]:
+    for xVal in xHP:
         meanHP = 0
         for k in graphs:
             meanHP += speedupGaloisSSSP_HP[k][xVal]
         meanHP /= len(graphs)
-        print(round(meanHP * 10)/ 10, end="&")
+        val = round(meanHP * 10)/ 10
+        print(val, end="&")
+        means_HP[xVal] = val
 
         variance = 0
         for k in graphs:
             variance += (speedupGaloisSSSP_HP[k][xVal] - meanHP)**2
         variance /= len(graphs)
-        print("(", round(variance * 10)/ 10, ")", end="&")
+        variance = round(variance * 10)/ 10
+        print(variance, end="&")
+        vars_HP[xVal] = variance
+
 
     print("\\\\")
+
+
+    print()
+    for x_ in xHP:
[…]

### 2020-09-18 documentation/paper/sections/overview/gemini.tex

@@
 The real bottleneck of distributed systems is not the communication itself, but the extra instructions, as well as memory references and a lower usage of multiple cores compared to the shared memory counterparts.
-The main reasons for this are (1) the use of hash maps to convert vertex IDs between local and global states, (2) the maintenance of vertex replicas, (3) the communication-bound apply phase in the GAS abstraction, and (4) the lack of dynamic scheduling \cite{Gemini}.
+There are four main reasons for this.
+The first reason is the use of hash maps to convert the vertex IDs between the global and the local state.
+The second reason is the maintenance of vertex replicas on the different systems.
+Another reason is the communication-bound apply phase in GAS abstraction.
+And the last reason is the lack of dynamic scheduling.
+
 Gemini tries to work around all the problems mentioned above, by implemening a message-based system from scratch and getting rid of the extra mapping layer between shared-memory computation and communication.

### 2020-09-18 documentation/paper/sections/overview/giraph.tex
kleine änderungen von Rückmeldung 3 im overview

@@
 This method receives messages sent in the previous superstep as well as its vertex value and the values of outgoing edges.
-With this data the values are modified and messages to other vertices are sent.
+With this data, the vertex values are modified and messages to other vertices are sent.
 Communication between vertices is only performed via messages, so a vertex has no direct access to values of other vertices.
 The only visible information is the set of attached edges and their weights.
-Supersteps are synchronized with barriers, meaning that all messages only get delivered in the following superstep and computation for the next superstep can only begin after every vertex has finished computing the current superstep.
+Supersteps are synchronized using barriers, meaning that all messages only get delivered in the following superstep and computation for the next superstep can only begin after every vertex has finished computing the current superstep.
 Edge and vertex values are retained across supersteps.
@@
 In order for Giraph a achieve scalability and parallelization, it is built on top of Apache Hadoop \cite{Giraph}.
-Hadoop is a MapReduce infrastructure providing a fault tolerant basis for large scale graph processing.
+Hadoop is a MapReduce infrastructure providing a fault tolerant basis for large scale data processing.
 Hadoop supplies a distributed file system (HDFS), on which all computations are performed.

### 2020-09-18 documentation/paper/sections/results/sssp.tex
kleine änderungen von Rückmeldung 3 in den sssp results, todo sind noch die result-Überschriften

@@
 
-\subsection{Single-source Shortest-paths}
-For the results of our SSSP runs, we first analyze the single-node and distributed performances separately before comparing the two.
-
+\subsection{Single-Source Shortest-Paths}
+In this section, we compare the different frameworks in their performance on the SSSP algorithm.
+We first analyze the frameworks on a single computation node and compare to the distributed setup after that.
 \subsubsection{Single-node}
@@
 Thus, the data points on the larger graphs for Giraph are missing in the figures and our evaluation.
-Also, both Ligra and Polymer failed on rMat28. Because of that, removed rMat28 in this section.
+Also, both Ligra and Polymer failed on rMat28.
 
@@
 excluding Giraph in the further comparisons), shows them to perform similar on the smaller graphs.
-There, the average calculation times of the four frameworks (Galois, Gemini, Ligra, Polymer) are 35ms on flickr, 77ms on orkut, 538ms on wikipedia.
+There, the average calculation times of the four frameworks (Galois, Gemini, Ligra, Polymer) are 35ms on flickr, 77ms on orkut and 538ms on wikipedia.
 The four frameworks are close to this average for the smaller graphs. Galois and Gemini being 24ms and 12ms faster than the average. Ligra and Polymer are slower by 27ms and 9ms.

### 2020-09-18 plots/makedata.py

@@ -83,8 +83,6 @@ for f in frameworks:
         fname += "-cpu-96thread" # select Galois thread count here
     k = 0 # graph index
     for g in graphs:
-        if g == 'rMat28':
-            continue
         for i in range(len(graph)):
             if g == graph[i] and fname == algo[i]:
                 tmpCalc.append(calcTime[i])

### 2020-09-18 plots/makeplots.py

@@ -5,9 +5,9 @@ from plotFunctions import *
 
 
 distributed = False
-singleNode = False
+singleNode = True
 
-sssp = False
+sssp = True
 bfs = False
 pr = False
 
@@ -16,7 +16,7 @@ execTimes = False
 overheads = False
 
 
-galoisSpeedup = True
+galoisSpeedup = False
 
 
 colors = {"Galois":"#1f77b4",
@@ -34,7 +34,7 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 	## CALC TIME
 	#SSSP
 	if sssp or calcTimes:
-		grouped_bar_plot(graphs[:-1], frameworks, calcTimeSSSP_singleNode, yErrs=yErrCalcSSSP_singleNode, yLabel='Calculation time (s)', saveToFile="singleNodeSSSP_calcTime.png", colors_=colors)
+		grouped_bar_plot(graphs, frameworks, calcTimeSSSP_singleNode, yErrs=yErrCalcSSSP_singleNode, yLabel='Calculation time (s)', saveToFile="singleNodeSSSP_calcTime.png", colors_=colors)
 
 	#BFS
 	if bfs or calcTimes:
@@ -50,7 +50,7 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 	## EXEC TIME
 	#SSSP
 	if sssp or execTimes:
-		grouped_bar_plot(graphs[:-1], frameworks, execTimeSSSP_singleNode, yErrs=yErrExecSSSP_singleNode, yLabel='Execution time (s)', saveToFile="singleNodeSSSP_execTime.png", colors_=colors)
+		grouped_bar_plot(graphs, frameworks, execTimeSSSP_singleNode, yErrs=yErrExecSSSP_singleNode, yLabel='Execution time (s)', saveToFile="singleNodeSSSP_execTime.png", colors_=colors)
 
 	#BFS
 	if bfs or execTimes:
@@ -68,9 +68,9 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 	## OVERHEAD
 	#SSSP
 	if sssp or overheads:
-		grouped_bar_plot(graphs[:-1], frameworks, overheadSSSP_singleNode, yLabel='Overhead time (s)', saveToFile="singleNodeSSSP_overheadTime.png", colors_=colors)
[…]

### 2020-09-18 documentation/paper/sections/preliminiaries.tex
congest model

@@
 The Bulk-Synchronous Parallel (BSP) model is a computation model developed by Leslie Valiant \cite{bsp}. It is commonly used in computation environments with large amounts of synchronous computation.
-
 This model describes components, a communication network between those components and a method of synchronization.
@@
 Pairs of components can only communicate using messages, thus remote memory access is also only possible in this way.
-The Messages have a user-defined form and should be as small as possible to keep the network traffic low. \todo{Congest Model, log limit in was?}
+The Messages have a user-defined form and should be as small as possible to keep the network traffic low.
+The \emph{Congest model} is a closely related model and furthermore describes the messages.
+There, the message length has to be logarithmic in the graph size \todo{cite}.
+
 Synchronization is realized through barriers for some or all processes.

### 2020-09-18 documentation/paper/sections/results/bfs.tex
kleine änderungen von Rückmeldung 3 in bfs

@@
 The calculation times provided in \autoref{fig:singleNodeBFS_calc} show Gemini and Ligra to be comparable in their performance. 
-Their computation times deviate less than 151ms on all graphs except wikipedia and rMat28, wight Ligra being the faster framework on most graphs.
+Their computation times deviate less than 151ms on all graphs except wikipedia and rMat28, with Ligra being the faster framework on most graphs.
 Ligra is between 2ms (twitter) and 151ms (friendster) faster than Gemini.
@@
 For the remaining frameworks, we compare to Ligra since it is generally slightly faster than Gemini.
-Giraph and Polymer are one to two orders of magnitude slower than both Ligra or Gemini. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
+Giraph and Polymer are one to two orders of magnitude slower than Ligra. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
 Giraph takes between 13$\times$ and 32$\times$ longer than Ligra on flickr, orkut or wikipedia.
@@
 There are 20\% (439M) more edges in friendster, yet the computation for the synthetic rMat27 takes 42\% longer.
-
 Galois calculation performance is comparable to Gemini or Ligra on the smaller three graphs (flickr, orkut and wikipedia).
 Only on the larger graphs is Galois slower than Gemini or Ligra, meanwhile Galois is always faster than Polymer by one order of magnitude.
+
+The execution time results are very similiar to our findings of SSSP (cf.
+\autoref{fig:singleNodeBFS_exec}).
+Just like on SSSP, Galois is the fastest framework and Ligra is second fastest on all graphs except flickr.
+Again, Gemini and Polymer are comparable in their performance and at the same time the slowest frameworks on all graphs except flickr, orkut and wikipedia.
+On those three graphs, Giraph is the slowest, by a difference of one to two orders of magnitude compared to Galois.
+On the other graphs, Polymer and Gemini are one order of magnitude slower than Galois.
 

### 2020-09-18 documentation/paper/sections/results/results.tex
änderung der Sections aber das ist noch todo

@@
 
-\section{Results}
+\subsection{Results}
+\todo{Section}
 
-\subsection{The frameworks during setup and benchmark}
+\subsubsection{The frameworks during setup and benchmark}
+\todo{umbenennen}
 We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.
@@
 While we don't expect this to have a significant performance impact on other tasks, it is still suboptimal.
-	%\item The Hadoop distributed file system (HDFS) ran us into disk space problems on multiple occasions.
-First, unless configured otherwise, the standard implementation replicates all data three times, distributed over all participating nodes.
-Second, deleting files on the HDFS does not immediately free up disk space because the files are moved to a \emph{recycling bin}-like location.
-
-	\item 
+	\item Giraph ran us into disk space problems on multiple occasions.
+First, deleting files on the Hadoop distributed file system (HDFS) does not immediately free up disk space because the files are moved to a \emph{recycling bin}-like location.
+Second, some log files that can easily be multiple gigabytes in size are stored outside of the HDFS and are never mentioned in the Giraph documentation.
 \end{enumerate}

### 2020-09-18 documentation/paper/sections/testingMethods.tex
hauptsächlich pr single node auswertung
kleine änderungen von Rückmeldung 3 und reasoning für algorithmen

@@
 
-Setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
+The setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
 All benchmark cases were initiated by our benchmark script available in our repository.
@@
 All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
-Both the rMat27 and rMat28 were created with a modified version (we changed the output format to EdgeList) of a graph generator provided by Ligra.
+Both the rMat27 and rMat28 were created with a modified version of a graph generator provided by Ligra (we changed the output format to EdgeList).
 \begin{table}
@@
 Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.
-For measuring the calculation time, came up with the following:
+This includes time to read the input graph, initialization and any other tasks other than the actual user-defined algorithm.
+Measuring the execution time is straight forward and was done using console time stamps.
+For measuring the calculation time, we came up with the following:
 \begin{itemize}
@@
 
+We chose SSSP and BFS because they are iterative traversal algorithms.
+Active vertices typically are locally concentrated in the graph.
+The results of these algorithms can give some insight on the behaviour of the framework with other, similar behaving algorithms.
+PageRank on the other hand is an algorithm that is very different to SSSP or BFS for that matter.
+With PR, there are many active vertices spread across the entire graph, enforcing different data handling strategies from the framework.
+
 In detail, the algorithms for each framework are:
+\todo{Algorithmen auflisten}
 \begin{itemize}

### 2020-09-18 documentation/paper/main.tex
citation for the congest model
Umsortieren der ganzen Results in Evaluation

@@
 
-\section*{TODOs}
-\begin{itemize}
-  \item Appendix für Setup Guides - md zu pdf
-\end{itemize}
-
 \section{Introduction}
@@
 
+\section{Evaluation}
+\todo{Einleitung}
+
 \input{sections/testingMethods}
@@
 \section*{Acknowledgments}
-%We would like to thank our supervisor Heiko Geppert for his continued guidance and support.
+We would like to thank our supervisor Heiko Geppert for his continued guidance and support.
 
@@
 
-%\bibliographystyle{IEEEtran}
-%\bibliography{literature}
+% \bibliographystyle{IEEEtran}
+% \bibliography{literature}
 
@@
   \url{https://doi.org/10.1145/79173.79181}
+\BIBentrySTDinterwordspacing
+
+\bibitem{congestModel}
+\BIBentryALTinterwordspacing
+S.~Forster and D.~Nanongkai, ``A faster distributed single-source shortest
+  paths algorithm,'' \emph{2018 IEEE 59th Annual Symposium on Foundations of
+  Computer Science (FOCS)}, Oct 2018. [Online]. Available:
+  \url{http://dx.doi.org/10.1109/FOCS.2018.00071}
 \BIBentrySTDinterwordspacing

### 2020-09-18 documentation/paper/sections/results/bfs.tex
Umsortieren der ganzen Results in Evaluation

@@
 
-\subsection{Breadth-first search}
+\subsubsection{Breadth-first search}
 As with our results for SSSP, for BFS we will begin with our single-node results before looking at the distributed scenario.
-\subsubsection{Single-Node}
+
+\todo{Einleitung, analog SSSP}
+
+\paragraph{Single-Node}
 \begin{figure*}
@@
 
-\subsubsection{Distributed}
+\paragraph{Distributed}
 \begin{figure*}
@@
 This makes Galois Push the clear winner for distributed BFS.
+
+\todo{Discussion ausarbeiten}

### 2020-09-18 documentation/paper/sections/results/pr.tex
Umsortieren der ganzen Results in Evaluation

@@
 
-\subsection{PageRank}
+\subsubsection{PageRank}
+\todo{Einleitung}
 
-\subsubsection{Single-Node}
+\paragraph{Single-Node}
 Ligra and Polymer support both regular and Delta-PageRank variants.
@@
 
-\subsubsection{Distributed}
+\paragraph{Distributed}
 The \autoref{fig:distributedPR} shows our results of PageRank on the distributed cluster.
@@
 \todo{Auswertung fehlt noch.}
+\todo{Discussion}

### 2020-09-18 documentation/paper/sections/results/results.tex
Umsortieren der ganzen Results in Evaluation

@@
 \subsection{Results}
-\todo{Section}
+\todo{Einleitung}
 
-\subsubsection{The frameworks during setup and benchmark}
-\todo{umbenennen}
+\subsubsection{Lessons Learned}
+
+\todo{umbenennen?}
+
 We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.
@@
 \input{sections/results/galoisSpeedup}
-
-% \begin{figure}
-% 	\includegraphics[width=\columnwidth]{../../plots/singleNodeSSSP_overheadTime.png}
-% 	\caption{Overhead SSSP single node}
-% 	\label{fig:singleNodeSSSP_overhead}
-% \end{figure}
-
-% \begin{figure}
-% 	\includegraphics[width=\columnwidth]{../../plots/singleNodeBFS_overheadTime.png}
-% 	\caption{Overhead BFS single node}
-% 	\label{fig:singleNodeBFS_overhead}
-% \end{figure}

### 2020-09-18 documentation/paper/sections/relatedWork.tex
related work init

@@
 \section{Related Work}
-\todo{Section}
+Almost every framework, that gets published in a paper compares itself to other, similar frameworks. 
+Hence there already exists some information on the relative performance between frameworks. 
+However, many of the most common frameworks today have been released many years ago. 
+Because of that two problems arise.
+First, the comparisons are often themselves very old. 
+And second, many other frameworks are no longer in use because they were replaced by something else.
+We are thus providing an objective benchmark on what changed since the initial release, with a comparison between the most commonly used frameworks today.
+
+The most recent publication on Galois was in 2013 \cite{Galois}, along with a comparison to GraphLab, PowerGraph and Ligra, which is in our testing lineup itself.
+The results of their comparison is mainly, that Galois outperforms Ligra on BFS and SSSP on most graphs.
+Meanwhile, their performance on PR is similar but very dependent on the graph.
+In some cases, Ligra outperforms Galois, on other graphs Galois is faster than Ligra.
+
+Gemini puts itself against Ligra, Galois, PowerGraph, PowerLyra and GraphX \cite{Gemini}.
+In a single-threaded scenario, Ligra is faster than both Galois and Gemini in SSSP and BFS applications. 
+On the other hand, Gemini beats the other two in PR.
+Furthermore, Gemini is presented to be one order of magnitude faster than PowerGraph, PowerLyra and GraphX on a multi-node computation unit.
+
+The publication on Ligra is mainly about the multicore behaviour of Ligra rather than a comparison to other frameworks \cite{Ligra}.
+
+Polymer provides a large set of comparisons to Galois, Ligra and X-Stream, all on a single computation node.
+In nearly all of their test cases, Polymer outperformed the other three frameworks \cite{Polymer}.

### 2020-09-18 documentation/paper/sections/overview/galois.tex
Preliminaries section description + Galois writeup improvements

@@
 The system reduces the complexity of writing parallel applications by providing implicitly parallel (unordered or partially ordered) set iterators.
-These iterators perform operations optimistically, detect arising conflicts and resolve these by invoking inverse methods accordingly.
-The tasks in the sets can be ordered, the tasks may be executed out of order, but the conflict resolution will ensure semantic equivalence to sequential strictly ordered execution.
+These iterators perform operations optimistically, detect arising conflicts and resolve them by invoking inverse methods accordingly.
+The tasks can be ordered, the ordering ensures a sequential strictly ordered semantic.
+Ordered tasks may still be executed out of order without affecting the ordered semantic due to the conflict resolution.
 
@@
 the amount of threads used or policies for splitting the graph.
-All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment
+All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment \cite{vertGalois}.
+
+%For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
+Galois includes applications for many graph analytics problems, among these are Single-Source Shorthest-Paths (SSSP), Breath-First Search (BFS) and PageRank (PR).
+For most of these applications Galois offers several different algorithms to perform these analytics problems and many setting options like the amount of threads used or policies for splitting the graph.
+All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment.
 

### 2020-09-18 documentation/paper/sections/preliminiaries.tex
citation for the congest model
Preliminaries section description + Galois writeup improvements

--- 2020-09-15 documentation/paper/sections/preliminiaries.tex
@@
 \todo{Section beschreiben}
+This section briefly explain the concepts and applications necessary, but not directly related to our work. 
+Initialy graphs and paths are defined, followed by explanations for various graph analysis applications.
+Afterwards a few models for compution on graphs or large data sets in general are explained as well as hugepages.
 

### 2020-09-18 documentation/paper/sections/overview/polymer.tex

@@
 
-Polymer is a NUMA-aware graph-analytics system that inherits the scatter- gather programming interface from Ligra \cite{Ligra}. 
+Polymer is a Non Uniform Memory Access (NUMA) aware graph-analytics system that inherits the scatter-gather programming interface from Ligra \cite{Ligra}.
 Key differences are the data layout and access strategies, Polymer implements.
@@
 Thus, all updates are applied on a single copy of application-defined data.
-
-Runtime states that are dynamically allocated in each iteration would however create overhead due to repeated construction of a virtual address space.
+Data such as the active vertices in each iteration are runtime states, that are dynamically allocated in each iteration.
+This allocation would however create overhead due to repeated construction of a virtual address space.
 These states are thus stored in a custom lock-less (i.e. avoiding contention) lookup table.
@@
 Polymer does not only optimize data locality but also its scheduling is custom.
-Time to synchronize threads on different cores increases dramatically with the growing number of involved sockets.
+The time to synchronize threads on different cores increases dramatically with the growing number of involved sockets.
 Inter-node synchronization takes one order of magnitude longer time than intra-node synchronization \cite{Polymer}.
@@
 
-Furthermore, Polymer switches between different data structures representing the runtime state, similiar to Ligra's behaviour.
+Furthermore, Polymer switches between different data structures representing the runtime state.
 The main deciding factor to switch is the amount of active vertices relative to an application-defined threshold.

### 2020-09-18 documentation/paper/sections/preliminiaries.tex

@@
 Pairs of components can only communicate using messages, thus remote memory access is also only possible in this way.
-The Messages have a user-defined form and should be as small as possible to keep the network traffic low. \todo{Congest Model, log limit in was?}
+The Messages have a user-defined form and should be as small as possible to keep the network traffic low.
+The \emph{Congest model} is a closely related model and furthermore describes the messages.
+There, the message length has to be logarithmic in the graph size \cite{congestModel}.
+
 Synchronization is realized through barriers for some or all processes.

### 2020-09-18 documentation/paper/sections/results/galoisSpeedup.tex

@@
 
-\subsection{Galois speedup}
+\subsubsection{Behaviour of Galois}
+
+\todo{Section umbenennen?}
+
 \label{sec:galois_speedup}
@@
 
-\subsubsection{Single-Source Shortest-Paths}
+\paragraph{Single-Source Shortest-Paths}
 \begin{figure*}
@@
 \end{figure*}
+\begin{table}
+\renewcommand{\arraystretch}{1.3}
+\centering
+\caption{Mean Speedups and Variances for SSSP With and Without HugePages}
+\label{tbl:ssspMeansVariances}
+\begin{tabular}{c@{\hskip 2\tabcolsep}rrc@{\hskip 2\tabcolsep}rr}
+\toprule
+&\multicolumn{2}{c}{\!\!\!$\mu$}&\enspace&\multicolumn{2}{c}{\!\!$\sigma^2$}\\
+\cmidrule(r){2-3}\cmidrule(r){5-6}
+{\bf\#Threads}&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}&\enspace&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}\\\midrule
+\bf1 & 1.0 & 1.0 &\enspace& 0.0 & 0.0 \\
+\bf2 & 1.6 & 1.6 &\enspace& 0.2 & 0.1 \\
+\bf4 & 2.5 & 2.5 &\enspace& 1.1 & 0.9 \\
+\bf8 & 4.5 & 3.4 &\enspace& 5.3 & 2.1 \\
+\bf16 & 6.7 & 5.8 &\enspace& 13.0 & 7.3 \\
+\bf32 & 9.6 & 7.0 &\enspace& 38.3 & 10.8 \\
+\bf48 & 10.3 & 10.7 &\enspace& 41.4 & 31.4 \\
+\bf96 & 10.2 & 10.8 &\enspace& 38.1 & 29.9 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 Starting with SSSP, we see an algorithm that benefits from many available threads this can be seen in \autoref{fig:galoisSpeedupSSSP}.
@@
 But the speedup is much more reliable regardless of the graph topology.
+\todo{Discussion ausarbeiten}
 
-\begin{table}
-\renewcommand{\arraystretch}{1.3}
-\centering
-\caption{Mean Speedups and Variances for SSSP With and Without HugePages}
-\label{tbl:ssspMeansVariances}
-% \begin{tabular}{ccccccccc}
-% \toprule
-% &\multicolumn{8}{c}{\bf Thread count}\\
-% \cmidrule(r){2-9}
-% &\multicolumn{2}{c}{\bf 1}&\multicolumn{2}{c}{\bf 2}&\multicolumn{2}{c}{\bf 4}&\multicolumn{2}{c}{\bf 8}\\
-% {\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
-% w/o&1.0&0.0&1.6&0.2&2.5&1.1&4.5&5.3\\
-% w/&1.0&0.0&1.6&0.1&2.5&0.9&3.4&2.1\\
-% \midrule
-% \addlinespace
-% %&\multicolumn{8}{c}{\bf Thread count}\\
-% %\cmidrule(r){2-9}
-% &\multicolumn{2}{c}{\bf 16}&\multicolumn{2}{c}{\bf 32}&\multicolumn{2}{c}{\bf 48}&\multicolumn{2}{c}{\bf 96}\\
-% {\bf HP*}&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$&$\mu$&$\sigma^2$\\\hline
-% w/o&6.7&13.0&9.6&38.3&10.3&41.4&10.2&38.1\\
-% w/&5.8&7.3&7.0&10.8&10.7&31.4&10.8&29.9\\\bottomrule
-% \multicolumn{9}{c}{* HugePages}
-% \end{tabular}
-\begin{tabular}{c@{\hskip 2\tabcolsep}rrc@{\hskip 2\tabcolsep}rr}
-\toprule
-&\multicolumn{2}{c}{\!\!\!$\mu$}&\enspace&\multicolumn{2}{c}{\!\!$\sigma^2$}\\
-\cmidrule(r){2-3}\cmidrule(r){5-6}
-{\bf\#Threads}&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}&\enspace&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}\\\midrule
-\bf1 & 1.0 & 1.0 &\enspace& 0.0 & 0.0 \\
-\bf2 & 1.6 & 1.6 &\enspace& 0.2 & 0.1 \\
-\bf4 & 2.5 & 2.5 &\enspace& 1.1 & 0.9 \\
-\bf8 & 4.5 & 3.4 &\enspace& 5.3 & 2.1 \\
-\bf16 & 6.7 & 5.8 &\enspace& 13.0 & 7.3 \\
-\bf32 & 9.6 & 7.0 &\enspace& 38.3 & 10.8 \\
-\bf48 & 10.3 & 10.7 &\enspace& 41.4 & 31.4 \\
-\bf96 & 10.2 & 10.8 &\enspace& 38.1 & 29.9 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsubsection{Breadth-First Search}
+\paragraph{Breadth-First Search}
 \begin{figure*}
@@
 \end{figure*}
+
 For our speedup results on BFS, \autoref{fig:galoisSpeedupBFS} shows the calculation time speedup of Galois' BFS with and without hugepages.
@@
 The other graphs hardly reach a speedup of 3$\times$, even at 96 threads.
+\todo{Discussion ausarbeiten}
 
-\subsubsection{PageRank Pull}
+\paragraph{PageRank Pull}
 \begin{figure*}
@@
 \end{figure*}
+
 We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.
@@
 Only with hugepages enabled do we observe a considerable increase in performance when using many threads, rather than a slowdown compared to the single threaded performance.
+\todo{Discussion ausarbeiten}
 
-\subsubsection{PageRank Push}
+\paragraph{PageRank Push}
 \begin{figure*}
@@
 \end{figure*}
+
 The speedup results on the Push variant show odd behaviour in the Galois implementation, both with and without hugepages.
@@
 Using hugepages shows very effective, while it does not completely remove the anomalies we observed, it definitely increases performance by a considerable margin.
+\todo{Discussion ausarbeiten}

### 2020-09-18 documentation/paper/sections/testingMethods.tex

@@
 
-\section{Evaluation}
+\subsection{Testing Methods}
 \todo{Einleitung}
 
-\subsection{Environment}
+\subsubsection{Environment}
 For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each.
@@
 Unless mentioned otherwise, we always show results of each framework utilizing 96 threads (i.e. the maximum on our machines) for the single-node evaluation.
+The complete benchmark log files and extracted raw results are available in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt}}.
 
-The complete benchmark log files and extracted raw results are available in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt} // das ist hier unschön.}.
-
-\subsection{Data Sets}
+\subsubsection{Data Sets}
 The graphs used in our testing can be seen in detail in \autoref{tbl:graphs}.
@@
 
-\subsection{Measurements}
+\subsubsection{Algorithms}
+The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark each framework with every graph.
+We always show the results of PageRank with a maximum of five iterations.
+For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included both in our evaluation.
+We chose SSSP and BFS because they are iterative traversal algorithms.
+Active vertices typically are locally concentrated in the graph.
+The results of these algorithms can give some insight on the behaviour of the framework with other, similar behaving algorithms.
+PageRank on the other hand is an algorithm that is very different to SSSP or BFS for that matter.
+With PR, there are many active vertices spread across the entire graph, enforcing different data handling strategies from the framework.
+
+In detail, the algorithms for each framework are:
+\todo{Algorithmen auflisten}
+\begin{itemize}
+	\item Galois supports all of our tested algorithms, with both a Push and a Pull variant for PageRank available.
+In the distributed scenario, there are Push and Pull versions for SSSP and BFS available as well.
+	\item Gemini \todo{}
+	\item Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
+For SSSP, slight variations had to be made to the default implementation, to allow us to use different start vertices.
+For PageRank the supplied implementation is used.
+	\item Ligra \todo{}
+	\item Polymer supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
+\end{itemize}
+
+\subsubsection{Measurements}
 For every framework, we measured the \emph{execution time} as the time from start to finish of the console command.
@@
 Later on, we provide the mean values of the individual times as well as the standard deviation where meaningful.
-
-\subsection{Algorithms}
-The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark framework with every graph.
-We always show the results of PageRank with a maximum of five iterations.
-For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included both in our evaluation.
-
-We chose SSSP and BFS because they are iterative traversal algorithms.
-Active vertices typically are locally concentrated in the graph.
-The results of these algorithms can give some insight on the behaviour of the framework with other, similar behaving algorithms.
-PageRank on the other hand is an algorithm that is very different to SSSP or BFS for that matter.
-With PR, there are many active vertices spread across the entire graph, enforcing different data handling strategies from the framework.
-
-In detail, the algorithms for each framework are:
-\todo{Algorithmen auflisten}
-\begin{itemize}
-	\item Galois supports all of our tested algorithms, with both a Push and a Pull variant for PageRank available.
-In the distributed scenario, there are Push and Pull versions for SSSP and BFS available as well.
-	\item Gemini \todo{}
-	\item Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
-For SSSP, slight variations had to be made to the default implementation, to allow us to use different start vertices.
-For PageRank the supplied implementation is used.
-	\item Ligra \todo{}
-	\item Polymer supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
-\end{itemize}

### 2020-09-19 documentation/paper/sections/results/pr.tex
pr distributed

@@
 \paragraph{Distributed}
-The \autoref{fig:distributedPR} shows our results of PageRank on the distributed cluster.
-First of all, Giraph was unable to complete the test because it required more than 250GB of RAM for rMat28, thus this result is missing.
-When comparing the calculation times in \autoref{fig:distributedPR_calc} to the execution times in \autoref{fig:distributedPR_exec}, we see similar behaviour of all frameworks.
-This means that unlike with SSSP or BFS, the calculation times and execution times are similar with respect to the relations of the frameworks to one another.
-More specifically, there are no overhead outliers like it was the case with Giraph on SSSP.
 \begin{figure*}
@@
 	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
+		\caption{Overhead}
+		\label{fig:distributedPR_overhead}
+	\end{subfigure}
+	\hfil
 	\caption{Average times for PR on the distributed cluster, black bars represent one standard deviation in our testing}
@@
 
-\todo{Auswertung fehlt noch.}
+Our benchmark results of PageRank on the distributed cluster can be seen in \autoref{fig:distributedPR}.
+First of all, Giraph was unable to complete the test on rMat28 because it ran out of memory, thus this result is missing.
+When comparing the calculation times in \autoref{fig:distributedPR_calc} to the execution times in \autoref{fig:distributedPR_exec}, we see similar behaviour of all frameworks.
+This means that unlike with SSSP or BFS, the calculation times and execution times are similar with respect to the relations of the frameworks to one another.
+
+Gemini has the shortest calculation times on all graphs (cf. \autoref{fig:distributedPR_calc}).
+The two Galois implementations are second on flickr, orkut, friendster and rMat28, with Giraph being second on the others.
+Generally, the calculation of Galois Push is anywhere from 6\% (flickr) to 46\% (orkut) faster than the Pull counterpart.
+
+This applies to the execution times in almost the same way (cf. \autoref{fig:distributedPR_exec}).
+Gemini is the fastest on all graphs except orkut, where both Galois implementations are faster. Galois Pull takes 13s, whereas Gemini requires 19.4s on orkut.
+Again, as for the calculation times, Galois is the second fastest framework on flickr, wikipedia, friendster and rMat28.
+And Galois Push has smaller execution times than the Pull version because the overhead times for both implementations are similar (cf. \autoref{fig:distributedPR_overhead}).
+
 \todo{Discussion}

### 2020-09-19 documentation/paper/sections/overview/gemini.tex
updatet gemini and push-pull + reorderd frameworks according to dependencies

@@
 It was developed with the goal to deliver a generally better performance through efficient communication.
-While most other distributed graph processing systems achieve very good results in the shared-memory area, they often deliver unsatisfactory results in distributed computing.
-Furthermore, a well optimized single-threaded implementation often outperforms a distributed system.
+While most other graph processing systems achieve very good results in the shared-memory area, they often deliver unsatisfactory results in distributed computing.
+Furthermore, a well optimized single-threaded implementation often outperforms a distributed system \cite{scalability}.
 Therefore it is necessary to not only focus on the performance of the computation but also of the performance of the communication.
-Gemini tries to bridge the gap between efficient shared-memory and scalable distributed systems.
+Gemini tries to bridge the gap between efficient shared-memory and scalable distributed systems \cite{Gemini}.
 To achieve this goal, Gemini, in contrast to the other frameworks discussed here, does not support shared-memory calculation, but chooses the distributed message-based approach from scratch.
@@
 
-Gemini tries to work around all the problems mentioned above, by implemening a message-based system from scratch and getting rid of the extra mapping layer between shared-memory computation and communication.
+Gemini tries to work around all the problems, by implemening a message-based system from scratch and getting rid of the extra mapping layer between shared-memory computation and communication.
 Therefore Ligra's push-pull computation model was adopted and applied to the distributed computation.

### 2020-09-19 documentation/paper/main.tex
some work on discussion/comparison

@@
 \newcommand*\OK{\ding{51}} % a checkmark-style sign
+\newcommand{\txtdagger}{\textsuperscript{\textdagger{}}}
 
@@
 
-\input{sections/relatedWork}
-
 \input{sections/preliminiaries}
@@
 \input{sections/overview/overview}
+
+\input{sections/relatedWork}
 
@@
 \section{Discussion}
-\todo{Hier zunächst mal ein kurzer Vergleich zwischen Distributed und Single Node wo sinnvoll, dann zwischen den Frameworks und so weiter..}
+\todo{unfertig}
+
+There are two main cases that should be regarded in a comparison of the frameworks.
+One, the case of individual calculations on a graph, i.e.
+for each calculation, the graph has to be loaded.
+For this case, frameworks with small execution times and small overhead are preferred.
+Case two on the other hand is the case of a running system, that performs multiple calculations on a single graph, without the need of reloading graph data with every calculation.
+In this case, frameworks with short calculation times should be preferred because the overhead time is only spent once on startup.
+
+Orthogonal to this, the graph size and topology of the system is a decinding factor.
+Large graphs often require a distributed cluster, because the graph is simply too large to fit in the RAM of just one machine.
+We observed this for example with Giraph on 4 of 7 graphs.
+Furthermore, some environments require a distributed framework setup regardless of any other factors.
+
+SSSP
+For graphs that are not too large for the RAM and individual test runs, Galois as the fastest single-node framework should be utilized.
+It provides excellent performance, being orders of magnitude faster than the competing frameworks on a single node.
+
+Especially for frameworks like Giraph, that require large amounts of memory, the distributed scenario allows them to complete their computations.
+
+BFS
+
+PR
 

### 2020-09-19 documentation/paper/sections/relatedWork.tex
some work on discussion/comparison

@@
 \section{Related Work}
+\todo{Der erste Abschnitt hier kann eigentlich so wie er ist auch in die Introduction.}
+
 Almost every framework, that gets published in a paper compares itself to other, similar frameworks. 
@@
 However, many of the most common frameworks today have been released many years ago. 
-Because of that two problems arise.
-First, the comparisons are often themselves very old. 
-And second, many other frameworks are no longer in use because they were replaced by something else.
+Because of that, two problems arise.
+First, the comparisons are based on very old versions of the frameworks, with no information on the current behaviour available. 
+And second, many of the other frameworks are no longer in use because they were replaced by something else, making the comparisons difficult to interpret.
 We are thus providing an objective benchmark on what changed since the initial release, with a comparison between the most commonly used frameworks today.
+
+It quickly becomes apparent that many of the comparisons provided
+
+The comparisons often include PowerGraph or Apache's Spark GraphX, two frameworks that we did not include in our testing.
 
@@
 The results of their comparison is mainly, that Galois outperforms Ligra on BFS and SSSP on most graphs.
-Meanwhile, their performance on PR is similar but very dependent on the graph.
+Meanwhile, performance of Ligra and Galois on PR is similar but very dependent on the graph.
 In some cases, Ligra outperforms Galois, on other graphs Galois is faster than Ligra.
@@
 Gemini puts itself against Ligra, Galois, PowerGraph, PowerLyra and GraphX \cite{Gemini}.
-In a single-threaded scenario, Ligra is faster than both Galois and Gemini in SSSP and BFS applications. 
-On the other hand, Gemini beats the other two in PR.
+In a single-threaded scenario, Ligra was shown to be faster than both Galois and Gemini in SSSP and BFS applications. 
+On the other hand, Gemini beat the other two in PR applications.
 Furthermore, Gemini is presented to be one order of magnitude faster than PowerGraph, PowerLyra and GraphX on a multi-node computation unit.

### 2020-09-19 documentation/paper/sections/results/results.tex
some work on discussion/comparison

@@
 
+\subsubsection{Comparison of a Single-Node to the Distributed Cluster}
+\begin{table}
+	\caption{Execution Times on a SingleNode vs. 5-Node Distributed Cluster}
+	\label{tbl:SSSP_execTimeComparison}
+	\renewcommand{\arraystretch}{1.2}
+	\centering
+	\begin{tabular}{ccr@{\tabskip 1 \tabcolsep}r
+	r@{\tabskip 1 \tabcolsep}r
+	r@{\tabskip 1 \tabcolsep}r}
+		\toprule
+		&&\multicolumn{2}{c}{\bf Galois}&\multicolumn{2}{c}{\bf Gemini}&\multicolumn{2}{c}{\bf Giraph}\\
+		\cmidrule{3-4}\cmidrule{5-6}\cmidrule{7-8}
+		&\bf Graph&1N&5N&1N&5N&1N&5N\\
+		\midrule
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}&flickr & 0.3 & 2.4 & 0.6 & 2.5 & 38.8 & 42.6 \\
+		& orkut & 0.8 & 4.2 & 13.9 & 17.3 & 135.9 & 60.4 \\
+		& wikipedia & 1.8 & 14.2 & 48.5 & 74.5 & 377.2 & 100.8 \\
+		& twitter & 10.8 & 40.0 & 383.0 & 391.1 & - & 349.9 \\
+		& rMat27 & 16.0 & 39.0 & 563.5 & 386.0 & - & 565.8 \\
+		& friendster & 14.4 & 58.5 & 742.6 & 568.9 & - & 444.0 \\
+		& rMat28 & 27.8 & 71.5 & 1236.7 & 792.0 & - & 1180.2 \\
+		\midrule
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}& flickr & 0.6 & 2.3 & 0.7 & 2.4 & 38.9 & 41.9 \\
+		& orkut & 0.9 & 4.0 & 12.5 & 13.8 & 128.2 & 58.1 \\
+		& wikipedia & 2.4 & 13.2 & 44.9 & 60.9 & 360.7 & 95.1 \\
+		& twitter & 14.2 & 37.9 & 355.0 & 293.2 & - & 322.3 \\
+		& rMat27 & 14.6 & 34.4 & 540.8 & 305.3 & - & 539.2 \\
+		& friendster & 12.8 & 47.6 & 708.4 & 450.2 & - & 412.3 \\
+		& rMat28 & 33.1 & 67.1 & 1178.7 & 634.3 & - & 1135.6 \\
+		\midrule
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR}}& flickr & 0.3\txtdagger & 2.6 & 0.7 & 2.2 & 45.6 & 44.1 \\
+		& orkut & 0.7\txtdagger & 8.0 & 13.1 & 19.4 & 561.3 & 106.2 \\
+		& wikipedia & 1.7\txtdagger & 206.5 & 44.4 & 54.4 & 1501.3 & 252.6 \\
+		& twitter & 8.7\txtdagger & 910.9 & 359.9 & 363.5 & - & 1200.2 \\
+		& rMat27 & 19.2\txtdagger & 1287.4 & 536.3 & 376.3 & - & 1369.5 \\
+		& friendster & 20.4\txtdagger & 593.3 & 716.5 & 591.1 & - & 1655.0 \\
+		& rMat28 & 46.0\txtdagger & 2540.2 & 1188.5 & 774.2 & - & - \\
+		\bottomrule
+		\multicolumn{8}{l}{(\txtdagger) Results of Galois Pull shown.}
+	\end{tabular}
+\end{table}
+Until now, we have only compared the frameworks with each other, under the same setup circumstances.
+Hence, a comparison of the same framework, in single-node versus distributed setup is the content of this section.
+Of course only frameworks that were tested in both setups are shown.
+We focus primarily on the results of the execution times, the data can be seen in \autoref{tbl:SSSP_execTimeComparison}.
+We show the results of Galois Push for the 5-node cluster in the table, since it is faster than Pull.
+
+There are some, rather large differences in the execution times of each framework. Galois is consistently faster on a single computation node compared to the distributed setup. 
+On both SSSP and BFS, the margin between the two setups is already noticeable.
+The distributed scenario requires from 2$\times$ (BFS, rMat28) to 7.8$\times$ (SSSP, wikipedia) the execution time of single-node Galois.
+For PR however, the difference is multiple orders of magnitude large. Distributed Galois requires from 8.6$\times$ to 121$\times$ more time than single-node Galois.
+
+Gemini's distributed calculation only requires more time on the smaller graphs (i.e.
+flickr to wikipedia).
+On the larger graphs, the added computation power can be leveraged to overcome the synchronization overhead of the distributed scenario.
+On rMat27, friendster and rMat28, Gemini is anywhere from 1.3$\times$ to 1.6$\times$ faster than on a single-node.
+Giraph already uses the distributed cluster efficiently on the smaller graphs as well.
+Only on flickr is the distributed version slower than a single node.
+The differences for Giraph on the other graphs are rather large, single-node execution time is 2.25$\times$ and 3.7$\times$ the execution time of the distributed version on orkut and wikipedia respectively.
+
 \input{sections/results/galoisSpeedup}

### 2020-09-19 plots/makedata.py

@@ -192,6 +192,11 @@ if False:
 
 
 
+
+
+
+
+
 #### BFS SINGLE NODE
 calcTimeBFS_singleNode = []
 yErrCalcBFS_singleNode = []
@@ -440,6 +445,59 @@ for f in dist_frameworks_pr:
     overheadPR_distributed.append(overhead)
     overheadPRNormalized_distributed.append(overheadNormalized)
 
+
+
+if False:
+    #GaloisPush = 0
+    #GaloisPull = 1
+    print([calcTimePR_distributed[0][i] / calcTimePR_distributed[1][i] for i in range(len(graphs))])
+
+    print(execTimePR_distributed[1][1], execTimePR_distributed[2][1])
+
+
+
+
+
+
+
+
+
+
+if True:
+    for i in range(len(graphs)):
+        # timeGaloisDist = round(execTimeSSSP_distributed[0][i] * 10)/10
+        # timeGeminiDist = round(execTimeSSSP_distributed[2][i] * 10)/10
+        # timeGiraphDist = round(execTimeSSSP_distributed[3][i] * 10)/10
+        # timeGalois = round(execTimeSSSP_singleNode[0][i] * 10)/10
+        # timeGemini = round(execTimeSSSP_singleNode[1][i] * 10)/10
+        # timeGiraph = round(execTimeSSSP_singleNode[2][i] * 10)/10
+
+        # timeGaloisDist = round(execTimeBFS_distributed[0][i] * 10)/10
+        # timeGeminiDist = round(execTimeBFS_distributed[2][i] * 10)/10
+        # timeGiraphDist = round(execTimeBFS_distributed[3][i] * 10)/10
+        # timeGalois = round(execTimeBFS_singleNode[0][i] * 10)/10
+        # timeGemini = round(execTimeBFS_singleNode[1][i] * 10)/10
+        # timeGiraph = round(execTimeBFS_singleNode[2][i] * 10)/10
+
+        timeGaloisDist = round(execTimePR_distributed[0][i] * 10)/10
+        timeGeminiDist = round(execTimePR_distributed[2][i] * 10)/10
+        timeGiraphDist = round(execTimePR_distributed[3][i] * 10)/10
+        timeGalois = round(execTimePR_singleNode[1][i] * 10)/10
+        timeGemini = round(execTimePR_singleNode[2][i] * 10)/10
+        timeGiraph = round(execTimePR_singleNode[3][i] * 10)/10
+        print(timeGaloisDist/timeGalois)
+
[…]

### 2020-09-19 plots/makeplots.py
removed some titles in the plots

@@ -7,9 +7,9 @@ from plotFunctions import *
 distributed = False
 singleNode = True
 
-sssp = True
+sssp = False
 bfs = False
-pr = False
+pr = True
 
 calcTimes = False
 execTimes = False
@@ -38,11 +38,11 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 	#BFS
 	if bfs or calcTimes:
-		grouped_bar_plot(graphs, frameworks, calcTimeBFS_singleNode, yErrs=yErrCalcBFS_singleNode, title='BFS single node', yLabel='Calculation time (s)', saveToFile="singleNodeBFS_calcTime.png", colors_=colors)
+		grouped_bar_plot(graphs, frameworks, calcTimeBFS_singleNode, yErrs=yErrCalcBFS_singleNode, yLabel='Calculation time (s)', saveToFile="singleNodeBFS_calcTime.png", colors_=colors)
 
 	#PR
 	if pr or calcTimes:
-		grouped_bar_plot(graphs, list(singleNode_frameworks_pr.values()), calcTimePR_singleNode, yErrs=yErrCalcPR_singleNode, title='PR single node (ugly)', yLabel='Calculation time (s)', saveToFile="singleNodePR_calcTime.png", colors_=colors)
+		grouped_bar_plot(graphs, list(singleNode_frameworks_pr.values()), calcTimePR_singleNode, yErrs=yErrCalcPR_singleNode, yLabel='Calculation time (s)', saveToFile="singleNodePR_calcTime.png", colors_=colors)
 
 
 
@@ -54,11 +54,11 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 	#BFS
 	if bfs or execTimes:
-		grouped_bar_plot(graphs, frameworks, execTimeBFS_singleNode, yErrs=yErrExecBFS_singleNode, title='BFS single node', yLabel='Execution time (s)', saveToFile="singleNodeBFS_execTime.png", colors_=colors)
+		grouped_bar_plot(graphs, frameworks, execTimeBFS_singleNode, yErrs=yErrExecBFS_singleNode, yLabel='Execution time (s)', saveToFile="singleNodeBFS_execTime.png", colors_=colors)
 
 	#PR
 	if pr or execTimes:
[…]

### 2020-09-19 documentation/paper/sections/results/results.tex
results/discussion

@@
 	\caption{Execution Times on a SingleNode vs. 5-Node Distributed Cluster}
-	\label{tbl:SSSP_execTimeComparison}
+	\label{tbl:execTimeComparison}
 	\renewcommand{\arraystretch}{1.2}
@@
 Of course only frameworks that were tested in both setups are shown.
-We focus primarily on the results of the execution times, the data can be seen in \autoref{tbl:SSSP_execTimeComparison}.
+We focus primarily on the results of the execution times, the data can be seen in \autoref{tbl:execTimeComparison}.
 We show the results of Galois Push for the 5-node cluster in the table, since it is faster than Pull.
@@
 flickr to wikipedia).
-On the larger graphs, the added computation power can be leveraged to overcome the synchronization overhead of the distributed scenario.
-On rMat27, friendster and rMat28, Gemini is anywhere from 1.3$\times$ to 1.6$\times$ faster than on a single-node.
+There, the distributed scenario is anywhere from 1.1$\times$ (BFS, orkut) to 4.2$\times$ (SSSP, flickr) slower.
+Twitter is the tipping point for Gemini's SSSP and PR algorithm.
+Here, execution times on the single-node and distributed cluster are within 2\% of each other.
+For BFS, the tipping point is anywhere between twitter and rMat27.
+Above that, i.e.
+on the larger graphs, the added computation power can be leveraged to overcome the synchronization overhead of the distributed scenario.
+Hence, on those graphs single-node Gemini becomes up to 1.8$\times$ (BFS, rMat28) slower than the distributed version.
+This relation is very dependent on the graph size.
+The smaller the graph, the faster single-node Gemini is in comparison.
+Analogously, the larger the graph, the larger the gap becomes, in favour of the distributed version.
+
+A very similiar behaviour can be observed for Giraph, but the relation tips in favour of the distributed version much sooner.
 Giraph already uses the distributed cluster efficiently on the smaller graphs as well.
-Only on flickr is the distributed version slower than a single node.
-The differences for Giraph on the other graphs are rather large, single-node execution time is 2.25$\times$ and 3.7$\times$ the execution time of the distributed version on orkut and wikipedia respectively.
+On flickr, the execution times of single-node vs distributed Giraph are already within 9\% of each other.
+On the larger graphs, the differences quickly increase, again in favour of the distributed scenario.
+Giraph's single-node version runs up to 5.9$\times$ (PR, wikipedia) slower than the distributed version.
+Keep in mind, that only a comparison between the three smallest graphs can be made here.
 

### 2020-09-19 plots/makedata.py

@@ -463,14 +463,14 @@ if False:
 
 
 
-if True:
+if False:
     for i in range(len(graphs)):
-        # timeGaloisDist = round(execTimeSSSP_distributed[0][i] * 10)/10
-        # timeGeminiDist = round(execTimeSSSP_distributed[2][i] * 10)/10
-        # timeGiraphDist = round(execTimeSSSP_distributed[3][i] * 10)/10
-        # timeGalois = round(execTimeSSSP_singleNode[0][i] * 10)/10
-        # timeGemini = round(execTimeSSSP_singleNode[1][i] * 10)/10
-        # timeGiraph = round(execTimeSSSP_singleNode[2][i] * 10)/10
+        timeGaloisDist = round(execTimeSSSP_distributed[0][i] * 10)/10
+        timeGeminiDist = round(execTimeSSSP_distributed[2][i] * 10)/10
+        timeGiraphDist = round(execTimeSSSP_distributed[3][i] * 10)/10
+        timeGalois = round(execTimeSSSP_singleNode[0][i] * 10)/10
+        timeGemini = round(execTimeSSSP_singleNode[1][i] * 10)/10
+        timeGiraph = round(execTimeSSSP_singleNode[2][i] * 10)/10
 
         # timeGaloisDist = round(execTimeBFS_distributed[0][i] * 10)/10
         # timeGeminiDist = round(execTimeBFS_distributed[2][i] * 10)/10
@@ -479,13 +479,13 @@ if True:
         # timeGemini = round(execTimeBFS_singleNode[1][i] * 10)/10
         # timeGiraph = round(execTimeBFS_singleNode[2][i] * 10)/10
 
-        timeGaloisDist = round(execTimePR_distributed[0][i] * 10)/10
-        timeGeminiDist = round(execTimePR_distributed[2][i] * 10)/10
-        timeGiraphDist = round(execTimePR_distributed[3][i] * 10)/10
-        timeGalois = round(execTimePR_singleNode[1][i] * 10)/10
-        timeGemini = round(execTimePR_singleNode[2][i] * 10)/10
-        timeGiraph = round(execTimePR_singleNode[3][i] * 10)/10
-        print(timeGaloisDist/timeGalois)
+        # timeGaloisDist = round(execTimePR_distributed[0][i] * 10)/10
+        # timeGeminiDist = round(execTimePR_distributed[2][i] * 10)/10
[…]

### 2020-09-19 documentation/paper/sections/overview/overview.tex
kleine fixes von der Rückmeldung und Polymer Overview
updatet gemini and push-pull + reorderd frameworks according to dependencies
framework overview introduction

--- 2020-08-29 documentation/paper/sections/overview/overview.tex
@@
 
-\section{Overview}
-We start with a short overview describing functionality and characteristics of each framework, followed by their required input data formats.
+\section{Framework Overview}
+The following paragraphs is a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
+All of the described frameworks are part of our testing.
+\todo
 
-//
 \subsection{Galois and Gluon}

### 2020-09-19 documentation/paper/sections/overview/gemini.tex

@@
 Gemini is fairly lightweight and has a clearly defined API between the core framework and the implementations of the individual algorithms.
-The five already implemented algorithms are Single-Source Shortest-Path (SSSP), Breath-First Search (BFS), PageRank (PR), Connected-Components (CC) and Biconnected-Components (BC).
+The five already implemented algorithms are Single-Source Shortest-Path (SSSP), Breath-First Search (BFS), PageRank (PR), Connected-Components (CC) and Betweenness-Centrality (BC).

### 2020-09-19 documentation/paper/sections/preliminiaries.tex

@@
 \subsection{Push and Pull Variants}
-Many parallel graph algorithms, as well as the three we consider here, are implemented by iterating over vertices, or edges \cite{tao}.
-The respective vertex or edge, which is considered in an iteration, is also called \emph{active}.
-Applying an operator to this \emph{active} vertex or edge is called an \emph{activity}.
-This operator considers always only a local so called \emph{neighborhood} to the \emph{active} vertex or edge.
-To this \emph{neighborhood} belong only vertices and edges, which are in the direct surrounding of the \emph{active} vertex or edge, which allows to parallelise the iteration relatively easy.
+Many parallel graph algorithms, as well as the three we consider here, are implemented by iterating over vertices \cite{tao}.
+The respective vertex, which is considered in an iteration and the operator is applied to, is also called \emph{active}.
+It's also possible to iterate over edges and consider them as \emph{active}, but in the following, whithout loss of generality, we will use the term \emph{active vertex}.
+This operator considers always only a local so called \emph{neighborhood} to the \emph{active} vertex.
+To this \emph{neighborhood} belong only vertices, which are in the direct surrounding of the \emph{active} vertex.
+This allows to parallelise the iteration relatively easy, caused by the locality of the operator.
 
 In general such an operator can change the whole structure of the \emph{neighbourhood}.
-But in the following we assume, that an operator changes only \emph{labels} of vertices or edges in the \emph{neighbourhood}.
-\emph{Labels} of edges also called \emph{weights}.
+But in the following we assume, that an operator changes only \emph{labels} of vertices in the \emph{neighbourhood}.
+\emph{Labels} of edges are also called \emph{weights}.
 Often this operator can be implemented in two different ways, called \emph{push style} or \emph{pull style}.
-A \emph{push-style} operator reads the \emph{label} of the \emph{active} vertex or edge and updates the \emph{label} of its \emph{neighborhood}.
-In contrast the \emph{pull-style} operator reads all values of its \emph{neighborhood} and updates the value of the \emph{active} vertex or edge.
-
+A \emph{push-style} operator reads the \emph{label} of the \emph{active} vertex and updates the \emph{label} of its \emph{neighborhood}.
+These operators are more efficient, if there are only a few \emph{active} vertices at the same time, or the \emph{neighborhoods} do not overlap, which can not be avoided in general.
+In contrast the \emph{pull-style} operator reads all values of its \emph{neighborhood} and updates the value of the \emph{active} vertex.
 \emph{Pull-style} operators need less synchronization in parallel implementations, because unlike \emph{push style} there is only one write and many read operations.
 Thus locks can be avoided.
-So \emph{push-style} operators are more efficient, if there are only a few \emph{active} vertices or edges at the same time, or the \emph{neighborhoods} do not overlap, which can not be avoided in general.
-On the other hand \emph{pull-style} operators are more efficient, if there are many \emph{active} vertices or edges at the same time.
+So these operators are more efficient, if there are many \emph{active} vertices at the same time.
 

### 2020-09-19 documentation/paper/main.tex
Ideen in Introduction rein gepastet

--- 2020-09-18 documentation/paper/main.tex
@@
 
-\section{Introduction}
-\todo{Section}
-% In recent years graph sizes have increased significantly [?] .
-% Thus performance and memory efficiency of the graph analysis applications is now more important than ever.
-
-% Many applications like
-
-% Wachsende graphen,
-% benötigt für folgende Applikationen,
-% verwenden graph algorithmen,
-% diese müssen schnell und verteilt laufen
-% dafür wurden Frameworks entwickelt...
-
-% und das noch schön formulieren
-
-% We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
-% The comparison is performed on both real world data sets and synthetic graphs.
-
-% To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
-
-%One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
-%This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.
-
-% To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.
-
-This paper makes the following contributions:
-\begin{itemize}
-  \item Comparison of several state-of-the-art graph processing frameworks
-  \item
-  \item
-  \item
-\end{itemize}
+\input{sections/introduction}
 
@@
 
+\bibitem{scalability}
+\BIBentryALTinterwordspacing
+F.~McSherry, M.~Isard, and D.~G. Murray, ``Scalability! but at what {COST}?''
+  in \emph{15th Workshop on Hot Topics in Operating Systems (HotOS
+  {XV})}.\hskip 1em plus 0.5em minus 0.4em\relax Kartause Ittingen,
+  Switzerland: {USENIX} Association, May 2015. [Online]. Available:
+  \url{https://www.usenix.org/conference/hotos15/workshop-program/presentation/mcsherry}
+\BIBentrySTDinterwordspacing
+
 \bibitem{konect}

### 2020-09-19 documentation/paper/sections/introduction.tex
Ideen in Introduction rein gepastet

%!TEX root=../../main.tex
\section{Introduction}

1. Einführung, warum ist Graph analysis wichtig\\
- Graph processing is gaining increasing attentions in both academic and industrial communities. (Gemini)\\
- Graphen werden immer wichtiger in vielen Domänen \todo{welche}\\
-Many machine learning, data mining and scientific computation can be modeled as graph-structured computation, resulting in a new application domain called graph analytics. (Polymer)\\
- In recent years graph sizes have increased significantly. Thus performance and memory efficiency of the graph analysis applications is now more important than ever. \todo{Quelle}\\
- Many applications like, Wachsende graphen, benötigt für folgende Applikationen, verwenden graph algorithmen, diese müssen schnell und verteilt laufen\\
- Graphen wachsen schneller als CPUs \todo{Quelle}\\
\\
2. Ausgangssituation und Problem\\
- dafür wurden Zahlreiche Framweworks entwickelt (Paper: Galois)\\
- Domänenspezifische Sprachen DSL (Paper: Galois)\\
- Es wurden auch Librarys und Technologien entwickelt (boost, hugepages, numa, ...)\\
- Um es zu erleichtern Graphprogramme zu schreiben (Paper: Galois)\\
- Größe der Graphen fordert Parallelität/Verteiltheit => Probleme  (Paper:Galois)\\
- Prolem der Vergleichbarkeit der Systeme\\
- Systeme entwickeln sich weiter und wurden nur zu deren Start verglichen\\
- Vergleiche wurden von Machern der Systeme erledigt\\
\\
3. Was wir zur Lösung des Problems beitragen\\
- ersten, die unabhängig alle Systeme Vergleichen 2020\\
- Wir vergleichen die Frameworks nach folgenden Gesichtspunkten:\\
(Benutzerfreundlichkeit, Performanz/Varianz, Weiterentwicklung)
- We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).\\
- The comparison is performed on both real world data sets and synthetic graphs.
To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms. To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.\\
- Comparison of several state-of-the-art graph processing frameworks\\
\\
4. Wie ist das Paper aufgebaut\\
- Section 2: Relatet Work (Wie haben die Macher der Systeme die Systeme verglichen)\\
- Section 3: Preliminaries (Was muss man an technischen Details wissen)\\
- Section 4: Framework Overview (Welche Frameworks und was macht diese aus)\\
- Section 5: Evaluaten (Was wurde gemessen?, Wie wurde gemessen?, Was sind die Ergebnisse?)\\
- Section 6: Discussion (Warum sehen die Ergebnisse so aus, wie sie aussehen?)\\
- Section 7: Concluseion (Was lernen wir daraus?)\\

### 2020-09-19 plots/makeplots.py
removed all titles from plots

@@ -5,11 +5,11 @@ from plotFunctions import *
 
 
 distributed = False
-singleNode = True
+singleNode = False
 
 sssp = False
 bfs = False
-pr = True
+pr = False
 
 calcTimes = False
 execTimes = False
@@ -70,7 +70,7 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 	if sssp or overheads:
 		grouped_bar_plot(graphs, frameworks, overheadSSSP_singleNode, yLabel='Overhead time (s)', saveToFile="singleNodeSSSP_overheadTime.png", colors_=colors)
 
-		grouped_bar_plot(graphs, frameworks, overheadSSSPNormalized_singleNode, title='SSSP single node', yLabel='Overhead time (s) (normalized)', saveToFile="singleNodeSSSP_overheadTimeNormalized.png", yScale='linear', colors_=colors)
+		grouped_bar_plot(graphs, frameworks, overheadSSSPNormalized_singleNode, yLabel='Overhead time (s) (normalized)', saveToFile="singleNodeSSSP_overheadTimeNormalized.png", yScale='linear', colors_=colors)
 
 	#BFS
 	if bfs or overheads:
@@ -93,9 +93,9 @@ if galoisSpeedup:
 
 	line_plot(graphs, x, speedupGaloisBFS, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeBFSGaloisThreads.png", ylim=[0,11])
 
-	line_plot(graphs, x, speedupGaloisPRPush, title='Speedup PR Push', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPushGaloisThreads.png", ylim=[0.85,2.5])
+	line_plot(graphs, x, speedupGaloisPRPush, xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPushGaloisThreads.png", ylim=[0.85,2.5])
 
-	line_plot(graphs, x, speedupGaloisPRPull, title='Speedup PR Pull', yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPullGaloisThreads.png", ylim=[0.5,2.6])
+	line_plot(graphs, x, speedupGaloisPRPull, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPullGaloisThreads.png", ylim=[0.5,2.6])
 
 
 ## GALOIS HP THREAD COUNT
[…]

### 2020-09-19 documentation/paper/sections/results/galoisSpeedup.tex
results/discussion
speedup nachmal gegengelesen, discussions angefangen

@@
 \subsubsection{Behaviour of Galois}
-
-\todo{Section umbenennen?}
-
 \label{sec:galois_speedup}
-This section is dedicated to analyzing the speedup behaviour of Galois under the change to two parameters.
+This section is dedicated to analyzing the speedup behaviour of Galois under two parameters.
 First we change the thread count Galois is using.
@@
 Thus, we show the calculation time on any thread count normalized by the calculation time in the single-threaded environment.
-Beginning with SSSP, followed by BFS and last the comparison for both PR Push and Pull.
+Beginning with SSSP, followed by BFS and finally the comparison for both PR Push and Pull.
 
@@
 \label{tbl:ssspMeansVariances}
-\begin{tabular}{c@{\hskip 2\tabcolsep}rrc@{\hskip 2\tabcolsep}rr}
+\begin{tabular}{c
+r@{\tabskip 1 \tabcolsep}r
+r@{\tabskip 1 \tabcolsep}r}
 \toprule
-&\multicolumn{2}{c}{\!\!\!$\mu$}&\enspace&\multicolumn{2}{c}{\!\!$\sigma^2$}\\
-\cmidrule(r){2-3}\cmidrule(r){5-6}
-{\bf\#Threads}&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}&\enspace&\multicolumn{1}{c}{w/}&\multicolumn{1}{c}{w/o}\\\midrule
-\bf1 & 1.0 & 1.0 &\enspace& 0.0 & 0.0 \\
-\bf2 & 1.6 & 1.6 &\enspace& 0.2 & 0.1 \\
-\bf4 & 2.5 & 2.5 &\enspace& 1.1 & 0.9 \\
-\bf8 & 4.5 & 3.4 &\enspace& 5.3 & 2.1 \\
-\bf16 & 6.7 & 5.8 &\enspace& 13.0 & 7.3 \\
-\bf32 & 9.6 & 7.0 &\enspace& 38.3 & 10.8 \\
-\bf48 & 10.3 & 10.7 &\enspace& 41.4 & 31.4 \\
-\bf96 & 10.2 & 10.8 &\enspace& 38.1 & 29.9 \\
+&\multicolumn{2}{c}{\!\!\!$\mu$}&\multicolumn{2}{c}{\!\!$\sigma^2$}\\
+\cmidrule(r){2-3}\cmidrule(r){4-5}
+{\bf\#Threads}&w/&w/o&w/&w/o\\\midrule
+\bf1 & 1.0 & 1.0 & 0.0 & 0.0 \\
+\bf2 & 1.6 & 1.6 & 0.2 & 0.1 \\
+\bf4 & 2.5 & 2.5 & 1.1 & 0.9 \\
+\bf8 & 4.5 & 3.4 & 5.3 & 2.1 \\
+\bf16 & 6.7 & 5.8 & 13.0 & 7.3 \\
+\bf32 & 9.6 & 7.0 & 38.3 & 10.8 \\
+\bf48 & 10.3 & 10.7 & 41.4 & 31.4 \\
+\bf96 & 10.2 & 10.8 & 38.1 & 29.9 \\
 \bottomrule
@@
 
-Starting with SSSP, we see an algorithm that benefits from many available threads this can be seen in \autoref{fig:galoisSpeedupSSSP}.
+Starting with SSSP, an algorithm that greatly benefits from many available threads. The results can be seen in \autoref{fig:galoisSpeedupSSSP}.
 We first look at the speedups without hugepages, seen in \autoref{fig:galoisSpeedupSSSP_noHP}.
@@
 Speedup above 16 threads is always less than double the speedup of 16 threads.
-The maximum measured speedups are 10$\times$ (96 threads) for wikipedia, 17$\times$ (96 threads) for twitter, 11$\times$ (96 threads) for rMat27, 16$\times$ (48 threads) for friendster and 19$\times$ (40 threads) for rMat28.
-In some cases increasing thread counts even prolongues calculation time.
+The maximum measured speedups are between 10$\times$ (96 threads, wikipedia) and 19$\times$ (40 threads, rMat28).
+In some cases, increasing the thread count can even decrease the speedup again.
 For example calculation on rMat28 is actually slower with 48 or 96 threads compared to 40 threads.
@@
 Performance on flickr can not be sped up at all, with speedup on flickr being very close to 1 for 1 to 8 threads and between 0.7$\times$ to 0.9$\times$ from 16 to 96 threads.
-Orkut reaches maximum speedup of 1.6$\times$ at 16 threads. However on orkut, the speedup is always greater or equal to 1.
+Orkut reaches maximum speedup of 1.6$\times$ at 16 threads. But on orkut, the speedup is always greater or equal to 1.
 
@@
 Here, the overall results are similar to the findings without hugepages.
-\autoref{tbl:ssspMeansVariances} shows the mean speedups and variances over the different graphs.
+\autoref{tbl:ssspMeansVariances} shows the mean speedups and variances over the different graphs with and without hugepages.
 We see that the mean speedup is either very similar or slightly reduced by the hugepages.
-But in all cases, the variance is significantly smaller when using hugepages.
+But in all cases with hugepages, the variance is significantly smaller.
 This proves a slightly smaller but more reliable speedup with hugepages.
@@
 On the other hand, twitter reached a speedup of 17$\times$ without hugepages and only 13$\times$ with.
-
 While the hugepages did not neccessarily help improve the speedup value, they helped bring the different graphs together. 
@@
 But the speedup is much more reliable regardless of the graph topology. 
-\todo{Discussion ausarbeiten}
+In this regard, hugepages can be beneficial when many different graphs are in use, because it is more likely a speedup can be achieved. 
+If only a few graphs are used, activating hugepages can decrease mutlithreaded performance, if the graphs behave for example like twitter above.
 
@@
 So while speedups are possible, not even remotely to the same degree as on SSSP.
-This in turn is not intuitive, one would assume that these two algorithms perform similarly.
+This, in turn is not intuitive, one would assume these two algorithms to perform similarly.
 Both algorithms are iterative traversal algorithms with comparable computation and synchronization complexity.
-This behaviour extends even to the case with hugepages (cf.
+But the worse speedup behaviour extends even to the case with hugepages (cf.
 \autoref{fig:galoisSpeedupBFS_HP}).
-While the results are generally better, still not to the same degree as SSSP.
+While the results are generally better, still not to the same degree as on SSSP.
 With hugepages, BFS reaches a maximum speedup of 10.5$\times$ on wikipedia.
-The two graphs with largest speedup, namely wikipedia and friendster roughly follow a line with slope 0.125.
+The two graphs with largest speedup, namely wikipedia and friendster roughly follow a line with slope $1/8$.
 So with every 8 threads, the speedup is increased by about 1.
-Orkut follows the same line up to about 32 threads, slowly flatting off above that.
-The other graphs hardly reach a speedup of 3$\times$, even at 96 threads.
-\todo{Discussion ausarbeiten}
+Orkut follows the same line up to about 32 threads, slowly flattening off above that.
+On the other graphs, a speedup of 3$\times$ is hardly reached, even at 96 threads.
+
+\todo{Discussion ausarbeiten.. Was ist denn hier die Essenz, das ist ja beides irgendwie super schlecht.}
 
@@
 The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 1.3$\times$ at 96 threads.
-All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases and if so only by a small margin.
-Computation on Orkut and Twitter reaches a speedup maximum of 12\%\ and 5\%\ at 4 threads, while being less or equal to 1 in all other cases.
+All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases, and if so only by a small margin.
+Computation on Orkut and Twitter reaches a speedup maximum of 1.1$\times$ and 1.05$\times$ at 4 threads, while being less or equal to 1 in all other cases.
 Speedup on the wikipedia graph is never grater than one.
-Friendster and rMat27 can be sped up by 6.5\%\ or 10\%\ respectively on 8 threads.
+Friendster and rMat27 can be sped up to 1.07$\times$ or 1.1$\times$ respectively on 8 threads.
 
-Most of this changes with the activation of hugepages, our data can be seen in \autoref{fig:galoisSpeedupPRPull_HP}. 
+Most of this changes drastically with the activation of hugepages, our data can be seen in \autoref{fig:galoisSpeedupPRPull_HP}. 
 Here actually all graphs except flickr reach a speedup greater than 1.5$\times$. Remember, that 1.6$\times$ was the maximum possible speedup without hugepages.
@@
 Twitter, rMat27, friendster and rMat28 all reach a speedup of around 2.5$\times$ at 96 threads.
-Thus, we conclude that hugepages are very much recommended when using Galois PageRank Pull or a similar algorithm.
-Only with hugepages enabled do we observe a considerable increase in performance when using many threads, rather than a slowdown compared to the single threaded performance.
+
+We conclude hugepages to be very much recommended for any pull based algorithm similar to PageRank (or PR itself).
+Only with hugepages enabled do we observe a considerable increase in performance and thus a benefit of utilizing many threads.
+Without hugepages tough, having many threads available is in most cases slowing down performance.
+
 \todo{Discussion ausarbeiten}
@@
 The speedup results on the Push variant show odd behaviour in the Galois implementation, both with and without hugepages.
-Especially without hugepages, there is a significant performance loss on 4, 24 and 40 threads that is far from the expected behaviour.
+Especially without hugepages though, there is a significant performance loss on 4, 24 and 40 threads that is far from the expected behaviour.
 This is most visible in \autoref{fig:galoisSpeedupPRPush_noHP}, we validated the shown results with multiple samples each.
@@
 Actually however, the system does not reach a speedup of more than 1.04$\times$ on any graph, with only rMat27 actually reaching a value greater than 1.
-On all other graphs, using 24 threads is anywhere from 3\% (flickr) to 9\%\ (wikipedia) slower than when using just one thread.
-We intially assumed that this is due to the missing hugepages, Galois is recommended to run with.
+On all other graphs, using 24 threads is anywhere from 3\% (flickr) to 9\%\ (wikipedia) slower than using just one thread.
+We intially assumed that this is due to the missing hugepages, Galois is recommended to be run with.
 This seems to be only partly true, since the results with hugepages are generally better, but still show some of this behaviour (cf.
 \autoref{fig:galoisSpeedupPRPush_HP}).
-There is only one significant anomaly at 32 threads, while the points where we observed the performance drop before look more \emph{reasonable}\todo{das ist nicht das wort, das ich suche..}.
+There is only one significant anomaly at 32 threads, while the points where we observed the performance drop without hugepages look more \emph{reasonable}\todo{das ist nicht das wort das ich suche..}.
 At 32 threads, speedup for 5 of 7 graphs drops to a value between 1.4$\times$ and 1.5$\times$.
-Those five graphs reach a speedup larger than 1.8$\times$ both at 16 and 48 threads (i.e.
+Yet those five graphs reach a speedup larger than 1.8$\times$ both at 16 and 48 threads (i.e.
 immediately before and after 32).
 
-If we disregard these anomalies however, we see a significant performance increase with hugepages compared to without them.
+Nevertheless, if we disregard these anomalies, we see a significant performance increase with hugepages compared to without them.
 Actually, all graphs show an improvement in their maximum reached spedups.
-The largest improvements are from 1.8$\times$ to 2.2$\times$ for flickr, 1.5$\times$ to 2.1$\times$ for rMat27, 1.5$\times$ to 2.4$\times$ for friendster and 1.4$\times$ to 2.2$\times$ for rMat28, just by switching hugepages on.
-Interestingly, with hugepages all graphs reach a speedup with 4 threads that is very similar, if not equal to the speedup at 96 threads.
+The largest improvements are from 1.5$\times$ to 2.1$\times$ for rMat27, 1.5$\times$ to 2.4$\times$ for friendster and 1.4$\times$ to 2.2$\times$ for rMat28, just by switching hugepages on.
+Interestingly, with hugepages all graphs reach a speedup at 4 threads that is very similar, if not equal to the speedup at 96 threads.
 The largest discrepancy between the two thread counts is rMat27, with a performance uplift from 1.96$\times$ to 2.14$\times$ at 4 and 96 threads.
@@
 Using hugepages shows very effective, while it does not completely remove the anomalies we observed, it definitely increases performance by a considerable margin.
+
+But, at least on our testing data, more than 4 threads did not change the performance significantly.
 \todo{Discussion ausarbeiten}

### 2020-09-20 documentation/paper/sections/discussion.tex
discussion

%!TEX root=../main.tex

\section{Discussion}
\todo{unfertig}

There are two main cases that should be regarded in a comparison of the frameworks. One, the case of individual calculations on a graph, i.e. for each calculation, the graph has to be loaded. This is called the \emph{research} case. For research, frameworks with small execution times and small overhead are preferred. Case two on the other hand is the case of a running system, that performs multiple calculations on a single graph, without the need of reloading graph data with every calculation. This is the \emph{production} case.
In production, frameworks with short calculation times should be preferred because the overhead time is only spent once on startup.

Orthogonal to this, the graph size and topology of the system is a deciding factor. Large graphs often require a distributed cluster, because the graph is simply too large to fit in the RAM of just one machine. We observed this for example with Giraph on multiple cases.
Furthermore, some environments require a distributed framework setup regardless of any other factors.

%### SSSP
%### BFS
For algorithms that perform a traversal through the graph, with a concentrated and relativly small amount of active vertices, our results for SSSP and BFS are most relevant.

%### general
For graphs that are not too large for single node RAM and when performing research work, Galois as the fastest single-node framework is recommended.
It provides excellent performance, being orders of magnitude faster than the competing frameworks.
In some cases, especially on PageRank-like applications, tests to determine whether a push- or pull-style implementation is most efficient might be necessary.
Because hugepages provide an improvement in performance when running calculations on many different graphs, using them in research systems is beneficial as well.

%## production
On production systems we expect it to be possible to determine what configuration is best, before putting the system in service.
There are multiple possibilities to be accounted for

hp on/off
thread count.. weniger wichtig

%### PR
There is still some testing to be performed, in order to find the reason for the performance anomalies we encountered.

%### Distributed
A distributed setup can have many advantages,

first, there is some data security, because many distributed file systems also support automatic data replication etc.

There is an argument to be made for systems like Hadoop.
A user defining an application does not need to know the topology of the system that runs the application later on.
Hadoop transparently splits the input data and distributes work among the worker nodes.
And Hadoop automatically handles node failures, a node failure does not immediately result in loss of data or a faulty computation result.
This is especially important on very large clusters, that can not easily be monitored by humans.
Furthermore, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
This shows the hadoop infrastructure to be working well. 
The Giraph/hadoop system is recommended for very large graphs on very large computation clusters.
Only a system like hadoop can reasonably be used on such large clusters.
Any of the other framewors will just result in large administration overhead due to node failures or other problems of distributed systems in general.

Especially for frameworks like Giraph, that require large amounts of memory, the distributed scenario allows them to complete their computations.

### 2020-09-20 documentation/paper/sections/testingMethods.tex
some work on discussion/comparison

@@
 \begin{itemize}
-	\item Galois supports all of our tested algorithms, with both a Push and a Pull variant for PageRank available.
+	\item Ligra supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delata Variant.
+	\item Polymer supports the same algoritms as ligra.
+	\item Gemini supports all of our tested algorithms and there are no setting options or specifications which implementations for the algorithms are used.
+	\item Galois supports all of our tested algorithms too, with both a Push and a Pull variant for PageRank available.
 In the distributed scenario, there are Push and Pull versions for SSSP and BFS available as well.
-	\item Gemini \todo{}
+It also supports multiple implementations of the shared-memory allgorithms.
+The default implementation of SSSP is deltaTile.
+A lot of setting options are avilable as well, but we're gone with the defaults.
 	\item Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
@@
 For PageRank the supplied implementation is used.
-	\item Ligra \todo{}
-	\item Polymer supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
 \end{itemize}

### 2020-09-20 documentation/paper/sections/overview/overview.tex
most introductions done

@@
 \section{Framework Overview}
-The following paragraphs is a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
+The following paragraphs are a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
 All of the described frameworks are part of our testing.
 \todo
-
-\subsection{Galois and Gluon}
-\input{sections/overview/galois}
-
-\subsection{Gemini}
-\input{sections/overview/gemini}
-
-\subsection{Giraph}
-\input{sections/overview/giraph}
 
@@
 
+\subsection{Gemini}
+\input{sections/overview/gemini}
+
+\subsection{Galois and Gluon}
+\input{sections/overview/galois}
+
+\subsection{Giraph}
+\input{sections/overview/giraph}
+
 \input{sections/graphFormats}

### 2020-09-20 documentation/paper/sections/preliminiaries.tex
most introductions done
various imporvements

@@
 \section{Preliminaries}
-\todo{Section beschreiben}
 This section briefly explain the concepts and applications necessary, but not directly related to our work.
@@
 The output is a set of vertices that are reachable from the start vertex.
-In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.
+In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists and otherwise false.
 It is called Breadth-First search because the algorithm searches in a path length-based way.
@@
 \subsection{Push and Pull Variants}
-Many parallel graph algorithms, as well as the three we consider here, are implemented by iterating over vertices \cite{tao}.
-The respective vertex, which is considered in an iteration and the operator is applied to, is also called \emph{active}.
-It's also possible to iterate over edges and consider them as \emph{active}, but in the following, whithout loss of generality, we will use the term \emph{active vertex}.
-This operator considers always only a local so called \emph{neighborhood} to the \emph{active} vertex.
+Many parallel graph algorithms, including the three we consider here, are implemented by iterating over vertices \cite{tao}.
+The vertex, to which the operator is applied, is called \emph{active}.
+It's possible to iterate over edges and consider them as \emph{active}, but in the following, whithout loss of generality, we will use the term \emph{active vertex}.
+This operator considers only a so called \emph{neighborhood} to the \emph{active} vertex.
 To this \emph{neighborhood} belong only vertices, which are in the direct surrounding of the \emph{active} vertex.
@@
 Synchronization is realized through barriers for some or all processes.
-BSP algorithms are performed in a series of global supersteps. These consist of three steps, beginng with the processors performing local computations concurrently.
+BSP algorithms are performed in a series of global supersteps. These consist of three steps, beginning with the processors performing local computations concurrently.
 This step can overlap with the second, the communication between components. Processes can exchange information to access remote data.

### 2020-09-20 documentation/paper/sections/results/galoisSpeedup.tex
galois time comparison init

@@
 This section is dedicated to analyzing the speedup behaviour of Galois under two parameters.
-First we change the thread count Galois is using.
-Second, a comparison between Galois using hugepages and Galois without hugepages is made.
-We compare the \emph{calculation speedups} of Galois on the different graphs.
+First we compare the calculation and execution times with and without hugepages on the maximum thread count.
+Second we change the thread count Galois is using, we compare the \emph{calculation speedups} of Galois on the different graphs.
 Thus, we show the calculation time on any thread count normalized by the calculation time in the single-threaded environment.
-Beginning with SSSP, followed by BFS and finally the comparison for both PR Push and Pull.
+This is done along with a comparison between Galois using hugepages and Galois without hugepages.
+We begin with the time comparisons, followed by the speedups. These begin with SSSP, second by BFS and finally the comparison for both PR Push and Pull.
+
+\paragraph{Time Comparison for Hugepages}
+\begin{table}
+	\caption{Times With and Without Hugepages}
+	\label{tbl:timeComparison_HPnoHP}
+	\renewcommand{\arraystretch}{1.2}
+	\centering
+	\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
+		\toprule
+		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
+		\cmidrule{3-4}\cmidrule{5-6}
+		&\bf Graph&w/o&w/&w/o&w/\\
+		\midrule
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
+		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5 \\
+		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1 \\
+		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1 \\
+		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4 \\
+		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5 \\
+		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1 \\
+		\midrule
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}
+		& flickr & 0.01 & \bf 0.00 & 0.6 & \bf 0.3 \\
+		& orkut & 0.05 & \bf 0.01 & 0.9 & \bf 0.6 \\
+		& wikipedia & 0.52 & \bf 0.12 & 2.4 & \bf 1.4 \\
+		& twitter & 7.52 & \bf 3.08 & 14.2 & \bf 6.9 \\
+		& rMat27 & 5.62 & \bf 4.79 & 14.6 & \bf 9.2 \\
+		& friendster & 3.78 & \bf 1.67 & 12.8 & \bf 7.2 \\
+		& rMat28 & 13.14 & \bf 10.25 & 33.1 & \bf 19.1 \\
+		\midrule
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Push}}
+		& flickr & - & - & 0.6 & \bf 0.3 \\
+		& orkut & - & - & 2.1 & \bf 1.2 \\
+		& wikipedia & - & - & 13.1 & \bf 11.5 \\
+		& twitter & - & - & \bf 74.9 & 201.3 \\
+		& rMat27 & - & - & 239.7 & \bf 178.1 \\
+		& friendster & - & - & 54.2 & \bf 44.4 \\
+		& rMat28 & - & - & 689.3 & \bf 340.1 \\
+		\midrule
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
+		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6 \\
+		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4 \\
+		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3 \\
+		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1 \\
+		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1 \\
+		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4 \\
+		\bottomrule
+		\multicolumn{6}{l}{(-) Result is less or equal to 1ms}
+	\end{tabular}
+\end{table}
 
@@
 \cmidrule(r){2-3}\cmidrule(r){4-5}
-{\bf\#Threads}&w/&w/o&w/&w/o\\\midrule
+{\bf\#Threads}&w/o&w/&w/o&w/\\\midrule
 \bf1 & 1.0 & 1.0 & 0.0 & 0.0 \\
@@
 
-\todo{Discussion ausarbeiten.. Was ist denn hier die Essenz, das ist ja beides irgendwie super schlecht.}
+\begin{table}
+	\caption{Maxium Reached Speedups With And Without Hugepages}
+	\label{tbl:maxReachedSpeedupSSSPBFS}
+	\renewcommand{\arraystretch}{1.3}
+	\centering
+	\begin{tabular}{cr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
+		\toprule
+		&\multicolumn{4}{c}{\bf Max Speedup}\\
+		&\multicolumn{2}{c}{\bf SSSP}&\multicolumn{2}{c}{\bf BFS}\\
+		\cmidrule{2-3}\cmidrule{4-5}
+		\bf Graph & w/o & w/&w/o&w/\\
+		\midrule
+		flickr & \bf 1.1 &  1.0 & 1.1 & \bf 1.2 \\
+		orkut & 1.6 & \bf 4.3 & 1.6 & \bf 5.7 \\
+		wikipedia & 10.3 & \bf 15.8 & 4.9 & \bf 10.5 \\
+		twitter & \bf 16.8 &  13.0 & 2.5 & \bf 2.7 \\
+		rMat27 & 14.0 & \bf 15.8 & \bf 4.2 & 2.9 \\
+		friendster & \bf 15.9 &  14.2 & 5.6 & \bf 8.3 \\
+		rMat28 & \bf 19.5 &  14.9 & \bf 4.7 & 2.8 \\
+		\bottomrule
+	\end{tabular}
+\end{table}
+One might assume the two algorithms SSSP and BFS to behave very similarly, yet they produce very different results.
+With SSSP, we observed a decrease in the variance of the speedups, while the average speedup decreased or stayed the same. This provided a more reliable speedup. 
+But hugepages did not necessarily increase the maximum possible speedup. For 4 of 7 graphs the larger speedup value is reached without hugepages (cf. \autoref{tbl:maxReachedSpeedupSSSPBFS}).
+However exactly this is happening with BFS, we measured a larger speedup with hugepages on all real-world graphs (i.e. 5 of 7).
+But as can also be seen in \autoref{fig:galoisSpeedupBFS_HP}, the variance of the speedup increases, making the possible speedup very dependent on the graph.
+
+While it is most likely beneficial to use hugepages in both cases, it is due to different reasons.
+On BFS, the maximum observed speedup increases on almost every graph with hugepages compared to without. So using hugepages should generally improve multithreaded performance on most graphs.
+SSSP on the other hand sometimes performes better without
+
+In applications similar to SSSP or BFS, the benefit of many threads can be on the one hand be increased, if the graph allows it.
 
@@
 Twitter, rMat27, friendster and rMat28 all reach a speedup of around 2.5$\times$ at 96 threads.
-
-We conclude hugepages to be very much recommended for any pull based algorithm similar to PageRank (or PR itself).
 Only with hugepages enabled do we observe a considerable increase in performance and thus a benefit of utilizing many threads.
-Without hugepages tough, having many threads available is in most cases slowing down performance.
-
-\todo{Discussion ausarbeiten}
+Without hugepages though, having many threads available is in most cases slowing down the performance.
+Hence, we see a great benefit in enabling hugepages, without any drawbacks resulting from them on this algorithm.
 
@@
 
-Using hugepages shows very effective, while it does not completely remove the anomalies we observed, it definitely increases performance by a considerable margin.
-
-But, at least on our testing data, more than 4 threads did not change the performance significantly.
-\todo{Discussion ausarbeiten}
+Using hugepages is again very effective and increases performance by a considerable margin.
+Though, they do not completely remove the performance drops on some thread counts we observed without hugepages, the abnormal behaviour is at least less severe.
+But even with hugepages, a thread count larger than 4 did often not improve performance at all.
+This is due to the great synchonization overhead of a push based implementation.
+Generally, the push based application requires many more write operations, with each requiring synchronization.
+So, on push based applications a large thread pool is not helpful, rather should the framework itself be optimized. In this optimization, hugepages are a great factor for Galois.

### 2020-09-20 plots/makedata.py

@@ -599,11 +599,16 @@ if False:
 
 
 
+
+
+
 ###SPEEDUP GALOIS HP
 speedupGaloisSSSP_HP = {}
 speedupGaloisBFS_HP = {}
 speedupGaloisPRPush_HP = {}
 speedupGaloisPRPull_HP = {}
+
+
 xHP = []
 for g in graphs:
     speedupGaloisSSSP_HP[g] = {}
@@ -727,3 +732,140 @@ if False:
 
     from plotFunctions import line_plot
     line_plot(["Means no HP", "Means HP", "Variances no HP", "Variances HP"], xHP, valuesForMeanSpeedup, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="meanSpeedup.png")
+
+
+if False:
+    for g in graphs:
+        maxSSSP_noHP = round(max([speedupGaloisSSSP[g][i] for i in x]) * 10) / 10
+        maxSSSP_HP = round(max([speedupGaloisSSSP_HP[g][i] for i in xHP]) * 10) / 10
+
+        maxBFS_noHP = round(max([speedupGaloisBFS[g][i] for i in x]) * 10) / 10
+        maxBFS_HP = round(max([speedupGaloisBFS_HP[g][i] for i in xHP]) * 10) / 10
+
+        print(g, end=" & ")
+
+        if maxSSSP_HP > maxSSSP_noHP:
+             print(maxSSSP_noHP, "& \\bf", maxSSSP_HP, end=" & ")
+        else:
+             print("\\bf", maxSSSP_noHP, "& ", maxSSSP_HP, end=" & ")
+        if maxBFS_HP > maxBFS_noHP:
+            print(maxBFS_noHP, "& \\bf", maxBFS_HP, "\\\\")
+        else:
+            print("\\bf", maxBFS_noHP, "&", maxBFS_HP, "\\\\")
+
+
+
+
+
+
+
+
+
+#### TIMES GALOIS HUGEPAGES
+calcTimeSSSP_HP = []
+calcTimeBFS_HP = []
+calcTimePRPull_HP = []
+calcTimePRPush_HP = []
+
+execTimeSSSP_HP = []
+execTimeBFS_HP = []
+execTimePRPull_HP = []
+execTimePRPush_HP = []
+
+for g in graphs:
+    for i in range(len(graph)):
+        if not "galois" in algo[i]:
+            continue
+        if not "-hp-" in algo[i]:
+            continue
+        if not "96thread" in algo[i]:
+            continue
+
[…]

### 2020-09-20 documentation/paper/sections/results/galoisSpeedup.tex
galois time comparison init

@@
 \begin{table}
-	\caption{Times With and Without Hugepages}
+	\caption{Times for 96 Threads With and Without Hugepages}
 	\label{tbl:timeComparison_HPnoHP}
@@
 \end{table}
+When comparing the performance with and without hugepages, the results really speak for themselves they can be seen in \autoref{tbl:timeComparison_HPnoHP}.
+It is obvious that enabling hugepages greatly improves performance in almost every case.
+This means that on every algorithm and almost all graphs, calculation time and execution tame can be reduced significantly with just the use of hugepages.
+The only exception is twitter on both PageRank implementations, where the execution time is smaller without hugepages (74.9s vs 201.3s for Push and 8.7s vs 9.3s for Pull with and without hugepages respectively).
+In all other cases (that's 26 of 28 test cases), the hugepage-version is significantly faster.
+The execution time is approximately halved on SSSP and BFS. And on PageRank, the execution time is reduced to 35\% to 87\% of the time without hugepages.
 

### 2020-09-21 documentation/paper/sections/discussion.tex
galois hugepages zeug.. und ein bisschen discussion

@@
 It provides excellent performance, being orders of magnitude faster than the competing frameworks.
-In some cases, especially on PageRank-like applications, tests to determine whether a push- or pull-style implementation is most efficient might be necessary.
-Because hugepages provide an improvement in performance when running calculations on many different graphs, using them in research systems is beneficial as well.
+And when it is possible to use hugepages, these improve performance even further, sometimes up to a factor of 2 faster compared to Galois without hugepages.
+Also, we have shown that most algorithms are able to utilize many threads in favour of much smaller calculation times.
+Graph-traversal algorithms like BFS or SSSP shown here are examples for such algorithms.
+In some cases, especially on PageRank-like applications, it might be necessary to perform some tests whether a push- or pull-style implementation is most efficient.
+However, especially when using push-style implementations, a large thread count is most likely not going to significantly improve performance over single threaded performance.
+The reason for this is the synchronization required for push applications.
+This is true not only for Galois but any framework.
 
@@
 There are multiple possibilities to be accounted for
-
-hp on/off
-thread count.. weniger wichtig
 

### 2020-09-21 documentation/paper/sections/results/galoisSpeedup.tex
galois hugepages zeug.. und ein bisschen discussion

@@
 
+Keep in mind, that we have shown the results of 96 threads in this section. The following sections compare the behavoiur of Galois on different numbers of threads.
+
 \paragraph{Single-Source Shortest-Paths}
@@
 While the hugepages did not neccessarily help improve the speedup value, they helped bring the different graphs together.
-Thus, the overall speedup benefit might not be as good as without hugepages. 
-But the speedup is much more reliable regardless of the graph topology. 
-In this regard, hugepages can be beneficial when many different graphs are in use, because it is more likely a speedup can be achieved. 
-If only a few graphs are used, activating hugepages can decrease mutlithreaded performance, if the graphs behave for example like twitter above.
 
@@
 
-While it is most likely beneficial to use hugepages in both cases, it is due to different reasons.
-On BFS, the maximum observed speedup increases on almost every graph with hugepages compared to without. So using hugepages should generally improve multithreaded performance on most graphs.
-SSSP on the other hand sometimes performes better without
+%Das ist irgendwie quatsch.
+% While it is beneficial to use hugepages on both algorithms, simply due to the general improvement in the needed times, the hugepages bring different multithreaded behaviour.
+% On BFS, the maximum observed speedup increases on almost every graph with hugepages compared to without. So with hugepages
 
-In applications similar to SSSP or BFS, the benefit of many threads can be on the one hand be increased, if the graph allows it.
+% In applications similar to SSSP or BFS, the benefit of many threads can be on the one hand be increased, if the graph allows it.
 
@@
 Only with hugepages enabled do we observe a considerable increase in performance and thus a benefit of utilizing many threads.
-Without hugepages though, having many threads available is in most cases slowing down the performance.
-Hence, we see a great benefit in enabling hugepages, without any drawbacks resulting from them on this algorithm.
+Without hugepages though, having many threads available is in most cases slowing down the performance. Hence, enabling hugepages is very much recommended to be able to use a machine with many cores.
 
@@
 \autoref{fig:galoisSpeedupPRPush_HP}).
-There is only one significant anomaly at 32 threads, while the points where we observed the performance drop without hugepages look more \emph{reasonable}\todo{das ist nicht das wort das ich suche..}.
+There is only one significant anomaly at 32 threads, while the points where we observed the performance drop without hugepages look normal.
 At 32 threads, speedup for 5 of 7 graphs drops to a value between 1.4$\times$ and 1.5$\times$.
@@
 
-Using hugepages is again very effective and increases performance by a considerable margin.
+Using hugepages is very effective and yet again increases performance by a considerable margin.
 Though, they do not completely remove the performance drops on some thread counts we observed without hugepages, the abnormal behaviour is at least less severe.

### 2020-09-21 documentation/paper/main.tex
discussion
most introductions done
hugepages section

@@
 \section{Evaluation}
-\todo{Einleitung}
+The main part of this paper is the evaluation, which consists of two parts.
+The first is a description and explanation of our testing environment and methodology.
+Followed by a lengthy presentation and discussion of the test results of the frameworks.
 
@@
 
-\bibitem{newman2010networks}
+\bibitem{Giraph}
 \BIBentryALTinterwordspacing
-M.~Newman, \emph{Networks: An Introduction}.\hskip 1em plus 0.5em minus
-  0.4em\relax OUP Oxford, 2010. [Online]. Available:
-  \url{https://books.google.de/books?id=-DgTDAAAQBAJ}
+{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
+  Available: \url{https://giraph.apache.org}
 \BIBentrySTDinterwordspacing
@@
 
+\bibitem{Galois}
+\BIBentryALTinterwordspacing
+D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
+  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
+  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
+  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
+  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
+\BIBentrySTDinterwordspacing
+
 \bibitem{Ligra}
@@
   \url{https://doi.org/10.1145/2442516.2442530}
+\BIBentrySTDinterwordspacing
+
+\bibitem{newman2010networks}
+\BIBentryALTinterwordspacing
+M.~Newman, \emph{Networks: An Introduction}.\hskip 1em plus 0.5em minus
+  0.4em\relax OUP Oxford, 2010. [Online]. Available:
+  \url{https://books.google.de/books?id=-DgTDAAAQBAJ}
 \BIBentrySTDinterwordspacing
@@
 
-\bibitem{Giraph}
-\BIBentryALTinterwordspacing
-{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
-  Available: \url{https://giraph.apache.org}
-\BIBentrySTDinterwordspacing
-
 \bibitem{graphx}
@@
 
+\bibitem{virtual_memory}
+\BIBentryALTinterwordspacing
+E.~Abrossimov, M.~Rozier, and M.~Shapiro, ``Generic virtual memory management
+  for operating system kernels,'' in \emph{Proceedings of the Twelfth ACM
+  Symposium on Operating Systems Principles}, ser. SOSP '89.\hskip 1em plus
+  0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
+  Machinery, 1989, p. 123–136. [Online]. Available:
+  \url{https://doi.org/10.1145/74850.74863}
+\BIBentrySTDinterwordspacing
+
 \bibitem{hugepages}
@@
 
-\bibitem{origGalois}
-\BIBentryALTinterwordspacing
-M.~Kulkarni, K.~Pingali, B.~Walter, G.~Ramanarayanan, K.~Bala, and L.~P. Chew,
-  ``Optimistic parallelism requires abstractions,'' in \emph{Proceedings of the
-  28th ACM SIGPLAN Conference on Programming Language Design and
-  Implementation}, ser. PLDI '07.\hskip 1em plus 0.5em minus 0.4em\relax New
-  York, NY, USA: Association for Computing Machinery, 2007, p. 211–222.
-  [Online]. Available: \url{https://doi.org/10.1145/1250734.1250759}
-\BIBentrySTDinterwordspacing
-
-\bibitem{Galois}
-\BIBentryALTinterwordspacing
-D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
-  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
-  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
-  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
-  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
-\BIBentrySTDinterwordspacing
-
-\bibitem{vertGalois}
-\BIBentryALTinterwordspacing
-R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
-  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
-  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
-  Conference on Programming Language Design and Implementation}, ser. PLDI
-  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
-  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
-  \url{https://doi.org/10.1145/3192366.3192404}
-\BIBentrySTDinterwordspacing
+\bibitem{superpages}
+M.~Gorman and P.~Healy, ``Performance characteristics of explicit superpage
+  support,'' in \emph{Computer Architecture}, A.~L. Varbanescu, A.~Molnos, and
+  R.~van Nieuwpoort, Eds.\hskip 1em plus 0.5em minus 0.4em\relax Berlin,
+  Heidelberg: Springer Berlin Heidelberg, 2012, pp. 293--310.
 
@@
   \url{https://www.usenix.org/conference/hotos15/workshop-program/presentation/mcsherry}
+\BIBentrySTDinterwordspacing
+
+\bibitem{origGalois}
+\BIBentryALTinterwordspacing
+M.~Kulkarni, K.~Pingali, B.~Walter, G.~Ramanarayanan, K.~Bala, and L.~P. Chew,
+  ``Optimistic parallelism requires abstractions,'' in \emph{Proceedings of the
+  28th ACM SIGPLAN Conference on Programming Language Design and
+  Implementation}, ser. PLDI '07.\hskip 1em plus 0.5em minus 0.4em\relax New
+  York, NY, USA: Association for Computing Machinery, 2007, p. 211–222.
+  [Online]. Available: \url{https://doi.org/10.1145/1250734.1250759}
+\BIBentrySTDinterwordspacing
+
+\bibitem{vertGalois}
+\BIBentryALTinterwordspacing
+R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
+  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
+  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
+  Conference on Programming Language Design and Implementation}, ser. PLDI
+  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
+  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
+  \url{https://doi.org/10.1145/3192366.3192404}
 \BIBentrySTDinterwordspacing

### 2020-09-21 documentation/paper/sections/preliminiaries.tex
hugepages section

@@
 \subsection{Hugepages}
-\todo{Section schreiben}
-\cite{hugepages}
+Most systems in use today use a so-called virtual memory management \cite{virtual_memory}.
+It is implemented in the kernel and represents an abstraction between the memory devices of a machine and the individual programs.
+Each program gets its own virtual leading address space.
+This simplifies the implementation of applications considerably and also increases security, since each program can only access its own virtual address space.
+The address areas are organized in so called pages, which in most cases are 4 KiB in size.
+In the Translation lookaside buffer (TBL) the most recent translations of virtual memory to physical memory are cached.
+
+Translating from the virtual to the physical address space significantly reduces the performance of today's data processing systems, since the size of the RAM grows much faster than the size of the TBL \cite{hugepages, superpages}.
+To counteract this problem, hugepages were developed, which are usually several mibibibytes in size.
+This minimizes the CPU time needed for table lookups, because there are much less TBL misses.
+This speedup is especially noticeable in applications that are very memory intensive, like graph processing systems.

### 2020-09-21 documentation/paper/sections/discussion.tex
discussion

@@
 
-%### SSSP
-%### BFS
 For algorithms that perform a traversal through the graph, with a concentrated and relativly small amount of active vertices, our results for SSSP and BFS are most relevant.
+If the problem is however of a nature, where many vertices are active, thus many vertices, widely spread across the graph performe some kind of calculation, the results for PR are most applicable.
+Further, some considerations whether a push or pull-style implemenation is better should be made.
+
+Generally, we expect pull-style implementations to be faster on a single-node, we observed this for example for PR with Galois.
+On a distributed cluster though, the synchronization time is not as severely impacting performance because communication over the network is needed, which is slower than local synchonization in any case.
+We already know push implementations to be more efficient on problems with only few active vertices or where the affected push-areas are not overlapping. 
+Because SSSP and BFS often have few active vertices with few overlapping vertices in the neighborhood, synchronization is often not impacting performance as much, making push the better implementation for those algorithms.
+PR on the other hand is expected to perform much better with pull implementations, simply due to less synchronization needed.
+Hence, a good guideline is pull-style on single-node systems or PR-like applications and push-style for SSSP and BFS, especially on distributed systems.
+
+While we often observed Galois Push to be slightly faster than the pull implementations on distributed systems, we expect pull implementations 
+Our data for PR on a single-node proves pull implemenations to be faster.
+
+%Furthermore, in every application case, a pull-style implementation is most likely to perform better than an equivalent push algorithm. This is simply due to the
 
@@
 %### Distributed
-A distributed setup can have many advantages,
+If the graph data becomes too large for one node or the computation cluster is distributed anyways, there are multiple possibilities.
 
-first, there is some data security, because many distributed file systems also support automatic data replication etc.
-
-There is an argument to be made for systems like Hadoop.
+First, there is an argument to be made for systems like Hadoop.
 A user defining an application does not need to know the topology of the system that runs the application later on.
@@
 Furthermore, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
-This shows the hadoop infrastructure to be working well. 
-The Giraph/hadoop system is recommended for very large graphs on very large computation clusters.
+This shows the hadoop infrastructure to be working well and is exactly what the distributed system should be used for anyways. 
 Only a system like hadoop can reasonably be used on such large clusters.
 Any of the other framewors will just result in large administration overhead due to node failures or other problems of distributed systems in general.
+Furthermore, the calculation times of Giraph proved to be very fast in comparison to the other distributed frameworks, with Giraph often being one of the fastest framework in computation time.
+Very short calculation times along with automatic handling of the computation cluster make Giraph using Hadoop a good choice for production systems.
+Pregel is a similarly fault-tolerant, scalable system but it is contrary to Giraph closed-source.
 
-Especially for frameworks like Giraph, that require large amounts of memory, the distributed scenario allows them to complete their computations.
+On research systems however, single computations are performed and node failures are not as common, because the system is most likely not in-use constantly.
+Hence, Hadoop handling all this in the background is a nice-to-have but most likely, a plain faster framework would be preferred.
+Furthermore, administration is most likely needed frequently because system reconfiguration is happening often?
+Because the system is most likely reconfigured frequently.
+
+Here, the fastest framework really depends on the use case. Graph traversal algorithms prefer
+
+Here, Galois is the fastest distributed framework for SSSP and BFS, while Gemini outperforms Galois on PR.
+
+%A distributed setup can have many advantages,
+
+%Especially for frameworks like Giraph, that require large amounts of memory, the distributed scenario allows them to complete their computations.
+
+%first, there is some data security, because many distributed file systems also support automatic data replication etc.

### 2020-09-21 documentation/paper/sections/discussion.tex
todos in results und discussion

@@
 PR on the other hand is expected to perform much better with pull implementations, simply due to less synchronization needed.
-Hence, a good guideline is pull-style on single-node systems or PR-like applications and push-style for SSSP and BFS, especially on distributed systems.
-
-While we often observed Galois Push to be slightly faster than the pull implementations on distributed systems, we expect pull implementations 
-Our data for PR on a single-node proves pull implemenations to be faster.
-
-%Furthermore, in every application case, a pull-style implementation is most likely to perform better than an equivalent push algorithm. This is simply due to the
+Hence, a good guideline is pull-style on single-node systems with PR-like applications and push-style for SSSP and BFS, especially on distributed systems. 
+Our data backs this up, we observed single-node PageRank to be fastest with Galois Pull, while Galois Push was the fastest application on distributed systems.
+In general however, some tests determining which implementation is faster can be beneficial.
 
@@
 There are multiple possibilities to be accounted for
-
-%### PR
-There is still some testing to be performed, in order to find the reason for the performance anomalies we encountered.
+\todo{unfertig}
 
@@
 Because the system is most likely reconfigured frequently.
-
+\todo{unfertig}
 Here, the fastest framework really depends on the use case. Graph traversal algorithms prefer

### 2020-09-21 documentation/paper/sections/overview/ligra.tex
Ligra writeup

@@
 Ligra is a lightweight graph processing framework for shared memory machines \cite{Ligra}. 
-It offers a vertex-centric programming interface which can be used to apply a function to each vertex or outgoing edge of a set of vertices in parallel.
-While doing so the framework can generate a set of active vertices for the next iteration.
-This abstraction makes it well suited for writing graph traversal algorithms.
+It offers a vertex-centric programming interface which consists of two routines. 
+The \texttt{VertexMap} applies a user defined function to each vertex and \texttt{EdgeMap} applies a user defined function to each outgoing edge of a set of vertices. 
+Active vertices can be represented by a \texttt{VertexSubset}. 
+This data structure can be passed to both of the routines and is maintained by them. 
+That makes Ligra well suited for expressing graph traversal algorithms.
 
-When mapping over edges Ligra optimizes algorithms by switching between a push-based and a pull-based approach based on the size of the set.
-When the size of the set is above a threshold a pull approach is used.
-If the threshold is not reached a push approach is used.
+When \texttt{EdgeMap} is used Ligra decides between a push and a pull style execution.
+The decisive factor is the number of active vertices and their outdegree.
+When this number is small a push style execution is used, otherwise a pull style execution is used.
+The default threshold is set to one twentieth of all edges.
+This hybrid approach can increase performance significantly \cite{hybridBFS}.
+When push and pull mode changes also the representation of the \texttt{VertexSubset} is changed.
+A push style execution corresponds to a sparse representation as array of the active vertex IDs, while a pull style execution corresponds to a dense representation as bitmap.

### 2020-09-21 documentation/paper/sections/discussion.tex
discussion

@@
 This is the \emph{production} case.
-In production, frameworks with short calculation times should be preferred because the overhead time is only spent once on startup.
+In production, frameworks with short calculation times should be preferred because the overhead time is only spent once on startup and amortizes quickly.
 
@@
 If the problem is however of a nature, where many vertices are active, thus many vertices, widely spread across the graph performe some kind of calculation, the results for PR are most applicable.
-Further, some considerations whether a push or pull-style implemenation is better should be made.
+Further, some considerations whether a push or pull-style implemenation is more efficient should be made.
+Generally, we expect pull-style implementations to be faster on a single-node, like for example PR with Galois.
+On a distributed cluster though, the synchronization time is not as severely impacting performance because communication over the network is needed, which is slower than local synchonization in any case.
 
-Generally, we expect pull-style implementations to be faster on a single-node, we observed this for example for PR with Galois.
-On a distributed cluster though, the synchronization time is not as severely impacting performance because communication over the network is needed, which is slower than local synchonization in any case.
 We already know push implementations to be more efficient on problems with only few active vertices or where the affected push-areas are not overlapping. 
-Because SSSP and BFS often have few active vertices with few overlapping vertices in the neighborhood, synchronization is often not impacting performance as much, making push the better implementation for those algorithms.
-PR on the other hand is expected to perform much better with pull implementations, simply due to less synchronization needed.
+Because SSSP and BFS often have few active vertices with few overlapping vertices (i.e.
+little synchronization needed), synchronization is often not impacting performance as much, making push often the better implementation for those algorithms.
+PageRank on the other hand has many active vertices and is thus expected to perform much better with pull implementations, because less synchronization is needed in that case.
 Hence, a good guideline is pull-style on single-node systems with PR-like applications and push-style for SSSP and BFS, especially on distributed systems. 
-Our data backs this up, we observed single-node PageRank to be fastest with Galois Pull, while Galois Push was the fastest application on distributed systems.
+Our data backs this up, we observed single-node PageRank to be fastest with Galois Pull, while the push based algorithms were fastest on the distributed systems.
 In general however, some tests determining which implementation is faster can be beneficial.
 
-%### general
-For graphs that are not too large for single node RAM and when performing research work, Galois as the fastest single-node framework is recommended.
+For graphs that are not too large for single node RAM 
+%and when performing research work
+, Galois as the fastest single-node framework is recommended.
 It provides excellent performance, being orders of magnitude faster than the competing frameworks.
-And when it is possible to use hugepages, these improve performance even further, sometimes up to a factor of 2 faster compared to Galois without hugepages.
+And when it is possible to use hugepages, these improve performance even further, sometimes up to a factor of 2 times faster compared to Galois without hugepages.
 Also, we have shown that most algorithms are able to utilize many threads in favour of much smaller calculation times.
 Graph-traversal algorithms like BFS or SSSP shown here are examples for such algorithms.
-In some cases, especially on PageRank-like applications, it might be necessary to perform some tests whether a push- or pull-style implementation is most efficient.
-However, especially when using push-style implementations, a large thread count is most likely not going to significantly improve performance over single threaded performance.
+Supplying a large thread pool for Galois is most likely improving performance on those algorithms.
+However, especially when using push-style algorithms, a large thread count is most likely not going to significantly improve performance over single threaded performance.
 The reason for this is the synchronization required for push applications.
-This is true not only for Galois but any framework.
+This is true not only for single-node Galois but any framework.
 
 %## production
-On production systems we expect it to be possible to determine what configuration is best, before putting the system in service.
-There are multiple possibilities to be accounted for
-\todo{unfertig}
+On production systems, we expect it to be possible to determine what configuration is best before putting the system in service.
+But not only the absolute runtime of each computation has to be taken into account. Production systems have to be very reliable because downtime often directly correlates to financial loss.
 
-%### Distributed
-If the graph data becomes too large for one node or the computation cluster is distributed anyways, there are multiple possibilities.
-
-First, there is an argument to be made for systems like Hadoop.
+This is often a reason for choosing a distributed system. Many nodes together are less prone to failure of the entire system.
+Thus, there is an argument to be made for systems like Hadoop.
+It automatically handles node failures, a node failure does not immediately result in loss of data or a faulty computation result.
+This is especially important on very large clusters, that can not easily be monitored by humans.
 A user defining an application does not need to know the topology of the system that runs the application later on.
 Hadoop transparently splits the input data and distributes work among the worker nodes.
-And Hadoop automatically handles node failures, a node failure does not immediately result in loss of data or a faulty computation result.
-This is especially important on very large clusters, that can not easily be monitored by humans.
 Furthermore, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
@@
 Very short calculation times along with automatic handling of the computation cluster make Giraph using Hadoop a good choice for production systems.
-Pregel is a similarly fault-tolerant, scalable system but it is contrary to Giraph closed-source.
+Pregel is a similarl, fault-tolerant and scalable system but it is contrary to Giraph closed-source.
+
+But not only distributed production systems are possible, of course.  Single-node systems are possible, as long as the graph data fits in local memory. 
+Further, horizontal scaling can for example be used to provide fault-tolerance of the entire system.
+While Giraph has very short computation times even on a single node, we would advise against it.
+Giraph requires a lot more memory than any of the other frameworks and ran out of memory on multiple occasions.
+Thus it should only really be considered on distributed setups.
+
+There are multiple possibilities to be accounted for
+\todo{unfertig}
+
+%### Distributed
+If the graph data becomes too large for one node or the computation cluster is distributed anyways, there are multiple possibilities.
 

### 2020-09-21 documentation/paper/sections/results/results.tex
most introductions done
galois time comparison init
todos in results und discussion

@@
 \subsection{Results}
-\todo{Einleitung}
+Ahead of the discussion of the results a few of the issues we encountered while testing are mentioned.
+Afterwards the benchmark results for SSSP, BFS, PR and the speedup of Galois given a varying number of threads are discussed in that order.
 
@@
 On a plus side, setup of the frameworks Polymer and Ligra was straight forward and did not require any special treatment.
+\todo{Gemini? Außerdem sollte hier noch ein bisschen was hin, die beiden Zeilen sehen ja nur traurig aus.}
 
@@
 \begin{table}
-	\caption{Execution Times on a SingleNode vs. 5-Node Distributed Cluster}
+	\caption{Execution Times in Seconds on a SingleNode vs. 5-Node Distributed Cluster}
 	\label{tbl:execTimeComparison}

### 2020-09-21 documentation/paper/main.tex
merge und besprechung

@@
 
+\input{sections/relatedWork}
+
 \section{Evaluation}
@@
 The first is a description and explanation of our testing environment and methodology.
-Followed by a lengthy presentation and discussion of the test results of the frameworks.
+Followed by a presentation and discussion of the test results of the frameworks.
 
@@
 
-\section{Discussion}
-\todo{Hier zunächst mal ein kurzer Vergleich zwischen Distributed und Single Node wo sinnvoll, dann zwischen den Frameworks und so weiter..}
+\input{sections/discussion}
 

### 2020-09-21 documentation/paper/sections/results/results.tex
merge und besprechung

@@
 
-\subsubsection{Lessons Learned}
-
-\todo{umbenennen?}
-
+\subsubsection{Encountered Issues}
 We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.
@@
 \begin{table}
-	\caption{Execution Times in Seconds on a SingleNode vs. 5-Node Distributed Cluster}
+	\caption{Execution Times in Seconds on a Single Node vs. 5-Node Distributed Cluster}
 	\label{tbl:execTimeComparison}
@@
 		\midrule
-		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}&flickr & 0.3 & 2.4 & 0.6 & 2.5 & 38.8 & 42.6 \\
-		& orkut & 0.8 & 4.2 & 13.9 & 17.3 & 135.9 & 60.4 \\
-		& wikipedia & 1.8 & 14.2 & 48.5 & 74.5 & 377.2 & 100.8 \\
-		& twitter & 10.8 & 40.0 & 383.0 & 391.1 & - & 349.9 \\
-		& rMat27 & 16.0 & 39.0 & 563.5 & 386.0 & - & 565.8 \\
-		& friendster & 14.4 & 58.5 & 742.6 & 568.9 & - & 444.0 \\
-		& rMat28 & 27.8 & 71.5 & 1236.7 & 792.0 & - & 1180.2 \\
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}&flickr & \bf 0.3 & 2.4 & \bf 0.6 & 2.5 & \bf 38.8 & 42.6 \\
+		& orkut & \bf 0.8 & 4.2 & \bf 13.9 & 17.3 & 135.9 & \bf 60.4 \\
+		& wikipedia & \bf 1.8 & 14.2 & \bf 48.5 & 74.5 & 377.2 & \bf 100.8 \\
+		& twitter & \bf 10.8 & 40.0 & \bf 383.0 & 391.1 & - & \bf 349.9 \\
+		& rMat27 & \bf 16.0 & 39.0 & 563.5 & \bf 386.0 & - & \bf 565.8 \\
+		& friendster & \bf 14.4 & 58.5 & 742.6 & \bf 568.9 & - & \bf 444.0 \\
+		& rMat28 & \bf 27.8 & 71.5 & 1236.7 & \bf 792.0 & - & \bf 1180.2 \\
 		\midrule
-		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}& flickr & 0.6 & 2.3 & 0.7 & 2.4 & 38.9 & 41.9 \\
-		& orkut & 0.9 & 4.0 & 12.5 & 13.8 & 128.2 & 58.1 \\
-		& wikipedia & 2.4 & 13.2 & 44.9 & 60.9 & 360.7 & 95.1 \\
-		& twitter & 14.2 & 37.9 & 355.0 & 293.2 & - & 322.3 \\
-		& rMat27 & 14.6 & 34.4 & 540.8 & 305.3 & - & 539.2 \\
-		& friendster & 12.8 & 47.6 & 708.4 & 450.2 & - & 412.3 \\
-		& rMat28 & 33.1 & 67.1 & 1178.7 & 634.3 & - & 1135.6 \\
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}& flickr & \bf 0.6 & 2.3 & \bf 0.7 & 2.4 & \bf 38.9 & 41.9 \\
+		& orkut & \bf 0.9 & 4.0 & \bf 12.5 & 13.8 & 128.2 & \bf 58.1 \\
+		& wikipedia & \bf 2.4 & 13.2 & \bf 44.9 & 60.9 & 360.7 & \bf 95.1 \\
+		& twitter & \bf 14.2 & 37.9 & 355.0 & \bf 293.2 & - & \bf 322.3 \\
+		& rMat27 & \bf 14.6 & 34.4 & 540.8 & \bf 305.3 & - & \bf 539.2 \\
+		& friendster & \bf 12.8 & 47.6 & 708.4 & \bf 450.2 & - & \bf 412.3 \\
+		& rMat28 & \bf 33.1 & 67.1 & 1178.7 & \bf 634.3 & - & \bf 1135.6 \\
 		\midrule
-		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR}}& flickr & 0.3\txtdagger & 2.6 & 0.7 & 2.2 & 45.6 & 44.1 \\
-		& orkut & 0.7\txtdagger & 8.0 & 13.1 & 19.4 & 561.3 & 106.2 \\
-		& wikipedia & 1.7\txtdagger & 206.5 & 44.4 & 54.4 & 1501.3 & 252.6 \\
-		& twitter & 8.7\txtdagger & 910.9 & 359.9 & 363.5 & - & 1200.2 \\
-		& rMat27 & 19.2\txtdagger & 1287.4 & 536.3 & 376.3 & - & 1369.5 \\
-		& friendster & 20.4\txtdagger & 593.3 & 716.5 & 591.1 & - & 1655.0 \\
-		& rMat28 & 46.0\txtdagger & 2540.2 & 1188.5 & 774.2 & - & - \\
+		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR}}& flickr & \bf 0.3\txtdagger & 2.6 & \bf 0.7 & 2.2 & 45.6 & \bf 44.1 \\
+		& orkut & \bf 0.7\txtdagger & 8.0 & \bf 13.1 & 19.4 & 561.3 & \bf 106.2 \\
+		& wikipedia & \bf 1.7\txtdagger & 206.5 & \bf 44.4 & 54.4 & 1501.3 & \bf 252.6 \\
+		& twitter & \bf 8.7\txtdagger & 910.9 & \bf 359.9 & 363.5 & - & \bf 1200.2 \\
+		& rMat27 & \bf 19.2\txtdagger & 1287.4 & 536.3 & \bf 376.3 & - & \bf 1369.5 \\
+		& friendster & \bf 20.4\txtdagger & 593.3 & 716.5 & \bf 591.1 & - & \bf 1655.0 \\
+		& rMat28 & \bf 46.0\txtdagger & 2540.2 & 1188.5 & \bf 774.2 & - & - \\
 		\bottomrule

### 2020-09-21 documentation/paper/sections/testingMethods.tex
merge und besprechung

@@
 \subsection{Testing Methods}
-\todo{Einleitung}
+The testing methods cover the test environment, including hardware and the setup of the frameworks, the graphs used, followed by the algorithms utilized and how the times are measured.
 
@@
 In detail, the algorithms for each framework are:
-\todo{Algorithmen auflisten}
 \begin{itemize}
-	\item Ligra supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delata Variant.
-	\item Polymer supports the same algoritms as ligra.
+	\item Ligra supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
+	\item Polymer supports the same algoritms as Ligra.
 	\item Gemini supports all of our tested algorithms and there are no setting options or specifications which implementations for the algorithms are used.

### 2020-09-21 documentation/paper/sections/results/sssp.tex
some work on discussion/comparison
merge und besprechung

@@
 
-\subsection{Single-Source Shortest-Paths}
+\subsubsection{Single-Source Shortest-Paths}
 In this section, we compare the different frameworks in their performance on the SSSP algorithm.
 We first analyze the frameworks on a single computation node and compare to the distributed setup after that.
-\subsubsection{Single-node}
+\paragraph{Single-node}
 \begin{figure*}
@@
 \begin{table}
-\renewcommand{\arraystretch}{1.2}
+\renewcommand{\arraystretch}{1.1}
 \tiny
@@
 \caption{Distributed SSSP Execution Times and Their Realation to Galois Push}
-\begin{tabular}{crrrrrrrr}
+\begin{tabular}{c@{\tabskip 1.4 \tabcolsep}r@{\tabskip 1 \tabcolsep}r@{\tabskip 1.4 \tabcolsep}r@{\tabskip 1 \tabcolsep}r@{\tabskip 1.4 \tabcolsep}r@{\tabskip 1 \tabcolsep}r@{\tabskip 1.4 \tabcolsep}r@{\tabskip 1 \tabcolsep}r}
 \toprule
@@
 \multicolumn{2}{c}{\bf Galois Push}&\multicolumn{2}{c}{\bf Galois Pull}&\multicolumn{2}{c}{\bf Gemini}&\multicolumn{2}{c}{\bf Giraph}\\
-&$\times$&(s)&$\times$&(s)&$\times$&(s)&$\times$&(s)\\\midrule
-flickr & 1.0 & (2.4) & 0.97 & (2.4) & 1.02 & (2.5) & 17.46 & (42.6)\\
-orkut & 1.0 & (4.2) & 1.28 & (5.4) & 4.12 & (17.3) & 14.38 & (60.4)\\
-wikipedia & 1.0 & (14.2) & 1.88 & (26.6) & 5.26 & (74.5) & 7.11 & (100.8) \\
-twitter & 1.0 & (40.0) & 3.56 & (142.2) & 9.79 & (391.1) & 8.76 & (349.9)\\
-rMat27 & 1.0 & (39.0) & 3.98 & (154.9) & 9.91 & (386.0) & 14.52 & (565.8)\\
-friendster & 1.0 & (58.5) & 2.32 & (136.0) & 9.72 & (568.9) & 7.59 & (444.0)\\
-rMat28 & 1.0 & (71.5) & 5.7 & (407.9) & 11.07 & (792.0) & 16.5 & (1180.2)\\\bottomrule
+\cmidrule{2-3}\cmidrule{4-5}\cmidrule{6-7}\cmidrule{8-9}
+&$\times$G.Push&time (s)&$\times$G.Push&time (s)&$\times$G.Push&time (s)&$\times$G.Push&time (s)\\\midrule
+flickr & 1.0 & 2.4 & 0.97 & 2.4 & 1.02 & 2.5 & 17.46 & 42.6\\
+orkut & 1.0 & 4.2 & 1.28 & 5.4 & 4.12 & 17.3 & 14.38 & 60.4\\
+wikipedia & 1.0 & 14.2 & 1.88 & 26.6 & 5.26 & 74.5 & 7.11 & 100.8 \\
+twitter & 1.0 & 40.0 & 3.56 & 142.2 & 9.79 & 391.1 & 8.76 & 349.9\\
+rMat27 & 1.0 & 39.0 & 3.98 & 154.9 & 9.91 & 386.0 & 14.52 & 565.8\\
+friendster & 1.0 & 58.5 & 2.32 & 136.0 & 9.72 & 568.9 & 7.59 & 444.0\\
+rMat28 & 1.0 & 71.5 & 5.7 & 407.9 & 11.07 & 792.0 & 16.5 & 1180.2\\\bottomrule
 \end{tabular}

### 2020-09-21 documentation/paper/sections/discussion.tex
merge und besprechung
discussion

@@
 \section{Discussion}
-\todo{unfertig}
-
 There are two main cases that should be regarded in a comparison of the frameworks.
@@
 for each calculation, the graph has to be loaded.
+Further, the algorithm can change frequently, requiring the framework to be relatively fast on different alogrithms.
 This is called the \emph{research} case.
-For research, frameworks with small execution times and small overhead are preferred.
+For research, frameworks with overall small execution times and small overhead are preferred.
 Case two on the other hand is the case of a running system, that performs multiple calculations on a single graph, without the need of reloading graph data with every calculation.
@@
 
-Orthogonal to this, the graph size and topology of the system is a deciding factor.
+Orthogonal to this, the graph size and topology of the system are deciding factors.
 Large graphs often require a distributed cluster, because the graph is simply too large to fit in the RAM of just one machine.
-We observed this for example with Giraph on multiple cases.
+We observed this for example with Giraph on multiple graphs.
 Furthermore, some environments require a distributed framework setup regardless of any other factors.
-
-For algorithms that perform a traversal through the graph, with a concentrated and relativly small amount of active vertices, our results for SSSP and BFS are most relevant.
-If the problem is however of a nature, where many vertices are active, thus many vertices, widely spread across the graph performe some kind of calculation, the results for PR are most applicable.
-Further, some considerations whether a push or pull-style implemenation is more efficient should be made.
-Generally, we expect pull-style implementations to be faster on a single-node, like for example PR with Galois.
-On a distributed cluster though, the synchronization time is not as severely impacting performance because communication over the network is needed, which is slower than local synchonization in any case.
-
-We already know push implementations to be more efficient on problems with only few active vertices or where the affected push-areas are not overlapping. 
-Because SSSP and BFS often have few active vertices with few overlapping vertices (i.e.
-little synchronization needed), synchronization is often not impacting performance as much, making push often the better implementation for those algorithms.
-PageRank on the other hand has many active vertices and is thus expected to perform much better with pull implementations, because less synchronization is needed in that case.
-Hence, a good guideline is pull-style on single-node systems with PR-like applications and push-style for SSSP and BFS, especially on distributed systems. 
-Our data backs this up, we observed single-node PageRank to be fastest with Galois Pull, while the push based algorithms were fastest on the distributed systems.
-In general however, some tests determining which implementation is faster can be beneficial.
-
-For graphs that are not too large for single node RAM 
-%and when performing research work
-, Galois as the fastest single-node framework is recommended.
-It provides excellent performance, being orders of magnitude faster than the competing frameworks.
-And when it is possible to use hugepages, these improve performance even further, sometimes up to a factor of 2 times faster compared to Galois without hugepages.
-Also, we have shown that most algorithms are able to utilize many threads in favour of much smaller calculation times.
-Graph-traversal algorithms like BFS or SSSP shown here are examples for such algorithms.
-Supplying a large thread pool for Galois is most likely improving performance on those algorithms.
-However, especially when using push-style algorithms, a large thread count is most likely not going to significantly improve performance over single threaded performance.
-The reason for this is the synchronization required for push applications.
-This is true not only for single-node Galois but any framework.
 
@@
 But not only the absolute runtime of each computation has to be taken into account. Production systems have to be very reliable because downtime often directly correlates to financial loss.
-
 This is often a reason for choosing a distributed system. Many nodes together are less prone to failure of the entire system.
@@
 
-But not only distributed production systems are possible, of course.  Single-node systems are possible, as long as the graph data fits in local memory. 
-Further, horizontal scaling can for example be used to provide fault-tolerance of the entire system.
-While Giraph has very short computation times even on a single node, we would advise against it.
+But not only distributed production systems are possible, of course. 
+For small graphs, i.e. those that are not too large for single node RAM, a single-node setup is usually faster than the same framework on a distributed cluster. 
+We have shown that the distributed cluster of the same framework becomes faster than the equivalent single-node setup only on large graphs.
+We have found around 2B edges, or roughly the size of the twitter graph to be a size at which Ligra and Giraph become faster on the distributed setup compared to single-node.
+If required, horizontal scaling can for example be used to provide fault-tolerance of the entire system.
+
+If, the deciding factor is only the calculation time, there are multiple options, with the deciding factors being the graph size and the application.
+For algorithms that perform a traversal through the graph, with a concentrated and relativly small amount of active vertices, our results for SSSP and BFS are most relevant.
+Here Galois is fastest on very small graphs (i.e. flickr), but is quickly overtaken by other frameworks.
+In the case of SSSP those are Gemini and Polymer and on BFS Ligra and again Gemini.
+While Gemini is not the fastest on either BFS or SSSP, it is still the safest bet on other algorithms.
+That is because Polymer is fast on SSSP but slowest on BFS and in turn Ligra is fastest on BFS but one of the slowest frameworks on SSSP.
+If the problem is however of a nature, where many vertices are active, thus many vertices, widely spread across the graph performe some kind of calculation, our results for PR are to be considered.
+Further, some considerations whether a push or pull-style implemenation is more efficient should be made.
+Galois is the fastest framework in either calculation or execution time, regardless of push or pull version.
 Giraph requires a lot more memory than any of the other frameworks and ran out of memory on multiple occasions. 
@@
 
-There are multiple possibilities to be accounted for
-\todo{unfertig}
+%## research
+For single-node research systems, a overall good execution time on \emph{any} application is necessary. 
+Here we recommend Galois as the fastest executing single-node framework overall.
+It provides excellent performance, being orders of magnitude faster than the competing frameworks.
+And when it is possible to use hugepages, these improve performance even further, sometimes up to a factor of 2 times faster compared to Galois without hugepages.
+Also, we have shown that most algorithms are able to utilize many threads in favour of much smaller calculation times.
+Graph-traversal algorithms like BFS or SSSP shown here are examples for such algorithms.
+Supplying a large thread pool for Galois is most likely improving performance on those algorithms.
+However, especially when using push-style algorithms, a large thread count is most likely not going to significantly improve performance over single threaded performance.
+The reason for this is the synchronization required for push applications.
+This is true not only for single-node Galois but any framework.
+On a distributed research system however, many different computations are performed and node failures are not as problematic, because the system is most likely reconfigured frequently anyways.
+Hence, Giraph as the go-to distributed system so far is not applicable here.
+Hadoop handling all the parallelization and fault-tolerance in the background is a nice-to-have in a production system but an overall faster framework would be preferred in research.
+Hence, Galois is once again the recommended framework. As mentioned before, it outperforms the other frameworks on almost any test case, even without hugepages.
 
-%### Distributed
-If the graph data becomes too large for one node or the computation cluster is distributed anyways, there are multiple possibilities.
+\todo{Das folgende muss noch irgendwie rein..}
 
-On research systems however, single computations are performed and node failures are not as common, because the system is most likely not in-use constantly.
-Hence, Hadoop handling all this in the background is a nice-to-have but most likely, a plain faster framework would be preferred.
-Furthermore, administration is most likely needed frequently because system reconfiguration is happening often?
-Because the system is most likely reconfigured frequently.
-\todo{unfertig}
-Here, the fastest framework really depends on the use case. Graph traversal algorithms prefer
-
-Here, Galois is the fastest distributed framework for SSSP and BFS, while Gemini outperforms Galois on PR.
-
-%A distributed setup can have many advantages,
-
-%Especially for frameworks like Giraph, that require large amounts of memory, the distributed scenario allows them to complete their computations.
-
-%first, there is some data security, because many distributed file systems also support automatic data replication etc.
+We already know push implementations to be more efficient on problems with only few active vertices or where the affected push-areas are not overlapping. 
+Because SSSP and BFS often have few active vertices with few overlapping vertices (i.e.
+little synchronization needed), synchronization is often not impacting performance as much, making push often the better implementation for those algorithms.
+PageRank on the other hand has many active vertices and is thus expected to perform much better with pull implementations, because less synchronization is needed in that case.
+Hence, a good guideline is pull-style on single-node systems with PR-like applications and push-style for SSSP and BFS, especially on distributed systems.
+The synchronization time on distributed systems is not as severely impacting performance because communication over the network is needed anyway, which is slower than local synchonization in any case.
+Our data backs this up, we observed single-node PageRank to be fastest with Galois Pull, while the push based algorithms were fastest on the distributed systems.
+In general however, some tests determining which implementation is faster can be beneficial.

### 2020-09-21 documentation/paper/sections/discussion.tex
discussion soweit fertig denke ich

@@
 There are two main cases that should be regarded in a comparison of the frameworks. 
-One, the case of individual calculations on a graph, i.e.
-for each calculation, the graph has to be loaded.
-Further, the algorithm can change frequently, requiring the framework to be relatively fast on different alogrithms.
+One, the case of individual calculations on a graph, i.e. for each calculation, the graph has to be loaded. 
+Further, the algorithm can change frequently, requiring the framework to be relatively fast on different algorithms.
 This is called the \emph{research} case.
@@
 We observed this for example with Giraph on multiple graphs.
+Giraph requires a lot more memory than any of the other frameworks and ran out of memory on multiple occasions. 
+Thus it should only really be considered on distributed setups.
 Furthermore, some environments require a distributed framework setup regardless of any other factors.
@@
 Thus, there is an argument to be made for systems like Hadoop.
-It automatically handles node failures, a node failure does not immediately result in loss of data or a faulty computation result.
+It automatically handles node failures, whith a node failure not immediately resulting in loss of data or faulty computation results.
 This is especially important on very large clusters, that can not easily be monitored by humans.
-A user defining an application does not need to know the topology of the system that runs the application later on.
+Further, a user defining an application does not need to know the topology of the system that runs the application later on.
 Hadoop transparently splits the input data and distributes work among the worker nodes.
-Furthermore, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
-This shows the hadoop infrastructure to be working well and is exactly what the distributed system should be used for anyways. 
+Additionally, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
+This shows the hadoop infrastructure to be working well with large graphs being exactly what the distributed system should be used for. 
 Only a system like hadoop can reasonably be used on such large clusters.
@@
 Very short calculation times along with automatic handling of the computation cluster make Giraph using Hadoop a good choice for production systems.
-Pregel is a similarl, fault-tolerant and scalable system but it is contrary to Giraph closed-source.
+Pregel is a similar, fault-tolerant and scalable system but it is contrary to Giraph closed-source.
 
@@
 
-If, the deciding factor is only the calculation time, there are multiple options, with the deciding factors being the graph size and the application.
+If only the calculation time is important, there are multiple options with the deciding factors being the graph size and the application.
 For algorithms that perform a traversal through the graph, with a concentrated and relativly small amount of active vertices, our results for SSSP and BFS are most relevant.
 Here Galois is fastest on very small graphs (i.e. flickr), but is quickly overtaken by other frameworks.
-In the case of SSSP those are Gemini and Polymer and on BFS Ligra and again Gemini.
+In the case of SSSP those are Gemini and Polymer and on BFS they are Ligra and again Gemini.
 While Gemini is not the fastest on either BFS or SSSP, it is still the safest bet on other algorithms.
 That is because Polymer is fast on SSSP but slowest on BFS and in turn Ligra is fastest on BFS but one of the slowest frameworks on SSSP.
-If the problem is however of a nature, where many vertices are active, thus many vertices, widely spread across the graph performe some kind of calculation, our results for PR are to be considered.
-Further, some considerations whether a push or pull-style implemenation is more efficient should be made.
-Galois is the fastest framework in either calculation or execution time, regardless of push or pull version.
-Giraph requires a lot more memory than any of the other frameworks and ran out of memory on multiple occasions. 
-Thus it should only really be considered on distributed setups.
+Gemini is second fastest on SSSP and BFS on the larger graphs.
+Also, it might be important to consider the implementation style of a custom application for either of the frameworks.
+Applications similar to BFS and SSSP are expected to perform better with a push style implementation. 
+This is because they are algorithms with only few active vertices or where the affected push areas are not overlapping. 
+Hence, only little synchronization is needed that could be impacting performance negatively.
+
+If the problem is however of a nature, where many vertices are active, thus many vertices, widely spread across the graph perform some kind of calculation, our results for PR are to be considered.
+Because PageRank has many active vertices it is expected to perform much better with pull implementations, because less synchronization is needed in that case.
+This should be kept in mind when developing custom applications.
+Galois is the fastest framework on PR in either calculation time or execution time, regardless of push or pull version.
+Though for custom applications, one style might be significantly faster than the other.
+
+A good guideline is pull-style on single-node systems with PR-like applications and push-style for SSSP and BFS, especially on distributed systems.
+The synchronization time on distributed systems is not as severely impacting performance because communication over the network is needed anyways.
+Communication over the network is usually slower than local synchonization.
+Our data backs this up, we observed single-node PageRank to be fastest with Galois Pull, while the push based algorithms were fastest on the distributed systems.
 
@@
 It provides excellent performance, being orders of magnitude faster than the competing frameworks.
-And when it is possible to use hugepages, these improve performance even further, sometimes up to a factor of 2 times faster compared to Galois without hugepages.
+And when it is possible to use hugepages, further improvements in performance can be expected.
+We frequently observed Galois with hugepages to be around 2 times faster compared to Galois without hugepages.
 Also, we have shown that most algorithms are able to utilize many threads in favour of much smaller calculation times.
@@
 Supplying a large thread pool for Galois is most likely improving performance on those algorithms.
-However, especially when using push-style algorithms, a large thread count is most likely not going to significantly improve performance over single threaded performance.
-The reason for this is the synchronization required for push applications.
-This is true not only for single-node Galois but any framework.
-On a distributed research system however, many different computations are performed and node failures are not as problematic, because the system is most likely reconfigured frequently anyways.
+However, especially when using push-style algorithms, a large thread count is most likely not going to significantly improve performance over single threaded results. 
+The reason for this is the synchronization required for push applications. This is true not only for single-node Galois but any framework.
+
+Distributed research systems do not have the same requirements as production systems.
+For example node failures are not as problematic, because the system is most likely reconfigured frequently anyways.
 Hence, Giraph as the go-to distributed system so far is not applicable here.
-Hadoop handling all the parallelization and fault-tolerance in the background is a nice-to-have in a production system but an overall faster framework would be preferred in research.
-Hence, Galois is once again the recommended framework. As mentioned before, it outperforms the other frameworks on almost any test case, even without hugepages.
-
-\todo{Das folgende muss noch irgendwie rein..}
-
-We already know push implementations to be more efficient on problems with only few active vertices or where the affected push-areas are not overlapping. 
-Because SSSP and BFS often have few active vertices with few overlapping vertices (i.e.
-little synchronization needed), synchronization is often not impacting performance as much, making push often the better implementation for those algorithms.
-PageRank on the other hand has many active vertices and is thus expected to perform much better with pull implementations, because less synchronization is needed in that case.
-Hence, a good guideline is pull-style on single-node systems with PR-like applications and push-style for SSSP and BFS, especially on distributed systems.
-The synchronization time on distributed systems is not as severely impacting performance because communication over the network is needed anyway, which is slower than local synchonization in any case.
-Our data backs this up, we observed single-node PageRank to be fastest with Galois Pull, while the push based algorithms were fastest on the distributed systems.
-In general however, some tests determining which implementation is faster can be beneficial.
+Hadoop handling all the parallelization and fault-tolerance in the background is a nice-to-have in a production system.
+But in research, one might not be willing to pay the price of increased execution times over other frameworks.
+A framework that is fast in many algorithms is required. 
+Galois is the fastest framework on SSSP and BFS, by a margin of around one order of magnitude to Gemini. 
+On PR however, Galois is only second fastest. It is about one order of magnitude behind Gemini on three of the seven graphs.
+Hence, the choice of framework really depends on the kind of work expected to run on the system.
+For graph-traversal or similar applications, Galois outperforms Gemini.
+For research with PR-like applications, Gemini would be a wise choice.

### 2020-09-22 documentation/paper/sections/introduction.tex
merge und besprechung
Anfang der Introduction schlecht ausformuliert

@@
-%!TEX root=../../main.tex
+%!TEX root=../main.tex
 \section{Introduction}
 
-1. Einführung, warum ist Graph analysis wichtig\\
-- Graph processing is gaining increasing attentions in both academic and industrial communities. (Gemini)\\
-- Graphen werden immer wichtiger in vielen Domänen \todo{welche}\\
--Many machine learning, data mining and scientific computation can be modeled as graph-structured computation, resulting in a new application domain called graph analytics. (Polymer)\\
-- In recent years graph sizes have increased significantly. Thus performance and memory efficiency of the graph analysis applications is now more important than ever. \todo{Quelle}\\
-- Many applications like, Wachsende graphen, benötigt für folgende Applikationen, verwenden graph algorithmen, diese müssen schnell und verteilt laufen\\
-- Graphen wachsen schneller als CPUs \todo{Quelle}\\
+In recent years, graph processing has become increasingly important, both in academic and industrial environments \cite{Gemini}.
+In many areas, such as machine learning and data mining, but also scientific computations, problems can be modeled and solved with the help of graphs \cite{Polymer}.
+Thus a new application domain called graph alnalytics was created.
+Some business models are based on working with graphs, such as marketing analysis, advertising, or even Google's search engine.
+The economic success of these business models depends on the efficiency of the graph algorithms.
+In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
+This makes the performance and memory efficency of graph analysis applications more important than ever.
+But also the parallelism and distribution of graph algorithms becomes more important with increasing graph sizes.
+
+Some frameworks were developed for this purpose.
+One of the best known is the (closed source) distributed graph processing system Pregel \cite{pregel}.
+Pregel, as well as some open source versions of it, like Giraph \cite{Giraph}, was built to handle large graphs reliable and fault tolerant on large MapReduce infrastructures.
+Due to the lack of cache locality, as these systems are written object oriented, they provide rather suboptimal performance.
+Because of this, some non-uniform memory access (NUMA) aware systems like Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
+\todo{Heiko fragen, ob es ok ist, dass in diesem Abschnitt einiges aus seinem Aushang stammt}
 \\
-2. Ausgangssituation und Problem\\
-- dafür wurden Zahlreiche Framweworks entwickelt (Paper: Galois)\\
-- Domänenspezifische Sprachen DSL (Paper: Galois)\\
-- Es wurden auch Librarys und Technologien entwickelt (boost, hugepages, numa, ...)\\
-- Um es zu erleichtern Graphprogramme zu schreiben (Paper: Galois)\\
-- Größe der Graphen fordert Parallelität/Verteiltheit => Probleme  (Paper:Galois)\\
 - Prolem der Vergleichbarkeit der Systeme\\
@@
 \\
+
+This paper makes the following contributions:
+\begin{itemize}
+  \item Comparison of several state-of-the-art graph processing frameworks
+  \item
+  \item
+  \item
+\end{itemize}
+
 4. Wie ist das Paper aufgebaut\\
@@
 - Section 7: Concluseion (Was lernen wir daraus?)\\
+
+% In recent years graph sizes have increased significantly [?] .
+% Thus performance and memory efficiency of the graph analysis applications is now more important than ever.
+
+% Many applications like
+
+% Wachsende graphen,
+% benötigt für folgende Applikationen,
+% verwenden graph algorithmen,
+% diese müssen schnell und verteilt laufen
+% dafür wurden Frameworks entwickelt...
+
+% und das noch schön formulieren
+
+% We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
+% The comparison is performed on both real world data sets and synthetic graphs.
+
+% To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
+
+%One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
+%This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.
+
+% To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.

### 2020-09-22 documentation/paper/sections/introduction.tex

@@
 Some business models are based on working with graphs, such as marketing analysis, advertising, or even Google's search engine.
-The economic success of these business models depends on the efficiency of the graph algorithms.
 In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
@@
 Some frameworks were developed for this purpose.
-One of the best known is the (closed source) distributed graph processing system Pregel \cite{pregel}.
-Pregel, as well as some open source versions of it, like Giraph \cite{Giraph}, was built to handle large graphs reliable and fault tolerant on large MapReduce infrastructures.
+One of the best known is the closed source distributed graph processing system Pregel \cite{pregel}.
+Pregel, as well as some open source versions of it, like Giraph \cite{Giraph} and GraphX \cite{graphx}, was built to handle large graphs reliable and fault tolerant on large MapReduce infrastructures.
 Due to the lack of cache locality, as these systems are written object oriented, they provide rather suboptimal performance.
 Because of this, some non-uniform memory access (NUMA) aware systems like Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
-\todo{Heiko fragen, ob es ok ist, dass in diesem Abschnitt einiges aus seinem Aushang stammt}
 \\
@@
 To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
-One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
-This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.
-To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.\\
 - Comparison of several state-of-the-art graph processing frameworks\\
@@
 - Section 7: Concluseion (Was lernen wir daraus?)\\
-
-% In recent years graph sizes have increased significantly [?] .
-% Thus performance and memory efficiency of the graph analysis applications is now more important than ever.
-
-% Many applications like
-
-% Wachsende graphen,
-% benötigt für folgende Applikationen,
-% verwenden graph algorithmen,
-% diese müssen schnell und verteilt laufen
-% dafür wurden Frameworks entwickelt...
-
-% und das noch schön formulieren
-
-% We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
-% The comparison is performed on both real world data sets and synthetic graphs.
-
-% To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
-
-%One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
-%This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.
-
-% To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.

### 2020-09-22 documentation/paper/sections/preliminiaries.tex
merge und besprechung

@@
 \section{Preliminaries}
-This section briefly explain the concepts and applications necessary, but not directly related to our work.
+This section briefly explains the concepts and applications necessary, but not directly related to our work.
 Initialy graphs and paths are defined, followed by explanations for various graph analysis applications.
@@
 \subsection{PageRank}
-\todo{Aufbau unschön, Formel fehlt}
 \emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}.
 The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.
+It is also used as a centrality metric.
 
-This represents a centrality metric of the vertices.
+The output of PageRank is a percentage for each vertex.
+This percentage, called the PageRank of the vertex, is the probability of arriving at this vertex when starting at a random vertex and then following random edges.
+With a small probability, the damping factor, a jump is made to a completely random vertex instead of following a edge.
+The PageRank can be calculated by iterating
+\begin{equation*}
+	\operatorname{PR}(v_i) = \frac{1 - d}{|V|} + d \sum_{v_j \in N^-(v_i)} \frac{\operatorname{PR}(v_j)}{\operatorname{deg^+}(v_j)}
+\end{equation*}
+for all edges $v_i \in V$.
+$d$ is the damping factor, $deg^+(v)$ is the number of outgoing edges the vertex $v$ has and $N^-(v)$ denotes the set of vertices that have edges to vertex $v$.
+
 The analogy is that the graph represents website pages of the Word Wide Web, that are hyperlinked between one another. 
-A website that is more important is likely to receive more links from other webiste.
-PageRank counts the number and quality of links to a page to estimate the importance of a page.
-The output of PageRank is a percentage for each vertex.
-This percentage, called the PageRank of the vertex, is the probability with which a web surfer starting at a random web page reaches this webpage (vertex).
-With a high probability the web surfer uses a random link from the web page they are currently on and with a smaller probability (called damping factor) they jump to a completely random web page.
+PageRank simulates a random surfer who is following links.
 
@@
 The PageRank score of a vertex is only updated if the relative change of the PageRank is larger than some user-defined delta.
-This effectively reduces the amount of vertices for which the PageRank has to be recalculated in following iterations.
+In the next iteration only neighbors of vertices whose PageRank has changed need to be recalculated, thus computation time can be saved.
 
@@
 
-In general such an operator can change the whole structure of the \emph{neighbourhood}.
+In general such an operator can change the whole structure of the \emph{neighborhood}.
 But in the following we assume, that an operator changes only \emph{labels} of vertices in the \emph{neighbourhood}.

### 2020-09-22 documentation/paper/sections/introduction.tex
Rest der Introduction schlecht ausformuliert

@@
 Because of this, some non-uniform memory access (NUMA) aware systems like Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
-\\
-- Prolem der Vergleichbarkeit der Systeme\\
-- Systeme entwickeln sich weiter und wurden nur zu deren Start verglichen\\
-- Vergleiche wurden von Machern der Systeme erledigt\\
-\\
-3. Was wir zur Lösung des Problems beitragen\\
-- ersten, die unabhängig alle Systeme Vergleichen 2020\\
-- Wir vergleichen die Frameworks nach folgenden Gesichtspunkten:\\
-(Benutzerfreundlichkeit, Performanz/Varianz, Weiterentwicklung)
-- We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).\\
-- The comparison is performed on both real world data sets and synthetic graphs.
-To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.
-- Comparison of several state-of-the-art graph processing frameworks\\
-\\
+These systems have some similarities, but many differences, which partly results in considerable performance differences.
+The systems were compared at their proposal.
+However, these comparisons were often incomplete, or are simply no longer relevant today, since most of the systems are several years old and some of them have evolved, or are still evolving.
+Other systems are simply no longer used today, which also reduces the significance of the comparisons.
 
-This paper makes the following contributions:
-\begin{itemize}
-  \item Comparison of several state-of-the-art graph processing frameworks
-  \item
-  \item
-  \item
-\end{itemize}
+Therefore we compared several state-of-the-art non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
+The comparison is performed on both real world data sets and synthetic graphs.
+To provide a comparison to a non-NUMA aware system, Giraph is included in the testing.
+Giraph itself has often been compared to other state-of-the-art systems like Pregel or GraphX.
+We also included a comparison of Galois with itself with and without hugepages \cite{hugepages}, and several runs with different numbers of threads to see how it scales.
+Where possible, several versions of the algorithms were used, like the push and pull variants.
 
-4. Wie ist das Paper aufgebaut\\
-- Section 2: Relatet Work (Wie haben die Macher der Systeme die Systeme verglichen)\\
-- Section 3: Preliminaries (Was muss man an technischen Details wissen)\\
-- Section 4: Framework Overview (Welche Frameworks und was macht diese aus)\\
-- Section 5: Evaluaten (Was wurde gemessen?, Wie wurde gemessen?, Was sind die Ergebnisse?)\\
-- Section 6: Discussion (Warum sehen die Ergebnisse so aus, wie sie aussehen?)\\
-- Section 7: Concluseion (Was lernen wir daraus?)\\
+The rest of this paper is organized as follows.
+In section 2 we first clarify what a graph is, how the algorithms we are looking at work and what else is necessary for further understanding.
+Then we give a short overview of the graph formats we are looking at in section 3 and of the graph formats used in section 4.
+Before we come to our results, we show in section 5 what has been done in this area so far.
+In Section 6, which is the core of the paper, we will present all our results and also briefly show how they were developed.
+After that, in section 7, we will discuss these results.
+And at the end in section 8 we will present our conclusion.

### 2020-09-22 documentation/paper/sections/preliminiaries.tex
Conclusion angefangen

@@
 \begin{equation*}
-	\operatorname{PR}(v_i) = \frac{1 - d}{|V|} + d \sum_{v_j \in N^-(v_i)} \frac{\operatorname{PR}(v_j)}{\operatorname{deg^+}(v_j)}
+	\operatorname{PR}(v_i) = \frac{1 - d}{|V|} + d \sum_{v_j \in \operatorname{N^-}(v_i)} \frac{\operatorname{PR}(v_j)}{\operatorname{deg^+}(v_j)}
 \end{equation*}
 for all edges $v_i \in V$.
-$d$ is the damping factor, $deg^+(v)$ is the number of outgoing edges the vertex $v$ has and $N^-(v)$ denotes the set of vertices that have edges to vertex $v$.
+$d$ is the damping factor, $\operatorname{deg^+}(v)$ is the number of outgoing edges the vertex $v$ has and $\operatorname{N^-}(v)$ denotes the set of vertices that have edges to vertex $v$.
 

### 2020-09-23 documentation/paper/sections/discussion.tex
kleiner Korrekturen an discussionn und introduction

@@
 Additionally, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
-This shows the hadoop infrastructure to be working well with large graphs being exactly what the distributed system should be used for. 
+This shows the hadoop infrastructure to be working well with large graphs, which is exactly what the distributed system should be used for. 
 Only a system like hadoop can reasonably be used on such large clusters.
@@
 We have shown that the distributed cluster of the same framework becomes faster than the equivalent single-node setup only on large graphs.
-We have found around 2B edges, or roughly the size of the twitter graph to be a size at which Ligra and Giraph become faster on the distributed setup compared to single-node.
+We have found around 2B edges, or roughly the size of the twitter graph to be a size at which Ligra and Giraph are faster on the distributed setup compared to single-node.
 If required, horizontal scaling can for example be used to provide fault-tolerance of the entire system.

### 2020-09-23 documentation/paper/sections/introduction.tex
kleiner Korrekturen an discussionn und introduction

@@
 In many areas, such as machine learning and data mining, but also scientific computations, problems can be modeled and solved with the help of graphs \cite{Polymer}.
-Thus a new application domain called graph alnalytics was created.
-Some business models are based on working with graphs, such as marketing analysis, advertising, or even Google's search engine.
+Thus a new application domain called graph analytics was created.
+Some business models are based on working with graphs, such as marketing analysis, advertising, or even Google's search engine \cite{pagerank}.
 In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
@@
 Some frameworks were developed for this purpose.
-One of the best known is the closed source distributed graph processing system Pregel \cite{pregel}.
-Pregel, as well as some open source versions of it, like Giraph \cite{Giraph} and GraphX \cite{graphx}, was built to handle large graphs reliable and fault tolerant on large MapReduce infrastructures.
-Due to the lack of cache locality, as these systems are written object oriented, they provide rather suboptimal performance.
-Because of this, some non-uniform memory access (NUMA) aware systems like Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
-These systems have some similarities, but many differences, which partly results in considerable performance differences.
-The systems were compared at their proposal.
-However, these comparisons were often incomplete, or are simply no longer relevant today, since most of the systems are several years old and some of them have evolved, or are still evolving.
-Other systems are simply no longer used today, which also reduces the significance of the comparisons.
+One of the most commonly known graph-processing systems is the closed-source and distributed system Pregel \cite{pregel}.
+Pregel, as well as some open source versions of it, like Giraph \cite{Giraph} and GraphX \cite{graphx}, was built to handle large graphs reliably and be fault-tolerant on large MapReduce infrastructures.
+Due to the lack of cache locality, as these systems are written object-oriented, they provide rather suboptimal performance.
+Because of this, many non-uniform memory access (NUMA) aware systems such as Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
+These systems have some similarities, but many differences. This results in considerable performance differences on many occasions.
+The systems were initially compared at their introduction.
+Because most of the systems are several years old and some of them have evolved, or are still evolving, these comparisons were often incomplete or are simply no longer relevant today. 
+Other systems are simply no longer used today, which also reduces the significance of some of the existing comparisons.
 
@@
 In section 2 we first clarify what a graph is, how the algorithms we are looking at work and what else is necessary for further understanding.
-Then we give a short overview of the graph formats we are looking at in section 3 and of the graph formats used in section 4.
+Then we give a short overview of the frameworks we comparing in section 3 and of the graph formats used in section 4.
 Before we come to our results, we show in section 5 what has been done in this area so far. 
 In Section 6, which is the core of the paper, we will present all our results and also briefly show how they were developed. 
-After that, in section 7, we will discuss these results.
-And at the end in section 8 we will present our conclusion.
+After that, in section 7, we will discuss these results. And at the end in section 8 we will present our conclusion.

### 2020-09-23 documentation/paper/sections/relatedWork.tex
related work

@@
 And second, many of the other frameworks are no longer in use because they were replaced by something else, making the comparisons difficult to interpret.
-We are thus providing an objective benchmark on what changed since the initial release, with a comparison between the most commonly used frameworks today.
+We are thus providing an objective benchmark on the current state of the most commonly used graph analysis frameworks.
 
@@
 
-The most recent publication on Galois was in 2013 \cite{Galois}, along with a comparison to GraphLab, PowerGraph and Ligra, which is in our testing lineup itself.
-The results of their comparison is mainly, that Galois outperforms Ligra on BFS and SSSP on most graphs.
+In 2013 \cite{Galois} was last compared to GraphLab, PowerGraph and Ligra in a shared memory setting.
+The results of their comparison are mainly, that Galois outperforms Ligra on BFS and SSSP on most graphs.
 Meanwhile, performance of Ligra and Galois on PR is similar but very dependent on the graph.
 In some cases, Ligra outperforms Galois, on other graphs Galois is faster than Ligra.
+
+About 5 years afterwards the capability of executing distributed was added to Galois through Gluon \cite{vertGalois}. 
+There distributed Galois is being compared to Gemini and distributed versions of Ligra and IrGl, where the distributed versions where created using Gluon. 
+IrGL is a graph analysis framework utilizing a GPU. 
+The results of the comparison were that Galois loads graphs faster, is faster than Gemini and Ligra at BFS and SSSP and depending on the graph either Ligra or Galois is the fastest at PR. 
+In most cases distributed IrGL was the fastest of the tested frameworks. 
+Galois and Ligra were shown to scale better on the number of nodes than Gemini.
 

### 2020-09-23 documentation/paper/sections/results/sssp.tex

@@
 
-\begin{table}
-\renewcommand{\arraystretch}{1.1}
-\tiny
+\begin{table*}
+\renewcommand{\arraystretch}{1.2}
+\renewcommand{\rot}[1]{\rotatebox{90}{#1}}
 \centering
@@
 \toprule
-\bf{Data Set}&
+&
 \multicolumn{2}{c}{\bf Galois Push}&\multicolumn{2}{c}{\bf Galois Pull}&\multicolumn{2}{c}{\bf Gemini}&\multicolumn{2}{c}{\bf Giraph}\\
 \cmidrule{2-3}\cmidrule{4-5}\cmidrule{6-7}\cmidrule{8-9}
-&$\times$G.Push&time (s)&$\times$G.Push&time (s)&$\times$G.Push&time (s)&$\times$G.Push&time (s)\\\midrule
+\bf{Data Set}&$\times$G.Push&time (s)&$\times$G.Push&time (s)&$\times$G.Push&time (s)&$\times$G.Push&time (s)\\\midrule
 flickr & 1.0 & 2.4 & 0.97 & 2.4 & 1.02 & 2.5 & 17.46 & 42.6\\
@@
 \label{tbl:ssspexec}
-\end{table}
+\end{table*}
 

### 2020-09-23 documentation/paper/sections/conclusion.tex

%!TEX root=../main.tex

\section{Conclusion}

We compared the performance of several graph processing frameworks including Ligra, Polymer, Gemini, Galois and Giraph. 
For comparison we used the three problems SSSP, BFS and PR on graphs of various sizes.
We measured the calculation time of the algorithm itself and the overall execution time, which also includes the startup time of the programm.
All of the frameworks were tested on a single compute node. Gemini, Galois and Giraph also support distributed computation and were therefore also tested on a five node cluster.
On a single node, different frameworks outperform others in computation time based on the problem under consideration.
For SSSP Polymer performed best while for BFS Ligra and Gemini best. PR was calculated by Galois the fastest.
When looking at the computation times of the distributed it stands out that the best time for each problem is beaten by most if not all of the single node times.
Comparing the distributed computation times SSSP and BFS results looked simular with Giraph being the fastest with the exception of syntetic graphs where Gemini was faster. The push version of Galois outperform the pull version but ultimativly was slower than Gemini. For distributed PR Gemini was the fastest by a margin.
In contrast, in terms of overall execution times, Galois almost always outperforms the other frameworks because the overhead is very small.
These benchmark results suggest that single node systems should be used whenever the graph fits into the RAM. 
If only individual calculations are performed on a graph, Galois is recommended due to its overall speed for a single node system. 
For a distributed system Galois is still a good choice if the algorithm is simular to SSSP and BFS. If the algorithm is more simular to PR Gemini would be a better choice.
In case of multiple calculations on a single graph, the graph doesn't need to be reloaded. Thus overall execution time is less important and only calculation time should to be considered.
As single node system, there is no single best framework. All choices excluding Giraph can be suitable here, depending on the exact circumstances. Gemini would be a safe bet being the second best framework often. 
In situations where a distributed system is required and the calculation time is important Giraph can be a good choice offering additional fault tolerance. Otherwise Gemini can be used.

\todo{Future work}

### 2020-09-23 documentation/paper/sections/conclusion.tex
kleine Korrekturen an Conclusion

@@
 We measured the calculation time of the algorithm itself and the overall execution time, which also includes the startup time of the programm.
-All of the frameworks were tested on a single compute node. Gemini, Galois and Giraph also support distributed computation and were therefore also tested on a five node cluster.
-On a single node, different frameworks outperform others in computation time based on the problem under consideration.
-For SSSP Polymer performed best while for BFS Ligra and Gemini best. PR was calculated by Galois the fastest.
-When looking at the computation times of the distributed it stands out that the best time for each problem is beaten by most if not all of the single node times.
-Comparing the distributed computation times SSSP and BFS results looked simular with Giraph being the fastest with the exception of syntetic graphs where Gemini was faster.
-The push version of Galois outperform the pull version but ultimativly was slower than Gemini.
-For distributed PR Gemini was the fastest by a margin.
+All of the frameworks were tested on a single compute node. Gemini, Galois and Giraph also support distributed computation and were therefore also tested on a five-node cluster.
+On a single node, the fastest framework in computation time is dependent on the problem under consideration.
+For SSSP, we found Polymer to perform best while for BFS Ligra and Gemini are fastest. PR was calculated by Galois the fastest.
+When looking at the computation times of the distributed systems, it stands out that the best time for each problem is beaten by most if not all of the single node times.
+Comparing the distributed computation times SSSP and BFS results looked simular with Giraph being the fastest with the exception of synthetic graphs where Gemini was faster.
+The push version of Galois outperforms the pull version but ultimativly was slower than Gemini.
+For distributed PR, Gemini was the fastest by a margin.
 In contrast, in terms of overall execution times, Galois almost always outperforms the other frameworks because the overhead is very small.
-These benchmark results suggest that single node systems should be used whenever the graph fits into the RAM. 
+These benchmark results suggest that single-node systems should be used whenever the graph fits into the RAM. 
 If only individual calculations are performed on a graph, Galois is recommended due to its overall speed for a single node system. 
-For a distributed system Galois is still a good choice if the algorithm is simular to SSSP and BFS. If the algorithm is more simular to PR Gemini would be a better choice.
-In case of multiple calculations on a single graph, the graph doesn't need to be reloaded. Thus overall execution time is less important and only calculation time should to be considered.
-As single node system, there is no single best framework.
+For a distributed system Galois is still a good choice if the algorithm is simular to SSSP and BFS. If the algorithm is more simular to PR, Gemini would be a better choice.
+In case of multiple calculations on a single graph, the graph does not need to be reloaded. Thus overall execution time is less important and only calculation time should to be considered.
+As single-node system, there is no single best framework.
 All choices excluding Giraph can be suitable here, depending on the exact circumstances.
 Gemini would be a safe bet being the second best framework often.
-In situations where a distributed system is required and the calculation time is important Giraph can be a good choice offering additional fault tolerance. Otherwise Gemini can be used.
+In situations where a distributed system is required and the calculation time is important, Gemini is faster than Giraph in most cases.
+However Giraph can be a good choice as well, offering additional features like node fault-tolerance.
 

### 2020-09-23 documentation/paper/main.tex
discussion
Conclusion angefangen
Abstract: Eigentlich alles gesagt, aber relativ kurz

@@
 \begin{abstract}
-\todo{Section}
+Graphs are growing fast and are becoming increasingly popular in many different domains.
+There are several different frameworks that make it easy to write graph algorithms.
+They often use different approaches and have significant performance differences on different graphs.
+In this paper we compare some state-of-the-art non-uniform memory access (NUMA) aware systems in terms of performance, both on real world and synthetic graphs.
+Therefore we consider the three algorithms single-source shortest-path (SSSP), breath-first search (BFS) and PageRrank (PR).
+We compare not only the time the algorithms indicate that they actually need to compute, but also the time they actually run and consume resources, and the resulting overhead time.
 \end{abstract}
@@
 
-\section{Conclusion}
-\todo{Section}
-
-\clearpage
+\input{sections/conclusion}
 

### 2020-09-23 documentation/paper/sections/conclusion.tex
Future work

@@
 All of the frameworks were tested on a single compute node. Gemini, Galois and Giraph also support distributed computation and were therefore also tested on a five-node cluster.
+
 On a single node, the fastest framework in computation time is dependent on the problem under consideration.
@@
 In contrast, in terms of overall execution times, Galois almost always outperforms the other frameworks because the overhead is very small.
+
 These benchmark results suggest that single-node systems should be used whenever the graph fits into the RAM. 
@@
 
-\todo{Future work}
+While comparing the frameworks we have noticed that the performance highly depends on the interaction between framework, algorithm and input graphs.
+Therefore there is room for further investigation here. These tests could incorporate new frameworks and new algorithms.
+Some of the tested systems offer a great range of settings and multiple implementations for the same problem. It would be iteresting to see a comparison there as well.
+At a later point in time, it is important to carry out such a comparison again because the frameworks are further developed and new ones are created.

### 2020-09-23 documentation/paper/sections/overview/galois.tex
various fixes

@@
 
-%Galois \cite{Galois} is a general purpose library designed for parallel programming.
-The Galois system supports fine grain tasks, allows for autonomous, speculative execution of these tasks and grants control over the task scheduling policies to the application.
-It also simplifies the implementation of parallel applications by providing an implicitly parallel unordered-set iterators.
-
 The graph analysis subsystem of Galois \cite{Galois} provides a library of scalable data structures and a topology aware priority scheduler, including optimizations for distributed execution. 
@@
 Every thread keeps a lazy cache of a portion of the global map, in order to reduce the strain on the global map. 
-Galois includes applications for many graph analytics problems, among these are SSSP, BFS and pagerank.
-For most of these applications Galois offers several different algorithms to perform these analytics problems and many options e.g.
-the amount of threads used or policies for splitting the graph.
+Galois includes applications for many graph analytics problems, among these are SSSP, BFS and PR. 
+For most of these applications Galois offers several different algorithms to solve these analytics problems and many options e.g. the amount of threads used or policies for splitting the graph. 
 All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment \cite{vertGalois}.
-
-%For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
-Galois includes applications for many graph analytics problems, among these are Single-Source Shorthest-Paths (SSSP), Breath-First Search (BFS) and PageRank (PR).
-For most of these applications Galois offers several different algorithms to perform these analytics problems and many setting options like the amount of threads used or policies for splitting the graph.
-All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment.
-
-%For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented.
-Galois includes applications for many graph analytics problems, among these are single-source shorthest-paths (sssp), breath-first-search (bfs) and pagerank.
-For most of these applications Galois offers several different algorithms to perform these analytics problems and many setting options like the amount of threads used or policies for splitting the graph.
-All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment.
 
@@
 Gluon is embedded in Galois, but can be integrated in other graph analysis frameworks as well \cite{vertGalois}.
-
-%The code of Gluon is embedded in Galois. It is possible to integrate Gluon in other frameworks \cite{vertGalois}.

### 2020-09-23 documentation/paper/sections/overview/gemini.tex
various fixes

@@
 It was developed with the goal to deliver a generally better performance through efficient communication.
-While most other graph processing systems achieve very good results in the shared-memory area, they often deliver unsatisfactory results in distributed computing.
+While most other graph processing systems achieve very good results in the shared-memory environment, they often deliver unsatisfactory results in distributed computing.
 Furthermore, a well optimized single-threaded implementation often outperforms a distributed system \cite{scalability}.
@@
 Gemini is fairly lightweight and has a clearly defined API between the core framework and the implementations of the individual algorithms.
-The five already implemented algorithms are Single-Source Shortest-Path (SSSP), Breath-First Search (BFS), PageRank (PR), Connected-Components (CC) and Betweenness-Centrality (BC).
+The five already implemented algorithms are Single-Source SSSP, BFS, PR, Connected-Components (CC) and Betweenness-Centrality (BC).

### 2020-09-23 documentation/paper/sections/overview/ligra.tex
various fixes

@@
 When \texttt{EdgeMap} is used Ligra decides between a push and a pull style execution.
-The decisive factor is the number of active vertices and their outdegree.
+The deciding factor is the number of active vertices and their outdegree.
 When this number is small a push style execution is used, otherwise a pull style execution is used.
@@
 This hybrid approach can increase performance significantly \cite{hybridBFS}.
-When push and pull mode changes also the representation of the \texttt{VertexSubset} is changed.
+When push and pull mode changes the representation of the \texttt{VertexSubset} is also changed.
 A push style execution corresponds to a sparse representation as array of the active vertex IDs, while a pull style execution corresponds to a dense representation as bitmap.

### 2020-09-23 documentation/paper/sections/overview/polymer.tex
various fixes

@@
 Vertices and Edges are partitioned and then allocated across the corresponding memory nodes of the threads, eliminating most remote memory accesses.
-However, some computation requires vertices to perform computations on edges that are not in the local NUMA-node.
+However, some computations require vertices to perform computations on edges that are not in the local NUMA-node.
 For this case, Polymer introduces lightweight vertex replicas that are used to initiate computation on remote edges.
@@
 
-Polymer does not only optimize data locality but also its scheduling is custom.
-The time to synchronize threads on different cores increases dramatically with the growing number of involved sockets.
+Polymer does not only optimize data locality but also utilizes custom scheduling.
+The time to synchronize threads on different cores increases dramatically with a growing number of involved sockets.
 Inter-node synchronization takes one order of magnitude longer time than intra-node synchronization \cite{Polymer}.
 Thus, Polymer implements a topology-aware hierarchical synchronization barrier.
-A group of threads on the same NUMA-node shares a partition of data.
+A group of threads on the same NUMA-node share a partition of data.
 This allows them to first only synchronize with threads on the same NUMA-node.
@@
 The leaves use bitmaps, which are efficient for a large proportion of active vertices.
-When only a small amount of vertices is active, the drawbacks of traversing through sparse bitmaps can be avoided by switching data structures.
+When only a small number of vertices are active, the drawbacks of traversing through sparse bitmaps can be avoided by switching data structures.
 

### 2020-09-23 documentation/paper/sections/preliminiaries.tex
various fixes

@@
 To counteract this problem, hugepages were developed, which are usually several mibibibytes in size.
-This minimizes the CPU time needed for table lookups, because there are much less TBL misses.
-This speedup is especially noticeable in applications that are very memory intensive, like graph processing systems.
+This minimizes the CPU time needed for table lookups, because there are fewer TBL misses.
+This speedup is especially noticeable in very memory intensive applications, like graph processing systems.

### 2020-09-23 documentation/paper/sections/relatedWork.tex
various fixes

@@
 
-It quickly becomes apparent that many of the comparisons provided
-
 The comparisons often include PowerGraph or Apache's Spark GraphX, two frameworks that we did not include in our testing.
 
-In 2013 \cite{Galois} was last compared to GraphLab, PowerGraph and Ligra in a shared memory setting.
+In 2013 \cite{Galois} Galois was last compared to GraphLab, PowerGraph and Ligra in a shared memory setting.
 The results of their comparison are mainly, that Galois outperforms Ligra on BFS and SSSP on most graphs.
@@
 About 5 years afterwards the capability of executing distributed was added to Galois through Gluon \cite{vertGalois}. 
-There distributed Galois is being compared to Gemini and distributed versions of Ligra and IrGl, where the distributed versions where created using Gluon. 
+There distributed Galois is being compared to Gemini and distributed versions of Ligra and IrGL, where the distributed versions where created using Gluon. 
 IrGL is a graph analysis framework utilizing a GPU. 

### 2020-09-23 documentation/paper/sections/results/galoisSpeedup.tex
merge und besprechung
discussion
various fixes

@@
 It is obvious that enabling hugepages greatly improves performance in almost every case.
-This means that on every algorithm and almost all graphs, calculation time and execution tame can be reduced significantly with just the use of hugepages.
+This means that on every algorithm and almost all graphs, calculation time and execution time can be reduced significantly with just the use of hugepages.
 The only exception is twitter on both PageRank implementations, where the execution time is smaller without hugepages (74.9s vs 201.3s for Push and 8.7s vs 9.3s for Pull with and without hugepages respectively).
@@
 
-Keep in mind, that we have shown the results of 96 threads in this section. The following sections compare the behavoiur of Galois on different numbers of threads.
+Keep in mind, that those were the results for 96 threads. The following sections compare the behavoiur of Galois on different numbers of threads.
 
@@
 \bf1 & 1.0 & 1.0 & 0.0 & 0.0 \\
-\bf2 & 1.6 & 1.6 & 0.2 & 0.1 \\
-\bf4 & 2.5 & 2.5 & 1.1 & 0.9 \\
-\bf8 & 4.5 & 3.4 & 5.3 & 2.1 \\
-\bf16 & 6.7 & 5.8 & 13.0 & 7.3 \\
-\bf32 & 9.6 & 7.0 & 38.3 & 10.8 \\
-\bf48 & 10.3 & 10.7 & 41.4 & 31.4 \\
-\bf96 & 10.2 & 10.8 & 38.1 & 29.9 \\
+\bf2 & 1.6 & 1.6 & 0.2 & \bf 0.1 \\
+\bf4 & 2.5 & 2.5 & 1.1 & \bf 0.9 \\
+\bf8 & \bf 4.5 & 3.4 & 5.3 & \bf 2.1 \\
+\bf16 & \bf 6.7 & 5.8 & 13.0 & \bf 7.3 \\
+\bf32 & \bf 9.6 & 7.0 & 38.3 & \bf 10.8 \\
+\bf48 & 10.3 & \bf 10.7 & 41.4 & \bf 31.4 \\
+\bf96 & 10.2 & \bf 10.8 & 38.1 & \bf 29.9 \\
 \bottomrule
@@
 For 40 threads, the speedup is nearly 19$\times$, on 48 threads 17$\times$ and with 96 threads only 15$\times$ compared to one thread.
-Small graphs, i.e. flickr and orkut neither benefit from more threads nor is the performance significantly held up by synchronization overhead.
+Small graphs, i.e. flickr and orkut neither benefit from more threads nor is the performance significantly reduced by the synchronization overhead.
 Performance on flickr can not be sped up at all, with speedup on flickr being very close to 1 for 1 to 8 threads and between 0.7$\times$ to 0.9$\times$ from 16 to 96 threads.

### 2020-09-23 documentation/paper/sections/results/results.tex
various fixes

@@
 
-	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code.
+	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and to modify the source code.
 Our changes can be found in one of our repositories\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.
@@
 On a plus side, setup of the frameworks Polymer and Ligra was straight forward and did not require any special treatment.
+Galois is also fairly easy to setup without hugepages and the process is well documented. 
+Setting up hugepages might require some research and a bit of trial and error to find the ideal way, since there are several ways to achive the same.
 \todo{Gemini? Außerdem sollte hier noch ein bisschen was hin, die beiden Zeilen sehen ja nur traurig aus.}

### 2020-09-23 documentation/paper/sections/testingMethods.tex
Conclusion angefangen
various fixes

@@
 All five frameworks are tested on a single server.
-Galois, Gemini and Giraph were benchmarked in on the distributed 5-node cluster as well.
+Galois, Gemini and Giraph were benchmarked on the distributed 5-node cluster as well.
 Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
@@
 	\item Ligra supports SSSP based on BellmanFord, BFS and two implementations of PageRank. The two implementations are a regular PR and a Delta Variant.
-	\item Polymer supports the same algoritms as Ligra.
-	\item Gemini supports all of our tested algorithms and there are no setting options or specifications which implementations for the algorithms are used.
+	\item Polymer supports the same algorithms as Ligra.
+	\item Gemini supports all of our tested algorithms. For SSSP a BellmanFord implementation is used.
 	\item Galois supports all of our tested algorithms too, with both a Push and a Pull variant for PageRank available.
@@
 The default implementation of SSSP is deltaTile.
-A lot of setting options are avilable as well, but we're gone with the defaults.
+A lot of options are available as well, but we have gone with the defaults.
 	\item Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
@@
 
-	This is not the most realiable way for measuring the calculation times.
+	This is not the most realiable way of measuring the calculation times.
 	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.

### 2020-09-23 documentation/paper/sections/discussion.tex

@@
 We have shown that the distributed cluster of the same framework becomes faster than the equivalent single-node setup only on large graphs.
-We have found around 2B edges, or roughly the size of the twitter graph to be a size at which Ligra and Giraph are faster on the distributed setup compared to single-node.
+We have found around 2B edges, or roughly the size of the twitter graph to be a size at which Gemini and Giraph are faster on the distributed setup compared to single-node.
 If required, horizontal scaling can for example be used to provide fault-tolerance of the entire system.
@@
 Here we recommend Galois as the fastest executing single-node framework overall.
-It provides excellent performance, being orders of magnitude faster than the competing frameworks.
+It provides excellent performance, being orders of magnitude faster than any of the competing frameworks.
 And when it is possible to use hugepages, further improvements in performance can be expected.

### 2020-09-23 documentation/paper/main.tex
kleine Korrekturen

@@
 In this paper we compare some state-of-the-art non-uniform memory access (NUMA) aware systems in terms of performance, both on real world and synthetic graphs.
-Therefore we consider the three algorithms single-source shortest-path (SSSP), breath-first search (BFS) and PageRrank (PR).
-We compare not only the time the algorithms indicate that they actually need to compute, but also the time they actually run and consume resources, and the resulting overhead time.
+Therefore, we consider the three algorithms Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
+We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, and the resulting overhead time.
 \end{abstract}

### 2020-09-23 documentation/paper/main.tex
kleine Korrekturen

@@
 \begin{IEEEkeywords}
-graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon, Gemini
+graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon, Gemini, SSSP, BFS, PR, benchmark
 \end{IEEEkeywords}
@@
 
+\bibitem{hybridBFS}
+S.~{Beamer}, K.~{Asanovic}, and D.~{Patterson}, ``Direction-optimizing
+  breadth-first search,'' in \emph{SC '12: Proceedings of the International
+  Conference on High Performance Computing, Networking, Storage and Analysis},
+  2012, pp. 1--10.
+
 \bibitem{Gemini}

### 2020-09-23 documentation/paper/sections/conclusion.tex
some more minor fixes

@@
 When looking at the computation times of the distributed systems, it stands out that the best time for each problem is beaten by most if not all of the single node times.
-Comparing the distributed computation times SSSP and BFS results looked simular with Giraph being the fastest with the exception of synthetic graphs where Gemini was faster.
-The push version of Galois outperforms the pull version but ultimativly was slower than Gemini.
+Comparing the distributed computation times SSSP and BFS results looked similar with Giraph being the fastest with the exception of synthetic graphs where Gemini was faster.
+The push version of Galois outperforms the pull version but ultimatively was slower than Gemini.
 For distributed PR, Gemini was the fastest by a margin.
@@
 If only individual calculations are performed on a graph, Galois is recommended due to its overall speed for a single node system. 
-For a distributed system Galois is still a good choice if the algorithm is simular to SSSP and BFS. If the algorithm is more simular to PR, Gemini would be a better choice.
+For a distributed system Galois is still a good choice if the algorithm is similar to SSSP and BFS. If the algorithm is more similar to PR, Gemini would be a better choice.
 In case of multiple calculations on a single graph, the graph does not need to be reloaded. Thus overall execution time is less important and only calculation time should to be considered.
@@
 All choices excluding Giraph can be suitable here, depending on the exact circumstances.
-Gemini would be a safe bet being the second best framework often.
+Gemini would be a safe bet, often being the second best framework.
 In situations where a distributed system is required and the calculation time is important, Gemini is faster than Giraph in most cases.
@@
 Therefore there is room for further investigation here. These tests could incorporate new frameworks and new algorithms.
-Some of the tested systems offer a great range of settings and multiple implementations for the same problem. It would be iteresting to see a comparison there as well.
-At a later point in time, it is important to carry out such a comparison again because the frameworks are further developed and new ones are created.
+Some of the tested systems offer a great range of settings and multiple implementations for the same problem. It would be interesting to see a comparison there as well.
+At a later point in time, it is important to repeat such a comparison, because the frameworks are further developed and new ones are created.

### 2020-09-23 documentation/paper/sections/discussion.tex
some more minor fixes

@@
 Thus, there is an argument to be made for systems like Hadoop.
-It automatically handles node failures, whith a node failure not immediately resulting in loss of data or faulty computation results.
+It automatically handles node failures, with a node failure not immediately resulting in loss of data or faulty computation results.
 This is especially important on very large clusters, that can not easily be monitored by humans.
@@
 If only the calculation time is important, there are multiple options with the deciding factors being the graph size and the application.
-For algorithms that perform a traversal through the graph, with a concentrated and relativly small amount of active vertices, our results for SSSP and BFS are most relevant.
+For algorithms that perform a traversal through the graph, with a concentrated and relatively small number of active vertices, our results for SSSP and BFS are most relevant.
 Here Galois is fastest on very small graphs (i.e. flickr), but is quickly overtaken by other frameworks.

### 2020-09-23 documentation/paper/sections/results/results.tex
kleine Korrekturen
some more minor fixes

@@
 
-\begin{enumerate}
+\begin{itemize}
 	\item During setup and benchmark of Gemini, we encountered several bugs in the cloned repository. These include non zero-terminated strings or even missing return statements.
@@
 Second, some log files that can easily be multiple gigabytes in size are stored outside of the HDFS and are never mentioned in the Giraph documentation.
-\end{enumerate}
-
+\end{itemize}
+\noindent
 On a plus side, setup of the frameworks Polymer and Ligra was straight forward and did not require any special treatment.
@@
 Setting up hugepages might require some research and a bit of trial and error to find the ideal way, since there are several ways to achive the same.
-\todo{Gemini? Außerdem sollte hier noch ein bisschen was hin, die beiden Zeilen sehen ja nur traurig aus.}
 

### 2020-09-24 documentation/paper/main.tex
Rückmeldung übernommen und todos eingefügt

@@
 \begin{abstract}
-Graphs are growing fast and are becoming increasingly popular in many different domains.
-There are several different frameworks that make it easy to write graph algorithms.
+Graphs are growing fast and are becoming increasingly popular in many different domains, like \todo{...}
+There are several different frameworks that help writing graph algorithms.
 They often use different approaches and have significant performance differences on different graphs.
+\todo{Giraph nicht NUMA aware, da unten raus nehmen..}
 In this paper we compare some state-of-the-art non-uniform memory access (NUMA) aware systems in terms of performance, both on real world and synthetic graphs.
-Therefore, we consider the three algorithms Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
+\todo{umformulieren}
+Therefore, we consider the three algorithms: Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
 We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, and the resulting overhead time.

### 2020-09-24 documentation/paper/sections/conclusion.tex
Rückmeldung übernommen und todos eingefügt

@@
 We compared the performance of several graph processing frameworks including Ligra, Polymer, Gemini, Galois and Giraph. 
-For comparison we used the three problems SSSP, BFS and PR on graphs of various sizes.
-We measured the calculation time of the algorithm itself and the overall execution time, which also includes the startup time of the programm.
+For comparison we used the three graph algorithms SSSP, BFS and PR on data sets of various sizes.
+We measured the calculation time of the algorithm itself and the overall execution time, which also includes the framework overhead.
 All of the frameworks were tested on a single compute node. Gemini, Galois and Giraph also support distributed computation and were therefore also tested on a five-node cluster.
 
+\todo{Ab hier kürzer fassen}
 On a single node, the fastest framework in computation time is dependent on the problem under consideration.

### 2020-09-24 documentation/paper/sections/discussion.tex
Rückmeldung übernommen und todos eingefügt

@@
 Further, a user defining an application does not need to know the topology of the system that runs the application later on.
-Hadoop transparently splits the input data and distributes work among the worker nodes.
-Additionally, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
+Hadoop splits the input data and distributes work among the worker nodes.
+Additionally, Giraph using hadoop shows a bias towards larger graphs. We observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
 This shows the hadoop infrastructure to be working well with large graphs, which is exactly what the distributed system should be used for. 
 Only a system like hadoop can reasonably be used on such large clusters.
-Any of the other framewors will just result in large administration overhead due to node failures or other problems of distributed systems in general.
+Any of the other evaluated frameworks will just result in large administration overhead due to node failures or other problems of distributed systems in general.
 Furthermore, the calculation times of Giraph proved to be very fast in comparison to the other distributed frameworks, with Giraph often being one of the fastest framework in computation time.
 Very short calculation times along with automatic handling of the computation cluster make Giraph using Hadoop a good choice for production systems.
-Pregel is a similar, fault-tolerant and scalable system but it is contrary to Giraph closed-source.
+GraphX and Pregel are similar, fault-tolerant and scalable systems. GraphX uses Apache Spark, extending on hadoop. Pregel is contrary to Giraph or GraphX closed-source.
 
@@
 We have shown that the distributed cluster of the same framework becomes faster than the equivalent single-node setup only on large graphs.
-We have found around 2B edges, or roughly the size of the twitter graph to be a size at which Gemini and Giraph are faster on the distributed setup compared to single-node.
+We have found that at around 2B edges (roughly the size of the twitter graph) Gemini and Giraph are faster on the distributed setup compared to single-node.
 If required, horizontal scaling can for example be used to provide fault-tolerance of the entire system.
@@
 In the case of SSSP those are Gemini and Polymer and on BFS they are Ligra and again Gemini.
-While Gemini is not the fastest on either BFS or SSSP, it is still the safest bet on other algorithms.
+While Gemini is not the fastest on either BFS or SSSP, it is still a good choice on other algorithms.
 That is because Polymer is fast on SSSP but slowest on BFS and in turn Ligra is fastest on BFS but one of the slowest frameworks on SSSP.

### 2020-09-24 documentation/paper/sections/introduction.tex
some more minor fixes
Rückmeldung übernommen und todos eingefügt

@@
 \section{Introduction}
-
+\todo{Hier gibts einige Todos, am besten nochmal nachschauen}
 In recent years, graph processing has become increasingly important, both in academic and industrial environments \cite{Gemini}.
@@
 Because most of the systems are several years old and some of them have evolved, or are still evolving, these comparisons were often incomplete or are simply no longer relevant today. 
-Other systems are simply no longer used today, which also reduces the significance of some of the existing comparisons.
+Other systems are no longer in use, which also reduces the significance of some of the existing comparisons.
 
@@
 
-The rest of this paper is organized as follows.
-In section 2 we first clarify what a graph is, how the algorithms we are looking at work and what else is necessary for further understanding.
-Then we give a short overview of the frameworks we comparing in section 3 and of the graph formats used in section 4.
-Before we come to our results, we show in section 5 what has been done in this area so far. 
-In Section 6, which is the core of the paper, we will present all our results and also briefly show how they were developed. 
-After that, in section 7, we will discuss these results. And at the end in section 8 we will present our conclusion.
+The rest of this paper is organized as follows:
+Section 2 is dedicated to all preliminaries, in order to keep this paper self-contained.
+After that, we give a short overview of each framework in our comparison.
+Section 4 explains the used graph formats. 
+Related work is discussed in section 5.
+In Section 6 we present our evaluation, followed by a discussion of these results in section 7. Finally we conclude the paper in section 8.

### 2020-09-24 documentation/paper/sections/overview/giraph.tex
various fixes
Rückmeldung übernommen und todos eingefügt

@@
 
-In order for Giraph a achieve scalability and parallelization, it is built on top of Apache Hadoop \cite{Giraph}.
+In order for Giraph to achieve scalability and parallelization, it is built on top of Apache Hadoop \cite{Giraph}.
 Hadoop is a MapReduce infrastructure providing a fault tolerant basis for large scale data processing.
 Hadoop supplies a distributed file system (HDFS), on which all computations are performed.
-Giraph is thus, even when only using a single node, running in a distributed manner.
+Giraph is thus, even when only using a single computation node, running in a distributed manner.
 Hence, expanding single-node processing to a multi-node cluster is seamless.

### 2020-09-24 documentation/paper/sections/overview/overview.tex
merge und besprechung
Rückmeldung übernommen und todos eingefügt

--- 2020-09-05 documentation/paper/sections/overview/overview.tex
@@
 \section{Framework Overview}
-We start with a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
-All of the following frameworks are part of our testing.
-
-\subsection{Galois and Gluon}
-\input{sections/overview/galois}
-
-\subsection{Gemini}
-\input{sections/overview/gemini}
-
-\subsection{Giraph}
-\input{sections/overview/giraph}
+The following paragraphs present a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
+All of the described frameworks are part of our evaluation.
 

### 2020-09-24 documentation/paper/sections/preliminiaries.tex
Rückmeldung übernommen und todos eingefügt

@@
 \end{equation*}
-and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E$, it holds $(x,y)=(y,x)$.
+and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E: (x,y)=(y,x)$ holds.
 The size of a graph is defined as the number of edges $|E|$ \cite{newman2010networks}.
@@
 all neighbors of the start vertex are checked before checking paths of length two and so on.
-The search algorithm, where the paths of maximum length are checked first is called Depth-First search.
 
@@
 \emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}.
-The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.
+The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses a modified version of PageRank to rank web pages in their search engine results.
 It is also used as a centrality metric.
@@
 This percentage, called the PageRank of the vertex, is the probability of arriving at this vertex when starting at a random vertex and then following random edges.
-With a small probability, the damping factor, a jump is made to a completely random vertex instead of following a edge.
+With a small probability, the damping factor, a jump is made to a completely random vertex instead of following an edge.
 The PageRank can be calculated by iterating
@@
 \subsection{Push and Pull Variants}
+\todo{kompliziert}
 Many parallel graph algorithms, including the three we consider here, are implemented by iterating over vertices \cite{tao}.
@@
 A user specifies just the two functions Map and Reduce.
-The system hides the details of parallelization, fault-tolerance, data distribution and load balancing away from the application logic.
+The system hides the details of parallelization: fault-tolerance, data distribution and load balancing away from the application logic.
 All of these features are automatically provided.
@@
 Translating from the virtual to the physical address space significantly reduces the performance of today's data processing systems, since the size of the RAM grows much faster than the size of the TBL \cite{hugepages, superpages}.
-To counteract this problem, hugepages were developed, which are usually several mibibibytes in size.
+To counteract this problem, hugepages were developed, which are usually several mibibytes in size.
 This minimizes the CPU time needed for table lookups, because there are fewer TBL misses.

### 2020-09-24 documentation/paper/sections/results/bfs.tex
some work on discussion/comparison
merge und besprechung
various fixes
Rückmeldung übernommen und todos eingefügt

@@
 As with our results for SSSP, for BFS we will begin with our single-node results before looking at the distributed scenario.
-
-\todo{Einleitung, analog SSSP}
 
@@
 	\end{subfigure}
-	\caption{Average times for BFS on a single computation node, black bars represent one standard deviation in our testing}
+	\caption{Average times for BFS on a single computation node}
 \end{figure*}
-Just like with SSSP, Giraph ran out of memory ($>$250 GB) on any graph larger than wikipedia.
-Also, Polymer failed to complete on rMat28.
-Ligra, that failed during SSSP however completed our benchmark for BFS.
+Giraph ran out of memory ($>$250 GB) on any graph larger than wikipedia. Also, Polymer failed to complete on rMat28. Ligra, that failed during SSSP however completed our benchmark for BFS.
 
@@
 
-For the remaining frameworks, we compare to Ligra since it is generally slightly faster than Gemini.
-Giraph and Polymer are one to two orders of magnitude slower than Ligra. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
+We compare the remaining frameworks to Ligra because it is generally slightly faster than Gemini.
+Giraph and Polymer are one to two orders of magnitude slower. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
 Giraph takes between 13$\times$ and 32$\times$ longer than Ligra on flickr, orkut or wikipedia.
-As we said, Polymer is comparable to Ligra on flickr, but for the other graphs, its computation times are even longer than those of Giraph.
-The upside is though, that Polymer managed -- contrary to Giraph -- to finish computation on some larger graphs.
+As already stated, Polymer is comparable to Ligra on flickr, but for the other graphs, its calculation times are even longer than those of Giraph.
+However, Polymer managed -- contrary to Giraph -- to finish computation on some larger graphs.
 Polymer takes between 55$\times$ and 530$\times$ longer than Ligra on the graphs larger than flickr.
-Especially interesting is here the difference between the times of rMat27 and friendster.
+Here, the difference between the times of rMat27 and friendster are especially interesting.
 There are 20\% (439M) more edges in friendster, yet the computation for the synthetic rMat27 takes 42\% longer.
@@
 The execution time results are very similiar to our findings of SSSP (cf.
-\autoref{fig:singleNodeBFS_exec}).
+\autoref{fig:singleNodeSSSP_exec}).
 Just like on SSSP, Galois is the fastest framework and Ligra is second fastest on all graphs except flickr.
@@
 	\hfil
-	% \begin{subfigure}{0.3\textwidth}
-	% 	\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
-	% 	\caption{Overhead time normalized by the graph size in million edges}
-	% 	\label{fig:distributedBFS_overheadNormalized}
-	% \end{subfigure}
-	% \hfil
-	\caption{Average times for BFS on the distributed cluster, black bars represent one standard deviation in our testing}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTime.png}
+		\caption{Overhead}
+		\label{fig:distributedBFS_overhead}
+	\end{subfigure}
+	\hfil
+	\caption{Average times for BFS on the distributed cluster}
 	\label{fig:distributedBFS}
@@
 This is expected since both are graph traversal algorithms starting in one source vertex.
-Hence, calculation complexity for each vertex and communication overhead is similar. All measurements can be seen in \autoref{fig:distributedBFS}.
+Hence, calculation complexity for each vertex and communication overhead is comparable. All measurements can be seen in \autoref{fig:distributedBFS}.
 
@@
 This makes Galois Push the clear winner for distributed BFS.
-
-\todo{Discussion ausarbeiten}

### 2020-09-24 documentation/paper/sections/results/galoisSpeedup.tex
kleine Korrekturen
Rückmeldung übernommen und todos eingefügt

@@
 \end{table}
-When comparing the performance with and without hugepages, the results really speak for themselves they can be seen in \autoref{tbl:timeComparison_HPnoHP}.
-It is obvious that enabling hugepages greatly improves performance in almost every case.
+When comparing the performance with and without hugepages, the results show that enabling hugepages greatly improves performance in almost every case, cf.
+\autoref{tbl:timeComparison_HPnoHP}.
 This means that on every algorithm and almost all graphs, calculation time and execution time can be reduced significantly with just the use of hugepages.
@@
 In all other cases (that's 26 of 28 test cases), the hugepage-version is significantly faster.
-The execution time is approximately halved on SSSP and BFS. And on PageRank, the execution time is reduced to 35\% to 87\% of the time without hugepages.
+The execution time is approximately halved on SSSP and BFS. 
+On PageRank, the execution time is reduced to a value in the range of 35\% to 87\% of the time without hugepages.
 
@@
 \bf1 & 1.0 & 1.0 & 0.0 & 0.0 \\
-\bf2 & 1.6 & 1.6 & 0.2 & \bf 0.1 \\
-\bf4 & 2.5 & 2.5 & 1.1 & \bf 0.9 \\
+\bf2 & \bf 1.6 & 1.6 & 0.2 & \bf 0.1 \\
+\bf4 & 2.5 & \bf 2.5 & 1.1 & \bf 0.9 \\
 \bf8 & \bf 4.5 & 3.4 & 5.3 & \bf 2.1 \\
@@
 
-Starting with SSSP, an algorithm that greatly benefits from many available threads. The results can be seen in \autoref{fig:galoisSpeedupSSSP}.
+Starting with SSSP, we observe Galois' SSSP implementation to be greatly benefitting from many available threads. The results can be seen in \autoref{fig:galoisSpeedupSSSP}.
 We first look at the speedups without hugepages, seen in \autoref{fig:galoisSpeedupSSSP_noHP}.
@@
 Behaviour on friendster is similarly good. Here speedup is 1.9$\times$ at 2 threads compared to one, 3.5$\times$ at 4, 6.1$\times$ at 8 threads and 9.7$\times$ at 16 threads.
-Anything above 16 threads however no longer helps decrease the computation time significantly on any graph.
+Anything above 16 threads however no longer helps decreasing the computation time significantly on any graph.
 Speedup above 16 threads is always less than double the speedup of 16 threads.
@@
 \end{table}
-One might assume the two algorithms SSSP and BFS to behave very similarly, yet they produce very different results.
+One might assume the two algorithms SSSP and BFS to behave very similarly, yet they produce very different results. \todo{Begründung, z.b. Frontier Set}
 With SSSP, we observed a decrease in the variance of the speedups, while the average speedup decreased or stayed the same. This provided a more reliable speedup. 
@@
 This maximum is reached at two threads, with speedup steadily declining above that.
-The rMat28 is the only other graph of one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 1.3$\times$ at 96 threads.
+The rMat28 is the only other graph one could say computation was sped up at large thread counts. Here we reached a maximum speedup of 1.3$\times$ at 96 threads.
 All 5 other graphs only reach a speedup greater or equal to 1 in just one or two cases, and if so only by a small margin.
-Computation on Orkut and Twitter reaches a speedup maximum of 1.1$\times$ and 1.05$\times$ at 4 threads, while being less or equal to 1 in all other cases.
-Speedup on the wikipedia graph is never grater than one.
-Friendster and rMat27 can be sped up to 1.07$\times$ or 1.1$\times$ respectively on 8 threads.
+% Computation on Orkut and Twitter reaches a speedup maximum of 1.1$\times$ and 1.05$\times$ at 4 threads, while being less or equal to 1 in all other cases.
+% Speedup on the wikipedia graph is never grater than one.
+% Friendster and rMat27 can be sped up to 1.07$\times$ or 1.1$\times$ respectively on 8 threads.
 
-Most of this changes drastically with the activation of hugepages, our data can be seen in \autoref{fig:galoisSpeedupPRPull_HP}. 
+Most of this changes drastically with the activation of hugepages. Our data can be seen in \autoref{fig:galoisSpeedupPRPull_HP}. 
 Here actually all graphs except flickr reach a speedup greater than 1.5$\times$. Remember, that 1.6$\times$ was the maximum possible speedup without hugepages.

### 2020-09-24 documentation/paper/sections/results/pr.tex
some work on discussion/comparison
pagerank introduction
merge und besprechung
Rückmeldung übernommen und todos eingefügt

@@
 \subsubsection{PageRank}
-\todo{Einleitung}
+In this section we compare the PageRank performance of the various frameworks. As usual we begin with the single-node performance and finish by discussing the distributed variants.
 
@@
 For the other graphs, the delta version is slower by a far greater margin of 13\% to 68\%.
+Because we only run five iterations of PR, the optimizations of the delta algorithm over the regular implementation are not yet visible.
 Hence, we only show the results of Ligra's regular PageRank implementation in our evaluation.
-For Polymer we found the delta version to be faster on all graphs except rMat28. Delta-PR is on average 15\%\ faster on the first six graphs, while only being 0.3\%\ slower on rMat28.
+Polymer's delta implementation on the other hand can use the optimizations on a much smaller number of iterations to its advantage.
+For Polymer we found the delta version to be faster on all graphs except rMat28.
+Delta-PR is on average 15\%\ faster on the first six graphs, while only being 0.3\%\ slower on rMat28.
 Thus, the following only shows Polymer's faster Delta-PR implementation.
@@
 	\hfil
-	\caption{Average times for PR on a single computation node, black bars represent one standard deviation in our testing}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTIme.png}
+		\caption{Overhead}
+		\label{fig:singleNodePR_overhead}
+	\end{subfigure}
+	\hfil
+	\caption{Average times for PR on a single computation node}
 	\label{fig:singleNodePR}
@@
 However, we are very suspicious of these results and thus exclude the calculation time for Galois Push in further comparisons.
-Because the execution time of Galois Push is always considerably longer than the execution time of Galois Pull (cf.
-\autoref{fig:singleNodePR_calc}).
-This leads us to believe that the output that we used for our measurements contains an error.
+We believe that the output we used for our measurements contains an error,
+because the \emph{execution} time of Galois Push is always considerably longer than that of Galois Pull (cf. \autoref{fig:singleNodePR_exec}).
 
 For the three graphs on which Giraph computed successfully, it is the slowest framework in both calculation and execution times.
-And that by a difference of three orders of magnitude in the calculation time and one to two orders of magnitude in execution time (cf \autoref{fig:singleNodePR}).
+Further, it is the slowest by three orders of magnitude in the calculation time and one to two orders of magnitude in the execution time (cf \autoref{fig:singleNodePR}).
 On the larger graphs (i.e.
@@
 Interestingly, the execution time for Ligra is at a maximum for twitter. The required time is steadily decreasing with increasing graph size.
+\todo{Begründung?}
 
@@
 	\hfil
-	\caption{Average times for PR on the distributed cluster, black bars represent one standard deviation in our testing}
+	\caption{Average times for PR on the distributed cluster}
 	\label{fig:distributedPR}
@@
 And Galois Push has smaller execution times than the Pull version because the overhead times for both implementations are similar (cf. \autoref{fig:distributedPR_overhead}).
-
-\todo{Discussion}

### 2020-09-24 documentation/paper/sections/results/sssp.tex
various fixes
Rückmeldung übernommen und todos eingefügt

@@
 \subsubsection{Single-Source Shortest-Paths}
-In this section, we compare the different frameworks in their performance on the SSSP algorithm.
-We first analyze the frameworks on a single computation node and compare to the distributed setup after that.
+In this section, we compare the different frameworks in their performance on the SSSP algorithm. We first analyze the frameworks on a single computation node and then compare to the distributed setup.
 \paragraph{Single-node}
@@
 	\end{subfigure}
-	\caption{Average times for SSSP on a single computation node, black bars represent one standard deviation in our testing.}
+	\caption{Average times for SSSP on a single computation node}
 	\label{fig:singleNodeSSSP}
@@
 
-Upon analyzing the calculation time (cf.
-\autoref{fig:singleNodeSSSP_calc}), most obvious is the fact that Giraph is at least one order of magnitude slower than any other framework.
-The only exception is Gemini, where Giraph is \emph{only} 4$\times$ slower on wikipedia.
-In general however, Giraph is on average 24$\times$ slower than the other frameworks on flickr, 17$\times$ on orkut and 12$\times$ on wikipedia.
-
+Upon analyzing the calculation time (cf. \autoref{fig:singleNodeSSSP_calc}), most obvious is the fact that Giraph is at least one order of magnitude slower than any other framework on most graphs.
 Comparing the other frameworks (i.e.
@@
 There, the average calculation times of the four frameworks (Galois, Gemini, Ligra, Polymer) are 35ms on flickr, 77ms on orkut and 538ms on wikipedia.
-The four frameworks are close to this average for the smaller graphs. Galois and Gemini being 24ms and 12ms faster than the average. Ligra and Polymer are slower by 27ms and 9ms.
+The four frameworks are close to this average for the smaller graphs. Galois and Gemini are slightly faster than the average, while Ligra and Polymer are slower.
 Orkut is the graph where all the four frameworks are within 20ms of their average, Galois being the outlier at 19ms slower than average.
 For wikipedia, Galois and Ligra are close to each other, while Gemini is 560ms (104\%) slower than the average of the four frameworks.
-And Polymer is the fastest framework on wikipedia, being 330ms (61\%) faster than the others.
-For the larger graphs, Polymer has the shortest computation time as well.
-Gemini is second, taking about 1.9$\times$ times the computation time of Polymer on twitter and friendster while being equally fast (8\% longer) on rMat27.
-Galois and Ligra however have much longer computation times compared to Polymer.
-Galois takes anywhere from 3.3$\times$ (twitter) to 4.2$\times$ (friendster) the computation time of Polymer.
-Ligra requires between 3.2$\times$ (friendster) and 9.8$\times$ (twitter) the computation time of Polymer on the larger graphs.
+Polymer is the fastest framework on wikipedia, being 330ms (61\%) faster than the others.
+For the larger graphs, Polymer has the shortest calculation time as well.
+Gemini is second, taking about 1.9$\times$ times the calculation time of Polymer on twitter and friendster while being equally fast (8\% longer) on rMat27.
+Galois and Ligra however have much longer calculation times compared to Polymer.
+Galois takes anywhere from 3.3$\times$ (twitter) to 4.2$\times$ (friendster) the calculation time of Polymer.
+Ligra requires between 3.2$\times$ (friendster) and 9.8$\times$ (twitter) the calculation time of Polymer on the larger graphs.
 This makes Ligra the slowest Framework on both twitter and rMat27.
 Galois is the slowest on friendster.
-However, while especially Galois is comparably slow to Polymer in the computation time, it is important to keep in mind that Polymer could not finish computation on rMat28.
-Meanwhile both Gemini and Galois managed just fine.
+However, while especially Galois is comparably slow to Polymer in the calculation time, it is important to keep in mind that Polymer could not finish computation on rMat28.
+Meanwhile both Gemini and Galois completed their calculations without any problems.
 
-The execution times show Galois to be the clear winner on most graphs (cf. \autoref{fig:singleNodeBFS_exec}).
-It has the smallest computation times on all graphs except flickr, where Galois is 25ms (8\%) slower than Ligra. 
+The execution times show Galois to be the clear winner on most graphs (cf. \autoref{fig:singleNodeSSSP_exec}).
+It has the smallest calculation times on all graphs except flickr, where Galois is 25ms (8\%) slower than Ligra. 
 On the other 5 graphs however, Galois has the smaller execution time by one or sometimes even two orders of magnitude. 
@@
 	\end{subfigure}
-	\caption{Average times for SSSP on the distributed cluster, black bars represent one standard deviation in our testing.}
+	\caption{Average times for SSSP on the distributed cluster}
 	\label{fig:distributedSSSP}
@@
 When taking a closer look at Giraph, it seems to not cope well with synthetic data sets.
-Analyzing the computation times in \autoref{fig:distributedSSSP_calc}, we see that it is the fastest framework on our real-world graphs.
+Analyzing the calculation times in \autoref{fig:distributedSSSP_calc}, we see that it is the fastest framework on our real-world graphs.
 And that with a considerable margin of other frameworks always taking at least 50\% longer (lower bound here is Gemini on flickr) up to Galois Pull needing 18$\times$ more time on wikipedia.
 On both synthetic graphs however, Giraph is actually the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.
-While Giraph's computation times are very competitive, when comparing the execution times in \autoref{fig:distributedSSSP_exec} we see that Giraph is actually the slowest framework on 5 out of 7 graphs.
+While Giraph's calculation times are very competitive, when comparing the execution times in \autoref{fig:distributedSSSP_exec} we see that Giraph is actually the slowest framework on 5 out of 7 graphs.
 For the other two, namely twitter and friendster, Giraph is second slowest with only Gemini taking longer to complete.

### 2020-09-24 documentation/paper/sections/testingMethods.tex
Rückmeldung übernommen und todos eingefügt

@@
 \subsection{Testing Methods}
-The testing methods cover the test environment, including hardware and the setup of the frameworks, the graphs used, followed by the algorithms utilized and how the times are measured.
+The testing methods cover the test environment, including hardware and the setup of the frameworks, the graphs used, followed by the algorithms utilized and our measurement procedures.
 
@@
 For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each.
+Thus, each machine has 96 cores, half of which are virtual.
 One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
@@
 We included a variaty of different graph sizes, from relatively small graphs like the flickr graph with 2 million edges up to an rMat28 with 4.2 billion edges.
-All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
+All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database associated with the Koblenz Network Collection (KONECT)\cite{konect}.
 Both the rMat27 and rMat28 were created with a modified version of a graph generator provided by Ligra (we changed the output format to EdgeList).
@@
 The default implementation of SSSP is deltaTile.
-A lot of options are available as well, but we have gone with the defaults.
+A lot of options are available as well, but we used the defaults.
 	\item Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.

### 2020-09-24 plots/makeplots.py
Plots für Galois Speedup vereinheitlicht, mit größerer Schriftart

@@ -16,7 +16,7 @@ execTimes = False
 overheads = False
 
 
-galoisSpeedup = False
+galoisSpeedup = True
 
 
 colors = {"Galois":"#1f77b4",
@@ -89,24 +89,24 @@ if singleNode and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 ## GALOIS THREAD COUNT
 if galoisSpeedup:
-	line_plot(graphs, x, speedupGaloisSSSP, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads.png", ylim=[0,20])
+	line_plot(graphs, x, speedupGaloisSSSP, yLabel='Avg. calc time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeSSSPGaloisThreads.png", ylim=[0,20])
 
-	line_plot(graphs, x, speedupGaloisBFS, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeBFSGaloisThreads.png", ylim=[0,11])
+	line_plot(graphs, x, speedupGaloisBFS, yLabel='Avg. calc time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodeBFSGaloisThreads.png", ylim=[0,11])
 
-	line_plot(graphs, x, speedupGaloisPRPush, xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPushGaloisThreads.png", ylim=[0.85,2.5])
+	line_plot(graphs, x, speedupGaloisPRPush, xLabel='Thread count', yLabel='Avg. calc time speedup', yScale='linear', saveToFile="singleNodePRPushGaloisThreads.png", ylim=[0.85,2.5])
 
-	line_plot(graphs, x, speedupGaloisPRPull, yLabel='Average calculation time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPullGaloisThreads.png", ylim=[0.5,2.6])
+	line_plot(graphs, x, speedupGaloisPRPull, yLabel='Avg. calc time speedup', xLabel='Thread count', yScale='linear', saveToFile="singleNodePRPullGaloisThreads.png", ylim=[0.5,2.6])
 
 
 ## GALOIS HP THREAD COUNT
 if galoisSpeedup:
[…]

### 2020-09-25 documentation/paper/main.tex

@@
 \usepackage{amssymb}
+
+\usepackage[
+  title={A Comparison of Graph Processing Systems},
+  author={Simon König (3344789)\\Leon Matzner (3315161)\\Felix Rollbühler (3310069)\\Jakob Schmid (3341630)\\},
+  type={Bachelor-Forschungsprojekt},
+  institute=ipvs,
+  course={Informatik},
+  examiner={Prof.\ Dr.\ rer.\ nat.\ Kurt Rothermel},
+  supervisor={Heiko Geppert,\ M.Sc.},
+  startdate={April 1, 2020},
+  enddate={October 1, 2020},
+  language=english
+]{scientific-thesis-cover}
 
@@
 \begin{document}
+\Coverpage
 
@@
 
+\Affirmation
+
 \end{document}

### 2020-09-25 documentation/paper/sections/preliminiaries.tex

@@
 
+\todo{choose paragraph version}
+Many parallel graph algorithms, including the three we consider here, are implemented by iterating over vertices \cite{tao}.
+A vertex to which an operator is applied to is called an \emph{active} vertex.
+Such an operator considers only the \emph{neighbourhood} of the \emph{active} vertex, where the \emph{neighbourhood} includes only adjacent vertices.
+The locality of the operator simplifies the parallelisation of iterations over vertices.
+
 In general such an operator can change the whole structure of the \emph{neighborhood}.

### 2020-09-25 documentation/paper/sections/results/bfs.tex
Grafiken umsortiert, verkleinert und zusammengebracht

@@
 \paragraph{Single-Node}
-\begin{figure*}
+\begin{figure*}[ht]
 	\begin{subfigure}{0.32\textwidth}
@@
 \end{figure*}
+\begin{figure*}[h]
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
+		\caption{Calculation time}
+		\label{fig:distributedBFS_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
+		\caption{Execution time}
+		\label{fig:distributedBFS_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTime.png}
+		\caption{Overhead}
+		\label{fig:distributedBFS_overhead}
+	\end{subfigure}
+	\hfil
+	\caption{Average times for BFS on the distributed cluster}
+	\label{fig:distributedBFS}
+\end{figure*}
 Giraph ran out of memory ($>$250 GB) on any graph larger than wikipedia. Also, Polymer failed to complete on rMat28. Ligra, that failed during SSSP however completed our benchmark for BFS.
@@
 The calculation times provided in \autoref{fig:singleNodeBFS_calc} show Gemini and Ligra to be comparable in their performance. 
-Their computation times deviate less than 151ms on all graphs except wikipedia and rMat28, with Ligra being the faster framework on most graphs.
+Their calculation times deviate less than 151ms on all graphs except wikipedia and rMat28, with Ligra being the faster framework on most graphs.
 Ligra is between 2ms (twitter) and 151ms (friendster) faster than Gemini.
@@
 \paragraph{Distributed}
-\begin{figure*}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
-		\caption{Calculation time}
-		\label{fig:distributedBFS_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
-		\caption{Execution time}
-		\label{fig:distributedBFS_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTime.png}
-		\caption{Overhead}
-		\label{fig:distributedBFS_overhead}
-	\end{subfigure}
-	\hfil
-	\caption{Average times for BFS on the distributed cluster}
-	\label{fig:distributedBFS}
-\end{figure*}
 For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case.

### 2020-09-25 documentation/paper/sections/results/galoisSpeedup.tex
Grafiken umsortiert, verkleinert und zusammengebracht

@@
 \paragraph{Single-Source Shortest-Paths}
-\begin{figure*}
+\begin{figure}
 	\hfil
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
@@
 	\label{fig:galoisSpeedupSSSP}
-\end{figure*}
+\end{figure}
 \begin{table}
@@
 Behaviour on friendster is similarly good. Here speedup is 1.9$\times$ at 2 threads compared to one, 3.5$\times$ at 4, 6.1$\times$ at 8 threads and 9.7$\times$ at 16 threads.
-Anything above 16 threads however no longer helps decreasing the computation time significantly on any graph.
+Anything above 16 threads however no longer helps decreasing the calculation time significantly on any graph.
 Speedup above 16 threads is always less than double the speedup of 16 threads.
@@
 \paragraph{Breadth-First Search}
-\begin{figure*}
+\begin{figure}
 	\hfil
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisThreads.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFSGaloisHPThreads.png}
@@
 	\label{fig:galoisSpeedupBFS}
-\end{figure*}
+\end{figure}
 
@@
 
-%Das ist irgendwie quatsch.
-% While it is beneficial to use hugepages on both algorithms, simply due to the general improvement in the needed times, the hugepages bring different multithreaded behaviour.
-% On BFS, the maximum observed speedup increases on almost every graph with hugepages compared to without. So with hugepages
-
-% In applications similar to SSSP or BFS, the benefit of many threads can be on the one hand be increased, if the graph allows it.
-
 \paragraph{PageRank Pull}
-\begin{figure*}
+\begin{figure}
 	\hfil
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
@@
 	\label{fig:galoisSpeedupPRPull}
-\end{figure*}
+\end{figure}
 
 We want to first take a look at the results for PageRank in Pull mode, seen in \autoref{fig:galoisSpeedupPRPull}.
-Without hugepages, computation time is hardly reduced on any graph other than flickr, where the reached maximum is 1.6$\times$ (cf.
+Without hugepages, calculation time is hardly reduced on any graph other than flickr, where the reached maximum is 1.6$\times$ (cf.
 \autoref{fig:galoisSpeedupPRPull_noHP}).
@@
 \paragraph{PageRank Push}
-\begin{figure*}
+\begin{figure}
 	\hfil
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisThreads.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.48\columnwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePRPushGaloisHPThreads.png}
@@
 	\label{fig:galoisSpeedupPRPush}
-\end{figure*}
+\end{figure}
 

### 2020-09-25 documentation/paper/sections/results/sssp.tex
Grafiken umsortiert, verkleinert und zusammengebracht

@@
 \paragraph{Single-node}
-\begin{figure*}
+\begin{figure*}[ht]
 	\begin{subfigure}{0.32\textwidth}
@@
 	\label{fig:singleNodeSSSP}
+\end{figure*}
+\begin{figure*}[h]
+	\centering
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
+		\caption{Calculation times}
+		\label{fig:distributedSSSP_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
+		\caption{Execution times}
+		\label{fig:distributedSSSP_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTime.png}
+		\caption{Overhead times}
+		\label{fig:distributedSSSP_overhead}
+	\end{subfigure}
+	\caption{Average times for SSSP on the distributed cluster}
+	\label{fig:distributedSSSP}
 \end{figure*}
@@
 \paragraph{Distributed}
-\begin{figure*}
-	\centering
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
-		\caption{Calculation times}
-		\label{fig:distributedSSSP_calc}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{Execution times}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTime.png}
-		\caption{Overhead times}
-		\label{fig:distributedSSSP_overhead}
-	\end{subfigure}
-	\caption{Average times for SSSP on the distributed cluster}
-	\label{fig:distributedSSSP}
-\end{figure*}
 On the distributed cluster, we find similar results as on the single node (cf. \autoref{fig:distributedSSSP}).
@@
 And that with a considerable margin of other frameworks always taking at least 50\% longer (lower bound here is Gemini on flickr) up to Galois Pull needing 18$\times$ more time on wikipedia.
-On both synthetic graphs however, Giraph is actually the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.
+On both synthetic graphs however, Giraph is actually the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the calculation time of Gemini on rMat27 or rMat28 respectively.
 While Giraph's calculation times are very competitive, when comparing the execution times in \autoref{fig:distributedSSSP_exec} we see that Giraph is actually the slowest framework on 5 out of 7 graphs.

### 2020-09-25 documentation/paper/sections/testingMethods.tex

@@
 
-\subsection{Testing Methods}
-The testing methods cover the test environment, including hardware and the setup of the frameworks, the graphs used, followed by the algorithms utilized and our measurement procedures.
+\subsection{Methodology}
+The testing methods cover the test environment, including hardware and the setup of the frameworks, the graphs used, followed by the algorithms utilized and our measuring procedures.
 

### 2020-09-25 documentation/paper/sections/results/galoisSpeedup.tex
bla

@@
 \end{table}
-One might assume the two algorithms SSSP and BFS to behave very similarly, yet they produce very different results. \todo{Begründung, z.b. Frontier Set}
+One might assume the two algorithms SSSP and BFS to behave very similarly, yet they produce very different results.
+While they both traverse the graph, the number of active vertices is generally higher on SSSP.
+This is because there can be multiple paths to each vertex.
+Thus, the work can be spread across more worker threads.
 With SSSP, we observed a decrease in the variance of the speedups, while the average speedup decreased or stayed the same. This provided a more reliable speedup. 

### 2020-09-25 documentation/paper/sections/introduction.tex

@@
 \section{Introduction}
-\todo{Hier gibts einige Todos, am besten nochmal nachschauen}
 In recent years, graph processing has become increasingly important, both in academic and industrial environments \cite{Gemini}.
-In many areas, such as machine learning and data mining, but also scientific computations, problems can be modeled and solved with the help of graphs \cite{Polymer}.
+In many areas, such as machine learning and data mining, but also scientific computations \todo{Quellen}, problems can be modeled and solved with the help of graphs \cite{Polymer}.
 Thus a new application domain called graph analytics was created.
-Some business models are based on working with graphs, such as marketing analysis, advertising, or even Google's search engine \cite{pagerank}.
+Some business models are based on working with graphs, such as marketing analysis, advertising \todo{Quellen}, or even Google's search engine \cite{pagerank}.
 In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
 This makes the performance and memory efficency of graph analysis applications more important than ever.
-But also the parallelism and distribution of graph algorithms becomes more important with increasing graph sizes.
+But also the parallelism and distribution of graph algorithms becomes more important with increasing graph sizes. \todo{Weiter gehen}
 
-Some frameworks were developed for this purpose.
-One of the most commonly known graph-processing systems is the closed-source and distributed system Pregel \cite{pregel}.
+Some frameworks were developed to facilitate the writing of parallel and distributed graph algorithms.
+One of the most commonly known graph-processing systems is the closed-source, distributed system Pregel \cite{pregel}.
 Pregel, as well as some open source versions of it, like Giraph \cite{Giraph} and GraphX \cite{graphx}, was built to handle large graphs reliably and be fault-tolerant on large MapReduce infrastructures.
@@
 Because of this, many non-uniform memory access (NUMA) aware systems such as Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
-These systems have some similarities, but many differences. This results in considerable performance differences on many occasions.
-The systems were initially compared at their introduction.
-Because most of the systems are several years old and some of them have evolved, or are still evolving, these comparisons were often incomplete or are simply no longer relevant today. 
+These systems have some similarities, but many differences \todo{Zu wem}. This results in considerable performance differences on many occasions.
+Each systems performance was evaluated at release time with some older graph processing frameworks.
+However most of the systems are several years old and some of them have evolved, or are still evolving, these comparisons were often incomplete or are simply outdated today.
 Other systems are no longer in use, which also reduces the significance of some of the existing comparisons.
 
-Therefore we compared several state-of-the-art non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
+Therefore we compared several state-of-the-art non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).\todo{Frameworks nennen}
 The comparison is performed on both real world data sets and synthetic graphs.
@@
 Giraph itself has often been compared to other state-of-the-art systems like Pregel or GraphX.
-We also included a comparison of Galois with itself with and without hugepages \cite{hugepages}, and several runs with different numbers of threads to see how it scales.
+We also included a comparison of Galois with and without hugepages \cite{hugepages}, and several runs with different numbers of threads to see how it scales.
 Where possible, several versions of the algorithms were used, like the push and pull variants.

### 2020-09-25 documentation/paper/sections/conclusion.tex
Conclusion gekürzt

@@
 
-\todo{Ab hier kürzer fassen}
 On a single node, the fastest framework in computation time is dependent on the problem under consideration.
 For SSSP, we found Polymer to perform best while for BFS Ligra and Gemini are fastest. PR was calculated by Galois the fastest.
+Depending on the exact circumstances all choices except Giraph can be suitable as single node system. 
 When looking at the computation times of the distributed systems, it stands out that the best time for each problem is beaten by most if not all of the single node times.
-Comparing the distributed computation times SSSP and BFS results looked similar with Giraph being the fastest with the exception of synthetic graphs where Gemini was faster.
+This suggests that single-node systems should be used whenever there is sufficent RAM.
+Comparing the distributed computation times SSSP and BFS results looked similar with Giraph being often the fastest, sometimes beaten by Gemini.
 The push version of Galois outperforms the pull version but ultimatively was slower than Gemini.
-For distributed PR, Gemini was the fastest by a margin.
+For PR, Gemini was the fastest by a margin.
+Overall, Gemini is the fastest distributed framework.
+However Giraph can be a good choice as well, outperforming Gemini at several occations on SSSP and BFS and offering additional features like node fault-tolerance.
 In contrast, in terms of overall execution times, Galois almost always outperforms the other frameworks because the overhead is very small.
-
-These benchmark results suggest that single-node systems should be used whenever the graph fits into the RAM. 
-If only individual calculations are performed on a graph, Galois is recommended due to its overall speed for a single node system. 
-For a distributed system Galois is still a good choice if the algorithm is similar to SSSP and BFS. If the algorithm is more similar to PR, Gemini would be a better choice.
-In case of multiple calculations on a single graph, the graph does not need to be reloaded. Thus overall execution time is less important and only calculation time should to be considered.
-As single-node system, there is no single best framework.
-All choices excluding Giraph can be suitable here, depending on the exact circumstances.
-Gemini would be a safe bet, often being the second best framework.
-In situations where a distributed system is required and the calculation time is important, Gemini is faster than Giraph in most cases.
-However Giraph can be a good choice as well, offering additional features like node fault-tolerance.
+This makes Galois perfect for individual calculations in a graph, where the loading time of the graph has a great influence.
 

### 2020-09-25 documentation/paper/sections/introduction.tex
introduction improvements

@@
 In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
-This makes the performance and memory efficency of graph analysis applications more important than ever.
-But also the parallelism and distribution of graph algorithms becomes more important with increasing graph sizes. \todo{Weiter gehen}
+This makes the performance and memory efficiency of graph analysis applications more important than ever.
+But also the parallelism and distribution of graph algorithms becomes more important with increasing graph sizes.
+Parallelism may be important to exploit in order to finish the computations in an acceptable time frame. 
+It is especially important to utilize distributed computation when the RAM of a single node is insufficient for the task.
 

### 2020-09-26 documentation/paper/main.tex
bla
introductin and abstract improvements

@@
 \begin{abstract}
-Graphs are growing fast and are becoming increasingly popular in many different domains, like \todo{...}
+Graphs are growing fast and are becoming increasingly popular. Many problems can be modeled and solved using graphs.
 There are several different frameworks that help writing graph algorithms.
 They often use different approaches and have significant performance differences on different graphs.
-\todo{Giraph nicht NUMA aware, da unten raus nehmen..}
-In this paper we compare some state-of-the-art non-uniform memory access (NUMA) aware systems in terms of performance, both on real world and synthetic graphs.
-\todo{umformulieren}
-Therefore, we consider the three algorithms: Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
-We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, and the resulting overhead time.
+In this paper we compare some state-of-the-art non-uniform memory access (NUMA) aware systems and Giraph in terms of performance, both on real world and synthetic graphs.
+Therefore, we evaluate the three algorithms Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
+We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, as well as the resulting overhead time.
 \end{abstract}
@@
 
+\clearpage
+\pagestyle{empty}
 \Affirmation

### 2020-09-26 documentation/paper/sections/introduction.tex
introductin and abstract improvements

@@
 Because of this, many non-uniform memory access (NUMA) aware systems such as Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
-These systems have some similarities, but many differences \todo{Zu wem}. This results in considerable performance differences on many occasions.
+The different NUMA aware systems have some similarities, but many differences and differ in the technologies they use.
+This results in considerable performance differences on many occasions.
 Each systems performance was evaluated at release time with some older graph processing frameworks.
@@
 
-Therefore we compared several state-of-the-art non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).\todo{Frameworks nennen}
+Therefore we compared the state-of-the-art NUMA aware systems Ligra, Polymer, Galois and Gemini in terms of their performance on three graph algorithms PageRank, SSSP, BFS.
 The comparison is performed on both real world data sets and synthetic graphs.
 To provide a comparison to a non-NUMA aware system, Giraph is included in the testing.
+It is like the NUMA aware systems open-source and is still being maintained.
 Giraph itself has often been compared to other state-of-the-art systems like Pregel or GraphX.

### 2020-09-26 documentation/paper/sections/overview/gemini.tex
gemini writeup improvements

@@
 The real bottleneck of distributed systems is not the communication itself, but the extra instructions, as well as memory references and a lower usage of multiple cores compared to the shared memory counterparts.
-There are four main reasons for this.
+There are three main reasons for this.
 The first reason is the use of hash maps to convert the vertex IDs between the global and the local state.
-The second reason is the maintenance of vertex replicas on the different systems.
-Another reason is the communication-bound apply phase in GAS abstraction.
+Another reason is the maintenance of vertex replicas on the different systems.
 And the last reason is the lack of dynamic scheduling.
@@
 Gemini is fairly lightweight and has a clearly defined API between the core framework and the implementations of the individual algorithms.
-The five already implemented algorithms are Single-Source SSSP, BFS, PR, Connected-Components (CC) and Betweenness-Centrality (BC).
+It also provides standard implementations for the following algorithms Single-Source SSSP, Breadth-First Search (BFS), PageRank (PR), Connected-Components (CC) and Betweenness-Centrality (BC).

### 2020-09-26 documentation/paper/main.tex
Abstract deutsch

@@
 
+\def\abstractname{Zusammenfassung}
+\begin{abstract}
+Graphen wachsen schnell und werden zunehmend beliebter, da viele Probleme mit ihnen modelliert und gelöst werden können.
+Es gibt verschieden Frameworks die beim schreiben von Graphalgorithmen helfen.
+Diese benutzen unterschiedliche Ansätze und haben Leistungsunterschiede bei verschiedenen Graphen.
+In dieser Arbeit vergleichen wir einige non-uniform memroy access (NUMA) fähige Systeme und Giraph hinsichtlich ihrer Leistung sowohl auf Graphen aus der realen Welt als auch auf syntetischen.
+Wir bewerten auf Grund von den drei Algorithmen Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) und PageRank (PR).
+Dabei vergleichen wir nicht nur die Zeit, die die Frameworks zum berechnen der Algorithmen brauchen, sondern auch die gesamte Zeit, die zum ausführen gebraucht wird.
+\end{abstract}
+
 \end{document}

### 2020-09-28 documentation/paper/sections/introduction.tex
PushPull simplified

@@
 In recent years, graph processing has become increasingly important, both in academic and industrial environments \cite{Gemini}.
-In many areas, such as machine learning and data mining, but also scientific computations \todo{Quellen}, problems can be modeled and solved with the help of graphs \cite{Polymer}.
+In many areas, such as machine learning and data mining, but also scientific computations problems can be modeled and solved with the help of graphs \cite{Polymer}.
 Thus a new application domain called graph analytics was created.
-Some business models are based on working with graphs, such as marketing analysis, advertising \todo{Quellen}, or even Google's search engine \cite{pagerank}.
+Some business models are based on working with graphs, such as marketing analysis, advertising or even Google's search engine \cite{pagerank}.
 In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.

### 2020-09-28 documentation/paper/sections/preliminiaries.tex
PushPull simplified

--- 2020-09-24 documentation/paper/sections/preliminiaries.tex
@@
 \subsection{Push and Pull Variants}
-\todo{kompliziert}
 Many parallel graph algorithms, including the three we consider here, are implemented by iterating over vertices \cite{tao}.
-The vertex, to which the operator is applied, is called \emph{active}.
-It's possible to iterate over edges and consider them as \emph{active}, but in the following, whithout loss of generality, we will use the term \emph{active vertex}.
-This operator considers only a so called \emph{neighborhood} to the \emph{active} vertex.
-To this \emph{neighborhood} belong only vertices, which are in the direct surrounding of the \emph{active} vertex.
-This allows to parallelise the iteration relatively easy, caused by the locality of the operator.
+To each of the vertices an operator is applied.
+This operator reads and writes the labels of the active vertex and of the vertices in the direct surrounding, also called neighbourhood, of the active vertex.
+It's also possible to iterate over edges, but in the following, whithout loss of generality, we will use the term active vertex.
 
-In general such an operator can change the whole structure of the \emph{neighborhood}.
-But in the following we assume, that an operator changes only \emph{labels} of vertices in the \emph{neighbourhood}.
-\emph{Labels} of edges are also called \emph{weights}.
 Often this operator can be implemented in two different ways, called \emph{push style} or \emph{pull style}.
-A \emph{push-style} operator reads the \emph{label} of the \emph{active} vertex and updates the \emph{label} of its \emph{neighborhood}.
-These operators are more efficient, if there are only a few \emph{active} vertices at the same time, or the \emph{neighborhoods} do not overlap, which can not be avoided in general.
-In contrast the \emph{pull-style} operator reads all values of its \emph{neighborhood} and updates the value of the \emph{active} vertex.
+A \emph{push-style} operator reads the label of the active vertex and updates the label of its neighborhood.
+These operators are more efficient, if there are only a few active vertices at the same time, or the neighborhoods do not overlap, which can not be avoided in general.
+In contrast the \emph{pull-style} operator reads all values of its neighborhood and updates the value of the active vertex.
 \emph{Pull-style} operators need less synchronization in parallel implementations, because unlike \emph{push style} there is only one write and many read operations.
 Thus locks can be avoided.
-So these operators are more efficient, if there are many \emph{active} vertices at the same time.
+So these operators are more efficient, if there are many active vertices at the same time.
 

### 2020-09-28 documentation/paper/sections/testingMethods.tex
PushPull simplified

@@
 \subsection{Methodology}
-The testing methods cover the test environment, including hardware and the setup of the frameworks, the graphs used, followed by the algorithms utilized and our measuring procedures.
+The methodology cover the test environment, including hardware and the setup of the frameworks, the graphs used, followed by the algorithms utilized and our measuring procedures.
 

### 2020-09-28 documentation/paper/sections/preliminiaries.tex
änderung push/pull style, und so

@@
 To each of the vertices an operator is applied.
-This operator reads and writes the labels of the active vertex and of the vertices in the direct surrounding, also called neighbourhood, of the active vertex.
-It's also possible to iterate over edges, but in the following, whithout loss of generality, we will use the term active vertex.
+This operator reads and writes the labels (i.e.
+vertex or edge weights) of the active vertex and of the vertices in the direct surrounding.
+Thse vertices are called the \emph{neighbourhood} of the active vertex.
+%It's also possible to iterate over edges, but in the following, whithout loss of generality, we will use the term active vertex.
 
 Often this operator can be implemented in two different ways, called \emph{push style} or \emph{pull style}.
-A \emph{push-style} operator reads the label of the active vertex and updates the label of its neighborhood.
-These operators are more efficient, if there are only a few active vertices at the same time, or the neighborhoods do not overlap, which can not be avoided in general.
+A \emph{push-style} operator reads the label of the active vertex and updates the labels of its neighborhood.
+These operators are more efficient, if there are only a few active vertices at the same time, or the neighborhoods do not overlap. Overlapping can not be avoided in general.
 In contrast the \emph{pull-style} operator reads all values of its neighborhood and updates the value of the active vertex.
 \emph{Pull-style} operators need less synchronization in parallel implementations, because unlike \emph{push style} there is only one write and many read operations.
-Thus locks can be avoided.
-So these operators are more efficient, if there are many active vertices at the same time.
+Thus many locks can be avoided.
+So, these operators are more efficient, if there are many active vertices at the same time.
 

### 2020-09-28 documentation/paper/sections/testingMethods.tex
umbenennen der environment zu testing environment

@@
 
-\subsubsection{Environment}
+\subsubsection{Testing Environment}
 For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each.

### 2020-09-28 documentation/paper/sections/overview/gemini.tex
minor improvements

@@
 Gemini is fairly lightweight and has a clearly defined API between the core framework and the implementations of the individual algorithms.
-It also provides standard implementations for the following algorithms Single-Source SSSP, Breadth-First Search (BFS), PageRank (PR), Connected-Components (CC) and Betweenness-Centrality (BC).
+It also provides standard implementations for the following algorithms SSSP, BFS, PR, Connected-Components (CC) and Betweenness-Centrality (BC).

### 2020-09-28 documentation/paper/sections/preliminiaries.tex
minor improvements

--- 2020-09-28 documentation/paper/sections/preliminiaries.tex
@@
 \subsection{Single-Source Shortest-Paths}
-Single-Source Shortest-Paths (SSSP) describes the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
+Single-Source Shortest-Paths (SSSP) is the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
 Input to the problem is a weighted graph $G=(V,E)$ and a start vertex $s\in V$. Output is the shortest possible distance from $s$ to each vertex in $V$.
@@
 Pairs of components can only communicate using messages, thus remote memory access is also only possible in this way.
-The Messages have a user-defined form and should be as small as possible to keep the network traffic low.
+The messages have a user-defined form and should be as small as possible to keep the network traffic low.
 The \emph{Congest model} is a closely related model and furthermore describes the messages.
@@
 The address areas are organized in so called pages, which in most cases are 4 KiB in size.
-In the Translation lookaside buffer (TBL) the most recent translations of virtual memory to physical memory are cached.
+In the Translation lookaside buffer (TLB) the most recent translations of virtual memory to physical memory are cached.
 
-Translating from the virtual to the physical address space significantly reduces the performance of today's data processing systems, since the size of the RAM grows much faster than the size of the TBL \cite{hugepages, superpages}.
+Translating from the virtual to the physical address space significantly reduces the performance of today's data processing systems, since the size of the RAM grows much faster than the size of the TLB \cite{hugepages, superpages}.
 To counteract this problem, hugepages were developed, which are usually several mibibytes in size.
-This minimizes the CPU time needed for table lookups, because there are fewer TBL misses.
+This minimizes the CPU time needed for table lookups, because there are fewer TLB misses.
 This speedup is especially noticeable in very memory intensive applications, like graph processing systems.

### 2020-09-29 documentation/paper/main.tex
ein weiteres Zitat

@@
 
-% \bibliographystyle{IEEEtran}
-% \bibliography{literature}
-
-% Alles ab hier kommt aus main.bbl
+%\bibliographystyle{IEEEtran}
+%\bibliography{literature}
 
@@
 
-\bibitem{Giraph}
+\bibitem{Gemini}
 \BIBentryALTinterwordspacing
-{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
-  Available: \url{https://giraph.apache.org}
+X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
+  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
+  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
+  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
+  301--316. [Online]. Available:
+  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
 \BIBentrySTDinterwordspacing
@@
 
-\bibitem{Galois}
+\bibitem{viralMarketing}
 \BIBentryALTinterwordspacing
-D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
-  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
-  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
-  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
-  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
+P.~Domingos and M.~Richardson, ``Mining the network value of customers,'' in
+  \emph{Proceedings of the Seventh ACM SIGKDD International Conference on
+  Knowledge Discovery and Data Mining}, ser. KDD '01.\hskip 1em plus 0.5em
+  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
+  2001, p. 57–66. [Online]. Available:
+  \url{https://doi.org/10.1145/502512.502525}
 \BIBentrySTDinterwordspacing
 
-\bibitem{Ligra}
+\bibitem{pagerank}
 \BIBentryALTinterwordspacing
-J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
-  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
-  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
-  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
-  Machinery, 2013, p. 135–146. [Online]. Available:
-  \url{https://doi.org/10.1145/2442516.2442530}
-\BIBentrySTDinterwordspacing
-
-\bibitem{newman2010networks}
-\BIBentryALTinterwordspacing
-M.~Newman, \emph{Networks: An Introduction}.\hskip 1em plus 0.5em minus
-  0.4em\relax OUP Oxford, 2010. [Online]. Available:
-  \url{https://books.google.de/books?id=-DgTDAAAQBAJ}
+S.~Brin and L.~Page, ``The anatomy of a large-scale hypertextual web search
+  engine,'' \emph{Computer Networks and ISDN Systems}, vol.~30, no.~1, pp. 107
+  -- 117, 1998, proceedings of the Seventh International World Wide Web
+  Conference. [Online]. Available:
+  \url{http://www.sciencedirect.com/science/article/pii/S016975529800110X}
 \BIBentrySTDinterwordspacing
@@
 
-\bibitem{pagerank}
+\bibitem{Giraph}
 \BIBentryALTinterwordspacing
-S.~Brin and L.~Page, ``The anatomy of a large-scale hypertextual web search
-  engine,'' \emph{Computer Networks and ISDN Systems}, vol.~30, no.~1, pp. 107
-  -- 117, 1998, proceedings of the Seventh International World Wide Web
-  Conference. [Online]. Available:
-  \url{http://www.sciencedirect.com/science/article/pii/S016975529800110X}
+{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
+  Available: \url{https://giraph.apache.org}
+\BIBentrySTDinterwordspacing
+
+\bibitem{graphx}
+J.~E. Gonzalez, R.~S. Xin, A.~Dave, D.~Crankshaw, M.~J. Franklin, and
+  I.~Stoica, ``Graphx: Graph processing in a distributed dataflow framework,''
+  in \emph{Proceedings of the 11th USENIX Conference on Operating Systems
+  Design and Implementation}, ser. OSDI'14.\hskip 1em plus 0.5em minus
+  0.4em\relax USA: USENIX Association, 2014, p. 599–613.
+
+\bibitem{Ligra}
+\BIBentryALTinterwordspacing
+J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
+  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
+  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
+  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
+  Machinery, 2013, p. 135–146. [Online]. Available:
+  \url{https://doi.org/10.1145/2442516.2442530}
+\BIBentrySTDinterwordspacing
+
+\bibitem{Galois}
+\BIBentryALTinterwordspacing
+D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
+  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
+  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
+  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
+  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
+\BIBentrySTDinterwordspacing
+
+\bibitem{hugepages}
+\BIBentryALTinterwordspacing
+A.~Panwar, A.~Prasad, and K.~Gopinath, ``Making huge pages actually useful,''
+  in \emph{Proceedings of the Twenty-Third International Conference on
+  Architectural Support for Programming Languages and Operating Systems}, ser.
+  ASPLOS '18.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA:
+  Association for Computing Machinery, 2018, p. 679–692. [Online]. Available:
+  \url{https://doi.org/10.1145/3173162.3173203}
+\BIBentrySTDinterwordspacing
+
+\bibitem{newman2010networks}
+\BIBentryALTinterwordspacing
+M.~Newman, \emph{Networks: An Introduction}.\hskip 1em plus 0.5em minus
+  0.4em\relax OUP Oxford, 2010. [Online]. Available:
+  \url{https://books.google.de/books?id=-DgTDAAAQBAJ}
 \BIBentrySTDinterwordspacing
@@
 
-\bibitem{graphx}
-J.~E. Gonzalez, R.~S. Xin, A.~Dave, D.~Crankshaw, M.~J. Franklin, and
-  I.~Stoica, ``Graphx: Graph processing in a distributed dataflow framework,''
-  in \emph{Proceedings of the 11th USENIX Conference on Operating Systems
-  Design and Implementation}, ser. OSDI'14.\hskip 1em plus 0.5em minus
-  0.4em\relax USA: USENIX Association, 2014, p. 599–613.
-
 \bibitem{powergraph}
@@
 
-\bibitem{hugepages}
-\BIBentryALTinterwordspacing
-A.~Panwar, A.~Prasad, and K.~Gopinath, ``Making huge pages actually useful,''
-  in \emph{Proceedings of the Twenty-Third International Conference on
-  Architectural Support for Programming Languages and Operating Systems}, ser.
-  ASPLOS '18.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA:
-  Association for Computing Machinery, 2018, p. 679–692. [Online]. Available:
-  \url{https://doi.org/10.1145/3173162.3173203}
-\BIBentrySTDinterwordspacing
-
 \bibitem{superpages}
@@
   2012, pp. 1--10.
-
-\bibitem{Gemini}
-\BIBentryALTinterwordspacing
-X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
-  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
-  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
-  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
-  301--316. [Online]. Available:
-  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
-\BIBentrySTDinterwordspacing
 
@@
 \BIBentrySTDinterwordspacing
-
 \end{thebibliography}

### 2020-09-30 documentation/paper/main.tex
Abgabe

@@
   enddate={October 1, 2020},
-  language=english
+  language=english,
+  setPageNumberToOne=true
 ]{scientific-thesis-cover}

### 2020-09-30 documentation/paper/sections/introduction.tex
ein weiteres Zitat
Abgabe

@@
 Thus a new application domain called graph analytics was created.
-Some business models are based on working with graphs, such as marketing analysis, advertising or even Google's search engine \cite{pagerank}.
+Some business models are based on working with graphs, such as viral marketing \cite{viralMarketing} or Google's search engine \cite{pagerank}.
 In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
@@
 Some frameworks were developed to facilitate the writing of parallel and distributed graph algorithms.
-One of the most commonly known graph-processing systems is the closed-source, distributed system Pregel \cite{pregel}.
+One of the most commonly known graph processing systems is the closed-source, distributed system Pregel \cite{pregel}.
 Pregel, as well as some open source versions of it, like Giraph \cite{Giraph} and GraphX \cite{graphx}, was built to handle large graphs reliably and be fault-tolerant on large MapReduce infrastructures.

### 2020-10-12 documentation/presentation/main.tex
Vortrag pdf

\documentclass{meetings}

\usepackage{makecell}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{positioning}

\setlength{\parskip}{0pt}
\raggedcolumns

\author{S. König, L. Matzner, F. Rollbühler and J. Schmid}
\date{Tuesday, 20\textsuperscript{th} October 2020}

\begin{document}

\section{Introduction}
Graphs are growing fast and are becoming increasingly popular. Many problems can be modeled and solved using graphs.

We compare some state-of-the-art non-uniform memory access (NUMA) aware systems and Giraph in terms of performance, both on real world and synthetic graphs.
Therefore, we evaluate the three algorithms Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, as well as the resulting overhead time.

\clearpage
\section{Overview}

\clearpage
\section{Preliminaries}
\begin{multicols}{2}
A \emph{weighted, directed graph} is the pair $G=(V,E,w)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set} with
	\begin{equation*}
	  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
	\end{equation*}
	and $w:E\rightarrow \mathbb R$ is a mapping of edge to a weight.

	\begin{tikzpicture}
		\node[circle, draw] (a) at (0,0) {A};
		\node[circle, draw] (b) at (2,0) {B};
		\node[circle, draw] (c) at (4,0) {C};
		\node[circle, draw] (d) at (0,2) {D};
		\node[circle, draw] (e) at (2,2) {E};

		\draw[-stealth, bend right, line width=1pt] (a) edge (b);
		\draw[-stealth, bend right, line width=1pt] (a) edge (c);

		\draw[-stealth, bend right, line width=1pt] (b) edge (a);
		\draw[-stealth, line width=1pt] (b) -- (d);
		\draw[-stealth, line width=1pt] (b) -- (e);

		\draw[-stealth, line width=1pt] (c) -- (e);

		\draw[-stealth, line width=1pt] (e) -- (a);
	\end{tikzpicture}

	\columnbreak

	Algorithms:
\begin{itemize}
	\item Single-Source Shortest-Paths (SSSP) is the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.

	\item Breadth-first search (BFS) is a search problem on a graph.

	\item PageRank (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph

\end{itemize}
\end{multicols}

\clearpage
\section{Push \& Pull}

\clearpage
\section{Hugepages}

SSSP, BFS, PR, Push/Pull, Hugepages

\clearpage
\section{Frameworks}

\clearpage
\section{Evaluation}
\begin{multicols}{2}
\noindent 5 Machines, with
\begin{itemize}
	\item 96 cores, of which 48 virtual
	\item 256 GB of RAM each, one machine only 128 GB
	\item Ubuntu 18.04.2 LTS
\end{itemize}
\noindent Measurements:
\begin{itemize}
	\item \textbf{execution time}: time from start to finish of the console command
	\item \textbf{calculation time}: time the framework actually executed the algorithm
	\item \textbf{overhead}: time difference between execution time and calculation time \newline (time to read the input graph, initialization, etc.)
\end{itemize}

\renewcommand{\arraystretch}{1.2}%
{%
\noindent
\centering
	\begin{tabular}{crr}
		\hline
		\bf{Graph}&\# Vertices (M)&\# Edges (M)\\\hline
		flickr&    		0.1&  2\\
		orkut&          3&    117\\
		wikipedia&      12&   378\\
		twitter&     	52&   1963\\
		rMat27&         63&   2147\\
		friendster&     68&   2586\\
		rMat28&         121&  4294\\
		\hline
	\end{tabular}
}%

each test case (graph, framework, algorithm) was run 10 times
\end{multicols}

\clearpage
\begin{multicols}{2}
\section{Production Case}
\begin{itemize}
	\item running system, that performs multiple calculations on a single graph
	\item without the need of reloading graph data with every calculation
	\item short calculation times should be preferred because the overhead time is only spent once on startup and amortizes quickly
\end{itemize}

\columnbreak
\section{Research Case}
\begin{itemize}
	\item individual calculations on a graph, i.e. for each calculation, the graph has to be loaded
	\item the algorithm can change frequently
	\item requiring the framework to be relatively fast on different algorithms
	\item overall small execution times and small overhead are preferred
\end{itemize}
\end{multicols}

Messfehlern bei Galois Calc was sagen

\clearpage
\section{Production Case Single Node}

\begin{figure}[h]
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
		\caption{SSSP}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
		\caption{BFS}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
		\caption{PR}
	\end{subfigure}
\end{figure}
\begin{itemize}
	\item Giraph is either very slow or requires too much RAM (>256 GB)
	\item On SSSP, Polymer is fastest, followed by Gemini on second place
	\item On BFS, Gemini and Ligra are comparable and fastest on the larger graphs
	\item On PR, Galois is fastest. But we exclude Galois Push because of possible measuring errors.
	\item Message-based approach can compete with shared-memory
\end{itemize}

\clearpage
\section{Production Case Distributed}

\begin{figure}[h]
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
		\caption{SSSP}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
		\caption{BFS}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
		\caption{PR}
	\end{subfigure}
\end{figure}
\begin{itemize}
	\item Giraph is fastest on SSSP and BFS on the real world graphs
	\item Giraph has problems with synthetic graphs
	\item Gemini is fastest on PR, with Giraph on second place
\end{itemize}

\section{Research Case Single Node}

\begin{figure}[h]
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
		\caption{SSSP}
		\label{fig:singleNodeSSSP_exec}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
		\caption{BFS}
		\label{fig:singleNodeSSSP_exec}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
		\caption{PR}
		\label{fig:singleNodeSSSP_exec}
	\end{subfigure}
\end{figure}
\begin{itemize}
	\item Giraph is either slowest or requires too much RAM (>256 GB)
	\item Galois is fastest in almost all cases, second fastest is Ligra
	\item Gemini and Polymer are comparably slow
\end{itemize}

\clearpage
\section{Research Case Distributed}

\begin{figure}[h]
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
		\caption{SSSP}
		\label{fig:distributedSSSP_exec}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
		\caption{BFS}
		\label{fig:distributedSSSP_exec}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
		\caption{PR}
		\label{fig:distributedSSSP_exec}
	\end{subfigure}
\end{figure}
\begin{itemize}
	\item Galois Push is faster than Pull in all cases
	\item Either Galois implementation is faster than any other frameworks on SSSP or BFS
	\item Gemini is fastest on PR and comparable to Giraph on SSSP and BFS
\end{itemize}

\section{Conclusion and Outlook}
Für jeden Fall ist das beste Framework sehr spezifisch.
Blabla.

Outlook:
Immer wieder solche Tests machen, viele der Frameworks ändern sich immer wieder weiter und so, nicht wahr.
\end{document}

### 2020-10-12 documentation/presentation/main.tex
Vortrag zeug

@@
 \usepackage{multicol}
+\usepackage{biblatex}
 \usepackage{tikz}
@@
 
+\bibliography{bib.bib}
+
 \author{S. König, L. Matzner, F. Rollbühler and J. Schmid}
@@
 \begin{document}
-
+% Ich weiß, dass es die Beamer-Klasse gibt.
+% Lasst mich doch alle in Ruhe!
+% Die hier hatte ich noch rumliegen, das muss jetzt reichen.
 \section{Introduction}
@@
 \begin{multicols}{2}
-A \emph{weighted, directed graph} is the pair $G=(V,E,w)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set} with
+\noindent
+	A \emph{weighted, directed graph} is the tuple $G=(V,E,w)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set} with
 	\begin{equation*}
@@
 
+	\begin{center}
 	\begin{tikzpicture}
-		\node[circle, draw] (a) at (0,0) {A};
-		\node[circle, draw] (b) at (2,0) {B};
+		\node[circle, draw] (a) at (1.7,0) {A};
+		\node[circle, draw] (b) at (0,2) {B};
 		\node[circle, draw] (c) at (4,0) {C};
-		\node[circle, draw] (d) at (0,2) {D};
-		\node[circle, draw] (e) at (2,2) {E};
+		\node[circle, draw] (d) at (0,4) {D};
+		\node[circle, draw] (e) at (2.5,2) {E};
 
-		\draw[-stealth, bend right, line width=1pt] (a) edge (b);
-		\draw[-stealth, bend right, line width=1pt] (a) edge (c);
+		\draw[->, bend right, line width=1pt] (a) edge (b);
+		\draw[->, bend right, line width=1pt] (a) edge (c);
 
-		\draw[-stealth, bend right, line width=1pt] (b) edge (a);
-		\draw[-stealth, line width=1pt] (b) -- (d);
-		\draw[-stealth, line width=1pt] (b) -- (e);
+		\draw[->, bend right, line width=1pt] (b) edge (a);
+		\draw[->, line width=1pt] (b) edge (d);
+		\draw[->, line width=1pt] (b) edge (e);
 
-		\draw[-stealth, line width=1pt] (c) -- (e);
+		\draw[->, line width=1pt] (c) edge (e);
 
-		\draw[-stealth, line width=1pt] (e) -- (a);
+		\draw[->, line width=1pt] (e) edge (a);
+
+		\node at (2.5,3) {$w(x)=1\ \forall x\in V$};
 	\end{tikzpicture}
+	\end{center}
 
@@
 
-	Algorithms:
+	\subsection{Algorithms}
+	\begin{description}
+		\item[Single-Source Shortest-Paths (SSSP):] find the shortest path from a starting vertex to every other vertex
+
+		\item[Breadth-first search (BFS):] find a node outgoing from a starting vertex, by increasing maximum hop count step-wise
+
+		\item[PageRank (PR):] link analysis algorithm; weighs vertices, measuring their relative importance
+	\end{description}
+\end{multicols}
+
+\clearpage
+\begin{multicols}{2}
+\section{Push Style}
+\begin{center}
+\begin{tikzpicture}
+	\node[circle, draw] (a) at (1.5,1.2) {\bf A};
+	\node[circle, draw] (b) at (0,0) {B};
+	\node[circle, draw] (c) at (3,0) {C};
+	\node[circle, draw] (d) at (1.5,3) {D};
+
+	\draw[->, line width=1pt] (a) edge node[midway, above] {1} (b);
+	\draw[->, line width=1pt] (a) edge node[midway, above] {1} (c);
+	\draw[->, line width=1pt] (a) edge node[midway, right] {1} (d);
+\end{tikzpicture}
+\end{center}
 \begin{itemize}
-	\item Single-Source Shortest-Paths (SSSP) is the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
+	\item reads active vertex, writes neighborhood 
+	\item more efficient, if only few active vertices at the same time
+	\item more efficient, if neighborhoods do not overlap
+\end{itemize}
 
-	\item Breadth-first search (BFS) is a search problem on a graph.
+\columnbreak
+\section{Pull Style}
+\begin{center}
+\begin{tikzpicture}
+	\node[circle, draw] (a) at (1.5,1.2) {\bf A};
+	\node[circle, draw] (b) at (0,0) {B};
+	\node[circle, draw] (c) at (3,0) {C};
+	\node[circle, draw] (d) at (1.5,3) {D};
 
-	\item PageRank (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph
-
+	\draw[->, line width=1pt] (b) edge node[midway, above] {3} (a);
+	\draw[->, line width=1pt] (c) edge node[midway, above] {2} (a);
+	\draw[->, line width=1pt] (d) edge node[midway, right] {6} (a);
+\end{tikzpicture}
+\end{center}
+\begin{itemize}
+	\item reads neighborhood, writes active vertex
+	\item[$\rightarrow$] only one write and many read operations
+	\item less synchronization in parallel implementations needed
+	\item more efficient, if many vertices active at the same time
 \end{itemize}
@@
 \clearpage
-\section{Push \& Pull}
-
-\clearpage
 \section{Hugepages}
-
-SSSP, BFS, PR, Push/Pull, Hugepages
 
@@
 Immer wieder solche Tests machen, viele der Frameworks ändern sich immer wieder weiter und so, nicht wahr.
+
+\clearpage
+\nocite{unserPaper}
+\printbibliography
 \end{document}

### 2020-10-13 documentation/presentation/main.tex
:-)

@@
 \usepackage{subcaption}
-\usepackage{multicol}
-\usepackage{biblatex}
+
 \usepackage{tikz}
@@
 \setlength{\parskip}{0pt}
+\setlength{\parindent}{0pt}
+\setlength{\columnsep}{1cm}
 \raggedcolumns
-
-\bibliography{bib.bib}
 
@@
 \section{Introduction}
+\begin{itemize}
+	\item 
+In the past years several graph processing systems emerged.
 Graphs are growing fast and are becoming increasingly popular. Many problems can be modeled and solved using graphs.
-
-We compare some state-of-the-art non-uniform memory access (NUMA) aware systems and Giraph in terms of performance, both on real world and synthetic graphs.
-Therefore, we evaluate the three algorithms Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
-We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, as well as the resulting overhead time.
+\item Comparison of non-uniform memory access (NUMA) aware systems and Giraph in their performance
+\begin{itemize}
+	\item on different graphs (real world and synthetic)
+\item and different algorithms (SSSP, BFS, PR)
+\end{itemize}
+\end{itemize}
 
@@
 
-	\subsection{Algorithms}
+	\subsection*{Algorithms}
 	\begin{description}
@@
 
-\clearpage
 \section{Hugepages}
 
-\clearpage
 \section{Frameworks}
 
-\clearpage
+\definecolor{Galois}{HTML}{1f77b4}
+\definecolor{Gemini}{HTML}{ff7f0e}
+\definecolor{Giraph}{HTML}{2ca02c}
+\definecolor{Ligra}{HTML}{d62728}
+\definecolor{Polymer}{HTML}{9467bd}
+
+\begin{multicols}{2}
+\begin{description}
+	\item[$\color{Galois}\blacksquare$ Galois] is a general purpose library designed for parallel programming
+	\begin{itemize}
+		\item Version 6.0 from 29\textsuperscript{th} June 2020 used 
+	\end{itemize}
+	\item[$\color{Gemini}\blacksquare$ Gemini] uses a distributed message-based approach from scratch
+	\begin{itemize}
+		\item Version from 2\textsuperscript{nd} November 2016 used
+		\item Version contains bugs that had to be fixed 
+	\end{itemize}
+	\item[$\color{Giraph}\blacksquare$ Giraph] is built on Apache Hadoop, a large scale data processing infrastructure
+	\begin{itemize}
+		\item Version 1.3 from 8\textsuperscript{th} May 2020 used
+		\item BFS is not natively supported 
+	\end{itemize}
+	\item[$\color{Ligra}\blacksquare$ Ligra] dynamically switches between push and pull style
+	\begin{itemize}
+		\item Version from 14\textsuperscript{th} August 2019
+	\end{itemize}
+	\item[$\color{Polymer}\blacksquare$ Polymer] optimizes data layout and memory access strategies
+	\begin{itemize}
+		\item Version from 28\textsuperscript{th} August 2018
+	\end{itemize}
+\end{description}
+\end{multicols}
+
 \section{Evaluation}
@@
 
+\columnbreak
+
+\begin{center}
 	\renewcommand{\arraystretch}{1.2}%
-{%
-\noindent
-\centering
 	\begin{tabular}{crr}
@@
 	\end{tabular}
-}%
+\end{center}
 
@@
 
-Messfehlern bei Galois Calc was sagen
-
-\clearpage
 \section{Production Case Single Node}
@@
 
-\clearpage
 \section{Production Case Distributed}
@@
 
-\clearpage
 \section{Research Case Distributed}
@@
 
+\section{Galois With Hugepages}
+\begin{multicols}{2}
+
+\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
+		\toprule
+		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
+		\cmidrule{3-4}\cmidrule{5-6}
+		&\bf Graph&w/o&w/&w/o&w/\\
+		\midrule
+\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
+		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5 \\
+		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1 \\
+		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1 \\
+		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4 \\
+		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5 \\
+		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1 \\
+		\midrule
+\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
+		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6 \\
+		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4 \\
+		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3 \\
+		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1 \\
+		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1 \\
+		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4 \\
+		\bottomrule
+\end{tabular}
+
+\columnbreak
+
+	\includegraphics[width=0.7\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+	SSSP
+
+	\includegraphics[width=0.7\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
+	PR Pull
+\end{multicols}
+
 \section{Conclusion and Outlook}
-Für jeden Fall ist das beste Framework sehr spezifisch.
-Blabla.
+\begin{itemize}
+	\item performance highly dependent on the framework, algorithm and data set
+	\begin{itemize}
+		\item Galois is almost always fastest in the research case; especially with hugepages
+		\item Giraph is good on SSSP or BFS in distributed production
+		\item Gemini is a good middleground for distributed PR and single node production
+	\end{itemize}
 
-Outlook:
-Immer wieder solche Tests machen, viele der Frameworks ändern sich immer wieder weiter und so, nicht wahr.
+	\item single node almost always preferrable, as long as RAM is sufficient
+\end{itemize}
 
-\clearpage
-\nocite{unserPaper}
-\printbibliography
+\subsection{Outlook}
+\begin{itemize}
+	\item[$\rightarrow$] incorporate new frameworks and new algorithms
+	\item[$\rightarrow$] great range of settings and multiple implementations for the same problem
+	\item[$\rightarrow$] At a later point in time, it is important to repeat such a comparison, because the frameworks are further developed and new ones are created.
+\end{itemize}
+
 \end{document}

### 2020-10-13 documentation/presentation/main.tex
:-)

@@
 \section{Hugepages}
+\begin{itemize}
+	\item most systems use virtual memory management
+\begin{itemize}
+	\item represents an abstraction to hardware memory
+	\item virtual memory is then organized in pages
+	\item translations of virtual memory to physical memory are cached, because every translation takes time
+\end{itemize}
+	\item typically, memory pages are 4 KiB in size
+	\item \textbf{hugepages} can be several MiB in size $\rightarrow$ reduce number of cache misses
+	\item especially noticeable in very memory intensive applications
+\end{itemize}
 

### 2020-10-14 documentation/presentation/main.tex
:-)

@@
 	\item \textbf{calculation time}: time the framework actually executed the algorithm
-	\item \textbf{overhead}: time difference between execution time and calculation time \newline (time to read the input graph, initialization, etc.)
+	\item \textbf{overhead}: time difference between execution time and calculation time \newline (time to read the input graph, initialization, etc.) -- brauchen wir das überhaupt?
 \end{itemize}
@@
 \section{Galois With Hugepages}
-\begin{multicols}{2}
-
-\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
+\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}rc}
 		\toprule
-		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
+		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}&\\
 		\cmidrule{3-4}\cmidrule{5-6}
-		&\bf Graph&w/o&w/&w/o&w/\\
+		&\bf Graph&w/o&w/&w/o&w/&\\
 		\midrule
 \multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
-		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5 \\
-		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1 \\
-		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1 \\
-		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4 \\
-		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5 \\
-		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1 \\
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 & 
+		\multirow{7}{8cm}{\includegraphics[width=5.5cm]{../../plots/singleNodeSSSPGaloisHPThreads.png}}
+		\\
+		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5 & \\
+		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1 & \\
+		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1 & \\
+		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4 & \\
+		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5 & \\
+		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1 & \\
 		\midrule
-\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
-		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6 \\
-		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4 \\
-		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3 \\
-		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1 \\
-		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1 \\
-		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4 \\
+
 		\bottomrule
@@
 
-\columnbreak
+\section{Galois With Hugepages}
+\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}rc}
+\toprule
+		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}&\\
+		\cmidrule{3-4}\cmidrule{5-6}
+		&\bf Graph&w/o&w/&w/o&w/&\\
+		\midrule
+	\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 & \\
+		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6 & \\
+		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4 & \\
+		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3 & \\
+		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1 & \\
+		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1 & \\
+		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4 & \\
+	\bottomrule
+\end{tabular}
 
-	\includegraphics[width=0.7\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
-	SSSP
+	\includegraphics[width=1cm]{../../plots/singleNodeSSSPGaloisHPThreads.png}
 
-	\includegraphics[width=0.7\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
-	PR Pull
-\end{multicols}
+	%\includegraphics[width=0.7\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
 

### 2020-10-14 documentation/presentation/main.tex
:-)

--- 2020-10-13 documentation/presentation/main.tex
@@
 \section{Introduction}
+\begin{center}
+	\begin{minipage}{0.6\linewidth}
+		\centering
+		Main Goal: \itshape  Comparison of five graph processing systems in their performance on different graphs and algorithms.
+	\end{minipage}
+\end{center}
+
+\begin{enumerate}
+	\item Preliminaries
 	\begin{itemize}
-	\item 
-In the past years several graph processing systems emerged.
-Graphs are growing fast and are becoming increasingly popular. Many problems can be modeled and solved using graphs.
-\item Comparison of non-uniform memory access (NUMA) aware systems and Giraph in their performance
+	\item Basics
+	\item Computation Styles
+	\item Hugepages
+	\end{itemize}
+	\item Frameworks
+	\item Evaluation
 	\begin{itemize}
-	\item on different graphs (real world and synthetic)
-\item and different algorithms (SSSP, BFS, PR)
+	\item Research vs. Production Case
+	\item Results for those Cases
+	\item Impact of Hugepages on Galois
 	\end{itemize}
-\end{itemize}
+	\item Conclusion
+\end{enumerate}
 
-\clearpage
-\section{Overview}
+% In the past years several graph processing systems emerged.
+% Graphs are growing fast and are becoming increasingly popular. Many problems can be modeled and solved using graphs.
 
@@
 	\end{itemize}
+
+	\columnbreak
+
 	\item[$\color{Giraph}\blacksquare$ Giraph] is built on Apache Hadoop, a large scale data processing infrastructure
@@
 		\item BFS is not natively supported 
+		\item state-of-the-art, but not NUMA-aware
 	\end{itemize}
@@
 \begin{multicols}{2}
-\noindent 5 Machines, with
+\noindent Machines: vsflash1-5, 
 \begin{itemize}
 	\item 96 cores, of which 48 virtual
-	\item 256 GB of RAM each, one machine only 128 GB
+	\item 256 GB of RAM each\footnote{\sffamily one machine only 128 GB}
 	\item Ubuntu 18.04.2 LTS
@@
 	\item \textbf{calculation time}: time the framework actually executed the algorithm
-	\item \textbf{overhead}: time difference between execution time and calculation time \newline (time to read the input graph, initialization, etc.)
+	%\item \textbf{overhead}: time difference between execution time and calculation time \newline (time to read the input graph, initialization, etc.) -- brauchen wir das überhaupt?
+	\item executed each test case 10 times
 \end{itemize}
@@
 \begin{center}
-	\renewcommand{\arraystretch}{1.2}%
-	\begin{tabular}{crr}
-		\hline
-		\bf{Graph}&\# Vertices (M)&\# Edges (M)\\\hline
-		flickr&    		0.1&  2\\
-		orkut&          3&    117\\
-		wikipedia&      12&   378\\
-		twitter&     	52&   1963\\
-		rMat27&         63&   2147\\
-		friendster&     68&   2586\\
-		rMat28&         121&  4294\\
-		\hline
-	\end{tabular}
+	\includegraphics[width=\linewidth]{graphsize.png}
 \end{center}
 
-each test case (graph, framework, algorithm) was run 10 times
 \end{multicols}
@@
 \begin{itemize}
-	\item running system, that performs multiple calculations on a single graph
-	\item without the need of reloading graph data with every calculation
-	\item short calculation times should be preferred because the overhead time is only spent once on startup and amortizes quickly
+	\item running system: multiple calculations on a single graph
+	\item graph data stays loaded between calculations
+	\item[$\rightarrow$] short calculation times should be preferred
 \end{itemize}
@@
 \begin{itemize}
-	\item individual calculations on a graph, i.e. for each calculation, the graph has to be loaded
-	\item the algorithm can change frequently
-	\item requiring the framework to be relatively fast on different algorithms
-	\item overall small execution times and small overhead are preferred
+	\item individual calculation cases: possibly new graph for each calculation
+	\item frequently changing algorithm
+	\item[$\rightarrow$] framework should be relatively fast on different algorithms
+	\item[$\rightarrow$] overall small execution times should be preferred
 \end{itemize}
@@
 \section{Galois With Hugepages}
-\begin{multicols}{2}
+\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
 
-\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
-		\toprule
 		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
@@
 \multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
-		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5 \\
-		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1 \\
-		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1 \\
-		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4 \\
-		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5 \\
-		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1 \\
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
+		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5\\
+		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1\\
+		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1\\
+		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4\\
+		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5\\
+		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1\\
+\end{tabular}
+\includegraphics[width=8cm]{../../plots/singleNodeSSSPGaloisThreads.png}
+\includegraphics[width=8cm]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+
+\section{Galois With Hugepages}
+\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}rc}
+		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
+		\cmidrule{3-4}\cmidrule{5-6}
+		&\bf Graph&w/o&w/&w/o&w/\\
 		\midrule
 	\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2 \\
-		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6 \\
-		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4 \\
-		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3 \\
-		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1 \\
-		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1 \\
-		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4 \\
-		\bottomrule
+		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
+		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6\\
+		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4\\
+		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3\\
+		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1\\
+		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1\\
+		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4\\
 \end{tabular}
-
-\columnbreak
-
-	\includegraphics[width=0.7\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
-	SSSP
-
-	\includegraphics[width=0.7\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
-	PR Pull
-\end{multicols}
+\includegraphics[width=0.5\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
+\includegraphics[width=0.5\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
 

### 2020-10-14 documentation/presentation/bla.py

@@ -1,0 +1,37 @@
+
+import numpy as np
+import matplotlib.patches as mpatches
+import matplotlib.pyplot as plt
+import matplotlib
+
+plt.style.use('./style.mplstyle')
+
+
+graphs = ('flickr', 'orkut', 'wikipedia', 'twitter', 'rMat27', 'friendster', 'rMat28')
+vals = [2, 117, 378, 1963, 2147, 2586, 4294];
+
+
+
+fig, ax = plt.subplots()
+
+
+ax.bar(graphs, vals, 0.4)
+	
+
+
+ax.set_xticklabels(graphs, rotation=-45)
+ax.set_yscale("linear")
+ax.set_title("")
+ax.set_ylabel("# of edges (Million)")
+
+saveToFile = "graphsize.png"
+if saveToFile != None:
+	plt.savefig(saveToFile)
+else:
+	plt.show()
+
+
+
+
+
+

### 2020-10-14 plots/makeplots.py

@@ -4,19 +4,19 @@ from plotFunctions import *
 
 
 
-distributed = False
+distributed = True
 singleNode = False
 
 sssp = False
 bfs = False
-pr = False
+pr = True
 
 calcTimes = False
 execTimes = False
 overheads = False
 
 
-galoisSpeedup = True
+galoisSpeedup = False
 
 
 colors = {"Galois":"#1f77b4",
@@ -141,7 +141,7 @@ if distributed and ( sssp or bfs or pr or calcTimes or execTimes or overheads):
 
 	#PR
 	if pr or calcTimes:
-		grouped_bar_plot(graphs, list(dist_frameworks_pr.values()), execTimePR_distributed, yErrs=yErrExecPR_distributed, yLabel='Calculation time (s)', saveToFile="distributedPR_execTime.png", colors_=colors)
+		grouped_bar_plot(graphs, list(dist_frameworks_pr.values()), execTimePR_distributed, yErrs=yErrExecPR_distributed, yLabel='Execution time (s)', saveToFile="distributedPR_execTime.png", colors_=colors)
 
 	## OVERHEAD
 	#SSSP

### 2020-10-14 documentation/presentation/main.tex
:-)

@@
 	\item Research vs. Production Case
-	\item Results for those Cases
+	\item Results
 	\item Impact of Hugepages on Galois
@@
 \begin{multicols}{2}
-\noindent
+	\subsection*{Graphs}
 	A \emph{weighted, directed graph} is the tuple $G=(V,E,w)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set} with
@@
 	\end{center}
-
 	\columnbreak
-
 	\subsection*{Algorithms}
@@
 \begin{multicols}{2}
-\noindent Machines: vsflash1-5, 
+\subsection*{Machines}
+vsflash1-5, 
 \begin{itemize}
@@
 \end{itemize}
-\noindent Measurements:
+\subsection*{Measurements}
 \begin{itemize}
@@
 
+\subsection*{Graphs}
+Both rMat graphs are synthetic, others are real-world data sets
 \begin{center}

### 2020-10-14 documentation/presentation/main.tex
hugepages und conclusion

@@
 
-\section{Galois With Hugepages}
-\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
+\newgeometry{paperwidth=28.1cm, paperheight=18cm, left=1cm, right=1cm, top=0.7cm,bottom=1cm,includefoot,heightrounded}
+\section{Galois With Hugepages on SSSP}
+\begin{figure}[h!]
+\hfil
+\begin{subfigure}{0.33\linewidth}
+	\sffamily
+	\small
+	\centering
+	\begin{tabular}{cr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
+		&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
+		\cmidrule{2-3}\cmidrule{4-5}
+		\bf Graph&w/o&w/&w/o&w/\\
+		\midrule
+		flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
+		orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5\\
+		wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1\\
+		twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1\\
+		rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4\\
+		friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5\\
+		rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1\\
+\end{tabular}
+\end{subfigure}
+\hfil
+\begin{subfigure}{0.3\linewidth}
+\centering
+\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
+	\caption{Without Hugepages}
+\end{subfigure}
+\begin{subfigure}{0.3\linewidth}
+\centering
+\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+	\caption{With Hugepages}
+\end{subfigure}
+\end{figure}
+\begin{itemize}
+	\item Hugepages further reduce both calculation and execution time
+	\item Speedups with hugepages are \enquote{more reliable}, i.e., more predictable
+\end{itemize}
 
-		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
-		\cmidrule{3-4}\cmidrule{5-6}
-		&\bf Graph&w/o&w/&w/o&w/\\
+\section{Galois With Hugepages on PR Pull}
+\begin{figure}[h!]
+\hfil
+\begin{subfigure}{0.33\linewidth}
+	\sffamily
+	\small
+	\centering
+	\begin{tabular}{cr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}rc}
+		&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
+		\cmidrule{2-3}\cmidrule{4-5}
+		\bf Graph&w/o&w/&w/o&w/\\
 		\midrule
-\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
-		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5\\
-		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1\\
-		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1\\
-		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4\\
-		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5\\
-		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1\\
+		flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
+		orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6\\
+		wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4\\
+		twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3\\
+		rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1\\
+		friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1\\
+		rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4\\
 \end{tabular}
-\includegraphics[width=8cm]{../../plots/singleNodeSSSPGaloisThreads.png}
-\includegraphics[width=8cm]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+\end{subfigure}
+\hfil
+	\begin{subfigure}{0.3\linewidth}
+	\centering
+\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
+		\caption{Without Hugepages}
+\end{subfigure}
+\begin{subfigure}{0.3\linewidth}
+	\centering
+\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
+		\caption{With Hugepages}
+	\end{subfigure}
+\end{figure}
+\begin{itemize}
+	\item Hugepages further reduce both calculation and execution time
+	\item Hugepages make speedup in multithreaded scenario possible
+	\item Speedup not to the same degree as with SSSP (15$\times$ vs. 2.5$\times$)
+\end{itemize}
 
-\section{Galois With Hugepages}
-\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}rc}
-		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
-		\cmidrule{3-4}\cmidrule{5-6}
-		&\bf Graph&w/o&w/&w/o&w/\\
-		\midrule
-	\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
-		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6\\
-		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4\\
-		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3\\
-		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1\\
-		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1\\
-		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4\\
-\end{tabular}
-\includegraphics[width=0.5\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
-\includegraphics[width=0.5\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
+\restoregeometry
 
 \section{Conclusion and Outlook}
+Generally: 1) performance highly dependent on the framework, algorithm and data set
+
+2) single node almost always preferrable, as long as RAM is sufficient
+\begin{multicols}{2}
+	\subsection{Production Case}
 	\begin{itemize}
-	\item performance highly dependent on the framework, algorithm and data set
-	\begin{itemize}
-		\item Galois is almost always fastest in the research case; especially with hugepages
-		\item Giraph is good on SSSP or BFS in distributed production
-		\item Gemini is a good middleground for distributed PR and single node production
+		\item Giraph is very fast on distributed systems (especially SSSP and BFS), with added benefit of Hadoop (fault-tolerance, automatic parallelization) 
+		\item Gemini (surprisingly) and Ligra are good options for single node 
 	\end{itemize}
 
-	\item single node almost always preferrable, as long as RAM is sufficient
+	\columnbreak
+	\subsection{Research Case}
+	\begin{itemize}
+		\item Galois is fastest in almost all cases; further improvements with hugepages possible
 	\end{itemize}
+\end{multicols}
 
@@
 	\item[$\rightarrow$] incorporate new frameworks and new algorithms
-	\item[$\rightarrow$] great range of settings and multiple implementations for the same problem
-	\item[$\rightarrow$] At a later point in time, it is important to repeat such a comparison, because the frameworks are further developed and new ones are created.
+	\item[$\rightarrow$] explore range of settings and other implementations
+	\item[$\rightarrow$] repeat similar tests in the future: frameworks are improved and new ones are introduced
 \end{itemize}

### 2020-10-16 documentation/presentation/background.tex
dinge

--- 2020-10-14 documentation/presentation/main.tex
@@
 \usepackage{subcaption}
-
 \usepackage{tikz}
@@
 
-\setlength{\parskip}{0pt}
-\setlength{\parindent}{0pt}
-\setlength{\columnsep}{1cm}
-\raggedcolumns
+\date{}
 
-\author{S. König, L. Matzner, F. Rollbühler and J. Schmid}
-\date{Tuesday, 20\textsuperscript{th} October 2020}
+\usepackage[pages=all]{background}
+
+\definecolor{left} {HTML}{001528}
+\backgroundsetup{
+scale=1,
+color=black,
+opacity=0.3,
+angle=0,
+contents={%
+  \begin{tikzpicture}
+  	\node [shading = axis,rectangle, left color=left, right color=white,shading angle=135, anchor=north, minimum width=\paperwidth, minimum height=\paperheight] (box) at (current page.north){};
+  \end{tikzpicture}
+  }%
+}
 
 \begin{document}
-% Ich weiß, dass es die Beamer-Klasse gibt.
-% Lasst mich doch alle in Ruhe!
-% Die hier hatte ich noch rumliegen, das muss jetzt reichen.
-\section{Introduction}
-\begin{center}
-	\begin{minipage}{0.6\linewidth}
-		\centering
-		Main Goal: \itshape  Comparison of five graph processing systems in their performance on different graphs and algorithms.
-	\end{minipage}
-\end{center}
-
-\begin{enumerate}
-	\item Preliminaries
-	\begin{itemize}
-	\item Basics
-	\item Computation Styles
-	\item Hugepages
-	\end{itemize}
-	\item Frameworks
-	\item Evaluation
-	\begin{itemize}
-	\item Research vs. Production Case
-	\item Results for those Cases
-	\item Impact of Hugepages on Galois
-	\end{itemize}
-	\item Conclusion
-\end{enumerate}
-
-% In the past years several graph processing systems emerged.
-% Graphs are growing fast and are becoming increasingly popular. Many problems can be modeled and solved using graphs.
-
-\clearpage
-\section{Preliminaries}
-\begin{multicols}{2}
-\noindent
-	A \emph{weighted, directed graph} is the tuple $G=(V,E,w)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set} with
-	\begin{equation*}
-	  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
-	\end{equation*}
-	and $w:E\rightarrow \mathbb R$ is a mapping of edge to a weight.
-
-	\begin{center}
-	\begin{tikzpicture}
-		\node[circle, draw] (a) at (1.7,0) {A};
-		\node[circle, draw] (b) at (0,2) {B};
-		\node[circle, draw] (c) at (4,0) {C};
-		\node[circle, draw] (d) at (0,4) {D};
-		\node[circle, draw] (e) at (2.5,2) {E};
-
-		\draw[->, bend right, line width=1pt] (a) edge (b);
-		\draw[->, bend right, line width=1pt] (a) edge (c);
-
-		\draw[->, bend right, line width=1pt] (b) edge (a);
-		\draw[->, line width=1pt] (b) edge (d);
-		\draw[->, line width=1pt] (b) edge (e);
-
-		\draw[->, line width=1pt] (c) edge (e);
-
-		\draw[->, line width=1pt] (e) edge (a);
-
-		\node at (2.5,3) {$w(x)=1\ \forall x\in V$};
-	\end{tikzpicture}
-	\end{center}
-
-	\columnbreak
-
-	\subsection*{Algorithms}
-	\begin{description}
-		\item[Single-Source Shortest-Paths (SSSP):] find the shortest path from a starting vertex to every other vertex
-
-		\item[Breadth-first search (BFS):] find a node outgoing from a starting vertex, by increasing maximum hop count step-wise
-
-		\item[PageRank (PR):] link analysis algorithm; weighs vertices, measuring their relative importance
-	\end{description}
-\end{multicols}
-
-\clearpage
-\begin{multicols}{2}
-\section{Push Style}
-\begin{center}
-\begin{tikzpicture}
-	\node[circle, draw] (a) at (1.5,1.2) {\bf A};
-	\node[circle, draw] (b) at (0,0) {B};
-	\node[circle, draw] (c) at (3,0) {C};
-	\node[circle, draw] (d) at (1.5,3) {D};
-
-	\draw[->, line width=1pt] (a) edge node[midway, above] {1} (b);
-	\draw[->, line width=1pt] (a) edge node[midway, above] {1} (c);
-	\draw[->, line width=1pt] (a) edge node[midway, right] {1} (d);
-\end{tikzpicture}
-\end{center}
-\begin{itemize}
-	\item reads active vertex, writes neighborhood 
-	\item more efficient, if only few active vertices at the same time
-	\item more efficient, if neighborhoods do not overlap
-\end{itemize}
-
-\columnbreak
-\section{Pull Style}
-\begin{center}
-\begin{tikzpicture}
-	\node[circle, draw] (a) at (1.5,1.2) {\bf A};
-	\node[circle, draw] (b) at (0,0) {B};
-	\node[circle, draw] (c) at (3,0) {C};
-	\node[circle, draw] (d) at (1.5,3) {D};
-
-	\draw[->, line width=1pt] (b) edge node[midway, above] {3} (a);
-	\draw[->, line width=1pt] (c) edge node[midway, above] {2} (a);
-	\draw[->, line width=1pt] (d) edge node[midway, right] {6} (a);
-\end{tikzpicture}
-\end{center}
-\begin{itemize}
-	\item reads neighborhood, writes active vertex
-	\item[$\rightarrow$] only one write and many read operations
-	\item less synchronization in parallel implementations needed
-	\item more efficient, if many vertices active at the same time
-\end{itemize}
-\end{multicols}
-
-\section{Hugepages}
-\begin{itemize}
-	\item most systems use virtual memory management
-\begin{itemize}
-	\item represents an abstraction to hardware memory
-	\item virtual memory is then organized in pages
-	\item translations of virtual memory to physical memory are cached, because every translation takes time
-\end{itemize}
-	\item typically, memory pages are 4 KiB in size
-	\item \textbf{hugepages} can be several MiB in size $\rightarrow$ reduce number of cache misses
-	\item especially noticeable in very memory intensive applications
-\end{itemize}
-
-\section{Frameworks}
-
-\definecolor{Galois}{HTML}{1f77b4}
-\definecolor{Gemini}{HTML}{ff7f0e}
-\definecolor{Giraph}{HTML}{2ca02c}
-\definecolor{Ligra}{HTML}{d62728}
-\definecolor{Polymer}{HTML}{9467bd}
-
-\begin{multicols}{2}
-\begin{description}
-	\item[$\color{Galois}\blacksquare$ Galois] is a general purpose library designed for parallel programming
-	\begin{itemize}
-		\item Version 6.0 from 29\textsuperscript{th} June 2020 used 
-	\end{itemize}
-	\item[$\color{Gemini}\blacksquare$ Gemini] uses a distributed message-based approach from scratch
-	\begin{itemize}
-		\item Version from 2\textsuperscript{nd} November 2016 used
-		\item Version contains bugs that had to be fixed 
-	\end{itemize}
-
-	\columnbreak
-
-	\item[$\color{Giraph}\blacksquare$ Giraph] is built on Apache Hadoop, a large scale data processing infrastructure
-	\begin{itemize}
-		\item Version 1.3 from 8\textsuperscript{th} May 2020 used
-		\item BFS is not natively supported 
-		\item state-of-the-art, but not NUMA-aware
-	\end{itemize}
-	\item[$\color{Ligra}\blacksquare$ Ligra] dynamically switches between push and pull style
-	\begin{itemize}
-		\item Version from 14\textsuperscript{th} August 2019
-	\end{itemize}
-	\item[$\color{Polymer}\blacksquare$ Polymer] optimizes data layout and memory access strategies
-	\begin{itemize}
-		\item Version from 28\textsuperscript{th} August 2018
-	\end{itemize}
-\end{description}
-\end{multicols}
-
-\section{Evaluation}
-\begin{multicols}{2}
-\noindent Machines: vsflash1-5, 
-\begin{itemize}
-	\item 96 cores, of which 48 virtual
-	\item 256 GB of RAM each\footnote{\sffamily one machine only 128 GB}
-	\item Ubuntu 18.04.2 LTS
-\end{itemize}
-\noindent Measurements:
-\begin{itemize}
-	\item \textbf{execution time}: time from start to finish of the console command
-	\item \textbf{calculation time}: time the framework actually executed the algorithm
-	%\item \textbf{overhead}: time difference between execution time and calculation time \newline (time to read the input graph, initialization, etc.) -- brauchen wir das überhaupt?
-	\item executed each test case 10 times
-\end{itemize}
-
-\columnbreak
-
-\begin{center}
-	\includegraphics[width=\linewidth]{graphsize.png}
-\end{center}
-
-\end{multicols}
-
-\clearpage
-\begin{multicols}{2}
-\section{Production Case}
-\begin{itemize}
-	\item running system: multiple calculations on a single graph
-	\item graph data stays loaded between calculations
-	\item[$\rightarrow$] short calculation times should be preferred
-\end{itemize}
-
-\columnbreak
-\section{Research Case}
-\begin{itemize}
-	\item individual calculation cases: possibly new graph for each calculation
-	\item frequently changing algorithm
-	\item[$\rightarrow$] framework should be relatively fast on different algorithms
-	\item[$\rightarrow$] overall small execution times should be preferred
-\end{itemize}
-\end{multicols}
-
-\section{Production Case Single Node}
-
-\begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
-		\caption{SSSP}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
-		\caption{BFS}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
-		\caption{PR}
-	\end{subfigure}
-\end{figure}
-\begin{itemize}
-	\item Giraph is either very slow or requires too much RAM (>256 GB)
-	\item On SSSP, Polymer is fastest, followed by Gemini on second place
-	\item On BFS, Gemini and Ligra are comparable and fastest on the larger graphs
-	\item On PR, Galois is fastest. But we exclude Galois Push because of possible measuring errors.
-	\item Message-based approach can compete with shared-memory
-\end{itemize}
-
-\section{Production Case Distributed}
-
-\begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
-		\caption{SSSP}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
-		\caption{BFS}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
-		\caption{PR}
-	\end{subfigure}
-\end{figure}
-\begin{itemize}
-	\item Giraph is fastest on SSSP and BFS on the real world graphs
-	\item Giraph has problems with synthetic graphs
-	\item Gemini is fastest on PR, with Giraph on second place
-\end{itemize}
-
-\section{Research Case Single Node}
-
-\begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
-		\caption{SSSP}
-		\label{fig:singleNodeSSSP_exec}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
-		\caption{BFS}
-		\label{fig:singleNodeSSSP_exec}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
-		\caption{PR}
-		\label{fig:singleNodeSSSP_exec}
-	\end{subfigure}
-\end{figure}
-\begin{itemize}
-	\item Giraph is either slowest or requires too much RAM (>256 GB)
-	\item Galois is fastest in almost all cases, second fastest is Ligra
-	\item Gemini and Polymer are comparably slow
-\end{itemize}
-
-\section{Research Case Distributed}
-
-\begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{SSSP}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
-		\caption{BFS}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
-		\caption{PR}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-\end{figure}
-\begin{itemize}
-	\item Galois Push is faster than Pull in all cases
-	\item Either Galois implementation is faster than any other frameworks on SSSP or BFS
-	\item Gemini is fastest on PR and comparable to Giraph on SSSP and BFS
-\end{itemize}
-
-\section{Galois With Hugepages}
-\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}r}
-
-		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
-		\cmidrule{3-4}\cmidrule{5-6}
-		&\bf Graph&w/o&w/&w/o&w/\\
-		\midrule
-\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
-		& orkut & 0.10 & \bf 0.02 & 0.8 & \bf 0.5\\
-		& wikipedia & 0.38 & \bf 0.11 & 1.8 & \bf 1.1\\
-		& twitter & 2.47 & \bf 0.94 & 10.8 & \bf 5.1\\
-		& rMat27 & 4.50 & \bf 1.39 & 16.0 & \bf 6.4\\
-		& friendster & 4.70 & \bf 1.78 & 14.4 & \bf 7.5\\
-		& rMat28 & 9.77 & \bf 3.34 & 27.8 & \bf 13.1\\
-\end{tabular}
-\includegraphics[width=8cm]{../../plots/singleNodeSSSPGaloisThreads.png}
-\includegraphics[width=8cm]{../../plots/singleNodeSSSPGaloisHPThreads.png}
-
-\section{Galois With Hugepages}
-\begin{tabular}{ccr@{\tabskip 2 \tabcolsep}rr@{\tabskip 2 \tabcolsep}rc}
-		&&\multicolumn{2}{c}{\bf Calc Time (s)}&\multicolumn{2}{c}{\bf Exec Time (s)}\\
-		\cmidrule{3-4}\cmidrule{5-6}
-		&\bf Graph&w/o&w/&w/o&w/\\
-		\midrule
-	\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR Pull}}
-		& flickr & 0.01 & \bf 0.01 & 0.3 & \bf 0.2\\
-		& orkut & 0.06 & \bf 0.02 & 0.7 & \bf 0.6\\
-		& wikipedia & 0.17 & \bf 0.03 & 1.7 & \bf 1.4\\
-		& twitter & 0.77 & \bf 0.11 & \bf 8.7 & 9.3\\
-		& rMat27 & 0.65 & \bf 0.13 & 19.2 & \bf 8.1\\
-		& friendster & 1.01 & \bf 0.14 & 20.4 & \bf 13.1\\
-		& rMat28 & 1.15 & \bf 0.24 & 46.0 & \bf 16.4\\
-\end{tabular}
-\includegraphics[width=0.5\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
-\includegraphics[width=0.5\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
-
-\section{Conclusion and Outlook}
-\begin{itemize}
-	\item performance highly dependent on the framework, algorithm and data set
-	\begin{itemize}
-		\item Galois is almost always fastest in the research case; especially with hugepages
-		\item Giraph is good on SSSP or BFS in distributed production
-		\item Gemini is a good middleground for distributed PR and single node production
-	\end{itemize}
-
-	\item single node almost always preferrable, as long as RAM is sufficient
-\end{itemize}
-
-\subsection{Outlook}
-\begin{itemize}
-	\item[$\rightarrow$] incorporate new frameworks and new algorithms
-	\item[$\rightarrow$] great range of settings and multiple implementations for the same problem
-	\item[$\rightarrow$] At a later point in time, it is important to repeat such a comparison, because the frameworks are further developed and new ones are created.
-\end{itemize}
-
+	\ 
 \end{document}

### 2020-10-16 documentation/presentation/main.tex
dinge

@@
 \usepackage{subcaption}
-
 \usepackage{tikz}
 \usetikzlibrary{positioning}
+
+\definecolor{heading}{HTML}{002d57}
+
+\usepackage[pages=all]{background}
+
+\backgroundsetup{
+scale=1,
+color=black,
+opacity=0.4,
+angle=0,
+contents={%
+	\includegraphics{background.pdf}
+  }%
+}
 
@@
 
+\newcommand{\todo}[1]{{\color{red}#1}}
+
 \begin{document}
@@
 % Die hier hatte ich noch rumliegen, das muss jetzt reichen.
-\section{Introduction}
+\section{Motivation and Goal}
+\sffamily
 \begin{center}
+
+	\vfill
+
+	\begin{itemize}
+		\item graph processing becomes increasingly important in academic and industrial environments
+		\item many problems modeled with graphs, e.\,g.,\xspace machine learning and data mining
+		\item many business models are based on graphs, e.\,g.,\xspace viral marketing or Google’s search engine
+		\item graph sizes increase to several billion edges 
+		\item[$\rightarrow$] performance, parallelism and distribution of graph algorithms becomes more important
+	\end{itemize}
+
+	\vfill
+
 	\begin{minipage}{0.6\linewidth}
@@
 	\end{minipage}
+
+	\vfill
 \end{center}
 
+\section{Overview}
+\vfill
 \begin{enumerate}
@@
 \end{enumerate}
+\vfill
 
@@
 
-		\node at (2.5,3) {$w(x)=1\ \forall x\in V$};
+		\node at (2.5,3) {$w(x)=1\ \forall x\in E$};
 	\end{tikzpicture}
@@
 	\item more efficient, if only few active vertices at the same time
-	\item more efficient, if neighborhoods do not overlap
+	\item more efficient, if neighborhoods of active vertices do not overlap
 \end{itemize}
@@
 \section{Hugepages}
+\vspace{2cm}
 \begin{itemize}
@@
 	\item represents an abstraction to hardware memory
-	\item virtual memory is then organized in pages
+	\item virtual memory is organized in pages
 	\item translations of virtual memory to physical memory are cached, because every translation takes time
@@
 
-\section{Frameworks}
+% \section{Frameworks}
 
@@
 
-\begin{multicols}{2}
-\begin{description}
-	\item[$\color{Galois}\blacksquare$ Galois] is a general purpose library designed for parallel programming
-	\begin{itemize}
-		\item Version 6.0 from 29\textsuperscript{th} June 2020 used 
-	\end{itemize}
-	\item[$\color{Gemini}\blacksquare$ Gemini] uses a distributed message-based approach from scratch
-	\begin{itemize}
-		\item Version from 2\textsuperscript{nd} November 2016 used
-		\item Version contains bugs that had to be fixed 
-	\end{itemize}
+% \begin{multicols}{2}
+% \begin{description}
+% 	\item[$\color{Galois}\blacksquare$ Galois] is a general purpose library designed for parallel programming
+% 	\begin{itemize}
+% 		\item Version 6.0 from 29\textsuperscript{th} June 2020 used 
+% 	\end{itemize}
+% 	\item[$\color{Gemini}\blacksquare$ Gemini] uses a distributed message-based approach from scratch
+% 	\begin{itemize}
+% 		\item Version from 2\textsuperscript{nd} November 2016 used
+% 		\item Version contains bugs that had to be fixed 
+% 	\end{itemize}
 
-	\columnbreak
+% 	\columnbreak
 
-	\item[$\color{Giraph}\blacksquare$ Giraph] is built on Apache Hadoop, a large scale data processing infrastructure
-	\begin{itemize}
-		\item Version 1.3 from 8\textsuperscript{th} May 2020 used
-		\item BFS is not natively supported 
-		\item state-of-the-art, but not NUMA-aware
-	\end{itemize}
-	\item[$\color{Ligra}\blacksquare$ Ligra] dynamically switches between push and pull style
-	\begin{itemize}
-		\item Version from 14\textsuperscript{th} August 2019
-	\end{itemize}
-	\item[$\color{Polymer}\blacksquare$ Polymer] optimizes data layout and memory access strategies
-	\begin{itemize}
-		\item Version from 28\textsuperscript{th} August 2018
-	\end{itemize}
-\end{description}
-\end{multicols}
+% 	\item[$\color{Giraph}\blacksquare$ Giraph] is built on Apache Hadoop, a large scale data processing infrastructure
+% 	\begin{itemize}
+% 		\item Version 1.3 from 8\textsuperscript{th} May 2020 used
+% 		\item BFS is not natively supported 
+% 		\item state-of-the-art, but not NUMA-aware
+% 	\end{itemize}
+% 	\item[$\color{Ligra}\blacksquare$ Ligra] dynamically switches between push and pull style
+% 	\begin{itemize}
+% 		\item Version from 14\textsuperscript{th} August 2019
+% 	\end{itemize}
+% 	\item[$\color{Polymer}\blacksquare$ Polymer] optimizes data layout and memory access strategies
+% 	\begin{itemize}
+% 		\item Version from 28\textsuperscript{th} August 2018
+% 	\end{itemize}
+% \end{description}
+% \end{multicols}
+
+% Features, version, numa, message passing/sharememory, bfs fehlt?
+
+\newgeometry{paperwidth=28.1cm, paperheight=18cm, left=2cm, right=2cm, top=0.7cm,bottom=1cm,includefoot,heightrounded}
+\section{Frameworks}
+\renewcommand{\arraystretch}{1.2}
+\begin{tabular}{lcccp{8cm}p{6cm}}
+	Framework&Version&NUMA&Dist.&\multicolumn{1}{c}{Features}&\multicolumn{1}{c}{Notes}\\
+	\toprule
+	$\color{Galois}\blacksquare$\ \bf Galois&29.06.2020&\checkmark&(\checkmark)&general purpose library designed for parallel programming&Distributed using Gluon\\\midrule
+	$\color{Gemini}\blacksquare$\ \bf Gemini&02.11.2016&\checkmark&\checkmark&distributed message-based approach from scratch&Version contains bugs that had to be fixed \\\midrule
+	$\color{Giraph}\blacksquare$\ \bf Giraph&08.05.2020&X&\checkmark&built on Apache Hadoop&BFS is not natively supported \\\midrule
+	$\color{Ligra}\blacksquare$\ \bf Ligra&14.08.2019&\checkmark&X&dynamically switches between push and pull style&\\\midrule
+	$\color{Polymer}\blacksquare$\ \bf Polymer&28.08.2018&\checkmark&X&optimizes data layout and memory access strategies&\\
+\end{tabular}
+
+\restoregeometry
 
@@
 \subsection*{Graphs}
-Both rMat graphs are synthetic, others are real-world data sets
+Both rMat graphs are synthetic, others are real-world data sets; Flickr: 24MB, rMat28: 76GB%
 \begin{center}
@@
 \end{center}
-
 \end{multicols}
 
-\clearpage
 \begin{multicols}{2}
 \section{Production Case}
+\vspace{1cm}
 \begin{itemize}
@@
 	\item[$\rightarrow$] short calculation times should be preferred
+	\item Not main focus of this presentation!\footnote{\sffamily see paper for details}
 \end{itemize}
@@
 \section{Research Case}
+\vspace{1cm}
 \begin{itemize}
@@
 
-\section{Production Case Single Node}
+% \section{Production Case Single Node}
 
-\begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
-		\caption{SSSP}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
-		\caption{BFS}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
-		\caption{PR}
-	\end{subfigure}
-\end{figure}
-\begin{itemize}
-	\item Giraph is either very slow or requires too much RAM (>256 GB)
-	\item On SSSP, Polymer is fastest, followed by Gemini on second place
-	\item On BFS, Gemini and Ligra are comparable and fastest on the larger graphs
-	\item On PR, Galois is fastest. But we exclude Galois Push because of possible measuring errors.
-	\item Message-based approach can compete with shared-memory
-\end{itemize}
+% \begin{figure}[h]
+% 	\begin{subfigure}{0.32\textwidth}
+% 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
+% 		\caption{SSSP}
+% 	\end{subfigure}
+% 	\begin{subfigure}{0.32\textwidth}
+% 		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
+% 		\caption{BFS}
+% 	\end{subfigure}
+% 	\begin{subfigure}{0.32\textwidth}
+% 		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
+% 		\caption{PR}
+% 	\end{subfigure}
+% \end{figure}
+% \begin{itemize}
+% 	\item Giraph is either very slow or requires too much RAM (>256 GB)
+% 	\item On SSSP, Polymer is fastest, followed by Gemini on second place
+% 	\item On BFS, Gemini and Ligra are comparable and fastest on the larger graphs
+% 	\item On PR, Galois Pull is fastest, we exclude Galois Push because of possible measuring errors.
+% 	\item \todo{Message-based approach can compete with shared-memory}
+% \end{itemize}
+% \todo{Grafiken Größer?}
 
@@
 \begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
-		\caption{BFS}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+	\hfil
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
@@
 \begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
-		\caption{BFS}
-		\label{fig:singleNodeSSSP_exec}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+	\hfil
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
@@
 \begin{figure}[h]
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
-		\caption{BFS}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+	\hfil
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
@@
 	\item Galois Push is faster than Pull in all cases
-	\item Either Galois implementation is faster than any other frameworks on SSSP or BFS
-	\item Gemini is fastest on PR and comparable to Giraph on SSSP and BFS
+	\item Both Galois implementations fastest on SSSP or BFS
+	\item Gemini is fastest on PR in almost all cases
 \end{itemize}
 
-\newgeometry{paperwidth=28.1cm, paperheight=18cm, left=1cm, right=1cm, top=0.7cm,bottom=1cm,includefoot,heightrounded}
-\section{Galois With Hugepages on SSSP}
+%\newgeometry{paperwidth=28.1cm, paperheight=18cm, left=1cm, right=1cm, top=0.7cm,bottom=1cm,includefoot,heightrounded}
+\section{Galois With Hugepages}
 \begin{figure}[h!]
-\hfil
-\begin{subfigure}{0.33\linewidth}
+\begin{subfigure}{0.4\textwidth}
 	\sffamily
@@
 \end{tabular}
+\caption{SSSP}
 \end{subfigure}
 \hfil
-\begin{subfigure}{0.3\linewidth}
-\centering
-\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
-	\caption{Without Hugepages}
-\end{subfigure}
-\begin{subfigure}{0.3\linewidth}
-\centering
-\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
-	\caption{With Hugepages}
-\end{subfigure}
-\end{figure}
-\begin{itemize}
-	\item Hugepages further reduce both calculation and execution time
-	\item Speedups with hugepages are \enquote{more reliable}, i.e., more predictable
-\end{itemize}
-
-\section{Galois With Hugepages on PR Pull}
-\begin{figure}[h!]
-\hfil
-\begin{subfigure}{0.33\linewidth}
+\begin{subfigure}{0.4\textwidth}
 \sffamily
@@
 \end{tabular}
+\caption{PR Pull}
+\end{subfigure}
+\end{figure}
+\vfill
+\begin{itemize}
+	\item Hugepages reduce both calculation and execution time on all algorithms
+	\item[$\rightarrow$] Execution times can be up to 3$\times$ shorter
+\end{itemize}
+\vfill
+
+\section{Multithreaded Speedup of Galois}
+\begin{figure}[h!]
+\begin{subfigure}{0.45\textwidth}
+	\centering
+\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+\caption{SSSP with Hugepages}
 \end{subfigure}
 \hfil
-	\begin{subfigure}{0.3\linewidth}
-	\centering
-\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
-		\caption{Without Hugepages}
-\end{subfigure}
-\begin{subfigure}{0.3\linewidth}
+\begin{subfigure}{0.45\textwidth}
 	\centering
 \includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
-		\caption{With Hugepages}
+		\caption{PR Pull with Hugepages}
 	\end{subfigure}
@@
 \begin{itemize}
-	\item Hugepages further reduce both calculation and execution time
-	\item Hugepages make speedup in multithreaded scenario possible
-	\item Speedup not to the same degree as with SSSP (15$\times$ vs. 2.5$\times$)
+	\item Speedups can be significant, with and without hugepages
+	\item Speedup of PR not to the same degree as on SSSP (2.5$\times$ vs. 15$\times$)
 \end{itemize}
@@
 	\begin{itemize}
-		\item Giraph is very fast on distributed systems (especially SSSP and BFS), with added benefit of Hadoop (fault-tolerance, automatic parallelization) 
-		\item Gemini (surprisingly) and Ligra are good options for single node 
+		\item Giraph is very fast on distributed systems (especially SSSP and BFS)
+		\item Gemini and Ligra are good options for single node 
 	\end{itemize}
@@
 	\item[$\rightarrow$] explore range of settings and other implementations
-	\item[$\rightarrow$] repeat similar tests in the future: frameworks are improved and new ones are introduced
+	\item[$\rightarrow$] repeat similar tests in the future: frameworks are updated and new ones are introduced
 \end{itemize}

### 2020-10-16 documentation/presentation/bla.py

@@ -26,7 +26,7 @@ ax.set_ylabel("# of edges (Million)")
 
 saveToFile = "graphsize.png"
 if saveToFile != None:
-	plt.savefig(saveToFile)
+	plt.savefig(saveToFile, transparent=True)
 else:
 	plt.show()
 

### 2020-10-16 plots/makeplots.py

@@ -4,19 +4,19 @@ from plotFunctions import *
 
 
 
-distributed = True
+distributed = False
 singleNode = False
 
 sssp = False
 bfs = False
-pr = True
+pr = False
 
 calcTimes = False
 execTimes = False
 overheads = False
 
 
-galoisSpeedup = False
+galoisSpeedup = True
 
 
 colors = {"Galois":"#1f77b4",

### 2020-10-16 plots/plotFunctions.py

@@ -47,7 +47,7 @@ def grouped_bar_plot(group_names, legend_names, values, colors_=None, yErrs=None
 	ax.legend(ps, legend_names)
 
 	if saveToFile != None:
-		plt.savefig(saveToFile)
+		plt.savefig(saveToFile, transparent=True)
 	else:
 		plt.show()
 
@@ -83,7 +83,7 @@ def line_plot(legend_names, x, values, title='', yLabel='', xLabel='', yScale=No
 	ax.legend(ps, legend_names)
 
 	if saveToFile != None:
-		plt.savefig(saveToFile)
+		plt.savefig(saveToFile, transparent=True)
 	else:
 		plt.show()
 

### 2020-10-19 documentation/presentation/main.tex
dinge
Kleine Änderungen an Präsentation

@@
 \section{Overview}
-\vfill
+{\large
 \begin{enumerate}
@@
 	\item Results
-	\item Impact of Hugepages on Galois
 	\end{itemize}
-	\item Conclusion
+	\item Conclusion and Outlook
 \end{enumerate}
-\vfill
-
+}
 % In the past years several graph processing systems emerged.
@@
 
-% \section{Frameworks}
-
 \definecolor{Galois}{HTML}{1f77b4}
@@
 
-% \begin{multicols}{2}
-% \begin{description}
-% 	\item[$\color{Galois}\blacksquare$ Galois] is a general purpose library designed for parallel programming
-% 	\begin{itemize}
-% 		\item Version 6.0 from 29\textsuperscript{th} June 2020 used 
-% 	\end{itemize}
-% 	\item[$\color{Gemini}\blacksquare$ Gemini] uses a distributed message-based approach from scratch
-% 	\begin{itemize}
-% 		\item Version from 2\textsuperscript{nd} November 2016 used
-% 		\item Version contains bugs that had to be fixed 
-% 	\end{itemize}
-
-% 	\columnbreak
-
-% 	\item[$\color{Giraph}\blacksquare$ Giraph] is built on Apache Hadoop, a large scale data processing infrastructure
-% 	\begin{itemize}
-% 		\item Version 1.3 from 8\textsuperscript{th} May 2020 used
-% 		\item BFS is not natively supported 
-% 		\item state-of-the-art, but not NUMA-aware
-% 	\end{itemize}
-% 	\item[$\color{Ligra}\blacksquare$ Ligra] dynamically switches between push and pull style
-% 	\begin{itemize}
-% 		\item Version from 14\textsuperscript{th} August 2019
-% 	\end{itemize}
-% 	\item[$\color{Polymer}\blacksquare$ Polymer] optimizes data layout and memory access strategies
-% 	\begin{itemize}
-% 		\item Version from 28\textsuperscript{th} August 2018
-% 	\end{itemize}
-% \end{description}
-% \end{multicols}
-
-% Features, version, numa, message passing/sharememory, bfs fehlt?
-
 \newgeometry{paperwidth=28.1cm, paperheight=18cm, left=2cm, right=2cm, top=0.7cm,bottom=1cm,includefoot,heightrounded}
 \section{Frameworks}
+\vfill
 \renewcommand{\arraystretch}{1.2}
@@
 	\toprule
-	$\color{Galois}\blacksquare$\ \bf Galois&29.06.2020&\checkmark&(\checkmark)&general purpose library designed for parallel programming&Distributed using Gluon\\\midrule
-	$\color{Gemini}\blacksquare$\ \bf Gemini&02.11.2016&\checkmark&\checkmark&distributed message-based approach from scratch&Version contains bugs that had to be fixed \\\midrule
+	$\color{Galois}\blacksquare$\ \bf Galois&29.06.2020&\checkmark&(\checkmark)&general purpose library designed for parallel programming, Hugepage support&distributed using Gluon\\\midrule
+	$\color{Gemini}\blacksquare$\ \bf Gemini&02.11.2016&\checkmark&\checkmark&distributed message-based approach from scratch&version contains bugs that had to be fixed \\\midrule
 	$\color{Giraph}\blacksquare$\ \bf Giraph&08.05.2020&X&\checkmark&built on Apache Hadoop&BFS is not natively supported \\\midrule
@@
 \end{tabular}
+\vfill
 

### 2020-10-20 documentation/presentation/main.tex
conclusion
Seitenzahlen
Folienreihenfolge geändert
Frameworks Folie wieder normal breit

@@
 }
+
+\rfoot{\makebox[\textwidth][r]{\sffamily\thepage}}
 
@@
 
-\newgeometry{paperwidth=28.1cm, paperheight=18cm, left=2cm, right=2cm, top=0.7cm,bottom=1cm,includefoot,heightrounded}
+%\newgeometry{paperwidth=28.1cm, paperheight=18cm, left=2cm, right=2cm, top=0.7cm,bottom=1cm,includefoot,heightrounded}
 \section{Frameworks}
@@
 \renewcommand{\arraystretch}{1.2}
-\begin{tabular}{lcccp{8cm}p{6cm}}
+\begin{tabular}{lcccp{7cm}p{5cm}}
 	Framework&Version&NUMA&Dist.&\multicolumn{1}{c}{Features}&\multicolumn{1}{c}{Notes}\\
@@
 
+\section{Research Case Distributed}
+
+\begin{figure}[h]
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
+		\caption{SSSP}
+		\label{fig:distributedSSSP_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
+		\caption{PR}
+		\label{fig:distributedSSSP_exec}
+	\end{subfigure}
+\end{figure}
+\begin{itemize}
+	\item Galois Push is faster than Pull in all cases
+	\item Both Galois implementations fastest on SSSP or BFS
+	\item Gemini is fastest on PR in almost all cases
+\end{itemize}
+
 \section{Research Case Single Node}
@@
 	\item Gemini and Polymer are comparably slow
-\end{itemize}
-
-\section{Research Case Distributed}
-
-\begin{figure}[h]
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
-		\caption{SSSP}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-	\hfil
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
-		\caption{PR}
-		\label{fig:distributedSSSP_exec}
-	\end{subfigure}
-\end{figure}
-\begin{itemize}
-	\item Galois Push is faster than Pull in all cases
-	\item Both Galois implementations fastest on SSSP or BFS
-	\item Gemini is fastest on PR in almost all cases
 \end{itemize}
@@
 		\item Giraph is very fast on distributed systems (especially SSSP and BFS)
+		\item Gemini is fast for distributed PR
 		\item Gemini and Ligra are good options for single node

### 2020-10-20 documentation/presentation/main.tex
.

@@
 % Die hier hatte ich noch rumliegen, das muss jetzt reichen.
+\thispagestyle{empty}
+\ 
+\vfill
+
+\begin{center}
+	\huge\sffamily\color{heading} Comparison of Graph Processing Systems
+
+	\vfill
+
+	\normalsize Tuesday, 20\textsuperscript{th} October 2020
+\end{center}
+
+\vfill
+
 \section{Motivation and Goal}
@@
 
+\section{Additional Data}
+\clearpage
+\begin{figure*}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
+		\caption{Calculation times}
+		\label{fig:singleNodeSSSP_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
+		\caption{Execution times}
+		\label{fig:singleNodeSSSP_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTime.png}
+		\caption{Overhead time}
+		\label{fig:singleNodeSSSP_overhead}
+	\end{subfigure}
+	\caption{Average times for SSSP on a single computation node}
+	\label{fig:singleNodeSSSP}
+\end{figure*}%
+\begin{figure*}
+	\centering
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
+		\caption{Calculation times}
+		\label{fig:distributedSSSP_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
+		\caption{Execution times}
+		\label{fig:distributedSSSP_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTime.png}
+		\caption{Overhead times}
+		\label{fig:distributedSSSP_overhead}
+	\end{subfigure}
+	\caption{Average times for SSSP on the distributed cluster}
+	\label{fig:distributedSSSP}
+\end{figure*}
+
+\clearpage
+\begin{figure*}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
+		\caption{Calculation time}
+		\label{fig:singleNodeBFS_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
+		\caption{Execution time}
+		\label{fig:singleNodeBFS_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTime.png}
+		\caption{Overhead time}
+		\label{fig:singleNodeBFS_overheadNormalized}
+	\end{subfigure}
+	\caption{Average times for BFS on a single computation node}
+\end{figure*}
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
+		\caption{Calculation time}
+		\label{fig:distributedBFS_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
+		\caption{Execution time}
+		\label{fig:distributedBFS_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTime.png}
+		\caption{Overhead}
+		\label{fig:distributedBFS_overhead}
+	\end{subfigure}
+	\hfil
+	\caption{Average times for BFS on the distributed cluster}
+	\label{fig:distributedBFS}
+\end{figure*}
+
+\clearpage
+
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_calcTime.png}
+		\caption{Calculation time}
+		\label{fig:singleNodePR_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_execTime.png}
+		\caption{Execution time}
+		\label{fig:singleNodePR_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePR_overheadTIme.png}
+		\caption{Overhead}
+		\label{fig:singleNodePR_overhead}
+	\end{subfigure}
+	\hfil
+	\caption{Average times for PR on a single computation node}
+	\label{fig:singleNodePR}
+\end{figure*}
+\begin{figure*}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_calcTime.png}
+		\caption{Calculation time}
+		\label{fig:distributedPR_calc}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_execTime.png}
+		\caption{Execution time}
+		\label{fig:distributedPR_exec}
+	\end{subfigure}
+	\hfil
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[width=\linewidth]{../../plots/distributedPR_overheadTime.png}
+		\caption{Overhead}
+		\label{fig:distributedPR_overhead}
+	\end{subfigure}
+	\hfil
+	\caption{Average times for PR on the distributed cluster}
+	\label{fig:distributedPR}
+\end{figure*}
+
+\clearpage
+
+\begin{figure}
+	\hfil
+	\begin{subfigure}{0.48\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
+		\caption{without hugepages}
+		\label{fig:galoisSpeedupSSSP_noHP}
+	\end{subfigure}
+	\begin{subfigure}{0.48\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisHPThreads.png}
+		\caption{with hugepages}
+		\label{fig:galoisSpeedupSSSP_HP}
+	\end{subfigure}
+	\hfil
+	\caption{Calculation time speedups on SSSP}
+	\label{fig:galoisSpeedupSSSP}
+\end{figure}
+
+\begin{figure}
+	\hfil
+	\begin{subfigure}{0.48\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisThreads.png}
+		\caption{without hugepages}
+		\label{fig:galoisSpeedupPRPull_noHP}
+	\end{subfigure}
+	\begin{subfigure}{0.48\columnwidth}
+		\includegraphics[width=\linewidth]{../../plots/singleNodePRPullGaloisHPThreads.png}
+		\caption{with hugepages}
+		\label{fig:galoisSpeedupPRPull_HP}
+	\end{subfigure}
+	\hfil
+	\caption{Calculation time speedups on PR Pull}
+	\label{fig:galoisSpeedupPRPull}
+\end{figure}
 \end{document}

### 2020-10-20 plots/makeplots.py

@@ -4,11 +4,11 @@ from plotFunctions import *
 
 
 
-distributed = False
-singleNode = False
+distributed = True
+singleNode = True
 
 sssp = False
-bfs = False
+bfs = True
 pr = False
 
 calcTimes = False
@@ -16,7 +16,7 @@ execTimes = False
 overheads = False
 
 
-galoisSpeedup = True
+galoisSpeedup = False
 
 
 colors = {"Galois":"#1f77b4",
